### 2009-02-05 JSS/Jss V16 Jan 2008.tex
Draft de Draft JSS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@37 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%% LyX 1.6.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[article]{jss}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[authoryear]{natbib}

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
%the following commands are used only for articles and codesnippets

\author{Régis Pouillot\\CFSAN \And 
        Marie Laure Delignette-Muller\\Université de Lyon \AND
        Jean-Baptiste Denis\\INRA}
\title{\pkg{fitdistrplus} and \pkg{mc2d}, Two Packages for Risk Assessment in \proglang{R}}

% the same as above, without any formatting
\Plainauthor{Régis Pouillot, Marie Laure Delignette Muller, Jean-Baptiste Denis}
\Plaintitle{fitdistrplus and mc2d, two packages for risk assessment in R} 
%if necessary, provide a short title
\Shorttitle{Package fitdistrplus and mc2d in R}

\Abstract{This is the abstract of the article.}
%at least one keyword is needed
\Keywords{keyword 1, keyword 2, \pkg{fitdistrplus}, \pkg{mc2d}, \proglang{R}}
%the same as above, without any formatting
\Plainkeywords{keyword 1, keyword 2, foo, R}

%The address of at least one author should be given in the following format
\Address{
  Régis Pouillot\\
  Center for Food Safety and Applied Nutrition\\
  FDA/HHS\\
  College Park, USA\\
  E-mail: \email{rpouillot@yahoo.fr}\\
  URL: \url{http://www.lyx.org}
}

%% need no \usepackage{Sweave.sty}

%you can add a telephone and fax number before the e-mail in the format
%Telephone: +12/3/4567-89
%Fax: +12/3/4567-89

%if you use Sweave,  include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

\makeatother

\begin{document}

\section[Introduction]{Introduction}

Ceci est un essai de bibliographie \citealt{CULLEN-FREY-1999}.

\subsection{Variability and uncertainty in the risk assessment framework }

According to international recommendations, a QRA should reflect the
{}``variability'' in the risk and calculate the {}``uncertainty''
associated with the risk estimate. The variability represents temporal,
geographical and/or individual heterogeneity of the risk for a given
population. The {}``uncertainty'' is understood as stemming from
a lack of perfect knowledge about the QRA model structure and associated
parameters%
\footnote{In the engineering risk community, these concepts are refered as \textquotedbl{}aleatoric
uncertainty\textquotedbl{} for \textquotedbl{}variability\textquotedbl{}
and \textquotedbl{}epistemic uncertainty\textquotedbl{} for \textquotedbl{}uncertainty\textquotedbl{}.%
}.

In order to estimate the natural {}``variability'' of the risk,
a Monte-Carlo simulation approach may be useful: the empirical distribution
of the risk within the population may be estimated from the mathematical
combination of distributions reflecting the variability of parameters
across the population.

A two-dimensional (or second-order) Monte-Carlo simulation was proposed
to estimate the {}``uncertainty'' in the risk estimates stemming
from parameter uncertainty. A two-dimensional Monte-Carlo simulation
is a Monte-Carlo simulation where the distributions reflecting \textquotedbl{}variability\textquotedbl{}
and the distributions representing \textquotedbl{}uncertainty\textquotedbl{}
are sampled separately in the simulation, so that \textquotedbl{}variability\textquotedbl{}
and \textquotedbl{}uncertainty\textquotedbl{} in the output may be
estinated separately.

\subsection{The need for a package for risk assessment}

Integrated

\subsection{An example}

\emph{Escherichia coli} O157:H7 in ground beef.

\section[Using fitdistrplus]{Using \pkg{fitdistrplus}}

To evaluate the intake distribution

\section[Using mc2d]{Using \pkg{mc2d}}

To derive the risk

\section{Conclusions}

\section{To be discarded}

\proglang{R} for the names of programming languages, \pkg{foo}
for software packages, and \code{some code} or \pkg{some code}
for code; $\E$  for expectations, $\VAR$  for variances, $\COV$
 for covariances, and $\Prob$  for probabilities.

Writing several lines of code:

\begin{CodeChunk}
\begin{Code}
first line of code
second line of code
third line of code
\end{Code}
\end{CodeChunk}

Alternatively, you can distinguish between input and output code:

\begin{CodeChunk}
\begin{CodeInput}
R> library(mc2d)
R> ndvar(1001)
\end{CodeInput}
\begin{CodeOutput}
[1] 1001 
\end{CodeOutput}
\begin{CodeInput}
ndunc(1001)
\end{CodeInput}
\begin{CodeOutput}
[1] 1001 
\end{CodeOutput}
\end{CodeChunk}

Or

\begin{CodeChunk}
\begin{CodeInput}
library(mc2d)
ndvar(1001)
\end{CodeInput}
\begin{CodeOutput}
[1] 1001 
\end{CodeOutput}
\begin{CodeInput}
ndunc(1001)
\end{CodeInput}
\begin{CodeOutput}
[1] 1001 
\end{CodeOutput}
\end{CodeChunk}

\section*{Acknowledgments}

Here you can write some acknowledgments.

\bibliographystyle{jss}
\bibliography{BibJss}

\end{document}

### 2009-02-06 pkg/mc2d/R/bern.R
New version of mc2d adapted to devel. No \code{} anymore
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@39 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -5,21 +5,21 @@ dbern <- function(x,prob=.5,log=FALSE)
 #KEYWORDS distribution
 #DESCRIPTION
 #Density, distribution function, quantile function and random generation
-#for the Bernoulli distribution with probability equals to \code{prob}.
+#for the Bernoulli distribution with probability equals to prob.
 #INPUTS
 #{x,q}<<vector of quantiles.>>
 #{p}<<vector of probabilities.>>
-#{n}<<number of observations. If \code{length(n) > 1}, the length is taken to be the number required.>>
+#{n}<<number of observations. If length(n) > 1, the length is taken to be the number required.>>
 #[INPUTS]
 #{prob}<<vector of probabilities of success of each trial.>>
-#{log, log.p}<<logical; if \code{TRUE}, probabilities \code{p} are given as \code{log(p)}.>>
-#{lower.tail}<<logical; if \code{TRUE} (default), probabilities are \code{P[X <= x]}, otherwise, \code{P[X > x]}.>>
+#{log, log.p}<<logical; if TRUE, probabilities p are given as log(p).>>
+#{lower.tail}<<logical; if TRUE (default), probabilities are P[X <= x], otherwise, P[X > x].>>
 #DETAILS
-#These fonctions use the corresponding functions from the \code{\link{binomial}} distribution with argument \code{size = 1}.
+#These fonctions use the corresponding functions from the \code{\link{binomial}} distribution with argument size = 1.
 #Thus, 1 is for success, 0 is for failure.
 #VALUE
-#\code{dbern} gives the density, \code{pbern} gives the distribution function,
-#\code{qbern} gives the quantile function, and \code{rbern} generates random deviates.
+#dbern gives the density, pbern gives the distribution function,
+#qbern gives the quantile function, and rbern generates random deviates.
 #EXAMPLE
 #rbern(n=10,prob=.5)
 #rbern(n=3,prob=c(0,.5,1))

### 2009-02-06 pkg/mc2d/R/empiricalD.R

@@ -6,7 +6,7 @@ dempiricalD <- function(x,values,prob=NULL,log=FALSE)
 #DESCRIPTION
 #Density, distribution function and random generation
 #for a discrete empirical distribution. This function is vectorized to accept
-#different sets of \code{values} or \code{prob}.
+#different sets of values or prob.
 #INPUTS
 #{x, q}<<Vector of quantiles.>>
 #{p}<<Vector of probabilities.>>
@@ -14,18 +14,18 @@ dempiricalD <- function(x,values,prob=NULL,log=FALSE)
 #{values}<<Vector or matrix of numerical values. See details.>>
 #[INPUTS]
 #{prob}<<Optionnal vector or matrix of count or probabilities. See details.>>
-#{log, log.p}<<logical; if \code{TRUE}, probabilities \code{p} are given as \code{log(p)}.>>
-#{lower.tail}<<logical; if \code{TRUE} (default), probabilities are \code{P[X <= x]}, otherwise, \code{P[X > x]}.>>
+#{log, log.p}<<logical; if TRUE, probabilities p are given as log(p).>>
+#{lower.tail}<<logical; if TRUE (default), probabilities are P[X <= x] otherwise, P[X > x].>>
 #DETAILS
-#If \code{prob} is missing, the discrete distribution is obtained directly from the vector of \code{values},
-#otherwise \code{prob} is used to weight the values. \code{prob} is normalized before use. Thus, \code{prob}
-#may be the count of each \code{values}. \code{prob} values should be non negative and their sum should not be 0.</>
-#\code{values} and/or \code{prob} may vary: in that case, \code{values} and/or \code{prob} should be sent
-#as matrixes, the first row being used for the first element of \code{x}, \code{q}, \code{p} or the first random value, the
-#second row for the second element of \code{x}, \code{q}, \code{p} or random value, ...
-#Recycling is permitted if the number of rows of \code{prob} and \code{values} are equal or
-#if the number of rows of \code{prob} and/or \code{values} are one.</>
-#\code{rempiricalD(n, values, prob)} with \code{values} and \code{prob} as vectors
+#If prob is missing, the discrete distribution is obtained directly from the vector of values,
+#otherwise prob is used to weight the values. prob is normalized before use. Thus, prob
+#may be the count of each values. prob values should be non negative and their sum should not be 0.</>
+#values and/or prob may vary: in that case, values and/or prob should be sent
[…]

### 2009-02-06 pkg/mc2d/R/evalmcmod.R

@@ -9,11 +9,11 @@ evalmcmod <- function(expr, nsv = ndvar(), nsu = ndunc(), seed = NULL)
 #[INPUTS]
 #{nsv}<<The number of simulations in the dimension of variability used in the evaluation.>>
 #{nsu}<<The number of simulations in the dimension of uncertainty used in the evaluation.>>
-#{seed}<<The random seed used for the evaluation. If \code{NULL} the \code{seed} is unchanged.>>
+#{seed}<<The random seed used for the evaluation. If NULL the seed is unchanged.>>
 #VALUE
-# The results of the evaluation. It should be a \code{mc} object.
+# The results of the evaluation. It should be a mc object.
 #DETAILS
-# The model is evaluated. The intermediate variables used to build the \code{mc} object are not stored.</>
+# The model is evaluated. The intermediate variables used to build the mc object are not stored.</>
 #NOTE
 #The seed is set at the beginning of the evaluation. Thus, the complete similarity
 #of two evaluations with similar seed is not certain, depending on the structure of your model.

### 2009-02-06 pkg/mc2d/R/is.mc.R

@@ -4,14 +4,14 @@ dimmcnode <- function(x)
 #DESCRIPTION
 # Provides the dimension (i.e. the number of simulations in the variability dimension,
 # the number of simulations in the uncertainty dimension and the 
-# maximum number of variates of a \code{mcnode} or a \code{mc} object.
+# maximum number of variates of a mcnode or a mc object.
 #KEYWORDS utilities
 #INPUTS
-#{x}<<a \code{mcnode} or a \code{mc} object.>>
+#{x}<<a mcnode or a mc object.>>
 #VALUE
-#A vector of three scalars: the dimension of variability (1 for \code{"0"} and \code{"U" mcnode}), 
-# the dimension of uncertainty (1 for \code{"0"} and \code{"V" mcnode}) and
-# the number of variates (the maximal number of variates for an \code{mc} object.
+#A vector of three scalars: the dimension of variability (1 for "0" and "U" mcnode), 
+# the dimension of uncertainty (1 for "0" and "V" mcnode) and
+# the number of variates (the maximal number of variates for an mc object.
 #NOTE
 #This function does not test if the object is correctly built. See \code{\link{is.mcnode}} and \code{\link{is.mc}} .
 #EXAMPLE
@@ -43,15 +43,15 @@ dimmc <- function(x)
 typemcnode <- function(x,index=FALSE)
 #TITLE Provides the Type of a mcnode Object
 #DESCRIPTION
-# Provide the type of a \code{mcnode} object.
+# Provide the type of a mcnode object.
 #KEYWORDS utilities
 #INPUTS
-#{x}<<a \code{mcnode} object>>
+#{x}<<a mcnode object>>
 #[INPUTS]
-#{index}<<if \code{TRUE} give the index of the type rather than the type.>>
+#{index}<<if TRUE give the index of the type rather than the type.>>
 #VALUE
-# \code{"0", "V","U" or "VU"} or the corresponding index if \code{index=TRUE}.</>
-#\code{NULL} if none of this element is found.
+# "0", "V","U" or "VU" or the corresponding index if index==TRUE.</>
+#NULL if none of this element is found.
 #NOTE
 #This function does not test if the object is correct. See \code{\link{is.mcnode}}.
 #EXAMPLE
@@ -73,19 +73,19 @@ typemcnode <- function(x,index=FALSE)
 is.mc <- function(x)
 #TITLE Tests mc and mcnode Objects
 #DESCRIPTION
[…]

### 2009-02-06 pkg/mc2d/R/is.na.R

@@ -3,12 +3,12 @@ is.na.mcnode <- function(x)
 #NAME NA.mcnode
 #TITLE Finite, Infinite, NA and NaN Numbers in mcnode.
 #DESCRIPTION
-# \code{is.na}, \code{is.nan}, \code{is.finite} and \code{is.infinite} return a logical \code{mcnode} 
-#of the same dimension as \code{x}.
+# is.na, is.nan, is.finite and is.infinite return a logical mcnode 
+#of the same dimension as x.
 #INPUTS
-#{x}<<A \code{mcnode} object.>>
+#{x}<<A mcnode object.>>
 #VALUE
-#A logical \code{mcnode} object.
+#A logical mcnode object.
 #SEE ALSO
 #\code{\link{is.finite}}, \code{\link{NA}}
 #KEYWORDS NA

### 2009-02-06 pkg/mc2d/R/mc.R

@@ -2,35 +2,35 @@
 mc <- function(..., name=NULL, devname=FALSE)
 #TITLE Monte Carlo Object
 #DESCRIPTION
-# Creates \code{mc} objects from \code{\link{mcnode}} or \code{mc} objects.
+# Creates mc objects from \code{\link{mcnode}} or mc objects.
 #KEYWORDS methods
 #INPUTS
-#{...}<<\code{mcnode} and/or \code{mc} object(s) to be gathered in a \code{mc} object
+#{...}<<mcnode and/or mc object(s) to be gathered in a mc object
 #separated by a coma.>>
 #[INPUTS]
-#{name}<<Vector of character of the same length of the final \code{mc} object.
+#{name}<<Vector of character of the same length of the final mc object.
 #If NULL, the name will be given from the name of the elements. >>
-#{devname}<<Develop the name from the name of the \code{mc} objects, if any.>>
-##{remove}<<If \code{TRUE}, original objects are removed from the parent environment.>>
+#{devname}<<Develop the name from the name of the mc objects, if any.>>
+##{remove}<<If TRUE, original objects are removed from the parent environment.>>
 #VALUE
-# An object of class \code{mc}.
+# An object of class mc.
 #DETAILS
-#A \code{mc} object is a list of \code{\link{mcnode}} objects.
-#\code{mcnode} objects must be of coherent dimensions.
+#A mc object is a list of \code{\link{mcnode}} objects.
+#mcnode objects must be of coherent dimensions.
 #
-#If one of the arguments is a \code{mc} object, the name of the elements of this \code{mc} object are used.
-#\code{devname = TRUE} will develop the name, using as a prefix the name of the \code{mc} object.</>
+#If one of the arguments is a mc object, the name of the elements of this mc object are used.
+#devname = TRUE will develop the name, using as a prefix the name of the mc object.</>
 #Finally, names are transformed to be unique.
 #SEE ALSO
-#\code{\link{mcnode}}, the basic element of a \code{mc} object.</>
-#To evaluate \code{mc} objects: \code{\link{mcmodel}}, \code{\link{evalmcmod}}, \code{\link{evalmccut}}</>
-#Informations about an \code{mc} object: \code{\link{is.mc}}, \code{\link{dimmc}}</>
[…]

### 2009-02-06 pkg/mc2d/R/mccontrol.R

@@ -8,14 +8,14 @@ ndvar <- function(n)
 #INPUTS
 #{n}<<Number of simulations.>>
 #DETAILS
-#\code{ndvar()} gets and \code{ndvar(n)} sets the default number of simulation in the 1D simulations
+#ndvar() gets and ndvar(n) sets the default number of simulation in the 1D simulations
 #or the number of simulation in the variability dimension in the 2D simulations.</>
-#\code{ndunc()} gets and \code{ndunc(n)} sets the number of simulations in the uncertainty dimension
+#ndunc() gets and ndunc(n) sets the number of simulations in the uncertainty dimension
 #in the 2D simulations.</>
-#\code{n} is rounded to its ceiling value.</>
-#The default values when loaded are 1001 for \code{ndvar} and 101 for \code{ndunc}.
+#n is rounded to its ceiling value.</>
+#The default values when loaded are 1001 for ndvar and 101 for ndunc.
 #VALUE
-#The current value, AFTER modification if \code{n} is present (!= \code{options}).
+#The current value, AFTER modification if n is present (!= \code{options()}).
 #EXAMPLE
 #(oldvar <- ndvar())
 #(oldunc <- ndunc())

### 2009-02-06 pkg/mc2d/R/mcdatanocontrol.R

@@ -2,8 +2,8 @@
 mcdatanocontrol <- function(data, type=c("V","U","VU","0"), nsv=ndvar(), nsu=ndunc(), nvariates=1, outm="each")
 #ISALIAS mcdata
 #DETAILS
-#\code{mcdatanocontrol} is a dangerous version of \code{mcnode} which forces the dimension
-#of data to be \code{(nsv x nsu x nvariates)} and gives the atributes and the class
+#mcdatanocontrol is a dangerous version of mcnode which forces the dimension
+#of data to be (nsv x nsu x nvariates) and gives the atributes and the class
 #without any control. This function is useful when your model is tested since
 #it is much more quicker.
 #

### 2009-02-06 pkg/mc2d/R/mcstoc.R

@@ -7,73 +7,73 @@ mcstoc <- function(func=runif, type=c("V","U","VU","0"), ..., nsv=ndvar(), nsu=n
 #INPUTS
 #{func}<<A function providing random data or its name as character.>>
 #[INPUTS]
-#{type}<<The type of \code{mcnode} to be built. By default, a \code{"V"} node. see \code{\link{mcnode}} for details.>>
-#{\dots}<<All other arguments but the size of the sample to be passed to \code{func}. These arguments
-#should be vectors or \code{mcnode}s (arrays prohibited).>>
+#{type}<<The type of mcnode to be built. By default, a "V" node. see \code{\link{mcnode}} for details.>>
+#{\dots}<<All other arguments but the size of the sample to be passed to func. These arguments
+#should be vectors or mcnodes (arrays prohibited).>>
 #{nsv}<<The number of simulations in the variability dimension.>>
 #{nsu}<<The number of simulations in the uncertainty dimension.>>
 #{nvariates}<<The number of variates of the output.>>
-#{outm}<<The  output of the \code{mcnode} for multivariates nodes. May be "each" (default)
+#{outm}<<The  output of the mcnode for multivariates nodes. May be "each" (default)
 #if an output should be provided for each variates considered independently, "none" for no output
 #or a vector of functions (as a character string) that will be applied on the variates dimension
-#before any output (ex: \code{"mean"}, \code{"median"}, \code{c("min","max")}). Each function should return 1
-#value when applied to 1 value (ex. do not use \code{"range"}).
-#Note that the \code{outm} attribute may be changed further using the \code{\link{outm}} function.>>
+#before any output (ex: "mean", "median", c("min","max")). Each function should return 1
+#value when applied to 1 value (ex. do not use "range").
+#Note that the outm attribute may be changed further using the \code{\link{outm}} function.>>
 #{nsample}<<The name of the parameter of the function giving the size of the vector.
-#By default, \code{n}, as in most of the random sampling distributions
-# of the \code{stats} library (with the exceptions of \code{rhyper} and \code{rwilcox} where \code{nsample="nn"} should be used).>>
[…]

### 2009-02-06 pkg/mc2d/R/outm.R

@@ -5,14 +5,14 @@ outm <- function(x,value="each",which.node=1)
 # Changes the output of Nodes
 #KEYWORDS misc
 #INPUTS
-#{x}<<A \code{mcnode} or a \code{mc} object.>>
+#{x}<<A mcnode or a mc object.>>
 #[INPUTS]
-#{value}<<The output of the \code{mcnode} for multivariates nodes. May be "each" (default)
+#{value}<<The output of the mcnode for multivariates nodes. May be "each" (default)
 #if output should be provided for each variates considered independently, "none" for no output
 #or a vector of name of function(s) (as a character string) that will be applied on the variates dimension
-#before any output (ex: \code{"mean"}, \code{"median"}, \code{c("min","max")}). The function should have no other arguments
-#and send one value per vector of values (ex. do not use \code{"range"}).>>
-#{which.node}<<which node should be changed in a \code{mc} object>>
+#before any output (ex: "mean", "median", c("min","max")). The function should have no other arguments
+#and send one value per vector of values (ex. do not use "range").>>
+#{which.node}<<which node should be changed in a mc object>>
 #EXAMPLE
 #data(total)
 #total$xVUM2
@@ -22,7 +22,7 @@ outm <- function(x,value="each",which.node=1)
 #summary(x)
 
 #VALUE
-#\code{x} with a modified \code{outm} attribute.
+#x with a modified outm attribute.
 #--------------------------------------------
 {
 

### 2009-02-06 pkg/mc2d/R/pert.R

@@ -5,8 +5,8 @@ dpert <- function(x,min=-1,mode=0,max=1,shape=4,log=FALSE)
 #KEYWORDS distribution
 #DESCRIPTION
 #Density, distribution function, quantile function and random generation
-#for the pert distribution with minimum equal to \code{min}, mode equal to \code{mode}
-#and maximum equal to \code{max}.
+#for the pert distribution with minimum equal to min, mode equal to mode
+#and maximum equal to max.
 #INPUTS
 #{x,q}<<Vector of quantiles.>>
 #{p}<<Vector of probabilities.>>
@@ -16,8 +16,8 @@ dpert <- function(x,min=-1,mode=0,max=1,shape=4,log=FALSE)
 #{mode}<<Vector of modes.>>
 #{max}<<Vector of maxima.>>
 #{shape}<<Vector of scaling parameters.>>
-#{log, log.p}<<Logical; if \code{TRUE}, probabilities \code{p} are given as \code{log(p)}.>>
-#{lower.tail}<<Logical; if \code{TRUE} (default), probabilities are \code{P[X <= x]}, otherwise, \code{P[X > x]}.>>
+#{log, log.p}<<Logical; if TRUE, probabilities p are given as log(p).>>
+#{lower.tail}<<Logical; if TRUE (default), probabilities are P[X <= x], otherwise, P[X > x].>>
 #DETAILS
 #The Pert distribution is a special case of the Beta distribution specified by the following parameters.
 #Given:
@@ -27,7 +27,7 @@ dpert <- function(x,min=-1,mode=0,max=1,shape=4,log=FALSE)
 #
 #\deqn{\alpha_{2}=\frac{\alpha_{1}\times (max-\mu)}{mu-min}}{shape2=shape1*(max - mu)/(mu - min)}
 #
-#on the domain \code{[min, max]}.</>
+#on the domain [min, max].</>
 #If \eqn{\mu=mode}{mu=mode}, \eqn{\alpha_{1}}{shape1} is set to \eqn{1+\nu/2}{1+shape/2}.
 #REFERENCE
 #Vose D. Risk Analysis - A Quantitative Guide (John Wiley & Sons, 2000).
@@ -39,8 +39,8 @@ dpert <- function(x,min=-1,mode=0,max=1,shape=4,log=FALSE)
 #SEE ALSO
 #\code{\link{Beta}}
 #VALUE
-#\code{dpert} gives the density, \code{ppert} gives the distribution function,
[…]

### 2009-02-06 pkg/mc2d/R/quantile.mc.R

@@ -2,18 +2,18 @@
 quantile.mc <- function(x, probs = seq(0, 1, 0.01), lim=c(0.025,0.975), na.rm=TRUE,  ...)
 #TITLE Quantiles of a mc Object
 #DESCRIPTION
-# Evaluates quantiles of a \code{mc} object. This function is used by \code{plot.mc}
+# Evaluates quantiles of a mc object. This function is used by plot.mc
 #KEYWORDS univar
 #INPUTS
-# {x}<<a \code{mc} objects>>
+# {x}<<a mc objects>>
 #[INPUTS]
 #{probs}<<the quantiles to be calculated>>
 #{na.rm}<<TRUE or FALSE>>
-#{lim}<<a vector of numbers (between 0 and 1) indicating the enveloppe. Maybe \code{NULL} or empty.>>
+#{lim}<<a vector of numbers (between 0 and 1) indicating the enveloppe. Maybe NULL or empty.>>
 #{\dots}<<For generic method consistancy.>>
 #DETAILS
 #The quantiles are evaluated in the variability dimension.
-#Then, the median, the mean and the \code{lim} quantiles are evaluated for each of these quantiles.
+#Then, the median, the mean and the lim quantiles are evaluated for each of these quantiles.
 #VALUE
 #A list of quantiles.
 #SEE ALSO

### 2009-02-06 pkg/mc2d/R/rtrunc.R

@@ -2,8 +2,8 @@
 rtrunc <- function(distr=runif, n, linf=-Inf, lsup=Inf,...)
 #TITLE Random Truncated Distributions
 #DESCRIPTION
-# Provides samples from classical \R distributions and \code{mc2d} specific
-# distributions truncated between \code{linf} and \code{lsup}.
+# Provides samples from classical \R distributions and \pkg{mc2d} specific
+# distributions truncated between linf and lsup.
 #KEYWORDS distribution
 #INPUTS
 #{distr}<<A function providing random data or its name as character.
@@ -13,13 +13,13 @@ rtrunc <- function(distr=runif, n, linf=-Inf, lsup=Inf,...)
 #[INPUTS]
 #{linf}<<A vector of lower bounds.>>
 #{lsup}<<A vector of upper bounds.>>
-#{\dots}<<All arguments to be passed to \code{pdistr} and \code{qdistr}.>>
+#{\dots}<<All arguments to be passed to pdistr and qdistr.>>
 #VALUE
-#A vector of \code{n} values.
+#A vector of n values.
 #DETAILS
-#The function 1) evaluates the \code{p} values corresponding to \code{linf} and \code{lsup} using \code{pdistr};
-#2) samples \code{n} values using \code{runif(n, min=pinf, max=psup)}, and 3) takes
-#the \code{n} corresponding quantiles from the specified distribution using \code{qdistr}.
+#The function 1) evaluates the p values corresponding to linf and lsup using pdistr;
+#2) samples n values using \code{runif(n, min=pinf, max=psup)}, and 3) takes
+#the n corresponding quantiles from the specified distribution using qdistr.
 #
 #All distributions (but sample) implemented in the stats library could be used.
 #The arguments in \dots should be named. Do not use 'log' or 'log.p' or 'lower.tail'.

### 2009-02-06 pkg/mc2d/R/summary.mc.R

@@ -2,11 +2,11 @@
 summary.mc <- function(object,probs = c(0,0.025,0.25,0.5,0.75,0.975,1),lim=c(0.025,0.975),...)
 #TITLE Summary of mcnode and mc Object
 #DESCRIPTION
-# Provides a summary of a \code{mcnode}, a \code{mc} or a \code{mccut} object.
+# Provides a summary of a mcnode, a mc or a mccut object.
 #KEYWORDS univar
 #INPUTS
-#{object}<<a \code{mcnode} or a \code{mc} object or a \code{mccut} object.>>
-#{x}<<A \code{summary.mc} object as provided by the \code{summary.mc} function.>>
+#{object}<<a mcnode or a mc object or a mccut object.>>
+#{x}<<A summary.mc object as provided by the summary.mc function.>>
 #[INPUTS]
 #{probs}<<A vector of values used for the quantile function (variability dimension).>>
 #{digits}<<Number of digits in the print.>>
@@ -15,14 +15,14 @@ summary.mc <- function(object,probs = c(0,0.025,0.25,0.5,0.75,0.975,1),lim=c(0.0
 #VALUE
 #a list.
 #DETAILS
-#The mean, the standard deviation and the \code{probs} quantiles will be evaluated in the variability dimension.
-#The median, the mean and the \code{lim} quantiles will then be evaluated on these statistics in the uncertainty dimension.
+#The mean, the standard deviation and the probs quantiles will be evaluated in the variability dimension.
+#The median, the mean and the lim quantiles will then be evaluated on these statistics in the uncertainty dimension.
 #
 #Multivariate nodes:
 #
-#If the \code{"outm"} attributes of the mcnode is "none", the node is not evaluated, if it is "each"
+#If the "outm" attributes of the mcnode is "none", the node is not evaluated, if it is "each"
 #the variates are evaluated one by one, if it is a function (e.g. "mean"), the function is applied on the
-#\code{nvariates} dimension before providing a classical output.
+#nvariates dimension before providing a classical output.
 #
 #SEE ALSO
 # \code{\link{mcnode}} for mcnode objects, \code{\link{mc}} for mc objects, \code{\link{mccut}} for mccut objects, \code{\link{quantile}}

### 2009-02-06 pkg/mc2d/R/Ops.mcnode.R
New version of mc2d adapted to devel. No \code{} anymore
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@39 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -2,40 +2,39 @@
 Ops.mcnode <- function(e1,e2)
 #TITLE Operations on mcnode Objects
 #DESCRIPTION
-# This function alters the way operations are performed on \code{mcnode} objects for a better consistancy of the theory.
+# This function alters the way operations are performed on mcnode objects for a better consistancy of the theory.
 #KEYWORDS utilities
 #INPUTS
-#{e1}<<An \code{mcnode} object, a vector or an array.>>
-#{e2}<<An optionnal \code{mcnode} object, a vector or a matrix with at least one of both objects as an \code{mcnode}.>>
+#{e1}<<An mcnode object, a vector or an array.>>
+#{e2}<<An optionnal mcnode object, a vector or a matrix with at least one of both objects as an mcnode.>>
 #VALUE
-#The results as a \code{mcnode} object.
+#The results as a mcnode object.
 #DETAILS
 #This method will be used for any of the Group \code{\link{Ops}} functions, i.e.:
-#{*}<<\code{"+"}, \code{"-"}, \code{"*"}, \code{"/"}, \code{"^"}, \code{"\%\%"},
-#\code{"\%/\%"}>>
-#{*}<<\code{"&", "|", "!"}>>
-#{*}<<\code{"==", "!=", "<", "<=", ">=", ">"}>>
+#{*}<<"+", "-", "*", "/", "^", "\%\%","\%/\%">>
+#{*}<<"&", "|", "!">>
+#{*}<<"==", "!=", "<", "<=", ">=", ">">>
 #
-#The rules are as following (illustrated with a \code{"+"} function and ignoring the \code{nvariates} dimension):
-#{*}<<\code{"0" + "0" = "0"};>>
-#{*}<<\code{"0" + "V" = "V"}: classical recycling of the scalar;>>
-#{*}<<\code{"0" + "U" = "U"}: classical recycling of the scalar;>>
-#{*}<<\code{"0" + "VU" = "VU"}: classical recycling of the scalar;>>
-#{*}<<\code{"V" + "V" = "V"}: if both of the same \code{(nsv)} dimension;>>
-#{*}<<\code{"V" + "U" = "VU"}: the \code{"U"} object will be recycled "by row". The \code{"V"} object will be recycled classically "by column";>>
-#{*}<<\code{"V" + "VU" = "VU"}: if the dimension of the \code{"V"} is \code{(nsv)} and the dimension of the \code{"VU"} is \code{(nsv x nsu)}. The \code{"V"} object will be recycled classically "by column";>>
[…]

### 2009-02-06 pkg/mc2d/R/betagen.R

@@ -5,19 +5,19 @@ dbetagen <- function(x,shape1,shape2,min=0,max=1,ncp=0,log=FALSE)
 #KEYWORDS distribution
 #DESCRIPTION
 #Density, distribution function, quantile function and random generation for the Beta distribution
-#defined on the \code{[min, max]} domain with parameters \code{shape1} and \code{shape2} (
-#and optional non-centrality parameter \code{ncp}).
+#defined on the [min, max] domain with parameters shape1 and shape2 (
+#and optional non-centrality parameter ncp).
 #INPUTS
 #{x,q}<<Vector of quantiles.>>
 #{p}<<Vector of probabilities.>>
-#{n}<<Number of observations. If \code{length(n) > 1}, the length is taken to be the number required.>>
+#{n}<<Number of observations. If length(n) > 1, the length is taken to be the number required.>>
 #{shape1, shape2}<<Positive parameters of the Beta distribution.>>
 #[INPUTS]
 #{min}<<Vector of minima.>>
 #{max}<<Vector of maxima.>>
 #{ncp}<<Non-centrality parameter of the Beta distribution.>>
-#{log, log.p}<<Logical; if \code{TRUE}, probabilities \code{p} are given as \code{log(p)}.>>
-#{lower.tail}<<Logical; if \code{TRUE} (default), probabilities are \code{P[X <= x]}, otherwise, \code{P[X > x]}.>>
+#{log, log.p}<<Logical; if TRUE, probabilities p are given as log(p).>>
+#{lower.tail}<<Logical; if TRUE (default), probabilities are P[X <= x], otherwise, P[X > x].>>
 #DETAILS
 #\deqn{x \sim  betagen(shape1, shape2, min, max, ncp)}{x ~ betagen(shape1, shape2, min, max, ncp)}
 #if
@@ -30,8 +30,8 @@ dbetagen <- function(x,shape1,shape2,min=0,max=1,ncp=0,log=FALSE)
 #SEE ALSO
 #\code{\link{Beta}}
 #VALUE
-#\code{dbetagen} gives the density, \code{pbetagen} gives the distribution function,
-#\code{qbetagen} gives the quantile function, and \code{rbetagen} generates random deviates.
+#dbetagen gives the density, pbetagen gives the distribution function,
+#qbetagen gives the quantile function, and rbetagen generates random deviates.
 #AUTHOR Regis Pouillot
 #CREATED 08-04-16
 #--------------------------------------------

### 2009-02-06 pkg/mc2d/R/converg.R

@@ -7,28 +7,28 @@ converg <- function(x, node=length(x), margin=c("var","unc"), nvariates=1, iter=
 #INPUTS
 #{x}<<A \code{\link{mcnode}} object, a \code{\link{mc}} object or a \code{\link{mccut}} object.>>
 #[INPUTS]
-#{node}<<The node to be considered in a \code{mc} object or a \code{mccut} object, displayed either as the order number or the name of the node.
-#By default: the last node of the object.The corresponding node should not be of type \code{"0"} in a \code{mc} object or 
-#of type \code{"0"} or \code{"V"} in a \code{mccut} object.>>
-#{margin}<<The margin used to plot the graph. \code{margin} is used only if the node is a \code{"VU" mcnode}.>>
-#{nvariates}<<The variates to be considered. \code{nvariates} is used only for multivariates nodes.>>
-#{iter}<<If \code{margin == "var"} and the node is a \code{"VU" mcnode}, \code{iter} specify
+#{node}<<The node to be considered in a mc object or a mccut object, displayed either as the order number or the name of the node.
+#By default: the last node of the object.The corresponding node should not be of type "0" in a mc object or 
+#of type "0" or "V" in a mccut object.>>
+#{margin}<<The margin used to plot the graph. margin is used only if the node is a "VU" mcnode.>>
+#{nvariates}<<The variates to be considered. nvariates is used only for multivariates nodes.>>
+#{iter}<<If margin == "var" and the node is a "VU" mcnode, iter specify
 #the iteration in the uncertainty dimension to be used for the graph.>>
 #{probs}<<The quantiles to be provided in the variability dimension.>>
 #{lim}<<The quantiles to be used in the uncertainty dimension.>>
-#{griddim}<<A vector of two integers, indicating the size of the grid of the graph. If \code{NULL}, the grid is calculated to produce a "nice" graph.>>
-#{log}<<If \code{TRUE}, the data will be log transformed.>>
+#{griddim}<<A vector of two integers, indicating the size of the grid of the graph. If NULL, the grid is calculated to produce a "nice" graph.>>
+#{log}<<If TRUE, the data will be log transformed.>>
 #NOTE
-#This function may be used on a \code{mccut} object only if a \code{summary.mc} function was used in the third block of the \code{\link{evalmccut}} call.
-#The values used as \code{probs} arguments in \code{converg} should have been used in the \code{summary.mc} function of this third block.
+#This function may be used on a mccut object only if a summary.mc function was used in the third block of the \code{\link{evalmccut}} call.
[…]

### 2009-02-06 pkg/mc2d/R/cornode.R

@@ -3,41 +3,41 @@ cornode <- function(...,target, outrank=FALSE, result=FALSE, seed=NULL)
 #TITLE Builds a Rank Correlation using the Iman and Connover Method.
 #KEYWORDS multivariate
 #DESCRIPTION
-#This function builds a rank correlation structure between columns of a matrix or between \code{mcnode} objects
+#This function builds a rank correlation structure between columns of a matrix or between mcnode objects
 #using the Iman and Connover method (1982).
 #INPUTS
-#{\ldots}<<A matrix (each of its \code{n} columns but the first one will be reordered)
-#or \code{n mcnode} objects (each elements but the first one will be reordered).>>
-#{target}<<A scalar (only if \code{n=2}) or a \code{(n x n)} matrix of correlation.>>
+#{\ldots}<<A matrix (each of its n columns but the first one will be reordered)
+#or n mcnode objects (each elements but the first one will be reordered).>>
+#{target}<<A scalar (only if n==2) or a (n x n) matrix of correlation.>>
 #[INPUTS]
 #{outrank}<<Should the order be returned?>>
 #{result}<<Should the correlation eventually obtained be printed?>>
-#{seed}<<The random seed used for building the correlation. If \code{NULL} the \code{seed} is unchanged.>>
+#{seed}<<The random seed used for building the correlation. If NULL the seed is unchanged.>>
 #DETAILS
 #The arguments should be named.
 #
-#The function accepts for \code{data} a matrix or:
-#{*}<<some \code{"V" mcnode} objects separated by a comma;>>
-#{*}<<some \code{"U" mcnode} objects separated by a comma;>>
-#{*}<<some \code{"VU" mcnode} objects separated by a comma. In that case, the structure is built columns by colums (the first column of each \code{"VU" mcnode}
+#The function accepts for data a matrix or:
+#{*}<<some "V" mcnode objects separated by a comma;>>
+#{*}<<some "U" mcnode objects separated by a comma;>>
+#{*}<<some "VU" mcnode objects separated by a comma. In that case, the structure is built columns by colums (the first column of each "VU" mcnode
 #will have a correlation structure, the second ones will have a correlation structure, ....).>>
-#{*}<<one \code{"V" mcnode} as a first element and some \code{"VU" mcnode} objects, separated by a comma.
-#In that case, the structure is built between the \code{"V" mcnode} and each column of the \code{"VU" mcnode} objects.
-#The correlation result (\code{result = TRUE}) is not provided in that case.>>
[…]

### 2009-02-06 pkg/mc2d/R/empiricalC.R

@@ -8,26 +8,26 @@ dempiricalC <- function(x, min, max, values, prob=rep(1,length(values)), log=FAL
 #INPUTS
 #{x, q}<<Vector of quantiles.>>
 #{p}<<Vector of probabilities.>>
-#{n}<<Number of random values. If \code{length(n) > 1}, the length is taken to be the number required.>>
+#{n}<<Number of random values. If length(n) > 1, the length is taken to be the number required.>>
 #{min}<<A finite minimal value.>>
 #{max}<<A finite maximal value.>>
 #{values}<<Vector of numerical values.>>
 #[INPUTS]
 #{prob}<<Optionnal vector of count or probabilities.>>
-#{log, log.p}<<logical; if \code{TRUE}, probabilities \code{p} are given as \code{log(p)}.>>
-#{lower.tail}<<logical; if \code{TRUE} (default), probabilities are \code{P[X <= x]}, otherwise, \code{P[X > x]}.>>
+#{log, log.p}<<logical; if TRUE, probabilities p are given as log(p).>>
+#{lower.tail}<<logical; if TRUE (default), probabilities are P[X <= x], otherwise, P[X > x].>>
 #DETAILS
-#Given \eqn{p_{i}}{p_i}, the distribution value for \eqn{x_{i}}{x_i} with \code{i} the rank \eqn{i = 0, 1, 2, \ldots, N+1},
+#Given \eqn{p_{i}}{p_i}, the distribution value for \eqn{x_{i}}{x_i} with i the rank \eqn{i = 0, 1, 2, \ldots, N+1},
 # \eqn{x_{0}=min}{x_0 = min} and \eqn{x_{N+1}=max}{x_(N+1) = max} the density is:
 # \deqn{f(x)=p_{i}+(\frac{x-x_{i}}{x_{i+1}-x_{i}})(p_{i+1}-p_{i})}{f(x) = p_i + (p_(i+1) - p_i)/(x_(i+1) - x_i) for x_i<=x<x_(i+1)}
-# The \code{p} values being normalized to give the distribution a unit area.
+# The p values being normalized to give the distribution a unit area.
 #
 #This function is not vectorized. It can not currently be use with varying parameters.
 #SEE ALSO
 #\code{\link{empiricalD}}
 #VALUE
-#\code{dempiricalC} gives the density, \code{pempiricalC} gives the distribution function,
-#\code{qempiricalC} gives the quantile function and \code{rempiricalC} generates random deviates.
[…]

### 2009-02-06 pkg/mc2d/R/evalmccut.R

@@ -3,73 +3,72 @@ evalmccut <- function(model, nsv = ndvar(), nsu = ndunc(), seed = NULL,  ind = "
 #TITLE Evaluates a Two-Dimensional Monte Carlo Model in a Loop.
 #NAME mccut
 #DESCRIPTION
-#\code{evalmccut} evaluates a Two-Dimensional Monte Carlo model using a loop on the uncertainty dimension.
+#evalmccut evaluates a Two-Dimensional Monte Carlo model using a loop on the uncertainty dimension.
 #Within each loop, it calculates statistics in the variability dimension and stores them 
 #for further analysis.
 #It allows to evaluate very high dimension models using (unlimited?) time instead of (limited) memory.</>
-#\code{mcmodelcut} builds a \code{mcmodelcut} object that can be sent to \code{evalmccut}.
+#mcmodelcut builds a mcmodelcut object that can be sent to evalmccut.
 #KEYWORDS methods
 #INPUTS
-#{model}<<a \code{mcmodelcut} object obtained using \code{mcmodelcut} function or (directly) a valid call including three blocks.
+#{model}<<a mcmodelcut object obtained using mcmodelcut function or (directly) a valid call including three blocks.
 #See Details and Examples for the structure of the call.>>
-#{x}<<a call or an expression (if \code{is.expr=TRUE}) including three blocks. See Details and Examples for the structure of the call.>>
+#{x}<<a call or an expression (if is.expr=TRUE) including three blocks. See Details and Examples for the structure of the call.>>
 #[INPUTS]
 #{nsv}<<The number of simulations for variability used in the evaluation.>>
 #{nsu}<<The number of simulations for uncertainty used in the evaluation.>>
-#{seed}<<The random seed used for the evaluation. If \code{NULL} the \code{seed} is unchanged.>>
-#{ind}<<The variable name used in \code{model} to refers to the uncertainty. see Details and Example.>>
-#{is.expr}<< \code{FALSE} to send a call,  \code{TRUE} to send an expression (see \code{\link{mcmodel}} examples)>>
+#{seed}<<The random seed used for the evaluation. If NULL the seed is unchanged.>>
+#{ind}<<The variable name used in model to refers to the uncertainty. see Details and Example.>>
+#{is.expr}<< FALSE to send a call,  TRUE to send an expression (see \code{\link{mcmodel}} examples)>>
 #{lim}<<A vector of values used for the quantile function (uncertainty dimension).>>
[…]

### 2009-02-06 pkg/mc2d/R/hist.mc.R

@@ -2,18 +2,18 @@
 hist.mc <- function(x, griddim = NULL, xlab = names(x),ylab = "Frequency", main = "",  ...)
 #TITLE Histogram of a Monte Carlo Simulation
 #DESCRIPTION
-# Shows histogram of a \code{mcnode} or a \code{mc} object.
+# Shows histogram of a mcnode or a mc object.
 #KEYWORDS hplot
 #INPUTS
-#{x}<<An \code{mcnode} or an \code{mc} object.>>
+#{x}<<An mcnode or an mc object.>>
 #[INPUTS]
-#{griddim}<<A vector of two integers, indicating the size of the grid of plots. If \code{NULL}, the grid is calculated to produce a "nice" graph.>>
-#{xlab}<<A vector of labels for the x-axis for drawn histograms (those whose \code{outm(x)!="none"}). May be recycled.>>
+#{griddim}<<A vector of two integers, indicating the size of the grid of plots. If NULL, the grid is calculated to produce a "nice" graph.>>
+#{xlab}<<A vector of labels for the x-axis for drawn histograms (those whose outm(x)!="none"). May be recycled.>>
 #{ylab}<<A vector of labels for the y-axis for drawn histograms. May be recycled.>>
 #{main}<<A vector of main title of histograms for drawn histograms. May be recycled.>>
 #{\dots}<<Other arguments to be passed to all calls of \code{hist}.>>
 #NOTE
-#For Two-dimensional \code{mc}, the histogram is based on all data (variability and uncertainty) pooled together.
+#For Two-dimensional mc, the histogram is based on all data (variability and uncertainty) pooled together.
 #EXAMPLE
 #data(total)
 #hist(xVUM3)

### 2009-02-06 pkg/mc2d/R/lhs.R

@@ -5,27 +5,27 @@ lhs <- function(distr=runif,nsv,nsu,nvariates=1,...)
 # Creates a Latin Hypercube Sample (LHS) of the specified distribution.
 #KEYWORDS design
 #INPUTS
-#{distr}<<The function for generating random sample or its name. If \code{distr} is "rdist",
+#{distr}<<The function for generating random sample or its name. If distr is "rdist",
 #the function "qdist" must be the quantile function of this distribution with argument
-#\code{p} as a vector of probabilities, as all univariates distributions of the \code{stat}
+#p as a vector of probabilities, as all univariates distributions of the \pkg{stat}
 #library.>>
 #{nsv}<<The number of raws of the final matrix.>>
 #{nsu}<<The number of columns of the final matrix>>
 #[INPUTS]
 #{nvariates}<<The number of variates>>
-#{\dots}<<All arguments to be passed to \code{distr} except the size of the sample.>>
+#{\dots}<<All arguments to be passed to distr except the size of the sample.>>
 #VALUE
-#A \code{nsv x nsu} matrix of random variates.
+#A (nsv x nsu) matrix of random variates.
 #NOTE
 #The resulting lhs is in fact a latin hypersquare sampling: the lhs is provided only in the first 2 dimensions.</>
 #It is not possible to send truncated distribution with \code{\link{rtrunc}}. Use \code{\link{mcstoc}} for
-#this purpose, with \code{lhs=TRUE} and \code{rtrunc=TRUE}.</>
+#this purpose, with lhs=TRUE and rtrunc=TRUE.</>
 #The \dots arguments will be recycled.
 #SEE ALSO
 #\code{\link{mcstoc}}
 #EXAMPLE
 #ceiling(lhs(runif,nsu=10,nsv=10)*10)
-#AUTHOR adapted from a code of Rob Carnell (library \code{lhs})
+#AUTHOR adapted from a code of Rob Carnell (library \pkg{lhs})
 #CREATED 08-01-25
 #--------------------------------------------
 {

### 2009-02-06 pkg/mc2d/R/mcapply.R

@@ -5,22 +5,22 @@
 mcapply <- function(x, margin=c("all","var","unc","variates"), fun, ...)
 #TITLE Apply Functions Over mc or mcnode Objects
 #DESCRIPTION
-# Apply a function on all values or over a given dimension of an \code{mcnode} object. May be used for all \code{mcnode} of an \code{mc} object.
+# Apply a function on all values or over a given dimension of an mcnode object. May be used for all mcnode of an mc object.
 #KEYWORDS misc
 #INPUTS
-#{x}<<A \code{mc} or a \code{mcnode} object.>>
+#{x}<<A mc or a mcnode object.>>
 #[INPUTS]
 #{margin}<<The dimension on which applying the function.
-#Maybe \code{"all"} (default) to apply the function on all values,
-#\code{"var"} to apply the function on the variability dimension,
-#\code{"unc"} to apply the function on the uncertainty dimension, or
-#\code{"variates"} to apply the function on the variates.>>
+#Maybe "all" (default) to apply the function on all values,
+#"var" to apply the function on the variability dimension,
+#"unc" to apply the function on the uncertainty dimension, or
+#"variates" to apply the function on the variates.>>
 #{fun}<<The function to be applied.
-#When applied to a vector of length \code{n}, \code{fun} should return a vector of length \code{n} or \code{1}.>>
-#{\dots}<<Optionnal arguments to \code{fun}.>>
+#When applied to a vector of length n, fun should return a vector of length n or 1.>>
+#{\dots}<<Optionnal arguments to fun.>>
 #VALUE
-#If \code{fun} returns a function of length \code{n} or if \code{margin="all"}, the returned \code{mcnode}s are of type and dimension of \code{x}.
-#In other cases, the type of \code{mcnode} is changed.
+#If fun returns a function of length n or if margin=="all", the returned mcnodes are of type and dimension of x.
+#In other cases, the type of mcnode is changed.
 #SEE ALSO
 #\code{\link{apply}}, \code{\link{mc}}, \code{\link{mcnode}}.
 #EXAMPLE

### 2009-02-06 pkg/mc2d/R/mcdata.R

@@ -3,87 +3,87 @@ mcdata <- function(data, type=c("V","U","VU","0"), nsv=ndvar(), nsu=ndunc(),nvar
 #TITLE Build mcnode Objects from Data or other mcnode Objects
 #NAME mcnode
 #DESCRIPTION
-# Creates a \code{mcnode} object from a vector, an array or a \code{mcnode}.
+# Creates a mcnode object from a vector, an array or a mcnode.
 #KEYWORDS methods
 #INPUTS
-#{data}<<The numeric/logical vector/matrix/array of data or the \code{mcnode} object.>>
+#{data}<<The numeric/logical vector/matrix/array of data or the mcnode object.>>
 #[INPUTS]
-#{type}<<The type of node to be built. By default, a \code{"V"} node.>>
-#{nsv}<<The variability dimension (\code{type="V"} or \code{type="VU"}) of the node.
+#{type}<<The type of node to be built. By default, a "V" node.>>
+#{nsv}<<The variability dimension (type="V" or type=="VU") of the node.
 #By default: the current value in \code{\link{mc.control}}>>
-#{nsu}<<The uncertainty dimension (\code{type="U"} or \code{type="VU"}) of the node.
+#{nsu}<<The uncertainty dimension (type=="U" or type=="VU") of the node.
 #By default: the current value in \code{\link{mc.control}}>>
 #{nvariates}<<The number of variates.
 #By default: 1>>
-#{outm}<<The output of the \code{mcnode} for multivariates nodes. May be "each" (default)
+#{outm}<<The output of the mcnode for multivariates nodes. May be "each" (default)
 #if output should be provided for each variates considered independently, "none" for no output
 #or a vector of name of function(s) (as a character string) that will be applied on the variates dimension
-#before any output (ex: \code{"mean"}, \code{"median"}, \code{c("min", "max")}). The function should have no other arguments
-#and send one value per vector of values (ex. do not use \code{"range"}).
-#Note that the \code{outm} attribute may be changed at any time using the \code{\link{outm}} function.>>
+#before any output (ex: "mean", "median", c("min", "max")). The function should have no other arguments
+#and send one value per vector of values (ex. do not use "range").
[…]

### 2009-02-06 pkg/mc2d/R/mcmodel.R

@@ -2,20 +2,20 @@
 mcmodel <- function(x, is.expr=FALSE)
 #TITLE Monte Carlo model
 #DESCRIPTION
-# Specify a \code{mcmodel}, without evaluating it, for a further evaluation using \code{\link{evalmcmod}}.
+# Specify a mcmodel, without evaluating it, for a further evaluation using \code{\link{evalmcmod}}.
 #KEYWORDS methods
 #INPUTS
 #{x}<<An \R call or an expression.>>
 #[INPUTS]
-#{is.expr}<< \code{FALSE} to send a call,  \code{TRUE} to send an expression (see Examples)>>
+#{is.expr}<< FALSE to send a call,  TRUE to send an expression (see Examples)>>
 #VALUE
-# an \R expression, with class \code{mcmodel}
+# an \R expression, with class mcmodel
 #DETAILS
-# The model should be put between \code{\{} and the last line should be of the form \code{mc(...)}.
+# The model should be put between "{" and the last line should be of the form mc(...).
 # Any reference to the number of simulation in the
-#dimension of variability should be done via \code{ndvar()} or (preferred) \code{nsv}.
+#dimension of variability should be done via ndvar() or (preferred) nsv.
 # Any reference to the number of simulations in the dimension of uncertainty
-# should be done via \code{ndunc()} or (preferred) \code{nsu}.
+# should be done via ndunc() or (preferred) nsu.
 #SEE ALSO
 #\code{\link{expression}}.</>
 #\code{\link{evalmcmod}} to evaluate the model.</>

### 2009-02-06 pkg/mc2d/R/mcprobtree.R

@@ -2,37 +2,37 @@
 mcprobtree <- function(mcswitch, mcvalues, type=c("V","U","VU","0"), nsv=ndvar(), nsu=ndunc(), nvariates=1, outm="each", seed=NULL)
 #TITLE Creates a Stochastic mcnode Object using a Probability Tree
 #DESCRIPTION
-#This function builds a \code{mcnode} as a mixture of \code{mcstoc} functions or \code{mcnode} objects.
+#This function builds a mcnode as a mixture of mcstoc functions or mcnode objects.
 #KEYWORDS methods
 #INPUTS
-#{mcswitch}<<A vector of probabilities/weights or a \code{mcnode} including the \code{mcstoc} functions/mcnodes to pick.>>
-#{mcvalues}<<A named list of \code{mcnode}, \code{mcdata} functions or \code{mcstoc} functions, or a combination of those objects.
-#Each element should lead to an \code{mcnode} of type \code{type} and of dimension \code{c(nsv x nsu x 1)} or \code{c(nsv x nsu x nvariates)} >>
+#{mcswitch}<<A vector of probabilities/weights or a mcnode including the mcstoc functions/mcnodes to pick.>>
+#{mcvalues}<<A named list of mcnode, mcdata functions or mcstoc functions, or a combination of those objects.
+#Each element should lead to an mcnode of type type and of dimension (nsv x nsu x 1) or (nsv x nsu x nvariates) >>
 #[INPUTS]
-#{type}<<The type of \code{mcnode} to be built. By default, a \code{"V"} node. see \code{\link{mcnode}} for details.>>
+#{type}<<The type of mcnode to be built. By default, a "V" node. see \code{\link{mcnode}} for details.>>
 #{nsv}<<The number of simulations in the variability dimension of the final node.>>
 #{nsu}<<The number of simulations in the uncertainty dimension of the final node.>>
-#{nvariates}<<The number of variates of the final \code{mcnode}.>>
-#{outm}<<The default output of the \code{mcnode} for multivariates nodes. see \code{\link{outm}}.>>
-#{seed}<<The random seed used for the evaluation. If \code{NULL} the \code{seed} is unchanged.>>
+#{nvariates}<<The number of variates of the final mcnode.>>
+#{outm}<<The default output of the mcnode for multivariates nodes. see \code{\link{outm}}.>>
+#{seed}<<The random seed used for the evaluation. If NULL the seed is unchanged.>>
 #VALUE
-#An \code{mcnode} object.
+#An mcnode object.
 #DETAILS
[…]

### 2009-02-06 pkg/mc2d/R/plot.mc.R

@@ -2,34 +2,34 @@
 plot.mc <- function(x, prec=0.01, stat = c("median","mean"), lim = c(0.025,0.975), na.rm=TRUE, griddim = NULL, xlab = NULL, ylab = "Fn(x)", main = "", draw = TRUE, ...)
 #TITLE Plots Results of a Monte Carlo Simulation
 #DESCRIPTION
-# Plots the empirical cumulative distribution function of a \code{mcnode} or a \code{mc} object ("0" and "V" nodes) or the
-#empirical cumulative distribution function of the estimate of a \code{mcnode} or \code{mc} object ("U" and "VU" nodes).
+# Plots the empirical cumulative distribution function of a mcnode or a mc object ("0" and "V" nodes) or the
+#empirical cumulative distribution function of the estimate of a mcnode or mc object ("U" and "VU" nodes).
 #KEYWORDS hplot
 #INPUTS
-# {x}<<a \code{mcnode} or a \code{mc} objects>>
+# {x}<<a mcnode or a mc objects>>
 #[INPUTS]
 #{prec}<<the precision of the plot. 0.01 will
 #provide an ecdf from the 0.00, 0.01, .02,  ..., 1.00 quantiles, 0.001 will provide
 #a 0.000, 0.001, 0.002, ..., 1.000 quantiles,... >>
-#{stat}<<the function used for estimates (2D \code{mc} or \code{mcnode}). By default the median.>>
-#{lim}<<a vector of numbers (between 0 and 1) indicating the enveloppe (2D \code{mc} or \code{mcnode}) . Maybe \code{NULL} or empty.>>
+#{stat}<<the function used for estimates (2D mc or mcnode). By default the median.>>
+#{lim}<<a vector of numbers (between 0 and 1) indicating the enveloppe (2D mc or mcnode) . Maybe NULL or empty.>>
 #{na.rm}<<Should NA values be discarded>>
-#{griddim}<<a vector of two integers, indicating the size of the grid of the graph. If \code{NULL}, the grid is calculated to produce a "nice" graph.>>
-#{xlab}<<vector of labels for the x-axis. If \code{NULL}, use the name of the node.>>
+#{griddim}<<a vector of two integers, indicating the size of the grid of the graph. If NULL, the grid is calculated to produce a "nice" graph.>>
+#{xlab}<<vector of labels for the x-axis. If NULL, use the name of the node.>>
 #{ylab}<<vector of labels for the y-axis.>>
[…]

### 2009-02-06 pkg/mc2d/R/plot.tornado.R

@@ -2,7 +2,7 @@
 plot.tornado <- function(x,which=1,name=NULL,stat=c("median","mean"),xlab="method",ylab="",...)
 #TITLE Draws a Tornado chart.
 #DESCRIPTION
-# Draws a Tornado chart as provided by \code{tornado}.
+# Draws a Tornado chart as provided by tornado.
 #KEYWORDS hplot
 #INPUTS
 #{x}<<A \code{\link{tornado}} object or a \code{\link{tornadounc}} object.>>
@@ -10,12 +10,12 @@ plot.tornado <- function(x,which=1,name=NULL,stat=c("median","mean"),xlab="metho
 #{which}<<Which output to print -for multivariates output-.>>
 #{name}<<Vector of name of input variables.
 #If NULL, the name will be given from the name of the elements.>>
-#{stat}<<The name of the statistics of the output to be considered. For a \code{tornado} object: "median" or "mean". 
-#For a \code{tornadounc} object: the value should match one row name of the \code{tornadounc} object.
-#Alternatively, for a \code{tornadounc} object, the number of the row may be used.>>
-#{xlab}<<Label of the x axis. if "method", use the correlation method used in the \code{tornado} object.>>
+#{stat}<<The name of the statistics of the output to be considered. For a tornado object: "median" or "mean". 
+#For a tornadounc object: the value should match one row name of the tornadounc object.
+#Alternatively, for a tornadounc object, the number of the row may be used.>>
+#{xlab}<<Label of the x axis. if "method", use the correlation method used in the tornado object.>>
 #{ylab}<<Label of the y axis.>>
-#{\dots}<<Further arguments to be passed to the \code{plot} function.>>
+#{\dots}<<Further arguments to be passed to the plot function.>>
 #VALUE
 # NULL
 #DETAILS

### 2009-02-06 pkg/mc2d/R/print.mc.R

@@ -2,10 +2,10 @@
 print.mc <- function(x, digits=3,...)
 #TITLE Prints a mcnode or a mc Object
 #DESCRIPTION
-# Print a description of the structure of the \code{mc} or the \code{mcnode} object.
+# Print a description of the structure of the mc or the mcnode object.
 #KEYWORDS print
 #INPUTS
-#{x}<<a \code{mcnode} or a \code{mc} object.>>
+#{x}<<a mcnode or a mc object.>>
 #[INPUTS]
 #{digits}<<Number of digits to be used.>>
 #{\dots}<<Further arguments to be passed to the print function.>>
@@ -13,8 +13,8 @@ print.mc <- function(x, digits=3,...)
 #An invisible data frame.
 #DETAILS
 #SEE ALSO
-# \code{\link{mcnode}} for \code{mcnode} objects.
-# \code{\link{mc}} for \code{mc} objects.
+# \code{\link{mcnode}} for mcnode objects.
+# \code{\link{mc}} for mc objects.
 #EXAMPLE
 #AUTHOR Regis Pouillot
 #CREATED 08-01-25

### 2009-02-06 pkg/mc2d/R/rdirichlet.R

@@ -10,11 +10,11 @@ ddirichlet <- function (x, alpha)
 #{n}<<Number of random vectors to generate. If length(n) > 1, the length is taken to be the number required.>>
 #DETAILS
 #The Dirichlet distribution is the multidimensional generalization of the beta distribution. 
-#The original code was adapted to provide a kind of "vectorization" used in multivariates \code{mcnode}.
+#The original code was adapted to provide a kind of "vectorization" used in multivariates mcnode.
 #VALUE
-#\code{ddirichlet} gives the density. \code{rdirichlet} returns a matrix with \code{n} rows, each containing a single Dirichlet random deviate.
+#ddirichlet gives the density. rdirichlet returns a matrix with n rows, each containing a single Dirichlet random deviate.
 #AUTHOR
-#Code is adapted from \code{MCMCpack}. It originates from Greg's Miscellaneous Functions (gregmisc).
+#Code is adapted from \pkg{MCMCpack}. It originates from Greg's Miscellaneous Functions (gregmisc).
 #SEE ALSO
 #\code{\link{Beta}}
 #EXAMPLE

### 2009-02-06 pkg/mc2d/R/rmultinomial.R

@@ -8,9 +8,9 @@ dmultinomial <- function (x, size = NULL, prob, log = FALSE)
 #{x}<<Vector of length K of integers in 0:size.>>
 #{n}<<Number of random vectors to draw.>>
 #{size}<<A vector of integers, say N, specifying the total number of objects that are put
-#into K boxes in the typical multinomial experiment. For \code{dmultinom}, it defaults to \code{sum(x)}.
-#The first element correspond to the vector \code{prob} or the first row of \code{prob}, ...>>
-#{prob}<<Numeric non-negative vector of length K, or matrix of size \code{(x x K)}
+#into K boxes in the typical multinomial experiment. For dmultinom, it defaults to sum(x).
+#The first element correspond to the vector prob or the first row of prob, ...>>
+#{prob}<<Numeric non-negative vector of length K, or matrix of size (x x K)
 #specifying the probability for the K classes; is internally normalized to sum 1.>>
 #{log}<<Logical; if TRUE, log probabilities are computed.>>
 #EXAMPLE

### 2009-02-06 pkg/mc2d/R/rmultinormal.R

@@ -9,20 +9,20 @@ rmultinormal <- function(n, mean, sigma , method=c("eigen", "svd", "chol"))
 #with varying vectors of means and varying covariance matrixes.
 #INPUTS
 #{n}<<Number of observations.>>
-#{mean}<<Vector of means (if unique for all \code{n}) or array of means (if varying according to \code{n}).>>
-#{sigma}<<Covariance vector corresponding to the coercion of the covariance matrix into a vector (if unique for all \code{n}) 
-#or array of covariance vectors (if varying according to \code{n}).>>
+#{mean}<<Vector of means (if unique for all n) or array of means (if varying according to n).>>
+#{sigma}<<Covariance vector corresponding to the coercion of the covariance matrix into a vector (if unique for all n) 
+#or array of covariance vectors (if varying according to n).>>
 #{method}<<Matrix decomposition used to determine the matrix root of sigma, possible methods are
 #eigenvalue decomposition ("eigen", default), singular value decomposition ("svd"), and Cholesky decomposition ("chol").>>
 #DETAILS
 #\code{rmvnorm(n, m, s)} is equivalent to \code{rmultinormal(n, m, as.vector(s))}.
 #
-#If \code{mean} and/or \code{sigma} is a matrix, 
-#the first random deviate will use the first row of \code{mean} and/or \code{sigma}, the second random
-#deviate will use the second row of \code{mean} and/or \code{sigma}, ...
+#If mean and/or sigma is a matrix, 
+#the first random deviate will use the first row of mean and/or sigma, the second random
+#deviate will use the second row of mean and/or sigma, ...
 #recycling being permitted by raw.
-#If \code{mean} is a vector of length \code{l} or is a matrix with \code{l} columns, \code{sigma}
-#should be a vector of length \code{l^2} or a matrix of number of \code{l^2} columns. 
+#If mean is a vector of length l or is a matrix with l columns, sigma
+#should be a vector of length (l^2) or a matrix of number of (l^2) columns. 
 #NOTE
 #The use of a varying sigma may be very time consumming.
 #EXAMPLE

### 2009-02-06 pkg/mc2d/R/tornado.R

@@ -1,21 +1,23 @@
 #<<BEGIN>>
 tornado <- function(mc,output = length(mc),use = "all.obs",	method=c("spearman","kendall","pearson"),lim=c(0.025,0.975))
 #TITLE Computes Correlation between Inputs and Output in a mc Object (tornado) in the Variability Dimension;
-#DESCRIPTION Provides statistics for a tornado chart. Evaluates correlations between output and inputs of a \code{mc} object.
+#DESCRIPTION Provides statistics for a tornado chart. Evaluates correlations between output and inputs of a mc object.
 #KEYWORDS univar
 #INPUTS
 #{mc}<<a \code{\link{mc}} object or a \code{\link{mccut}} object.>>
-#{x}<<A \code{tornado} object as provided by the \code{tornado} function.>>
+#{x}<<A tornado object as provided by the tornado function.>>
 #[INPUTS]
-#{output}<<(for \code{mc} objects only). The rank or the name of the output to be considered. By default: the last element of the \code{mc}.>>
-#{use}<<(for \code{mc} objects only). An optional character string giving a method for computing covariances in the presence of missing values. This must be (an abbreviation of) one of the strings "all.obs", "complete.obs" or "pairwise.complete.obs" (see \code{\link{cor}}).>>
-#{method}<<(for \code{mc} objects only). A character string indicating which correlation coefficient (or covariance) is to be computed. One of "spearman" (default), "kendall" or "pearson", can be abbreviated (see \code{\link{cor}}). Warning : the default is not the same in \code{\link{cor}}.>>
+#{output}<<(for mc objects only). The rank or the name of the output to be considered. By default: the last element of the mc.>>
+#{use}<<(for mc objects only). An optional character string giving a method for computing covariances in the presence of missing values. 
+#This must be (an abbreviation of) one of the strings "all.obs", "complete.obs" or "pairwise.complete.obs" (see \code{\link{cor}}).>>
+#{method}<<(for mc objects only). A character string indicating which correlation coefficient (or covariance) is to be computed. 
+#One of "spearman" (default), "kendall" or "pearson", can be abbreviated (see \code{\link{cor}}). Warning : the default is not the same in \code{\link{cor}}.>>
 #{lim}<<A vector of quantiles used to compute the credible interval in two-dimensional models.>>
 #{\dots}<<Further arguments to be passed to the final print function.>>
 #DETAILS
[…]

### 2009-02-06 pkg/mc2d/R/tornadounc.mc.R

@@ -5,41 +5,43 @@ tornadounc.mc <- function(mc,output = length(mc), quant=c(0.5,0.75,0.975),use = 
 #TITLE Computes Correlation between Inputs and Output in a mc Object (tornado) in the Uncertainty Dimension
 #KEYWORDS univar
 #NAME tornadounc
-#DESCRIPTION Provides statistics for a tornado chart. Evaluates correlations between output and inputs of a \code{mc} object in the uncertainty dimension.
+#DESCRIPTION Provides statistics for a tornado chart. Evaluates correlations between output and inputs of a mc object in the uncertainty dimension.
 #INPUTS
-#{mc}<<a \code{mc} object.>>
-#{x}<<a \code{tornadounc} object.>>
+#{mc}<<a mc object.>>
+#{x}<<a tornadounc object.>>
 #[INPUTS]
-#{output}<<The rank or the name of the output to be considered. Should be a \code{"VU"} or a \code{"U" type mcnode}. By default: the last element of \code{mc}.>>
+#{output}<<The rank or the name of the output to be considered. Should be a "VU" or a "U" type mcnode. By default: the last element of mc.>>
 #{quant}<<The vector of quantiles used in the variability dimension.>>
-#{use}<<An optional character string giving a method for computing covariances in the presence of missing values. This must be (an abbreviation of) one of the strings "all.obs", "complete.obs" or "pairwise.complete.obs" (see \code{\link{cor}}).>>
-#{method}<<A character string indicating which correlation coefficient (or covariance) is to be computed. One of "spearman" (default), "kendall" or "pearson", can be abbreviated (see \code{\link{cor}}). Warning : "pearson" is the default for \code{\link{cor}}).>>
+#{use}<<An optional character string giving a method for computing covariances in the presence of missing values. 
+#This must be (an abbreviation of) one of the strings "all.obs", "complete.obs" or "pairwise.complete.obs" (see \code{\link{cor}}).>>
+#{method}<<A character string indicating which correlation coefficient (or covariance) is to be computed. One of "spearman" (default),
+# "kendall" or "pearson", can be abbreviated (see \code{\link{cor}}). Warning : "pearson" is the default for \code{\link{cor}}).>>
 #{\dots}<<Further arguments to be passed to the final print function.>>
 #DETAILS
 # The \code{tornadounc.mc} function computes the spearman's rho statistic between 
-#{*}<<values (\code{"U" type mcnode}) or statistics calculated in the variability dimension (\code{"VU" type mcnode}) of inputs and>>
[…]

### 2009-02-06 pkg/mc2d/R/triang.R

@@ -5,8 +5,8 @@ dtriang <- function(x,min=-1,mode=0,max=1,log=FALSE)
 #KEYWORDS distribution
 #DESCRIPTION
 #Density, distribution function, quantile function and random generation
-#for the triangular distribution with minimum equal to \code{min}, mode equal \code{mode}
-#and maximum equal to \code{max}.
+#for the triangular distribution with minimum equal to min, mode equal mode
+#and maximum equal to max.
 #INPUTS
 #{x,q}<<vector of quantiles.>>
 #{p}<<vector of probabilities.>>
@@ -15,11 +15,11 @@ dtriang <- function(x,min=-1,mode=0,max=1,log=FALSE)
 #{min}<<vector of minima.>>
 #{mode}<<vector of modes.>>
 #{max}<<vector of maxima.>>
-#{log, log.p}<<logical; if \code{TRUE}, probabilities \code{p} are given as \code{log(p)}.>>
-#{lower.tail}<<logical; if \code{TRUE} (default), probabilities are \code{P[X <= x]}, otherwise, \code{P[X > x]}.>>
+#{log, log.p}<<logical; if TRUE, probabilities p are given as log(p).>>
+#{lower.tail}<<logical; if TRUE (default), probabilities are P[X <= x], otherwise, P[X > x].>>
 #VALUE
-#\code{dtriang} gives the density, \code{ptriang} gives the distribution function,
-#\code{qtriang} gives the quantile function, and \code{rtriang} generates random deviates.
+#dtriang gives the density, ptriang gives the distribution function,
+#qtriang gives the quantile function, and rtriang generates random deviates.
 #AUTHOR Regis Pouillot
 #EXAMPLE
 #curve(dtriang(x,min=3,mode=5,max=10), from = 2, to = 11)

### 2009-02-06 pkg/mc2d/R/unmc.R

@@ -2,16 +2,16 @@
 unmc <- function(x, drop=TRUE)
 #TITLE Unclasses the mc or the mcnode Object
 #DESCRIPTION
-#Unclasses the \code{mc} object in a list of arrays 
-#or the \code{mcnode} object in an array.  
+#Unclasses the mc object in a list of arrays 
+#or the mcnode object in an array.  
 #KEYWORDS manip
 #INPUTS
-#{x}<<A \code{mc} or a \code{mcnode} object.>>
+#{x}<<A mc or a mcnode object.>>
 #[INPUTS]
 #{drop}<<Should the dimensions of size 1 be dropped (see \code{\link{drop}}).>>
 #VALUE
-#if x is an \code{mc} object: a list of arrays. If \code{drop=TRUE}, a list of vectors, matrixes and arrays.
-#if x is an \code{mcnode} object: an array. If \code{drop=TRUE}, a vector, matrix or array.
+#if x is an mc object: a list of arrays. If drop=TRUE, a list of vectors, matrixes and arrays.
+#if x is an mcnode object: an array. If drop=TRUE, a vector, matrix or array.
 #EXAMPLE
 #data(total)
 ### A vector

### 2009-08-28 pkg/mc2d/inst/doc/docmcEnglish.Rnw
Try to provide a True vignette.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@58 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[10pt,english]{article}
\VignetteIndexEntry
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[letterpaper]{geometry}
\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
\setlength{\parskip}{\medskipamount}
\setlength{\parindent}{0pt}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{setspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\usepackage{noweb}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.

\usepackage{babel}

\begin{document}

\title{The {}``\texttt{mc2d}'' package. }

\author{R. POUILLOT, M.-L. DELIGNETTE-MULLER, D.L. KELLY \& J.-B. DENIS}

\maketitle
\tableofcontents{}

\noindent \begin{center}
\vfill{}

\par\end{center}

This documentation is intended for readers with:
\begin{itemize}
\item A medium level of experience in R. Please refer to the Manual {}``An
Introduction to R'' available with R distribution if needed;
\item Some knowledge about Monte-Carlo simulation (its basic principles
and its utility) and about Quantitative Risk Assessment (QRA). 
\end{itemize}
This documentation will not describe all arguments of the functions.
The definitive reference remains the documentation associated with
the package.

\section{Introduction}

\subsection{What is \texttt{mc2d}?}

{}``\texttt{mc2d}'' means Two-Dimensional Monte-Carlo (\emph{{}``Monte-Carlo
à Deux Dimensions''}). This package :
\begin{itemize}
\item provides additional probability distributions;
\item provides tools to construct One-Dimensional and Two-Dimensional Monte-Carlo
Simulations;
\item provides tools to analyse One-Dimensional and Two-Dimensional Monte-Carlo
Simulations.
\end{itemize}
In a previous version, some tools to fit parametric distributions
to data were included. Because these functions are useful for other
purposes, they have been moved to a separate package called \texttt{fitdistrplus}.
Both the \texttt{mc2d} and the \texttt{fitdistrplus} packages are
available at the URL \url{https://r-forge.r-project.org/projects/riskassessment/}.

\texttt{mc2d} was built for QRA in the Food Safety domain but it can
be used in other QRA domains.

\subsection{What is Two-Dimensional Monte-Carlo Simulation (briefly)?}

The following text and Figure \ref{fig:Mc2d} are adapted from \cite{POUILLOT-2004}
and \cite{POUILLOT-2007} where this method was used. The principal
reference for Two-Dimensional Monte-Carlo simulation remains \cite{CULLEN-FREY-1999}.

According to international recommendations, a QRA should reflect the
{}``variability'' in the risk and calculate the {}``uncertainty''
associated with the risk estimate. The {}``variability'' represents
temporal, geographical and/or individual heterogeneity of the risk
for a given population. The {}``uncertainty'' is understood as stemming
from a lack of perfect knowledge about the QRA model structure and
associated parameters%
\footnote{In the engineering risk community, these concepts are refered as \textquotedbl{}aleatoric
uncertainty\textquotedbl{} for \textquotedbl{}variability\textquotedbl{}
and \textquotedbl{}epistemic uncertainty\textquotedbl{} for \textquotedbl{}uncertainty\textquotedbl{}.%
}.

In order to estimate the natural {}``variability'' of the risk,
a Monte-Carlo simulation approach may be useful: the empirical distribution
of the risk within the population may be estimated from the mathematical
combination of distributions reflecting the variability of parameters
across the population.

A two-dimensional (or second-order) Monte-Carlo simulation was proposed
to estimate the {}``uncertainty'' in the risk estimates stemming
from parameter uncertainty \cite{CULLEN-FREY-1999}. A two-dimensional
Monte-Carlo simulation is a Monte-Carlo simulation where the distributions
reflecting \textquotedbl{}variability\textquotedbl{} and the distributions
representing \textquotedbl{}uncertainty\textquotedbl{} are sampled
separately in the simulation, so that \textquotedbl{}variability\textquotedbl{}
and \textquotedbl{}uncertainty\textquotedbl{} in the output may be
estinated separately. It may be described as following (see Figure
\ref{fig:Mc2d}): 
\begin{enumerate}
\item The parameters of the model should be divided into three categories:
the parameters whose distributions reflect {}``variability only'',
hereinafter denoted as {}``variable parameters'', the parameters
whose distributions reflect {}``uncertainty only'', denoted as {}``uncertain
parameters'' and the parameters whose distributions reflect both
uncertainty and variability. For this latter category, a hierarchical
structure, using {}``hyper-parameters'', should be specified: if
a parameter is both uncertain and variable, one should be able to
specify an empirical or parametric distribution representing variability.
This distribution is conditional upon other parameters for which there
is some associated uncertainty. As an example, one should be able
to specify a relationship such as $X\left|a,b\right.\sim N\left(a,b\right)$,
where the specified normal distribution represents variability in
$x$ conditional upon parameters $a$ and $b$. Hyperdistributions,
such as $a\sim Unif\left(l_{a},u_{a}\right)$and $b\sim Unif\left(l_{b},u_{b}\right)$,
represent the uncertainty in the parameters $a$ and $b$;
\item A set of uncertain parameters are randomly sampled from their respective
distributions; 
\item The QRA is performed using a classical (one-dimensional) Monte-Carlo
simulation of size $N_{v}$, treating the uncertain parameters as
fixed. This QRA takes into account the variability in all variable
parameters, and leads to an empirical density function reflecting
the variability of exposure/risk across the population, conditional
upon the uncertain parameters. Various statistics (\emph{e.g.} the
mean, the standard deviation, percentiles) of the resulting empirical
density function are evaluated and stored; 
\item Steps 2) and 3) are performed a large number ($N_{u}$) of times ,
leading to $N_{u}$ sets of statistics; 
\item As output, the $50^{\text{th}}$ percentile (median) of each statistic
is used as a point estimate of this statistic; the $2.5^{\text{th}}$
and $97.5^{\text{th}}$ percentiles of each statistic are used to
establish a 95\% credible interval (CI95) of this statistic. The median
of the $N_{u}$ estimated values for each of the 101 estimated percentiles
allows us to display a {}``variability cumulative distribution''
via a graph. This curve is surrounded by the 2.5th and 97.5th percentiles
obtained from the $N_{u}$ estimates of each of the 101 percentiles.
\end{enumerate}
%
\begin{figure}
\noindent \begin{centering}
\includegraphics[scale=0.5]{mc2dSaumon}
\par\end{centering}

\caption{\label{fig:Mc2d}Shematic Representation of a Two-Dimensional Monte-Carlo
Simulation.}

\end{figure}

\textquotedbl{}\texttt{mc2d}\textquotedbl{} is a set of R functions
that will help to develop such two-dimensional Monte-Carlo simulations.
The main difference from the procedure described above is that \texttt{mc2d}
uses arrays of (at least) two dimensions to derive the results: the
first dimension will reflect variability, the second will reflect
uncertainty. This document will not develop the method further, but
will illustrate the practical application of \texttt{mc2d}, using
a fictitious example.

\subsection{\label{sec:Exemple1}A basic example}

\textbf{Quantitative Risk Assessment:} \textbf{\emph{Escherichia coli}}
\textbf{O157:H7 infection linked to the consumption of frozen ground
beef in <3 year old children.}
\begin{itemize}
\item We assume that, in a given batch of ground beef, \emph{E. coli} O157:H7
are randomly distributed with a mean concentration of $c=10$ bacteria
(cfu) per gram of product;
\item We assume that no bacterial growth occurs in storage, since the product
is kept frozen until it is cooked, just before consumption;
\item 2.7\% of consumers cook their beef {}``rare'', 37.3\% {}``medium''
and 60.0\% {}``well done''; 
\item The following bacterial inactivation $i$ is associated with these
cooking practices:

\begin{itemize}
\item No inactivation for {}``rare'' cooking;
\item 1/5 surviving bacteria for a {}``medium'' cooking;
\item 1/50 surviving bacteria for a {}``well done'' cooking.
\end{itemize}
\item The variability in steak serving sizes $s$ for <3 year children was
estimated in a consumption survey. The {}``best fit'' to the data
was a gamma distribution with parameters: $shape=3.93,\: rate=0.0806$.
\item The dose-response relationship, describing the probability of illness,
$P,$ according to the dose is a one-hit model. The probability of
illness per hit $r$ is assumed to be constant with $r=0.001$.
\end{itemize}
The question is: {}``\emph{What is the distribution of the risk of
illness in the population that consumed the contaminated lot?}''

This distribution will be estimated using Monte-Carlo simulations
performed with R via the {}``\texttt{mc2d}'' package. First, the
model will be developed in a one dimensional framework. Then, in order
to include some uncertainties in the model, it will be derived in
a two dimensional framework.

\subsubsection{One Dimensional Monte-Carlo Simulation}

As a first step, we assume that no uncertainty exists in our model.
All distributions represent variability only. The model may be written
as:\begin{eqnarray*}
c & = & 10.\\
i & \sim & emp\left(\left\{ 1,1/5,1/50\right\} ,\left\{ 0.027,0.373,0.600\right\} \right)\\
s & \sim & gamma\left(3.93,0.0806\right)\\
n & \sim & Poisson\left(c\times i\times s\right)\\
P & = & 1-\left(1-0.001\right)^{n}\end{eqnarray*}

where $emp\left(\boldsymbol{X},\boldsymbol{P}\right)$ is an empirical
distribution wherein each value $X_{i}$ is associated with a probability
$P_{i}$.We will use a \textquotedbl{}classical\textquotedbl{} one
dimensional Monte-Carlo simulation, with 1,000 iterations. Using the
{}``\texttt{mc2d}'' package, the model may be written as:

<<sh0, echo=false>>=
set.seed(666)
@

<<sh1, echo=true>>=
library(mc2d)
ndvar(1000)
conc <- 10
cook <- mcstoc(rempiricalD, values=c(1,1/5,1/50), prob=c(0.027,0.373,0.600))
serving <- mcstoc(rgamma,shape=3.93,rate=0.0806)
expo <- conc * cook * serving
dose <- mcstoc(rpois,lambda=expo)
r <- 0.001
risk <- 1-(1-r)^dose
EC1 <- mc(cook,serving,expo,dose,risk)
print(EC1)
summary(EC1)
@

This One-Dimensional Monte-Carlo simulation provides an estimate of
the mean risk (approximately 5\%), as well as some quantiles of the
risk distribution (2.5\% of the population has a risk of illness greater
than 20.3\%).

\subsubsection{Two dimensional Monte-Carlo Simulation}

Assume now that:
\begin{itemize}
\item The mean concentration of bacteria in the batch is not known with
certainty, but was only a point estimate. Microbiologists think that
the uncertainty around this estimate can be represented via a normal
distribution with parameters $\mu=10$ and $\sigma=2$;
\item Epidemiological studies suggest that the $r$ parameter is also uncertain.
The uncertainty around the mean value of 0.001 can be represented
with a uniform distribution between 0.0005 and 0.0015. 
\end{itemize}
The model could then be written as:\begin{eqnarray*}
c & \sim & N\left(10,2\right)\\
i & \sim & emp\left(\left\{ 1,1/5,1/50\right\} ,\left\{ 0.027,0.373,0.600\right\} \right)\\
s & \sim & gamma\left(3.93,0.0806\right)\\
n & \sim & Poisson\left(c\times i\times s\right)\\
r & \sim & Unif\left(0.0005,0.0015\right)\\
P & = & 1-\left(1-r\right)^{n}\end{eqnarray*}

Note that the distributions of $r$ and $c$ represent uncertainty,
while the distributions of $i$ and $s$ represent variability. $n$,
which is a function of $c$, $i$ and $s,$ will be both variable
\emph{and} uncertain.

We will use a two-dimensional Monte-Carlo simulation, with 1,000 iterations
in the variability dimension and 100 iterations in the uncertainty
dimension. Using the {}``\texttt{mc2d}'' package, the model may
be written as:

<<sh2, echo=true>>=
ndunc(100)
conc <- mcstoc(rnorm,type="U",mean=10,sd=2)
cook <- mcstoc(rempiricalD, type="V",values=c(1,1/5,1/50), prob=c(0.027,0.373,0.600))
serving <- mcstoc(rgamma,type="V",shape=3.93,rate=0.0806)
expo <- conc * cook * serving
dose <- mcstoc(rpois,type="VU",lambda=expo)
r <- mcstoc(runif,type="U",min=0.0005,max=0.0015)
risk <- 1-(1-r)^dose
EC2 <- mc(conc,cook,serving,expo,dose,r,risk)
print(EC2,digits=2)
summary(EC2)
@

Note that the syntax is similar to the earlier model. However, a {}``\texttt{type}''
argument is provided for each distribution, indicating whether the
parameter distribution represents uncertainty (\texttt{type}=\textquotedbl{}\texttt{U}\textquotedbl{}),
variability (\texttt{type}=\textquotedbl{}V\textquotedbl{}), or both
(\texttt{type}=\textquotedbl{}\texttt{VU}\textquotedbl{}).

The summary provides estimates of the variability distributions (in
rows) but with a measure of their uncertainty, linked to the uncertainty
around \texttt{conc} and \texttt{r}. The estimate of the mean risk
is now uncertain. The median of the 100 simulations leads to a \textquotedbl{}best
estimate\textquotedbl{}of 0.0445, with a 95\% \textquotedbl{}credible
interval\textquotedbl{} of {[}0.191, 0.0730{]}.

\section{Basic Principles and Functions}

A typical session of R using {}``\texttt{mc2d}'' is as follows:
\begin{itemize}
\item From data, expert knowledge, \emph{etc.} an empirical or parametric
distribution is chosen for each {}``parent'' parameter. For developing
an empirical distribution from data , the {}``\texttt{fitdistrplus}''
package is recommended;
\item For each parameter, an \texttt{mcnode} object is constructed (key
functions: \texttt{mcdata, mcstoc}); 
\item Various \texttt{mcnode} objects are grouped into an \texttt{mc} object
(key function: \texttt{mc}).
\item The \texttt{mc} object is studied through summaries, graphs, and sensitivity
analysis (key functions: \texttt{summary.mc}, \texttt{plot.mc}, \texttt{tornado},
\texttt{tornadounc}).
\end{itemize}

\subsection{Preliminary Step}

The {}``\texttt{mc2d}'' library should be loaded at the beginning
of your R session (\textquotedbl{}\texttt{library(mc2d)}\textquotedbl{}).

The default size of the Monte-Carlo Simulation should be defined using
the \texttt{ndvar()} function (dimension of variability) and the \texttt{ndunc()}
function (dimension of uncertainty).

\subsection{The \texttt{mcnode} Object as an Elementary Object.}

\subsubsection{\texttt{mcnode} Object Structure}

An \texttt{mcnode} object is the basic element of an \texttt{mc} object.
It is an array of dimension $\left(nsv\times nsu\times nvariates\right)$
where $nsv$ is the dimension of variability, $nsu$ is the dimension
of uncertainty and $nvariates$ is the number of variates of the \texttt{mcnode}%
\footnote{In this section, we will only consider \texttt{mcnode}s with $nvariates=1$.%
}. Four types of \texttt{mcnode} exist: 
\begin{itemize}
\item \textquotedbl{}\texttt{V}\textquotedbl{} \texttt{mcnode}, for \textquotedbl{}\emph{Variability}\textquotedbl{},
is an array of dimension $\left(nsv\times1\times nvariates\right)$.
The distribution represents variability in the parameter; 
\item \textquotedbl{}\texttt{U}\textquotedbl{} \texttt{mcnode}, for \textquotedbl{}\emph{Uncertainty}\textquotedbl{},
is an array of dimension $\left(1\times nsu\times nvariates\right)$.
The distribution represents uncertainty in the parameter. 
\item \textquotedbl{}\texttt{VU}\textquotedbl{} \texttt{mcnode}, for \textquotedbl{}\emph{Variability
and Uncertainty}\textquotedbl{}, is an array of dimension $\left(nsv\times nsu\times nvariates\right)$.
The distribution represents both variability (in the first dimension)
and uncertainty (in the second dimension) in the parameter. 
\item Additionally, a \textquotedbl{}\texttt{0}\textquotedbl{} \texttt{mcnode}
is also defined . \textquotedbl{}\texttt{0}\textquotedbl{} stands
for \textquotedbl{}Neither Variability or Uncertainty\textquotedbl{}.
Such nodes are arrays of dimension $\left(1\times1\times nvariates\right)$.
No uncertainty or variability is considered for these nodes. A \textquotedbl{}\texttt{0}\textquotedbl{}
\texttt{mcnode} is not necessary in the univariate context (use a
scalar instead) but is useful in constructing multivariate nodes (See
section \ref{sec:Multivar}). 
\end{itemize}
%
\begin{figure}
\noindent \begin{centering}
\includegraphics[scale=0.4]{Illustration}
\par\end{centering}

\caption{\label{fig:Structure}Structure of the various \texttt{mcnode} objects.}

\end{figure}

There are 5 ways to construct an \texttt{mcnode} object:
\begin{enumerate}
\item The \texttt{mcstoc} function constructs an \texttt{mcnode} from random
number generating functions;
\item The \texttt{mcdata} function constructs an \texttt{mcnode} from data
sets; 
\item \texttt{An mcnode} can be constructed directly from operations on
\texttt{mcnode} objects;
\item \texttt{mcprobtree} is a special function that constructs an \texttt{mcnode}
from other \texttt{mcnodes} using a probability tree;
\item Some functions, such as {}``\texttt{==}'' or {}``\texttt{>}''
, \texttt{is.na}, \texttt{is.finite} generate a new \texttt{mcnode}
when applied to an existing \texttt{mcnode}.
\end{enumerate}

\subsubsection{\label{sub:mcstoc}The \texttt{\textmd{mcstoc}} function}

The \texttt{mcstoc} function is written as%
\footnote{as is standard in R, most arguments have logical default values and
will be infrequently modified.%
}:

\begin{singlespace}
\noindent \begin{flushleft}
\texttt{mcstoc(func=runif, type=c(\textquotedbl{}V\textquotedbl{},
\textquotedbl{}U\textquotedbl{}, \textquotedbl{}VU\textquotedbl{},
\textquotedbl{}0\textquotedbl{}), ..., nsv=ndvar(), nsu=ndunc(), nvariates=1,
outm=\textquotedbl{}each\textquotedbl{}, nsample=\textquotedbl{}n\textquotedbl{},
seed=NULL, rtrunc=FALSE, linf=-Inf, lsup=Inf, lhs=FALSE)}
\par\end{flushleft}
\end{singlespace}
\begin{itemize}
\item \texttt{func} is a function providing random data or its name as a
character. The table \ref{tab:Distributions} provides available distributions
from the \texttt{stats} and the \texttt{mc2d} libraries that can be
used in \texttt{mcstoc}; 
\item \texttt{type} is the type of requested \texttt{mcnode}. By default,
\texttt{mcstoc} constructs a {}``\texttt{V}'' \texttt{mcnode};
\item \ldots{} are the arguments to be passed to the function \texttt{func},
with the exception of the argument providing the size of the sample.
This latter is calculated by the function according to \texttt{func},
\texttt{type}, \texttt{nsv,} \texttt{nsu} and \texttt{nvariates}.
If the name of the argument specifying the size of the sample is not
\texttt{n} (\emph{e.g.} functions \texttt{rhyper} and \texttt{rwilcox},
see table \ref{tab:Distributions}), the name of this parameter should
be provided in the \texttt{nsample} argument. \emph{Note that all
of the following arguments should be named};
\item \texttt{nsv} and \texttt{nsu} are the number of samples needed in
the variability and uncertainty dimension, respectively. By default,
these values are the ones provided by \texttt{ndvar()} and \texttt{ndunc()},
respectively;
\item \texttt{nvariates} is the desired number of variates in the \texttt{mcnode};
\item \texttt{outm} is the default output for multivariate nodes;
\item \texttt{seed} optionally specifies a seed for the random number generator;
\item \texttt{rtrunc} allows truncation of a distribution between \texttt{linf}
and \texttt{lsup}. This function is not valid for every distribution
(see table \ref{tab:Distributions}). See the \texttt{rtrunc} function
help for further details;
\item \texttt{lhs} allows Latin hypercube sampling of the node . This function
is not valid for every distribution (see table \ref{tab:Distributions}).
See the \texttt{lhs} function help for further details. 
\end{itemize}
%
\begin{table}
\caption{\label{tab:Distributions}Available distributions}

\centering{}\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
Package & Distribution & function & \multicolumn{1}{c|}{Parameter \texttt{n}} & Other Parameters & trunc & lhs\tabularnewline
\hline
\hline 
\texttt{stats} & beta & \texttt{rbeta} & \texttt{n} & \texttt{shape1, shape2, ncp} & Y & Y\tabularnewline
\hline
 & binomial & \texttt{rbinom} & \texttt{n} & \texttt{size, prob} & Y & Y\tabularnewline
\hline 
 & Cauchy & \texttt{rcauchy} & \texttt{n} & \texttt{location, scale} & Y & Y\tabularnewline
\hline 
 & chi-squared & \texttt{rchisq} & \texttt{n} & \texttt{df, ncp} & Y & Y\tabularnewline
\hline 
 & exponential & \texttt{rexp} & \texttt{n} & \texttt{rate} & Y & Y\tabularnewline
\hline 
 & F & \texttt{rf} & \texttt{n} & \texttt{df1, df2, ncp} & Y & Y\tabularnewline
\hline 
 & gamma & \texttt{rgamma} & \texttt{n} & \texttt{shape, rate (or scale)} & Y & Y\tabularnewline
\hline 
 & geometric & \texttt{rgeom} & \texttt{n} & \texttt{prob} & Y & Y\tabularnewline
\hline 
 & hypergeometric & \texttt{rhyper} & \texttt{nn} & \texttt{m, n, k} & Y & Y\tabularnewline
\hline 
 & lognormal & \texttt{rlnorm} & \texttt{n} & \texttt{meanlog, sdlog} & Y & Y\tabularnewline
\hline 
 & logistic & \texttt{rlogis} & \texttt{n} & \texttt{location, scale} & Y & Y\tabularnewline
\hline 
 & negative binomial & \texttt{rnbinom} & \texttt{n} & \texttt{size, prob (or mu)} & Y & Y\tabularnewline
\hline 
 & normal & \texttt{rnorm} & \texttt{n} & \texttt{mean, sd} & Y & Y\tabularnewline
\hline 
 & Poisson & \texttt{rpois} & \texttt{n} & \texttt{lambda} & Y & Y\tabularnewline
\hline 
 & Student's t & \texttt{rt} & \texttt{n} & \texttt{df, ncp} & Y & Y\tabularnewline
\hline 
 & uniform & \texttt{runif} & \texttt{n} & \texttt{min, max} & Y & Y\tabularnewline
\hline 
 & Weibull & \texttt{rweibull} & \texttt{n} & \texttt{shape, scale} & Y & Y\tabularnewline
\hline
 & Wilcoxon & \texttt{rwilcox} & \texttt{nn} & \texttt{m,n} & Y & Y\tabularnewline
\hline 
\texttt{mc2d} & Bernoulli & \texttt{rbern} & \texttt{n} & \texttt{prob} & Y & Y\tabularnewline
\hline 
 & empirical discrete & \texttt{rempiricalD} & \texttt{n} & \texttt{values, prob} & Y & Y\tabularnewline
\hline 
 & empirical continuous & \texttt{rempiricalC} & n & \texttt{min, max, values, prob} & Y & Y\tabularnewline
\hline
 & PERT & \texttt{rpert} & \texttt{n} & \texttt{min, mode, max, shape} & Y & Y\tabularnewline
\hline
 & triangular & \texttt{rtriang} & \texttt{n} & \texttt{min, mode, max} & Y & Y\tabularnewline
\hline
 & generalised beta & \texttt{rbetagen} & \texttt{n} & \texttt{shape1,shape2,min,max,ncp} & Y & Y\tabularnewline
\hline
 & multinomial & \texttt{rmultinomial} & \texttt{n} & \texttt{n, size, prob} & N & N\tabularnewline
\hline
 & Dirichlet & \texttt{rdirichlet} & \texttt{n} & \texttt{alpha} & N & N\tabularnewline
\hline
 & multivariate normal & \texttt{rmultinormal} & \texttt{n} & \texttt{mean, sigma} & N & N\tabularnewline
\hline
\end{tabular}
\end{table}

In our basic example, \texttt{mcstoc} was used to specify \texttt{conc}
(a normal distribution), \texttt{cook} (an empirical discrete distribution),
\texttt{serving} (a gamma distribution), and \texttt{dose} (a Poisson
distribution). Note that the argument \texttt{lambda} of the Poisson
distribution (node \texttt{dose}) is itself an \texttt{mcnode}.

<<sh3, echo=true, eval=FALSE>>=
conc <- mcstoc(rnorm,type="U",mean=10,sd=2)
cook <- mcstoc(rempiricalD, type="V",values=c(1,1/5,1/50), prob=c(0.027,0.373,0.600))
serving <- mcstoc(rgamma,type="V",shape=3.93,rate=0.0806)
...
dose <- mcstoc(rpois,type="VU",lambda=expo)
r <- mcstoc(runif,type="U",min=0.0005,max=0.0015)
...
@

A normal distribution with parameters $mean=2$, $sd=3$, truncated
on the interval {[}1.5, 2{]}, with samples generated via Latin hypercube
sampling could be written%
\footnote{Note that the mean and the standard deviation of the untruncatednormal
distribution are not preserved in the truncated distribution.%
}:

<<sh3bis, echo=true, eval=true>>=
x <- mcstoc(rnorm, mean=2, sd=3, rtrunc=TRUE, linf=1.5, lsup=2, lhs=TRUE)
summary(x)
@

For convenience in using \texttt{mcstoc}, the following additional
distributions have been implemented: the Bernoulli distribution (\texttt{rbern}),
the empirical discrete distribution (\texttt{rempiricalD}), the PERT
distribution (\texttt{rpert})\cite{VOSE-2000}, the triangular distribution
(\texttt{rtriang}), the Dirichlet distribution (\texttt{rdirichlet})
and the multivariate normal distribution (\texttt{rmultinormal}).
The multinomial distribution has been adapted (vectorized): \texttt{rmultinomial}
(library \texttt{mc2d}) should be used in place of \texttt{rmultinom}
(library \texttt{stats}). The empirical discrete (\emph{e.g.} for
bootstrap), the Dirichlet, the multinomial and the multivariate normal
may be used with uncertain and/or variable parameters by specifying
multivariate nodes. See section \ref{sec:Multivar}.

\subsubsection{\label{sub:mcnode}The \texttt{\textmd{mcdata}} function}

Another way to construct an \texttt{mcnode} object is \emph{via} the
\texttt{mcdata} function, when data are available.

\begin{singlespace}
\noindent \begin{flushleft}
\texttt{mcdata(data, type=c(\textquotedbl{}V\textquotedbl{}, \textquotedbl{}U\textquotedbl{},
\textquotedbl{}VU\textquotedbl{}, \textquotedbl{}0\textquotedbl{}),
nsv=ndvar(), nsu=ndunc(), nvariates=1, outm=\textquotedbl{}each\textquotedbl{})}
\par\end{flushleft}
\end{singlespace}

\noindent \begin{flushleft}
See the documentation associated with this function to see the size/type
of data that can be used to construct an \texttt{mcnode}. The following
example places a \texttt{TRUE} value in a \textquotedbl{}\texttt{U}\textquotedbl{}
node in half of the simulations:
\par\end{flushleft}

<<sh3, echo=true>>=
nu <- ndunc()
tmp <- (1:nu) > (nu/2)
mcdata(tmp,type="U")
@

\subsubsection{Operations on an \texttt{mcnode}}

\texttt{mcnode}s can be automatically constructed using operations
on other \texttt{mcnode}s. Rules are used to transfer uncertainty
and variability coherently within the model. Logically, the rules
are as follows (illustrated here with a {}``+'')%
\footnote{These rules are not the standard R rules for recycling.%
}:
\begin{itemize}
\item \textquotedbl{}\texttt{0}\textquotedbl{} + \textquotedbl{}\texttt{0}\textquotedbl{}
= \textquotedbl{}\texttt{0}\textquotedbl{};
\item \textquotedbl{}\texttt{0}\textquotedbl{} + \textquotedbl{}\texttt{V}\textquotedbl{}
= \textquotedbl{}\texttt{V}\textquotedbl{}
\item \textquotedbl{}\texttt{0}\textquotedbl{} + \textquotedbl{}\texttt{U}\textquotedbl{}
= \textquotedbl{}\texttt{U}\textquotedbl{};
\item \textquotedbl{}\texttt{0}\textquotedbl{} + \textquotedbl{}\texttt{VU}\textquotedbl{}
= \textquotedbl{}\texttt{VU}\textquotedbl{};
\item \textquotedbl{}\texttt{V}\textquotedbl{} + \textquotedbl{}\texttt{V}\textquotedbl{}
= \textquotedbl{}\texttt{V}\textquotedbl{};
\item \textquotedbl{}\texttt{V}\textquotedbl{} + \textquotedbl{}\texttt{U}\textquotedbl{}
= \textquotedbl{}\texttt{VU}\textquotedbl{}: the {}``\texttt{U}''
mcnode is recycled by row, the \textquotedbl{}\texttt{V}\textquotedbl{}
mcnode is recycled in the standard manner by column;
\item \textquotedbl{}\texttt{V}\textquotedbl{} + \textquotedbl{}\texttt{VU}\textquotedbl{}
= \textquotedbl{}\texttt{VU}\textquotedbl{}: the {}``\texttt{V}''
mcnode is recycled in the standard manner by column;
\item \textquotedbl{}\texttt{U}\textquotedbl{} + \textquotedbl{}\texttt{U}\textquotedbl{}
= \textquotedbl{}\texttt{U}\textquotedbl{};
\item \textquotedbl{}\texttt{U}\textquotedbl{} + \textquotedbl{}\texttt{VU}\textquotedbl{}
= \textquotedbl{}\texttt{VU}\textquotedbl{}: the {}``\texttt{U}''
mcnode is recycled by row;
\item \textquotedbl{}\texttt{VU}\textquotedbl{} + \textquotedbl{}\texttt{VU}\textquotedbl{}
= \textquotedbl{}\texttt{VU}\textquotedbl{}
\end{itemize}
Thus, in our example:

<<sh4, echo=true, eval=false>>=
...
expo <- conc * cook * serving
...
risk <- 1-(1-r)^dose
@

\texttt{expo} is a function of a {}``\texttt{U}'' and two {}``\texttt{V}''
\texttt{mcnode}s: it is a {}``\texttt{VU}'' \texttt{mcnode} with
variability in the row dimension and uncertainty in the column dimension
. \texttt{risk} is a function of a {}``\texttt{U}'' and a {}``\texttt{VU}''
\texttt{mcnode}: it is therefore a {}``\texttt{VU}'' \texttt{mcnode}.

\subsubsection{\label{sub:The-mcprobtree-function}The \texttt{\textmd{mcprobtree}}
function}

The \texttt{mcprobtree} function can be used if a {}``probability
tree'' is needed to construct an \texttt{mcnode}. Assume that the
distribution representing the uncertainty on \texttt{conc} was not
itself certain, and that the microbiologists suggest that they are
75\% sure that $conc\sim N\left(10,2\right)$ but that they are 25\%
sure that $conc\sim U\left(8,12\right)$. This could be written using
\texttt{mcprobtree} as%
\footnote{two alternatives for \texttt{whichdist} are \texttt{whichdist <- mcstoc(rempiricalD,
type=\textquotedbl{}U\textquotedbl{}, values=c(0,1), prob=c(75,25))}
or \texttt{whichdist <- mcstoc(rbern,type=\textquotedbl{}U\textquotedbl{},prob=0.25)}%
}:

<<sh4, echo=true, eval=true>>=
conc1 <- mcstoc(rnorm,type="U",mean=10,sd=2)
conc2 <- mcstoc(runif,type="U",min=8,max=12)
whichdist <- c(0.75,0.25)
concbis <- mcprobtree(whichdist,list("0"=conc1,"1"=conc2),type="U")
@

\texttt{mcprobtree} can also be used to generate samples from a mixture
distribution for variability .

\subsubsection{Other functions for constructing an \texttt{mcnode}}

The functions {}``\texttt{==}'', ''\texttt{<}'', ''\texttt{<}='',
''\texttt{>=}'', {}``\texttt{>}'', generate an \texttt{mcnode}
when applied to another \texttt{mcnode}.

Special functions \texttt{is.na(x)}, \texttt{is.nan(x)}, \texttt{is.finite(x)},
\texttt{is.infinite(x)} are implemented to test if any values are
\texttt{NA} (missing data), \texttt{NaN} ({}``\emph{Not A Number}''),
or finite .

<<sh5, echo=true, eval=true>>=
cook < 1
tmp <- log(mcstoc(runif,min=-1,max=1))
tmp
is.na(tmp)
@

\subsubsection{Specifying a correlation between \texttt{mcnodes}}

Structural links between sets of parameters may be very important
in QRA. In \texttt{mc2d}, a Spearman rank correlation structure for
2 or more nodes may be specified with the \texttt{cornode} function.
This function uses the method of Iman \& Conover to generate correlated
samples \cite{IMAN-CONOVER-1982}. Assume that a study suggests that
people who consume rare ground beef alsoconsume larger serving sizes.
We could specify this relation using:

<<sh5bis, echo=true, eval=true>>=
cornode(cook,serving,target=0.5,result=TRUE) 
@

Note that the resulting correlation (around 0.4) is obviously an approximation
to the desired value of 0.5, because a discrete distribution (\texttt{cook}:
3 categories) is correlated with a continuous distribution (\texttt{serving}).

It is possible to create such correlations between {}``V'' nodes,
between {}`` U'' nodes, between {}``VU'' nodes, or between one
{}``V'' node and multiple {}``VU'' nodes.

The use of a multivariate normal distribution (\texttt{rmultinormal})
is another way to specify correlations among nodes, assuming that
the individual nodes are normally distributed.

\subsection{The \texttt{mc} Object}

Once the \texttt{mcnode} objects are constructed, one should group
them into a single object in order to analyse the Monte-Carlo results.
The {}``\texttt{mc}'' object is a list of \texttt{mcnode}s. There
are three ways to construct an \texttt{mc} object: using the \texttt{mc}
function, using the \texttt{evalmcmod} function, or within the \texttt{evalmccut}
function.

\subsubsection{The \texttt{mc} Function}

\noindent \texttt{mc(..., name=NULL, devname=FALSE)}

\ldots{} are \texttt{mcnodes} or \texttt{mc} objects to be gathered
into an \texttt{mc} object. \texttt{mc} value is an \texttt{mc} object
with specific methods, e.g. \texttt{print} or \texttt{summary}. In
our example, we used:

<<sh6, echo=true, eval=false>>=
...
EC2 <- mc(conc,cook,serving,expo,dose,r,risk)
print(EC2)
summary(EC2)
@

\subsubsection{The \texttt{mcmodel} and the \texttt{evalmcmod} Functions}

A model may be written in one step using \texttt{mcmodel} (just a
wrapper of your model in a function), and then evaluated using \texttt{evalmcmod}.
These functions may be used once your model is correct and has been
tested using a small number of iterations. For our example:

<<sh10, echo=true, eval=true>>=
modelEC3 <- mcmodel({
  conc <- mcstoc(rnorm,type="U",mean=10,sd=2)
  cook <- mcstoc(rempiricalD, type="V",values=c(1,1/5,1/50),
    prob=c(0.027,0.373,0.600))
  serving <- mcstoc(rgamma,type="V",shape=3.93,rate=0.0806)
  r <- mcstoc(runif,type="U",min=0.0005,max=0.0015)
  expo <- conc * cook * serving
  dose <- mcstoc(rpois,type="VU",lambda=expo)
  risk <- 1-(1-r)^dose
  mc(conc,cook,serving,expo,dose,r,risk)
})
modelEC3
@

Note that:
\begin{itemize}
\item the model is wrapped between {}``\texttt{\{}'' and {}``\texttt{\}}'';
\item any (valid) R code may be placed in the model%
\footnote{If needed, it is possible to make reference to the simulation dimensions
using \texttt{ndvar()} and/or \texttt{ndunc()}.%
};
\item The model should end with an \texttt{mc()} function\texttt{.}
\end{itemize}
The model is then evaluated using the \texttt{evalmcmod} function:

\noindent \begin{flushleft}
\texttt{evalmcmod(expr, nsv=ndvar(), nsu=ndunc(), seed=NULL)}
\par\end{flushleft}

One can re-run the model with various dimensions or random seeds in
one line:

<<sh11, echo=true, eval=false>>=
EC3 <- evalmcmod(modelEC3,nsv=100,nsu=10,seed=666)
EC4 <- evalmcmod(modelEC3,nsv=100,nsu=1000,seed=666)
@

\subsubsection{The \texttt{mcmodelcut} and the \texttt{evalmccut} Functions}

If evaluating a high-dimensional model, R may exceed its memory limit.
\texttt{evalmccut} evaluates a 2-dimensional Monte-Carlo model (written
with the \texttt{mcmodelcut} function) using a loop, and calculates
and stores statistics in the uncertainty dimension for further analysis.
Readers should refer to the corresponding documentation for further
details. Our example would be written as:

<<sh12, echo=true, eval=FALSE>>=
modEC4 <- mcmodelcut({
{cook <- mcstoc(rempiricalD, type = "V", values = c(0, 1/5, 1/50), 
                prob = c(0.027, 0.373, 0.6))
 serving <- mcstoc(rgamma, type = "V", shape = 3.93, rate = 0.0806)       
 conc <- mcstoc(rnorm, type = "U", mean = 10, sd = 2)       
 r <- mcstoc(runif, type = "U", min = 5e-04, max = 0.0015)     
}
## Second block 
{expo <- conc * cook * serving       
 dose <- mcstoc(rpois, type = "VU", lambda = expo)
 risk <- 1 - (1 - r)^dose
 res <- mc(zero,conc, cook, serving, expo, dose, r, risk)     }
## Third block 
{list(
  sum = summary(res),
  plot = plot(res, draw=FALSE),
  minmax = lapply(res,range),
  tor=tornado(res),
  et =  sapply(res,sd))
}
})
evalmccut(modEC4, nsv = 10001, nsu = 101, seed = 666)
@

Note that the use of a \texttt{tornado} function in the model should
be avoided as it slows the \texttt{evalmccut} function considerably.
The \texttt{tornado} function will be rewritten in the near future
to improve its performance.

\subsection{Analysing an \texttt{mc} Object}

As a reminder, the \texttt{print} function provides a very basic summary
of the \texttt{mc} object. It has a \texttt{digits} argument (default:
3). Obviously, other more informative functions are provided in the
\texttt{mc2d} package.

\subsubsection{The \texttt{summary} Function}

The \texttt{summary} function provides statistics on an \texttt{mc}
object:

\noindent \begin{flushleft}
\texttt{summary(object, probs=c(0,0.025,0.25,0.5,0.75,0.975,1), lim=c(0.025,0.975),
...)}
\par\end{flushleft}

The mean, the standard deviation and the quantiles provided in the
\texttt{probs} arguments are evaluated on the variability dimension.
Then, the median and the quantiles provided in the \texttt{lim} argument
are evaluated on these statistics. Of course, these arguments should
be changed if other quantiles are needed.

<<sh14, echo=true, eval=true>>=
tmp <- summary(EC2,probs=c(0.995,0.999),digits=12)
tmp$risk
@

\subsubsection{The \texttt{hist} Function}

The \texttt{hist} provides a histogram of the different \texttt{mcnodes}
making up the \texttt{mc} object (cf. Figure \ref{fig:Function-hist}).

\noindent \begin{flushleft}
\texttt{hist (x, griddim = NULL, xlab = names(x), ylab = \textquotedbl{}Frequency\textquotedbl{},
main = \textquotedbl{}\textquotedbl{}, ...)}
\par\end{flushleft}

In the current version, uncertainty and variability distributions
are collapsed. Thus, the resulting histogram may be meaningless.

<<sh15, echo=true, eval=false, fig=false>>=
hist(EC2)
@

%
\begin{figure}
\caption{\label{fig:Function-hist}Function \texttt{hist}.}

<<sh16, echo=false, eval=true, fig=true>>=
hist(EC2)
@
\end{figure}

\subsubsection{The \texttt{plot} function}

The \texttt{plot} function provides a graph of the empirical distribution
function of the estimate (mean or median) of the quantiles.

\noindent \begin{flushleft}
\texttt{plot (x, prec = 0.01, stat = c(\textquotedbl{}median\textquotedbl{},
\textquotedbl{}mean\textquotedbl{}), lim = c(0.025,0.975), na.rm =
TRUE, griddim = NULL, xlab = NULL, ylab = \textquotedbl{}Fn(x)\textquotedbl{},
main = \textquotedbl{}\textquotedbl{}, draw = TRUE, ...)}
\par\end{flushleft}

For our example, see Figure \ref{fig:Function-plot}, a default graph.

<<sh17, echo=true, eval=false, fig=false>>=
plot(EC2)
@

%
\begin{figure}
\caption{\label{fig:Function-plot}\texttt{plot} Function .}

<<sh18, echo=false, eval=true, fig=true>>=
plot(EC2)
@
\end{figure}

Note that \texttt{mcnode} objects have the same methods \texttt{print},
\texttt{summary}, \texttt{plot}, and \texttt{hist}.

\subsubsection{The \texttt{tornado} function}

The \texttt{tornado} function calculates the Spearman (default) rank
correlation between nodes of the \texttt{mc} object.

\noindent \begin{flushleft}
\texttt{tornado(x, output=length(x), use=\textquotedbl{}all.obs\textquotedbl{},
method=c(\textquotedbl{}spearman\textquotedbl{}, \textquotedbl{}kendall\textquotedbl{},\textquotedbl{}pearson\textquotedbl{}),
lim=c(0.025, 0.975))}
\par\end{flushleft}

where \texttt{output} is the \texttt{mcnode} (name or rank) of the
output (default: the last \texttt{mcnode}). Missing data are treated
using the \texttt{use} arguments (see the reference documentation).
\texttt{tornado} creates a \texttt{tornado} object with a \texttt{plot}
method (\emph{cf.} Figure \ref{fig:Function-plottor}).

<<sh19, echo=true, eval=true, fig=false>>=
torEC2 <- tornado(EC2)
plot(torEC2)
@

%
\begin{figure}
\caption{\label{fig:Function-plottor}\texttt{plot}.\texttt{tornado} Function
.}

<<sh19b, echo=false, eval=true, fig=true>>=
plot(torEC2)
@
\end{figure}

\subsubsection{The \texttt{tornadounc} function}

The \texttt{tornadounc} function examines the impact of the uncertainty
on the estimate of an output. It calculates the Spearman (default)
rank correlation between statistics of the \texttt{mc} object in the
variability dimension.

\noindent \begin{flushleft}
\texttt{tornadounc(mc,output = length(mc), quant=c(0.5,0.75,0.975),
use = \textquotedbl{}all.obs\textquotedbl{}, method=c(\textquotedbl{}spearman\textquotedbl{},\textquotedbl{}kendall\textquotedbl{},\textquotedbl{}pearson\textquotedbl{}),
...)}
\par\end{flushleft}

The \texttt{quant} argument indicates which quantiles should be used
in the variability dimension. \texttt{tornadounc} creates a \texttt{tornadounc}
object with a \texttt{plot} method

<<sh19c, echo=true, eval=true, fig=false>>=
tornadounc(EC2, output="risk", quant=.99)
@

The output shows the impact of the uncertain nodes (type \textquotedbl{}U\textquotedbl{}
nodes) and some statistics (mean, median and, here, the $99^{\text{th}}$percentile)
calculated on the variability dimension (type \textquotedbl{}V\textquotedbl{}
and type \textquotedbl{}VU\textquotedbl{} nodes) of some output statistics
.

\subsection{Other Functions and \texttt{mc} Objects}

\texttt{mc} objects are simply lists of three dimensional arrays;
within each array, values in a given column represent variability
in the parameter.

Knowing the structure of the \texttt{mc} and the structure of the
\texttt{mcnode} objects, it is straightforward to apply any R function
to these objects. The \textquotedbl{}\texttt{\$}\textquotedbl{} function
is helpful for extracting an \texttt{mcnode} from an \texttt{mc} object.
The \texttt{unmc} function removes all attributes, classes, and dimensions
equal to one, providing a list of vectors, matrices and/or arrays.

Here is a (silly) example building a linear model (in fact \texttt{ndunc()}
linear models) between the \texttt{risk} and the \texttt{dose} within
each uncertainty dimension and estimating some statistics for the
coefficients. This example is here only to illustrate that the entire
spectrum of R functionality is available for your analysis.

<<sh19ter, echo=true, eval=true>>=
tmp <- unmc(EC2, drop=TRUE)
dimu <- ncol(tmp$risk)
coef <- sapply(1:dimu, function(x) lm(tmp$risk[,x] ~ tmp$dose[,x])$coef) 
apply(coef,1,summary)
@

\section{\label{sec:Multivar}Multivariate Nodes}

The dimension \texttt{nvariates} is the third dimension of the \texttt{mcnode}.
One can ignore it while using \texttt{mc2d} . Nevertheless, its use
is mandatory to handle some multivariate distributions, and it may
be useful in other circumstances. Constructing multivariate nodes
is straightforward. We note that the following code:

<<sh19ter, echo=true, eval=false>>=
mcstoc(runif, nvariates=3, min=c(1,2,3),max=4)
@

will logically not provide a node with 3 variates, each having a different
limit. The recycling rule says that \texttt{c(1, 2, 3)} will be used
in the first dimension, i.e. the variability dimension. Use instead:

<<sh19quatr, echo=true, eval=true>>=
lim <- mcdata(c(1,2,3), type="0", nvariates=3)
mcstoc(runif, nvariates=3, min=lim,max=4)
@

\subsection{Multivariate Nodes for Multivariate Distributions}

The basic usage of multivariate nodes (and the reason why they have
been implemented) is for multivariate distributions such as the Dirichlet
distribution, the multinomial distribution, the multivariate normal
distribution and, possibly, the empirical distribution

As an example, assume that 3-member families buy 500 g of ground beef.
The proportions of steak eaten by the baby, his older brother and
his mother follow a Dirichlet (uncertainty) distribution with (vector)
parameter $\alpha=(2,3,5)$. We want to derive the distribution (variability)
of steak eaten by 500 babies sampled from these 500 families.

<<sh20, echo=true, eval=true>>=
(p <- mcstoc(rdirichlet, type="U", nsu=100, nvariates=3, alpha=c(2,3,5)))
s <- mcstoc(rmultinomial,type="VU", nsv=500, nsu=100, nvariates=3,size=500, prob=p)
summary(s)
@

Assume that each member of these families eats a {}``normal'' distribution
(variability) of steak with mean 100, 150 and 250 g. There is a positive
correlation between the servings of the children, and a negative one
with the serving of the mother. We want to derive the distribution
(variability) of steak eaten by 500 babies.

<<sh21, echo=true, eval=true>>=
(x <- mcstoc(rmultinormal,type="V", nvariates=3, mean=c(100,150,250), 
                              sigma=c(10,2,-5,2,10,-5,-5,-5,10))) 
cor(x[,1,])
@

In this example, \texttt{mean} could be variable or uncertain, as
well as \texttt{sigma}%
\footnote{Caution: the use of a varying \texttt{sigma} can make the analysis
very slow.%
}. You could have used, for an uncertain mean:

<<sh21, echo=true, eval=true>>=
m <- mcdata(c(100,150,250), type="0", nvariates=3)
mun <- mcstoc(rnorm, type="U", nvariates=3, mean=m, sd=20)
x <- mcstoc(rmultinormal, type="VU", nvariates=3, mean=mun, sigma=c(10,2,-5,2,10,-5,-5,-5,10))
cor(x[,1,])
@

The correlation is preserved, but the mean of each category is uncertain.

Finally, multivariate nodes may be useful to derive a nonparametric
bootstrap. Assume that, based on a study, you obtained 6 individuals
who eat 100 g, 12 individuals who eat 150 g, 6 individuals who eat
170 g and 6 individuals who eat 200 g of ground beef. You want to
use a nonparametric bootstrap to derive uncertainty \cite{CULLEN-FREY-1999},
and then select samples from the empirical distribution.

<<sh22, echo=true, eval=true>>=
(x <- mcstoc(rempiricalD, type="U", outm=c('min','mean','max'), nvariates=30, 
              values=c(100,150,170,200),prob=c(6,12,6,6)))
mcstoc(rempiricalD,type="VU", values=x)
@

Printing the statistics of the 30 variates of \texttt{x} is of no
interest. Instead, we use the {}``\texttt{outm}'' option, which
allows us to specify which output we want ({}``\texttt{none}'' for
none, {}``\texttt{each}'', the default, for a series of statistics
for each variate, or, as in the example, a vector of functions that
are applied over all the 30 variates).

\subsection{Multivariate Nodes as a {}``Third Dimension'' for Multiple Options
in a Model}

The recycling rules in \texttt{mc2d} regarding the \texttt{nvariate}
dimension are as follows: the recycling will be done from \texttt{nvariates=1}
to \texttt{nvariates=n} with $n>1$. This allows you to use multivariates
nodes as a third dimension, in case you want to test various alternatives.

Assume, as in section \ref{sub:The-mcprobtree-function}, that the
distribution representing uncertainty in \texttt{conc} was not certain,
and that the microbiologists suggest that $conc\sim N\left(10,2\right)$
is possible, but that $conc\sim U\left(8,12\right)$ is also possible.
We can \emph{i}) build a {}``bivariate'' node reflecting these two
options; ii) transfer these options into the final risk estimate.
We obtain a bivariate node for the risk, one using the first hypothesis,
the second the second hypothesis.

<<sh23, echo=true>>=
conc1 <- mcstoc(rnorm, type="U", mean=10, sd=2)
conc2 <- mcstoc(runif, type="U", min=8, max=12)
conc <- mcdata(c(conc1,conc2),type="U",nvariates=2)

cook <- mcstoc(rempiricalD, type="V", values=c(1,1/5,1/50), prob=c(0.027,0.373,0.600))
serving <- mcstoc(rgamma,type="V",shape=3.93,rate=0.0806)
expo <- conc * cook * serving
dose <- mcstoc(rpois,type="VU",nvariates=2,lambda=expo)
r <- mcstoc(runif,type="U",min=0.0005,max=0.0015)
risk <- 1-(1-r)^dose
EC5 <- mc(conc,cook,serving,expo,dose,r,risk)
summary(EC5)
@

(Do not forget to transfer the number of variates you want in \texttt{mcstoc}...
(see the definition of \texttt{dose}). \texttt{mc2d} cannot guess...)

\subsection{Multivariate Nodes as a {}``Third Dimension'' for Multiple Vectors/Contaminants}

The recycling rules in \texttt{mc2d} also allow you to use multivariate
nodes as a third dimension for multiple vectors/Contaminants.

Assume in our ground beef example that we have two contaminants: one
has a mean concentration that follows an uncertainty distribution
$conc\sim N\left(10,2\right)$, the second one follows $conc\sim N\left(14,2\right)$.
We can \emph{i}) build a {}``bivariate'' node reflecting these two
concentrations%
\footnote{Note that we could simulate a correlation between both contaminants
using a multivariate normal distribution.%
} ; \emph{ii}) transfer these options into the final dose; \emph{iii})
sum the dose over the variates (using \texttt{mcapply}). The behavior
of contaminants is transferred in the model.

<<sh24, echo=true>>=
mconc <- mcdata(c(10,14), type="0", nvariates=2)
conc <- mcstoc(rnorm, nvariates=2, type="U", mean=mconc, sd=2)

cook <- mcstoc(rempiricalD, type="V", values=c(1,1/5,1/50), prob=c(0.027,0.373,0.600))
serving <- mcstoc(rgamma,type="V",shape=3.93,rate=0.0806)
expo <- conc * cook * serving
dose <- mcstoc(rpois,type="VU",nvariates=2,lambda=expo)
dosetot <- mcapply(dose, margin="variates", fun=sum)
r <- mcstoc(runif,type="U",min=0.0005,max=0.0015)
risk <- 1-(1-r)^dosetot
EC6 <- mc(conc,cook,serving,expo,dose,dosetot,r,risk)
summary(EC6)
@

As a conclusion, this \textquotedbl{}third\textquotedbl{} dimension
is highly flexible...

\section{Another Example: A QRA of Waterborne Cryptosporidiosis in France}

This example is adapted from \cite{POUILLOT-2004}. The aim is to
evaluate the risk of infection with \emph{Cryptosporidium} \emph{parvum}
from consumption of tap water, given that $n$ oocysts /100 l. have
been observed in a storage reservoir.

\subsection{Tap Water Consumption Model}

<<Crypto1, echo=false, eval=true>>=
library(mc2d)
inca <- structure(c(0, 22.08, 60, 64.4, 72, 82.8, 90, 96, 100, 110, 120,  137.5, 144, 150, 160, 162.5, 165, 180, 182.5, 184, 192, 192.5,  200, 216, 220, 225, 230, 240, 250, 264, 270, 276, 288, 290, 300,  304, 312.8, 320, 322, 325, 330, 336, 340, 350, 360, 370, 375,  380, 384, 390, 400, 414, 420, 425, 430, 432, 432.5, 440, 442,  450, 456, 460, 460.8, 464, 470, 470.4, 480, 490, 500, 504, 510,  510.4, 516, 520, 525, 525.6, 528, 530, 540, 544, 550, 552, 560,  562, 565, 570, 576, 580, 582.5, 584, 585.6, 590, 596, 600, 606,  610, 614, 620, 625, 630, 635.4, 640, 648, 650, 656.2, 660, 664.4,  670, 670.4, 672, 675, 680, 682, 690, 696, 700, 710, 716, 720,  730, 730.4, 740, 744, 750, 756, 760, 774.8, 780, 784, 792, 796,  800, 810, 820, 828, 830, 840, 850, 850.4, 860, 864, 866.4, 870,  880, 890, 900, 908, 910, 915.2, 920, 930, 936, 950, 960, 970,  980, 984, 986.4, 990, 996, 1000, 1015.2, 1020, 1028, 1032, 1036,  1040, 1042, 1050, 1070, 1075, 1078.8, 1080, 1090, 1096, 1100,  1110, 1120, 1126.4, 1128, 1130, 1140, 1148, 1150, 1152, 1160,  1170, 1175, 1176.2, 1190, 1196, 1200, 1214, 1220, 1230, 1240,  1248, 1250, 1260, 1276, 1280, 1290, 1296, 1300, 1320, 1322, 1330,  1340, 1350, 1360, 1370, 1386.4, 1400, 1410, 1414, 1420, 1430,  1440, 1446, 1450, 1460, 1480, 1500, 1520, 1530, 1550, 1560, 1600,  1650, 1680, 1696, 1700, 1710, 1720, 1750, 1760, 1800, 1830, 1840,  1850, 1900, 1920, 1936, 1954, 1980, 1990, 2000, 2014, 2050, 2100,  2200, 2220, 2248, 2250, 2276, 2300, 2310, 2340, 2400, 2550, 2568,  2700, 2720, 2784, 2820, 2876, 3000, 3100, 3108, 3200, 2578, 7,  1, 8, 14, 3, 1, 1, 10, 1, 250, 1, 2, 120, 8, 6, 1, 5, 3, 12,  5, 5, 375, 2, 8, 7, 41, 408, 53, 4, 24, 7, 3, 2, 217, 1, 1, 44,  9, 1, 31, 1, 1, 17, 294, 5, 3, 9, 3, 12, 525, 5, 23, 1, 3, 4,  1, 28, 3, 154, 2, 5, 1, 2, 6, 1, 299, 8, 148, 1, 2, 1, 1, 8,  3, 1, 2, 14, 20, 1, 18, 2, 20, 6, 1, 8, 2, 8, 1, 1, 1, 4, 1,  487, 3, 5, 1, 7, 1, 5, 1, 24, 3, 17, 1, 42, 1, 2, 1, 1, 1, 16,  1, 3, 1, 30, 4, 1, 183, 4, 1, 5, 1, 141, 1, 14, 1, 12, 1, 2,  1, 206, 6, 2, 1, 4, 92, 10, 1, 5, 1, 3, 5, 5, 2, 87, 1, 1, 1,  5, 5, 4, 4, 78, 1, 3, 2, 1, 16, 1, 133, 1, 5, 1, 1, 1, 4, 1,  43, 1, 1, 1, 30, 1, 1, 7, 2, 6, 1, 1, 3, 3, 1, 10, 1, 5, 1, 1,  1, 1, 1, 159, 2, 1, 1, 10, 1, 16, 4, 2, 5, 3, 1, 3, 11, 4, 1,  2, 12, 5, 1, 1, 44, 3, 2, 1, 2, 17, 1, 4, 1, 1, 17, 1, 1, 3,  4, 18, 14, 4, 1, 2, 1, 1, 1, 2, 12, 1, 2, 1, 1, 1, 1, 1, 3, 1,  20, 1, 1, 1, 7, 1, 1, 3, 1, 2, 2, 1, 6, 1, 1, 1, 1, 1, 1, 1,  2, 1, 1, 2), .Dim = c(270L, 2L))
inca <- rep(inca[,1],inca[,2])/1000 
@

We have raw data of daily consumption of tap water from 1,180 tap
water consumers (var \texttt{inca}, see Figure \ref{fig:Function-inca}).
We could choose to use this empirical distribution to evaluate the
variability in the tap water consumption:

%
\begin{figure}
\caption{\label{fig:Function-inca}Histogram of daily tap water intake}

<<Crypto2, echo=false, eval=true, fig=true>>=
hist(inca, xlab="l/day",freq=FALSE,main="") 
@
\end{figure}

<<Crypto3, echo=true>>=
ndvar(1001);ndunc(1001)
mcstoc(rempiricalD,type="V",values=inca)
@

but we will use the \textquotedbl{}\texttt{fitdistrplus}\textquotedbl{}
library. \texttt{inca} includes a lot of \texttt{0} nodes, corresponding
to days when individuals do not drink tap water (possibly they drink
bottled water on those days). We could try a mixture of distributions,
with \textquotedbl{}\texttt{0}\textquotedbl{}and \textquotedbl{}non-\texttt{0}\textquotedbl{}
data.

<<Crypto4, echo=true, figure=false>>=
library(fitdistrplus) 
pzero <- sum(inca==0)/length(inca)
inca_non_0 <- inca[inca!=0] 
descdist(inca_non_0)
@

%
\begin{figure}
\caption{\label{fig:Function-descdist}Graph from the \texttt{descdist} function. }

<<Crypto5, echo=false, eval=true, fig=true>>=
descdist(inca_non_0)
@
\end{figure}

Following the \texttt{descdist} function (See figure \ref{fig:Function-descdist}),
let us try the lognormal distribution.

<<Crypto6, echo=true, figure=false>>=
Adj_water <- fitdist(inca_non_0,"lnorm",method="mle")
meanlog <- Adj_water$est[1]
sdlog   <- Adj_water$est[2]
summary(Adj_water)
plot(Adj_water)
@

%
\begin{figure}
\caption{\label{fig:Function-mledist}Graph from the \texttt{fitdist} function. }

<<Crypto7, echo=false, eval=true, fig=true>>=
plot(Adj_water)
@
\end{figure}

Not so bad (See Figure \ref{fig:Function-mledist}), and better than
a gamma distribution (results not shown). We can now rebuild our mixture.
We could consider uncertainty around the maximum likelihood estimates
using the \texttt{bootdist} function of the \texttt{fitdistrplus}
package, using something like:

<<Crypto7bis, echo=true, eval=false>>=
Boot <- bootdist(Ajust_lnorm, bootmethod="param", niter=ndunc())
Mean_conso <- mcdata(Boot$estim$meanlog, type="U")
Sd_conso <- mcdata(Boot$estim$sdlog, type="U")
conso1 <- mcstoc(rlnorm, type="VU", meanlog= Mean_conso, sdlog= Sd_conso)
@

But for simplicity, we will not consider uncertainty around the estimates.

We will use the \texttt{mcprobtree} function to construct a mixture
of \textquotedbl{}0\textquotedbl{}and \textquotedbl{}non-0\textquotedbl{}
distributions:

<<Crypto8, echo=true, figure=false>>=
conso0 <- mcdata(0,type="V")
conso1 <- mcstoc(rlnorm, type="V", meanlog=meanlog, sdlog=sdlog)
v <- mcprobtree(c(pzero,1-pzero), list("0"=conso0,"1"=conso1), type = "V")
summary(v)
@

\subsection{The Dose-Response Model}

We propose a boostrap from data (\texttt{datDR}) derived from \cite{CHAPPELL-1996}.
We first define a function \textquotedbl{}\texttt{DR}\textquotedbl{}
with an \texttt{n} argument for the size of the sample to draw. This
function may then be used in a \texttt{mcstoc} function:

<<Crypto9, echo=true, figure=false>>=
datDR <- list(   dose=c(30,100,300,500,1000,10000,100000,1000000),
                 pi=c(2,4,2,5,2,3,1,1),
                 ni=c(5,8,3,6,2,3,1,1))

estDR <- function(pos,ref){
   -glm(cbind(ref$ni-pos,pos) ~ ref$dose + 0, binomial(link="log"))$coefficients}

ml <- 1-exp(-estDR(datDR$pi, datDR) * datDR$dose)

DR <- function(n){
   boot <- matrix(rbinom(length(datDR$dose)*n,datDR$ni,ml),nrow=length(datDR$dose))
   apply(boot,2,estDR,ref=datDR)}
r <- mcstoc(DR, type="U")
summary(r)
@

\subsection{The Model}

Deriving the final model is straightforward. We construct the \texttt{mcnode}
corresponding to the recovery rate (Uncertainty,\texttt{ Rr}), the
probability for an oocyst to be infective (Variability, \texttt{w}):

<<Crypto10, echo=true, figure=false>>=
Rr <- mcstoc(rbeta, type="U", shape1=2.65, shape2=3.64)
w <- mcstoc(rbeta, type="V", shape1=2.6, shape2=3.4) 
@

Given that $O_{o}=2$ oocysts are observed in 100 l of water, the
expected number of oocysts in the sample is \texttt{l}:

<<Crypto11, echo=true, figure=false>>=
Oo <- 2
l <- (Oo + mcstoc(rnbinom, type="U", size=Oo+1, prob=Rr))/100
@

The expected number of oocysts drunk by the individuals is \texttt{Or}
and the risk ($\times10000$) is estimated by:

<<Crypto12, echo=true, figure=false>>=
Or <- l * v * w
P <- 10000 * (1-exp(-r*Or))
summary(P)
@

This result can be compared (roughly since there is some differences
in the modelsvariability) to the results shown in Table 2 in \cite{POUILLOT-2004}.

Improvement: the results for $O_{o}=\left\{ 0,1,2,5,10,20,50,100,1000\right\} $
can be obtained in one step using:

<<Crypto13, echo=true, eval=false, figure=false>>=
Oo <- mcdata(c(0,1,2,5,10,20,50,100,1000),type="0",nvariates=9) 
@

\section*{As a Conclusion}

We think and hope that {}``\texttt{mc2d}'' could help risk assessors
to constuct and analyse their models, and that it may help in developing
\textquotedbl{}two-dimensional\textquotedbl{} simulations. Nevertheless,
\textquotedbl{}\texttt{mc2d}\textquotedbl{} is currently under development:

\noindent \begin{center}
\emph{CHECK YOUR MODEL CAREFULLY AND EXAMINE RESULTS TO DETECT BUGS}
\par\end{center}

and, if you would like to improve it, join us at

\noindent \begin{center}
\emph{\url{http://riskassessment.r-forge.r-project.org/}}
\par\end{center}

Please refer any comments or bugs to \url{rpouillot@yahoo.fr}.

\bibliographystyle{plain}
\bibliography{Bibtex}

\end{document}

### 2010-02-03 pkg/mc2d/inst/doc/docmcEnglish.rnw
Test for a Vignette (once again)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@84 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2009-08-28 pkg/mc2d/inst/doc/docmcEnglish.Rnw
@@
-\documentclass[10pt,english]{article}
-\VignetteIndexEntry
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% mc2d: Tools for Two-Dimensional Monte-Carlo Simulations
+%%%
+%%%         Sweave vignette file
+%%%
+
+\documentclass[10pt, english]{article}
+
+%\usepackage[letterpaper]{geometry}
 \usepackage[T1]{fontenc}
 \usepackage[latin9]{inputenc}
-\usepackage[letterpaper]{geometry}
-\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
-\setlength{\parskip}{\medskipamount}
-\setlength{\parindent}{0pt}
-\usepackage{url}
 \usepackage{amsmath}
 \usepackage{graphicx}
-\usepackage{setspace}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-%% Because html converters don't know tabularnewline
-\providecommand{\tabularnewline}{\\}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\usepackage{noweb}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-
-\usepackage{babel}
+% sweave commands for vignette
+%\VignetteIndexEntry{Tools for Two-Dimensional Monte-Carlo Simulations}
+%\VignettePackage{mc2d}
+%\VignetteDepends{mvtnorm}
+
+\title{The \texttt{mc2d} package.}
+\author{R. POUILLOT, M.-L. DELIGNETTE-MULLER, D.L. KELLY \& J.-B. DENIS}
 
 \begin{document}
-
-\title{The {}``\texttt{mc2d}'' package. }
-
-\author{R. POUILLOT, M.-L. DELIGNETTE-MULLER, D.L. KELLY \& J.-B. DENIS}
-
 \maketitle
-\tableofcontents{}
-
-\noindent \begin{center}
-\vfill{}
-
-\par\end{center}
 
@@
 \begin{itemize}
-\item A medium level of experience in R. Please refer to the Manual {}``An
-Introduction to R'' available with R distribution if needed;
+\item A medium level of experience in R. Please refer to the Manual An
+Introduction to R available with R distribution if needed;
 \item Some knowledge about Monte-Carlo simulation (its basic principles
@@
 \end{itemize}
+
 This documentation will not describe all arguments of the functions.
@@
 
+\tableofcontents{}
+
 \section{Introduction}
-
 \subsection{What is \texttt{mc2d}?}
 
-{}``\texttt{mc2d}'' means Two-Dimensional Monte-Carlo (\emph{{}``Monte-Carlo
-à Deux Dimensions''}). This package :
+\texttt{mc2d} means Two-Dimensional Monte-Carlo (\emph{Monte-Carlo
+à Deux Dimensions}). This package :
 \begin{itemize}
@@
 \end{itemize}
+
 In a previous version, some tools to fit parametric distributions
@@
 purposes, they have been moved to a separate package called \texttt{fitdistrplus}.
-Both the \texttt{mc2d} and the \texttt{fitdistrplus} packages are
-available at the URL \url{https://r-forge.r-project.org/projects/riskassessment/}.
 
@@
 According to international recommendations, a QRA should reflect the
-{}``variability'' in the risk and calculate the {}``uncertainty''
-associated with the risk estimate. The {}``variability'' represents
+variability in the risk and calculate the uncertainty
+associated with the risk estimate. The variability represents
 temporal, geographical and/or individual heterogeneity of the risk
-for a given population. The {}``uncertainty'' is understood as stemming
+for a given population. The uncertainty is understood as stemming
 from a lack of perfect knowledge about the QRA model structure and
 associated parameters%
-\footnote{In the engineering risk community, these concepts are refered as \textquotedbl{}aleatoric
-uncertainty\textquotedbl{} for \textquotedbl{}variability\textquotedbl{}
-and \textquotedbl{}epistemic uncertainty\textquotedbl{} for \textquotedbl{}uncertainty\textquotedbl{}.%
+\footnote{In the engineering risk community, these concepts are refered as aleatoric
+uncertainty for variability
+and epistemic uncertainty for uncertainty.%
 }.
 
-In order to estimate the natural {}``variability'' of the risk,
+In order to estimate the natural variability of the risk,
 a Monte-Carlo simulation approach may be useful: the empirical distribution
@@
 A two-dimensional (or second-order) Monte-Carlo simulation was proposed
-to estimate the {}``uncertainty'' in the risk estimates stemming
+to estimate the uncertainty in the risk estimates stemming
 from parameter uncertainty \cite{CULLEN-FREY-1999}. A two-dimensional
 Monte-Carlo simulation is a Monte-Carlo simulation where the distributions
-reflecting \textquotedbl{}variability\textquotedbl{} and the distributions
-representing \textquotedbl{}uncertainty\textquotedbl{} are sampled
-separately in the simulation, so that \textquotedbl{}variability\textquotedbl{}
-and \textquotedbl{}uncertainty\textquotedbl{} in the output may be
+reflecting variability and the distributions
+representing uncertainty are sampled
+separately in the simulation, so that variability
+and uncertainty in the output may be
 estinated separately. It may be described as following (see Figure
 \ref{fig:Mc2d}):
+
 \begin{enumerate}
 \item The parameters of the model should be divided into three categories:
-the parameters whose distributions reflect {}``variability only'',
-hereinafter denoted as {}``variable parameters'', the parameters
-whose distributions reflect {}``uncertainty only'', denoted as {}``uncertain
-parameters'' and the parameters whose distributions reflect both
+the parameters whose distributions reflect variability only,
+hereinafter denoted as variable parameters, the parameters
+whose distributions reflect uncertainty only, denoted as uncertain
+parameters and the parameters whose distributions reflect both
 uncertainty and variability. For this latter category, a hierarchical
-structure, using {}``hyper-parameters'', should be specified: if
+structure, using hyper-parameters, should be specified: if
 a parameter is both uncertain and variable, one should be able to
@@
 is some associated uncertainty. As an example, one should be able
-to specify a relationship such as $X\left|a,b\right.\sim N\left(a,b\right)$,
+to specify a relationship such as
+\[ X\left|a,b\right.\sim N\left(a,b\right) \]
 where the specified normal distribution represents variability in
 $x$ conditional upon parameters $a$ and $b$. Hyperdistributions,
-such as $a\sim Unif\left(l_{a},u_{a}\right)$and $b\sim Unif\left(l_{b},u_{b}\right)$,
+such as
+\[ a\sim Unif\left(l_{a},u_{a}\right) \]
+and
+\[ b\sim Unif\left(l_{b},u_{b}\right) \]
 represent the uncertainty in the parameters $a$ and $b$;
@@
 leading to $N_{u}$ sets of statistics; 
-\item As output, the $50^{\text{th}}$ percentile (median) of each statistic
-is used as a point estimate of this statistic; the $2.5^{\text{th}}$
-and $97.5^{\text{th}}$ percentiles of each statistic are used to
+\item As output, the $50^{th}$ percentile (median) of each statistic
+is used as a point estimate of this statistic; the $2.5^{th}$
+and $97.5^{th}$ percentiles of each statistic are used to
 establish a 95\% credible interval (CI95) of this statistic. The median
 of the $N_{u}$ estimated values for each of the 101 estimated percentiles
-allows us to display a {}``variability cumulative distribution''
+allows us to display a variability cumulative distribution
 via a graph. This curve is surrounded by the 2.5th and 97.5th percentiles
@@
 \noindent \begin{centering}
-\includegraphics[scale=0.5]{mc2dSaumon}
+\includegraphics[angle=270, width=1\textwidth]{mc2dsaumon}
 \par\end{centering}
@@
 
-\textquotedbl{}\texttt{mc2d}\textquotedbl{} is a set of R functions
+\texttt{mc2d} is a set of R functions
 that will help to develop such two-dimensional Monte-Carlo simulations.
@@
 
-\textbf{Quantitative Risk Assessment:} \textbf{\emph{Escherichia coli}}
-\textbf{O157:H7 infection linked to the consumption of frozen ground
+\textbf{Quantitative Risk Assessment: \emph{Escherichia coli}
+O157:H7 infection linked to the consumption of frozen ground
 beef in <3 year old children.}
@@
 is kept frozen until it is cooked, just before consumption;
-\item 2.7\% of consumers cook their beef {}``rare'', 37.3\% {}``medium''
-and 60.0\% {}``well done''; 
+\item 2.7\% of consumers cook their beef rare, 37.3\% medium
+and 60.0\% well done; 
 \item The following bacterial inactivation $i$ is associated with these
@@
 \begin{itemize}
-\item No inactivation for {}``rare'' cooking;
-\item 1/5 surviving bacteria for a {}``medium'' cooking;
-\item 1/50 surviving bacteria for a {}``well done'' cooking.
+\item No inactivation for rare cooking;
+\item 1/5 surviving bacteria for a medium cooking;
+\item 1/50 surviving bacteria for a well done cooking.
 \end{itemize}
 \item The variability in steak serving sizes $s$ for <3 year children was
-estimated in a consumption survey. The {}``best fit'' to the data
+estimated in a consumption survey. The best fit to the data
 was a gamma distribution with parameters: $shape=3.93,\: rate=0.0806$.
@@
 \end{itemize}
-The question is: {}``\emph{What is the distribution of the risk of
-illness in the population that consumed the contaminated lot?}''
+The question is: \emph{What is the distribution of the risk of
+illness in the population that consumed the contaminated lot?}
 
 This distribution will be estimated using Monte-Carlo simulations
-performed with R via the {}``\texttt{mc2d}'' package. First, the
+performed with R via the \texttt{mc2d} package. First, the
 model will be developed in a one dimensional framework. Then, in order
@@
 All distributions represent variability only. The model may be written
-as:\begin{eqnarray*}
+as:
+\begin{eqnarray*}
 c & = & 10.\\
@@
 n & \sim & Poisson\left(c\times i\times s\right)\\
-P & = & 1-\left(1-0.001\right)^{n}\end{eqnarray*}
-
-where $emp\left(\boldsymbol{X},\boldsymbol{P}\right)$ is an empirical
+P & = & 1-\left(1-0.001\right)^{n}
+\end{eqnarray*}
+
+where $emp\left(X,P\right)$ is an empirical
 distribution wherein each value $X_{i}$ is associated with a probability
-$P_{i}$.We will use a \textquotedbl{}classical\textquotedbl{} one
+$P_{i}$.We will use a classical one
 dimensional Monte-Carlo simulation, with 1,000 iterations. Using the
-{}``\texttt{mc2d}'' package, the model may be written as:
+\texttt{mc2d} package, the model may be written as:
 
@@
 
-<<sh1, echo=true>>=
+%true
+<<sh1, eval=true, echo=true>>=
 library(mc2d)
@@
 in the variability dimension and 100 iterations in the uncertainty
-dimension. Using the {}``\texttt{mc2d}'' package, the model may
+dimension. Using the \texttt{mc2d} package, the model may
 be written as:
 
-<<sh2, echo=true>>=
+%true
+<<sh2, eval=true, echo=true>>=
 ndunc(100)
@@
 
-Note that the syntax is similar to the earlier model. However, a {}``\texttt{type}''
+Note that the syntax is similar to the earlier model. However, a \texttt{type}
 argument is provided for each distribution, indicating whether the
-parameter distribution represents uncertainty (\texttt{type}=\textquotedbl{}\texttt{U}\textquotedbl{}),
-variability (\texttt{type}=\textquotedbl{}V\textquotedbl{}), or both
-(\texttt{type}=\textquotedbl{}\texttt{VU}\textquotedbl{}).
+parameter distribution represents uncertainty (\texttt{type}=\texttt{U}),
+variability (\texttt{type}=V), or both
+(\texttt{type}=\texttt{VU}).
 
@@
 around \texttt{conc} and \texttt{r}. The estimate of the mean risk
-is now uncertain. The median of the 100 simulations leads to a \textquotedbl{}best
-estimate\textquotedbl{}of 0.0445, with a 95\% \textquotedbl{}credible
-interval\textquotedbl{} of {[}0.191, 0.0730{]}.
+is now uncertain. The median of the 100 simulations leads to a best
+estimateof 0.0445, with a 95\% credible
+interval of {[}0.191, 0.0730{]}.
 
@@
 
-A typical session of R using {}``\texttt{mc2d}'' is as follows:
+A typical session of R using \texttt{mc2d} is as follows:
 \begin{itemize}
 \item From data, expert knowledge, \emph{etc.} an empirical or parametric
-distribution is chosen for each {}``parent'' parameter. For developing
-an empirical distribution from data , the {}``\texttt{fitdistrplus}''
+distribution is chosen for each parent parameter. For developing
+an empirical distribution from data , the \texttt{fitdistrplus}
 package is recommended;
@@
 
-The {}``\texttt{mc2d}'' library should be loaded at the beginning
-of your R session (\textquotedbl{}\texttt{library(mc2d)}\textquotedbl{}).
+The \texttt{mc2d} library should be loaded at the beginning
+of your R session (\texttt{library(mc2d)}).
 
@@
 An \texttt{mcnode} object is the basic element of an \texttt{mc} object.
-It is an array of dimension $\left(nsv\times nsu\times nvariates\right)$
+It is an array of dimension \texttt{nsv} * \texttt{nsu} * \texttt{nvariates}
 where $nsv$ is the dimension of variability, $nsu$ is the dimension
@@
 \begin{itemize}
-\item \textquotedbl{}\texttt{V}\textquotedbl{} \texttt{mcnode}, for \textquotedbl{}\emph{Variability}\textquotedbl{},
+\item \texttt{V} \texttt{mcnode}, for \emph{Variability},
 is an array of dimension $\left(nsv\times1\times nvariates\right)$.
 The distribution represents variability in the parameter; 
-\item \textquotedbl{}\texttt{U}\textquotedbl{} \texttt{mcnode}, for \textquotedbl{}\emph{Uncertainty}\textquotedbl{},
+\item \texttt{U} \texttt{mcnode}, for \emph{Uncertainty},
 is an array of dimension $\left(1\times nsu\times nvariates\right)$.
 The distribution represents uncertainty in the parameter. 
-\item \textquotedbl{}\texttt{VU}\textquotedbl{} \texttt{mcnode}, for \textquotedbl{}\emph{Variability
-and Uncertainty}\textquotedbl{}, is an array of dimension $\left(nsv\times nsu\times nvariates\right)$.
+\item \texttt{VU} \texttt{mcnode}, for \emph{Variability
+and Uncertainty}, is an array of dimension $\left(nsv\times nsu\times nvariates\right)$.
 The distribution represents both variability (in the first dimension)
 and uncertainty (in the second dimension) in the parameter. 
-\item Additionally, a \textquotedbl{}\texttt{0}\textquotedbl{} \texttt{mcnode}
-is also defined . \textquotedbl{}\texttt{0}\textquotedbl{} stands
-for \textquotedbl{}Neither Variability or Uncertainty\textquotedbl{}.
+\item Additionally, a \texttt{0} \texttt{mcnode}
+is also defined . \texttt{0} stands
+for Neither Variability or Uncertainty.
 Such nodes are arrays of dimension $\left(1\times1\times nvariates\right)$.
-No uncertainty or variability is considered for these nodes. A \textquotedbl{}\texttt{0}\textquotedbl{}
+No uncertainty or variability is considered for these nodes. A \texttt{0}
 \texttt{mcnode} is not necessary in the univariate context (use a
@@
 \noindent \begin{centering}
-\includegraphics[scale=0.4]{Illustration}
+\includegraphics[angle=270, width=.5\textwidth]{Illustration}
 \par\end{centering}
@@
 \caption{\label{fig:Structure}Structure of the various \texttt{mcnode} objects.}
-
 \end{figure}
@@
 from other \texttt{mcnodes} using a probability tree;
-\item Some functions, such as {}``\texttt{==}'' or {}``\texttt{>}''
+\item Some functions, such as \texttt{==} or \texttt{>}
 , \texttt{is.na}, \texttt{is.finite} generate a new \texttt{mcnode}
@@
 
-\begin{singlespace}
+%\begin{singlespace}
 \noindent \begin{flushleft}
-\texttt{mcstoc(func=runif, type=c(\textquotedbl{}V\textquotedbl{},
-\textquotedbl{}U\textquotedbl{}, \textquotedbl{}VU\textquotedbl{},
-\textquotedbl{}0\textquotedbl{}), ..., nsv=ndvar(), nsu=ndunc(), nvariates=1,
-outm=\textquotedbl{}each\textquotedbl{}, nsample=\textquotedbl{}n\textquotedbl{},
+\texttt{mcstoc(func=runif, type=c(V,
+U, VU,
+0), ..., nsv=ndvar(), nsu=ndunc(), nvariates=1,
+outm=each, nsample=n,
 seed=NULL, rtrunc=FALSE, linf=-Inf, lsup=Inf, lhs=FALSE)}
 \par\end{flushleft}
-\end{singlespace}
+%\end{singlespace}
 \begin{itemize}
@@
 \item \texttt{type} is the type of requested \texttt{mcnode}. By default,
-\texttt{mcstoc} constructs a {}``\texttt{V}'' \texttt{mcnode};
+\texttt{mcstoc} constructs a \texttt{V} \texttt{mcnode};
 \item \ldots{} are the arguments to be passed to the function \texttt{func},
@@
 
+%false
 <<sh3, echo=true, eval=FALSE>>=
@@
 
+%true
 <<sh3bis, echo=true, eval=true>>=
@@
 
-\begin{singlespace}
+%\begin{singlespace}
 \noindent \begin{flushleft}
-\texttt{mcdata(data, type=c(\textquotedbl{}V\textquotedbl{}, \textquotedbl{}U\textquotedbl{},
-\textquotedbl{}VU\textquotedbl{}, \textquotedbl{}0\textquotedbl{}),
-nsv=ndvar(), nsu=ndunc(), nvariates=1, outm=\textquotedbl{}each\textquotedbl{})}
+\texttt{mcdata(data, type=c(V, U,
+VU, 0),
+nsv=ndvar(), nsu=ndunc(), nvariates=1, outm=each)}
 \par\end{flushleft}
-\end{singlespace}
+%\end{singlespace}
 
@@
 of data that can be used to construct an \texttt{mcnode}. The following
-example places a \texttt{TRUE} value in a \textquotedbl{}\texttt{U}\textquotedbl{}
+example places a \texttt{TRUE} value in a \texttt{U}
 node in half of the simulations:
@@
 
-<<sh3, echo=true>>=
+%true
+<<sh3, echo=true, eval=true>>=
 nu <- ndunc()
@@
 and variability coherently within the model. Logically, the rules
-are as follows (illustrated here with a {}``+'')%
+are as follows (illustrated here with a +)%
 \footnote{These rules are not the standard R rules for recycling.%
@@
 \begin{itemize}
-\item \textquotedbl{}\texttt{0}\textquotedbl{} + \textquotedbl{}\texttt{0}\textquotedbl{}
-= \textquotedbl{}\texttt{0}\textquotedbl{};
-\item \textquotedbl{}\texttt{0}\textquotedbl{} + \textquotedbl{}\texttt{V}\textquotedbl{}
-= \textquotedbl{}\texttt{V}\textquotedbl{}
-\item \textquotedbl{}\texttt{0}\textquotedbl{} + \textquotedbl{}\texttt{U}\textquotedbl{}
-= \textquotedbl{}\texttt{U}\textquotedbl{};
-\item \textquotedbl{}\texttt{0}\textquotedbl{} + \textquotedbl{}\texttt{VU}\textquotedbl{}
-= \textquotedbl{}\texttt{VU}\textquotedbl{};
-\item \textquotedbl{}\texttt{V}\textquotedbl{} + \textquotedbl{}\texttt{V}\textquotedbl{}
-= \textquotedbl{}\texttt{V}\textquotedbl{};
-\item \textquotedbl{}\texttt{V}\textquotedbl{} + \textquotedbl{}\texttt{U}\textquotedbl{}
-= \textquotedbl{}\texttt{VU}\textquotedbl{}: the {}``\texttt{U}''
-mcnode is recycled by row, the \textquotedbl{}\texttt{V}\textquotedbl{}
+\item \texttt{0} + \texttt{0}
+= \texttt{0};
+\item \texttt{0} + \texttt{V}
+= \texttt{V}
+\item \texttt{0} + \texttt{U}
+= \texttt{U};
+\item \texttt{0} + \texttt{VU}
+= \texttt{VU};
+\item \texttt{V} + \texttt{V}
+= \texttt{V};
+\item \texttt{V} + \texttt{U}
+= \texttt{VU}: the \texttt{U}
+mcnode is recycled by row, the \texttt{V}
 mcnode is recycled in the standard manner by column;
-\item \textquotedbl{}\texttt{V}\textquotedbl{} + \textquotedbl{}\texttt{VU}\textquotedbl{}
-= \textquotedbl{}\texttt{VU}\textquotedbl{}: the {}``\texttt{V}''
+\item \texttt{V} + \texttt{VU}
+= \texttt{VU}: the \texttt{V}
 mcnode is recycled in the standard manner by column;
-\item \textquotedbl{}\texttt{U}\textquotedbl{} + \textquotedbl{}\texttt{U}\textquotedbl{}
-= \textquotedbl{}\texttt{U}\textquotedbl{};
-\item \textquotedbl{}\texttt{U}\textquotedbl{} + \textquotedbl{}\texttt{VU}\textquotedbl{}
-= \textquotedbl{}\texttt{VU}\textquotedbl{}: the {}``\texttt{U}''
+\item \texttt{U} + \texttt{U}
+= \texttt{U};
+\item \texttt{U} + \texttt{VU}
+= \texttt{VU}: the \texttt{U}
 mcnode is recycled by row;
-\item \textquotedbl{}\texttt{VU}\textquotedbl{} + \textquotedbl{}\texttt{VU}\textquotedbl{}
-= \textquotedbl{}\texttt{VU}\textquotedbl{}
+\item \texttt{VU} + \texttt{VU}
+= \texttt{VU}
 \end{itemize}
@@
 
+%false
 <<sh4, echo=true, eval=false>>=
@@
 
-\texttt{expo} is a function of a {}``\texttt{U}'' and two {}``\texttt{V}''
-\texttt{mcnode}s: it is a {}``\texttt{VU}'' \texttt{mcnode} with
+\texttt{expo} is a function of a \texttt{U} and two \texttt{V}
+\texttt{mcnode}s: it is a \texttt{VU} \texttt{mcnode} with
 variability in the row dimension and uncertainty in the column dimension
-. \texttt{risk} is a function of a {}``\texttt{U}'' and a {}``\texttt{VU}''
-\texttt{mcnode}: it is therefore a {}``\texttt{VU}'' \texttt{mcnode}.
+. \texttt{risk} is a function of a \texttt{U} and a \texttt{VU}
+\texttt{mcnode}: it is therefore a \texttt{VU} \texttt{mcnode}.
 
@@
 
-The \texttt{mcprobtree} function can be used if a {}``probability
-tree'' is needed to construct an \texttt{mcnode}. Assume that the
+The \texttt{mcprobtree} function can be used if a probability
+tree is needed to construct an \texttt{mcnode}. Assume that the
 distribution representing the uncertainty on \texttt{conc} was not
@@
 \footnote{two alternatives for \texttt{whichdist} are \texttt{whichdist <- mcstoc(rempiricalD,
-type=\textquotedbl{}U\textquotedbl{}, values=c(0,1), prob=c(75,25))}
-or \texttt{whichdist <- mcstoc(rbern,type=\textquotedbl{}U\textquotedbl{},prob=0.25)}%
+type=U, values=c(0,1), prob=c(75,25))}
+or \texttt{whichdist <- mcstoc(rbern,type=U,prob=0.25)}%
 }:
 
+%true
 <<sh4, echo=true, eval=true>>=
@@
 
-The functions {}``\texttt{==}'', ''\texttt{<}'', ''\texttt{<}='',
-''\texttt{>=}'', {}``\texttt{>}'', generate an \texttt{mcnode}
+The functions \texttt{==}, \texttt{<}, \texttt{<}=,
+\texttt{>=}, \texttt{>}, generate an \texttt{mcnode}
 when applied to another \texttt{mcnode}.
@@
 \texttt{is.infinite(x)} are implemented to test if any values are
-\texttt{NA} (missing data), \texttt{NaN} ({}``\emph{Not A Number}''),
+\texttt{NA} (missing data), \texttt{NaN} (\emph{Not A Number}),
 or finite .
 
+%true
 <<sh5, echo=true, eval=true>>=
@@
 
+%true
 <<sh5bis, echo=true, eval=true>>=
@@
 
-It is possible to create such correlations between {}``V'' nodes,
-between {}`` U'' nodes, between {}``VU'' nodes, or between one
-{}``V'' node and multiple {}``VU'' nodes.
+It is possible to create such correlations between V nodes,
+between  U nodes, between VU nodes, or between one
+V node and multiple VU nodes.
 
@@
 them into a single object in order to analyse the Monte-Carlo results.
-The {}``\texttt{mc}'' object is a list of \texttt{mcnode}s. There
+The \texttt{mc} object is a list of \texttt{mcnode}s. There
 are three ways to construct an \texttt{mc} object: using the \texttt{mc}
@@
 
+%false
 <<sh6, echo=true, eval=false>>=
@@
 
+%true
 <<sh10, echo=true, eval=true>>=
@@
 \begin{itemize}
-\item the model is wrapped between {}``\texttt{\{}'' and {}``\texttt{\}}'';
+\item the model is wrapped between \texttt{\{} and \texttt{\}};
 \item any (valid) R code may be placed in the model%
@@
 
+%false
 <<sh11, echo=true, eval=false>>=
@@
 
-<<sh12, echo=true, eval=FALSE>>=
+%false
+<<sh12, echo=true, eval=false>>=
 modEC4 <- mcmodelcut({
@@
 
+%true
 <<sh14, echo=true, eval=true>>=
@@
 \noindent \begin{flushleft}
-\texttt{hist (x, griddim = NULL, xlab = names(x), ylab = \textquotedbl{}Frequency\textquotedbl{},
-main = \textquotedbl{}\textquotedbl{}, ...)}
+\texttt{hist (x, griddim = NULL, xlab = names(x), ylab = Frequency,
+main = , ...)}
 \par\end{flushleft}
@@
 
+%false
 <<sh15, echo=true, eval=false, fig=false>>=
@@
 
+%true
 <<sh16, echo=false, eval=true, fig=true>>=
@@
 \noindent \begin{flushleft}
-\texttt{plot (x, prec = 0.01, stat = c(\textquotedbl{}median\textquotedbl{},
-\textquotedbl{}mean\textquotedbl{}), lim = c(0.025,0.975), na.rm =
-TRUE, griddim = NULL, xlab = NULL, ylab = \textquotedbl{}Fn(x)\textquotedbl{},
-main = \textquotedbl{}\textquotedbl{}, draw = TRUE, ...)}
+\texttt{plot (x, prec = 0.01, stat = c(median,
+mean), lim = c(0.025,0.975), na.rm =
+TRUE, griddim = NULL, xlab = NULL, ylab = Fn(x),
+main = , draw = TRUE, ...)}
 \par\end{flushleft}
@@
 
+%false
 <<sh17, echo=true, eval=false, fig=false>>=
@@
 
+%true
 <<sh18, echo=false, eval=true, fig=true>>=
@@
 \noindent \begin{flushleft}
-\texttt{tornado(x, output=length(x), use=\textquotedbl{}all.obs\textquotedbl{},
-method=c(\textquotedbl{}spearman\textquotedbl{}, \textquotedbl{}kendall\textquotedbl{},\textquotedbl{}pearson\textquotedbl{}),
+\texttt{tornado(x, output=length(x), use=all.obs,
+method=c(spearman, kendall,pearson),
 lim=c(0.025, 0.975))}
@@
 
+%true
 <<sh19, echo=true, eval=true, fig=false>>=
@@
 
+%true
 <<sh19b, echo=false, eval=true, fig=true>>=
@@
 \texttt{tornadounc(mc,output = length(mc), quant=c(0.5,0.75,0.975),
-use = \textquotedbl{}all.obs\textquotedbl{}, method=c(\textquotedbl{}spearman\textquotedbl{},\textquotedbl{}kendall\textquotedbl{},\textquotedbl{}pearson\textquotedbl{}),
+use = all.obs, method=c(spearman,kendall,pearson),
 ...)}
@@
 
+%true
 <<sh19c, echo=true, eval=true, fig=false>>=
@@
 
-The output shows the impact of the uncertain nodes (type \textquotedbl{}U\textquotedbl{}
-nodes) and some statistics (mean, median and, here, the $99^{\text{th}}$percentile)
-calculated on the variability dimension (type \textquotedbl{}V\textquotedbl{}
-and type \textquotedbl{}VU\textquotedbl{} nodes) of some output statistics
+The output shows the impact of the uncertain nodes (type U
+nodes) and some statistics (mean, median and, here, the $99^{th}$percentile)
+calculated on the variability dimension (type V
+and type VU nodes) of some output statistics
 .
@@
 \texttt{mcnode} objects, it is straightforward to apply any R function
-to these objects. The \textquotedbl{}\texttt{\$}\textquotedbl{} function
+to these objects. The \texttt{\$} function
 is helpful for extracting an \texttt{mcnode} from an \texttt{mc} object.
@@
 
+%true
 <<sh19ter, echo=true, eval=true>>=
@@
 
-<<sh19ter, echo=true, eval=false>>=
+%true
+<<sh19ter, echo=true, eval=true>>=
 mcstoc(runif, nvariates=3, min=c(1,2,3),max=4)
@@
 
+%true
 <<sh19quatr, echo=true, eval=true>>=
@@
 
+%true
 <<sh20, echo=true, eval=true>>=
@@
 
-Assume that each member of these families eats a {}``normal'' distribution
+Assume that each member of these families eats a normal distribution
 (variability) of steak with mean 100, 150 and 250 g. There is a positive
@@
 
+%true
 <<sh21, echo=true, eval=true>>=
@@
 Printing the statistics of the 30 variates of \texttt{x} is of no
-interest. Instead, we use the {}``\texttt{outm}'' option, which
-allows us to specify which output we want ({}``\texttt{none}'' for
-none, {}``\texttt{each}'', the default, for a series of statistics
+interest. Instead, we use the \texttt{outm} option, which
+allows us to specify which output we want (\texttt{none} for
+none, \texttt{each}, the default, for a series of statistics
 for each variate, or, as in the example, a vector of functions that
@@
 
-\subsection{Multivariate Nodes as a {}``Third Dimension'' for Multiple Options
+\subsection{Multivariate Nodes as a Third Dimension for Multiple Options
 in a Model}
@@
 is possible, but that $conc\sim U\left(8,12\right)$ is also possible.
-We can \emph{i}) build a {}``bivariate'' node reflecting these two
+We can \emph{i}) build a bivariate node reflecting these two
 options; ii) transfer these options into the final risk estimate.
@@
 
-<<sh23, echo=true>>=
+%true
+<<sh23, echo=true, eval=true>>=
 conc1 <- mcstoc(rnorm, type="U", mean=10, sd=2)
@@
 
-\subsection{Multivariate Nodes as a {}``Third Dimension'' for Multiple Vectors/Contaminants}
+\subsection{Multivariate Nodes as a Third Dimension for Multiple Vectors/Contaminants}
 
@@
 $conc\sim N\left(10,2\right)$, the second one follows $conc\sim N\left(14,2\right)$.
-We can \emph{i}) build a {}``bivariate'' node reflecting these two
+We can \emph{i}) build a bivariate node reflecting these two
 concentrations%
@@
 
-<<sh24, echo=true>>=
+%true
+<<sh24, echo=true, eval=true>>=
 mconc <- mcdata(c(10,14), type="0", nvariates=2)
@@
 
-As a conclusion, this \textquotedbl{}third\textquotedbl{} dimension
+As a conclusion, this third dimension
 is highly flexible...
@@
 
+%true
 <<Crypto1, echo=false, eval=true>>=
@@
 
+%true
 <<Crypto2, echo=false, eval=true, fig=true>>=
@@
 
-<<Crypto3, echo=true>>=
+%true
+<<Crypto3, echo=true, eval=true>>=
 ndvar(1001);ndunc(1001)
@@
 
-but we will use the \textquotedbl{}\texttt{fitdistrplus}\textquotedbl{}
+but we will use the \texttt{fitdistrplus}
 library. \texttt{inca} includes a lot of \texttt{0} nodes, corresponding
@@
 bottled water on those days). We could try a mixture of distributions,
-with \textquotedbl{}\texttt{0}\textquotedbl{}and \textquotedbl{}non-\texttt{0}\textquotedbl{}
+with \texttt{0}and non-\texttt{0}
 data.
 
-<<Crypto4, echo=true, figure=false>>=
+%true
+<<Crypto4, echo=true, figure=false, eval=true>>=
 library(fitdistrplus) 
@@
 
-<<Crypto6, echo=true, figure=false>>=
+<<Crypto6, echo=true, figure=false, eval=true>>=
 Adj_water <- fitdist(inca_non_0,"lnorm",method="mle")
@@
 We will use the \texttt{mcprobtree} function to construct a mixture
-of \textquotedbl{}0\textquotedbl{}and \textquotedbl{}non-0\textquotedbl{}
+of 0and non-0
 distributions:
 
-<<Crypto8, echo=true, figure=false>>=
+<<Crypto8, echo=true, figure=false, eval=true>>=
 conso0 <- mcdata(0,type="V")
@@
 We propose a boostrap from data (\texttt{datDR}) derived from \cite{CHAPPELL-1996}.
-We first define a function \textquotedbl{}\texttt{DR}\textquotedbl{}
+We first define a function \texttt{DR}
 with an \texttt{n} argument for the size of the sample to draw. This
@@
 
-<<Crypto9, echo=true, figure=false>>=
+<<Crypto9, echo=true, figure=false, eval=true>>=
 datDR <- list(   dose=c(30,100,300,500,1000,10000,100000,1000000),
@@
 
-<<Crypto10, echo=true, figure=false>>=
+<<Crypto10, echo=true, figure=false, eval=true>>=
 Rr <- mcstoc(rbeta, type="U", shape1=2.65, shape2=3.64)
@@
 
-<<Crypto11, echo=true, figure=false>>=
+<<Crypto11, echo=true, figure=false, eval=true>>=
 Oo <- 2
@@
 
-<<Crypto12, echo=true, figure=false>>=
+<<Crypto12, echo=true, figure=false, eval=true>>=
 Or <- l * v * w
@@
 
-We think and hope that {}``\texttt{mc2d}'' could help risk assessors
+We think and hope that \texttt{mc2d} could help risk assessors
 to constuct and analyse their models, and that it may help in developing
-\textquotedbl{}two-dimensional\textquotedbl{} simulations. Nevertheless,
-\textquotedbl{}\texttt{mc2d}\textquotedbl{} is currently under development:
+two-dimensional simulations. Nevertheless,
+\texttt{mc2d} is currently under development:
 
@@
 \noindent \begin{center}
-\emph{\url{http://riskassessment.r-forge.r-project.org/}}
+http://riskassessment.r-forge.r-project.org/
 \par\end{center}
 
-Please refer any comments or bugs to \url{rpouillot@yahoo.fr}.
+Please refer any comments or bugs to rpouillot@yahoo.fr.
 
 \bibliographystyle{plain}
-\bibliography{Bibtex}
+\bibliography{docmcEnglish}
 

### 2010-02-13 pkg/mc2d/R/plot.mc.R
Some improvements on mc2d
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@85 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,5 +1,5 @@
 #<<BEGIN>>
-plot.mc <- function(x, prec=0.01, stat = c("median","mean"), lim = c(0.025,0.975), na.rm=TRUE, griddim = NULL, xlab = NULL, ylab = "Fn(x)", main = "", draw = TRUE, ...)
+plot.mc <- function(x, prec=0.001, stat = c("median","mean"), lim = c(0.025,0.25, 0.75, 0.975), na.rm=TRUE, griddim = NULL, xlab = NULL, ylab = "Fn(x)", main = "", draw = TRUE, paint=TRUE,...)
 #TITLE Plots Results of a Monte Carlo Simulation
 #DESCRIPTION
 # Plots the empirical cumulative distribution function of a \samp{mcnode} or a \samp{mc} object ("0" and "V" nodes) or the
@@ -8,9 +8,8 @@ plot.mc <- function(x, prec=0.01, stat = c("median","mean"), lim = c(0.025,0.975
 #INPUTS
 # {x}<<a \samp{mcnode} or a \samp{mc} objects>>
 #[INPUTS]
-#{prec}<<the precision of the plot. 0.01 will
-#provide an ecdf from the 0.00, 0.01, .02,  ..., 1.00 quantiles, 0.001 will provide
-#a 0.000, 0.001, 0.002, ..., 1.000 quantiles,... >>
+#{prec}<<the precision of the plot. 0.001 will
+#provide an ecdf from the 0.000, 0.001, .002,  ..., 1.000 quantiles.>>
 #{stat}<<the function used for estimates (2D \samp{mc} or \samp{mcnode}). By default the median.>>
 #{lim}<<a vector of numbers (between 0 and 1) indicating the enveloppe (2D \samp{mc} or \samp{mcnode}) . Maybe \samp{NULL} or empty.>>
 #{na.rm}<<Should NA values be discarded>>
@@ -19,7 +18,8 @@ plot.mc <- function(x, prec=0.01, stat = c("median","mean"), lim = c(0.025,0.975
 #{ylab}<<vector of labels for the y-axis.>>
 #{main}<<vector of main titles of the graph.>>
 #{draw}<<Should the plot be drawn?>>
-#{\dots}<<further arguments to be passed to \samp{plot.ecdf}.>>
+#{paint}<<Should the enveloppes be filled?>>
+#{\dots}<<further arguments to be passed to \samp{plot.stepfun}.>>
 #DETAILS
[…]

### 2010-02-13 pkg/mc2d/R/mcratio.R
Some improvements on mc2d
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@85 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,105 @@
+#<<BEGIN>>
+mcratio <- function(x, pcentral = .5, pvar = .975, punc = .975, na.rm=FALSE)
+#TITLE Ratio of uncertainty and the variability
+#DESCRIPTION
+# Provides measures of variability, uncertainty, and both combined for an 
+#\samp{mc} or an \samp{mcnode} object.
+#KEYWORDS distribution
+#INPUTS
+#{x}<<an \samp{mc} or an \samp{mcnode} object>>
+#[INPUTS]
+#{pcentral}<<the quantile for the central tendency.>>         .
+#{pvar}<<the quantile for the measure of variability.>>
+#{punc}<<the quantile for the measure of uncertainty.>>
+#{na.rm}<<a logical value indicating whether NA values should be stripped before the computation proceeds.>>
+#VALUE
+#A matrix. 
+#DETAILS
+#The function evaluates three ratios for each \samp{mcnode}. Given:
+#{A}<<the \samp{(100 * pcentral)}th percentile of uncertainty (by default the median) for the \samp{(100 * pcentral)}th percentile of variability>>
+#{B}<<the \samp{(100 * pcentral)}th percentile of uncertainty for the \samp{(100 * pvar)}th percentile of variability>>
+#{C}<<the \samp{(100 * punc)}th percentile of uncertainty for the \samp{(100 * pcentral)}th percentile of variability>>
+#{D}<<the \samp{(100 * punc)}th percentile of uncertainty for the \samp{(100 * pvar)}th percentile of variability>>
+
+#The following ratio are estimated
+#{*}<<Variability Ratio: B / A>>
+#{*}<<Uncertainty Ratio: C / A>>
+#{*}<<Overall Uncertainty Ratio: D / A>>
+
+#For multivariate nodes, the statistics are evaluate on each dimension or on statistics according to the corresponding \samp{outm} value. 
+
+#EXAMPLE
+#data(total)
+#mcratio(total, na.rm=TRUE)
+#AUTHOR Regis Pouillot
+#REFERENCE
+#Ozkaynak, H., Frey, H.C., Burke, J., Pinder, R.W. (2009) "Analysis of coupled model uncertainties in source-to-dose modeling
[…]

### 2010-02-13 pkg/mc2d/R/mcstoc.R

@@ -218,17 +218,18 @@ mcstoc <- function(func=runif, type=c("V","U","VU","0"), ..., nsv=ndvar(), nsu=n
 
        func <- function(...){
           argsd <- list(...)
-          linf <- argsd$linf
-          lsup <- argsd$lsup
-          if(any(pmin(linf,lsup) != linf)) stop("linf should be <= lsup")  #not min since vectors should be recycled
           nnfin <- argsd[[nsample]]
+          linf <- if(length(argsd$linf) <= nnfin) argsd$linf else rep(argsd$linf, length.out=nnfin)      # solve a problem when linf was multivariate
+          lsup <- if(length(argsd$lsup) <= nnfin) argsd$lsup else rep(argsd$lsup, length.out=nnfin)      # solve a problem when linf was multivariate
+          if(any(pmin(linf,lsup) != linf)) stop("linf should be <= lsup")  #not min since vectors should be recycled
           argsd$linf <- argsd$lsup <- argsd[[nsample]] <- NULL
           
           pinf <- do.call(pfun,c(list(q=linf),argsd),quote=TRUE)
           psup <- do.call(pfun,c(list(q=lsup),argsd),quote=TRUE)
+           
 
           if(!lhs) lesp <- runif(nnfin,min=pinf,max=psup)
-          else     lesp <- lhs(distr="runif",nsv=dimf[1],nsu=dimf[2], nvariates=dimf[3], min=pinf,max=psup) 
+          else     lesp <- lhs(distr="runif",nsv=dimf[1],nsu=dimf[2], nvariates=dimf[3], min=pinf, max=psup) 
 
           data <- do.call(qfun,c(list(p=lesp),argsd))
           data[pinf==0 & data > lsup] <- NaN          #ex: rtrunc("lnorm",10,linf=-2,lsup=-1)
@@ -236,7 +237,7 @@ mcstoc <- function(func=runif, type=c("V","U","VU","0"), ..., nsv=ndvar(), nsu=n
           data[is.na(linf) | is.na(lsup)] <- NaN      #ex: rtrunc("norm",10,sd=-2)
           return(data)}
     }
-    else func <- function(...) {                    # LHS only
+    else func <- function(...) {                      # LHS only
           argsd <- list(...)
           argsd[[nsample]] <- NULL
           lesp <- lhs(distr="runif",nsv=dimf[1],nsu=dimf[2],nvariates=dimf[3],min=0,max=1)
[…]

### 2010-02-13 pkg/mc2d/R/plot.tornado.R

@@ -26,7 +26,12 @@ plot.tornado <- function(x,which=1,name=NULL,stat=c("median","mean"),xlab="metho
 #EXAMPLE
 #AUTHOR Regis Pouillot
 #CREATED 07-08-01
-#REVISED 07-08-01
+#EXAMPLE
+#data(ec)
+#x <- evalmcmod(ec$modEC2, nsv=100, nsu=100, seed=666)
+#tor <- tornado(x,7)
+#plot(tor)
+#REVISED 10-02-10
 #--------------------------------------------
 #
 {
@@ -45,15 +50,18 @@ plot.tornado <- function(x,which=1,name=NULL,stat=c("median","mean"),xlab="metho
   stat <- ifelse(stat=="mean" && nr!=1, 2 ,1)
   val <- val[,order(abs(val[stat,])),drop=FALSE]
 	if(nr==1){
-		segments(0,1:nc,val,1:nc,lwd=4)
+		segments(0,1:nc,val,1:nc,lwd=2)
     }
 
   else {
+		segments(0,1:nc,val[stat,],1:nc,lwd=2,col="grey")
     points(val[stat,],1:nc)
     if(nr>3){
         val <- apply(val[3:nr,],2,range)
-        segments(val[1,],1:nc,val[2,],1:nc,lwd=4)
-        }
+        segments(val[1,],1:nc,val[2,],1:nc,lwd=2)
+        points(val[1,],1:nc, pch=3)
+        points(val[2,],1:nc, pch=3)
+            }
     }
 
 		abline(v=0)

### 2010-02-15 pkg/mc2d/R/mcstoc.R
Ooops, two small errors
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@86 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -25,7 +25,7 @@ mcstoc <- function(func=runif, type=c("V","U","VU","0"), ..., nsv=ndvar(), nsu=n
 #{seed}<<The random seed used for the evaluation. If \samp{NULL} the \samp{seed} is unchanged.>>
 #{rtrunc}<<Should the distribution be truncated? See \code{\link{rtrunc}}.>>
 #{linf}<<If truncated: lower limit. May be a scalar, an array or a mcnode.>>
-#{lsup}<<If truncated: upper limit. May be a scalar, an array or a mcnode.>>
+#{lsup}<<If truncated: upper limit. May be a scalar, an array or a mcnode. \samp{lsup} should be pairwise strictly greater then \samp{linf}>>
 #{lhs}<<Should a Random Latin Hypercube Sampling be used? see \code{\link{lhs}}>>
 #VALUE
 #An \samp{mcnode} object.
@@ -61,7 +61,7 @@ mcstoc <- function(func=runif, type=c("V","U","VU","0"), ..., nsv=ndvar(), nsu=n
 #See examples.
 #
 #\samp{trunc=TRUE} is valid for univariates distributions only.
-#The distribution will be truncated on \samp{[linf, lsup]}.
+#The distribution will be truncated on \samp{(linf, lsup]}.
 #The function 'func' should have a 'q' form (with first argument 'p') and a 'p' form, as
 #all current random generator of the \samp{stats} library.
 #Example : 'rnorm' (has a 'qnorm' and a 'pnorm' form), 'rbeta', 'rbinom', 'rgamma', ...
@@ -221,9 +221,10 @@ mcstoc <- function(func=runif, type=c("V","U","VU","0"), ..., nsv=ndvar(), nsu=n
           nnfin <- argsd[[nsample]]
           linf <- if(length(argsd$linf) <= nnfin) argsd$linf else rep(argsd$linf, length.out=nnfin)      # solve a problem when linf was multivariate
           lsup <- if(length(argsd$lsup) <= nnfin) argsd$lsup else rep(argsd$lsup, length.out=nnfin)      # solve a problem when linf was multivariate
-          if(any(pmin(linf,lsup) != linf)) stop("linf should be <= lsup")  #not min since vectors should be recycled
+          lmax <- max(length(linf),length(lsup))
+          if(any(rep(linf, length.out = lmax) >= rep(lsup, length.out = lmax))) stop("linf should be < lsup")  #recycle vectors
[…]

### 2010-02-15 pkg/mc2d/R/mcratio.R
Ooops, two small errors
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@86 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -82,8 +82,9 @@ mcratio <- function(x, pcentral = .5, pvar = .975, punc = .975, na.rm=FALSE)
     if(is.list(x)) return(lapply(x,LESSTAT,prob=prob,wdim=wdim))
     return(apply(x,wdim,quantile,prob = prob,na.rm=na.rm))    
         }
-  res   <- mapply(LESSTAT, x,  MoreArgs = list(prob=c(central,percvar), wdim=2),SIMPLIFY=FALSE)
-  res   <- mapply(LESSTAT, res,MoreArgs = list(prob=c(central,percunc), wdim=1),SIMPLIFY=FALSE)
+        
+  res   <- mapply(LESSTAT, x,  MoreArgs = list(prob=c(pcentral,pvar), wdim=2),SIMPLIFY=FALSE)
+  res   <- mapply(LESSTAT, res,MoreArgs = list(prob=c(pcentral,punc), wdim=1),SIMPLIFY=FALSE)
 
   LESRATIOS <- function(x){
     if(is.list(x)) return(lapply(x,LESRATIOS))

### 2010-03-02 pkg/mc2d/R/mcstoc.R
Solve some bugs
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@89 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -180,14 +180,18 @@ mcstoc <- function(func=runif, type=c("V","U","VU","0"), ..., nsv=ndvar(), nsu=n
             if(!(type=="VU" || typethismc=="0" || typethismc==type)) stop("Incompatible type of nodes") # incompatible node
 
             dimm <- dim(argsd)
-            if( (typethismc == "V" && dimm[1] != nsv) ||
-                (typethismc == "U" && dimm[2] != nsu) ||
-                (typethismc == "VU" && (dimm[1] != nsv || dimm[2] != nsu)))
-                   stop("Nodes of incompatible dimensions")                      # incompatible dimension
+            if ((typethismc == "V" && dimm[1] != nsv) || 
+                (typethismc == "U" && dimm[2] != nsu) || 
+                (typethismc == "VU" && (dimm[1] != nsv || dimm[2] != nsu))) 
+                stop("Nodes of incompatible dimensions")
+ # incompatible dimension
 
             if(maxdim3 > 1){  #at least one multivariate node as parameter, need recycling on the third dimension
               if(typethismc=="U") argsd <- apply(argsd, 3, matrix, nrow=maxdim1, ncol=maxdim2, byrow=TRUE)  # recycling U as matrix (maxdim1*maxdim2) x nvariates
-                else argsd <- apply(argsd, 3 ,matrix, nrow=maxdim1, ncol=maxdim2)                           # recycling 0, V, VU as matrix (maxdim1*maxdim2) x nvariates
+                else {
+                  if(maxdim1 ==1 && maxdim2 ==1) argsd <- matrix(argsd, nrow=1) # Very special case to be added 
+                  else argsd <- apply(argsd, 3, matrix, nrow = maxdim1, ncol = maxdim2)
+                  }        # recycling 0, V, VU as matrix (maxdim1*maxdim2) x nvariates
             }                           
 
             else { dim(argsd) <- NULL    # as vector
@@ -228,14 +232,14 @@ mcstoc <- function(func=runif, type=c("V","U","VU","0"), ..., nsv=ndvar(), nsu=n
        func <- function(...){
           argsd <- list(...)
           nnfin <- argsd[[nsample]]
-          linf <- if(length(argsd$linf) <= nnfin) argsd$linf else rep(argsd$linf, length.out=nnfin)      # solve a problem when linf was multivariate
[…]

### 2010-03-02 pkg/mc2d/R/plot.tornado.R

@@ -51,16 +51,17 @@ plot.tornado <- function(x,which=1,name=NULL,stat=c("median","mean"),xlab="metho
   val <- val[,order(abs(val[stat,])),drop=FALSE]
 	if(nr==1){
 		segments(0,1:nc,val,1:nc,lwd=2)
+    points(val,1:nc, pch=4, lwd=2)
     }
 
   else {
 		segments(0,1:nc,val[stat,],1:nc,lwd=2,col="grey")
-    points(val[stat,],1:nc)
+    points(val[stat,],1:nc, pch=4, lwd=2)
     if(nr>3){
         val <- apply(val[3:nr,],2,range)
         segments(val[1,],1:nc,val[2,],1:nc,lwd=2)
-        points(val[1,],1:nc, pch=3)
-        points(val[2,],1:nc, pch=3)
+        points(val[1,],1:nc, pch="|", lwd=2)
+        points(val[2,],1:nc, pch="|", lwd=2)
             }
     }
 

### 2010-03-02 pkg/mc2d/inst/doc/docmcEnglish.rnw
Some improvements on mc2d
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@85 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Solve some bugs
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@89 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage{graphicx}
+\usepackage[scale=0.8, centering]{geometry}
 
@@
 \begin{document}
+
+\newpage
 \maketitle
+\newpage
 
@@
 \begin{itemize}
-\item A medium level of experience in R. Please refer to the Manual An
-Introduction to R available with R distribution if needed;
+\item A medium level of experience in R. Please refer to the Manual \emph{An
+Introduction to R} available with R distribution if needed;
 \item Some knowledge about Monte-Carlo simulation (its basic principles
-and its utility) and about Quantitative Risk Assessment (QRA). 
+and its use) and about Quantitative Risk Assessment (QRA). 
 \end{itemize}
@@
 The definitive reference remains the documentation associated with
-the package.
+the package. It is recommended to try the examples while reading the 
+explanations.
 
@@
 \texttt{mc2d} means Two-Dimensional Monte-Carlo (\emph{Monte-Carlo
-à Deux Dimensions}). This package :
+à 2 Dimensions}). This package provides:
 \begin{itemize}
-\item provides additional probability distributions;
-\item provides tools to construct One-Dimensional and Two-Dimensional Monte-Carlo
+\item additional probability distributions;
+\item tools to construct One-Dimensional and Two-Dimensional Monte-Carlo
 Simulations;
-\item provides tools to analyse One-Dimensional and Two-Dimensional Monte-Carlo
+\item tools to analyse One-Dimensional and Two-Dimensional Monte-Carlo
 Simulations.
@@
 In a previous version, some tools to fit parametric distributions
-to data were included. Because these functions are useful for other
+to data were included. Because these functions can be useful for other
 purposes, they have been moved to a separate package called \texttt{fitdistrplus}.
@@
 \texttt{mc2d} was built for QRA in the Food Safety domain but it can
-be used in other QRA domains.
+be used in other frameworks.
 
@@
 According to international recommendations, a QRA should reflect the
-variability in the risk and calculate the uncertainty
+variability in the risk and take into account the uncertainty
 associated with the risk estimate. The variability represents
-temporal, geographical and/or individual heterogeneity of the risk
+for instance temporal, geographical and/or individual heterogeneity of the risk
 for a given population. The uncertainty is understood as stemming
@@
 
-In order to estimate the natural variability of the risk,
+In order to reflect the natural variability of a modelled risk,
 a Monte-Carlo simulation approach may be useful: the empirical distribution
-of the risk within the population may be estimated from the mathematical
+of the risk within the population may be obtained from the mathematical
 combination of distributions reflecting the variability of parameters
@@
 A two-dimensional (or second-order) Monte-Carlo simulation was proposed
-to estimate the uncertainty in the risk estimates stemming
+to superimpose the uncertainty in the risk estimates stemming
 from parameter uncertainty \cite{CULLEN-FREY-1999}. A two-dimensional
@@
 and uncertainty in the output may be
-estinated separately. It may be described as following (see Figure
+assessed separately. It may be described as following (see Figure
 \ref{fig:Mc2d}):
@@
 \begin{enumerate}
-\item The parameters of the model should be divided into three categories:
-the parameters whose distributions reflect variability only,
-hereinafter denoted as variable parameters, the parameters
-whose distributions reflect uncertainty only, denoted as uncertain
-parameters and the parameters whose distributions reflect both
-uncertainty and variability. For this latter category, a hierarchical
-structure, using hyper-parameters, should be specified: if
+\item The parameters of the model should be divided into four categories:
+the fix parameters, the parameters whose distributions reflect variability only,
+hereinafter denoted as \emph{variable parameters}, the parameters
+whose distributions reflect uncertainty only, denoted as \emph{uncertain
+parameters} and the parameters whose distributions reflect both
+uncertainty and variability, denoted as \emph{variable and uncertain parameters}.
+For this latter category, a hierarchical
+structure, using hyper-parameters, has to be specified: if
 a parameter is both uncertain and variable, one should be able to
 specify an empirical or parametric distribution representing variability.
-This distribution is conditional upon other parameters for which there
-is some associated uncertainty. As an example, one should be able
+This distribution is conditional upon other parameters to which is
+associated the uncertainty. As an example, one should be able
 to specify a relationship such as
-\[ X\left|a,b\right.\sim N\left(a,b\right) \]
+\[ R\left|a,b\right.\sim N\left(a,b\right) \]
 where the specified normal distribution represents variability in
-$x$ conditional upon parameters $a$ and $b$. Hyperdistributions,
+$r$ conditional upon parameters $a$ and $b$. Hyperdistributions,
 such as
@@
 \[ b\sim Unif\left(l_{b},u_{b}\right) \]
-represent the uncertainty in the parameters $a$ and $b$;
+represent the uncertainty in the parameters $a$ and $b$ with uniform distribution;
 \item A set of uncertain parameters are randomly sampled from their respective
 distributions; 
-\item The QRA is performed using a classical (one-dimensional) Monte-Carlo
+\item The model is evaluated using a classical (one-dimensional) Monte-Carlo
 simulation of size $N_{v}$, treating the uncertain parameters as
@@
 \item Steps 2) and 3) are performed a large number ($N_{u}$) of times ,
-leading to $N_{u}$ sets of statistics; 
+leading to $N_{u}$ sets of statistics. The uncertain parameters are generated
+with respect to their uncertain distributions; 
 \item As output, the $50^{th}$ percentile (median) of each statistic
@@
 \end{enumerate}
-%
+
 \begin{figure}
@@
 \texttt{mc2d} is a set of R functions
-that will help to develop such two-dimensional Monte-Carlo simulations.
-The main difference from the procedure described above is that \texttt{mc2d}
+that will help to implement such two-dimensional Monte-Carlo simulations.
+The main point to understand is that \texttt{mc2d}
 uses arrays of (at least) two dimensions to derive the results: the
@@
 beef in <3 year old children.}
+
+The model is built assuming that:
 \begin{itemize}
-\item We assume that, in a given batch of ground beef, \emph{E. coli} O157:H7
+\item in a given batch of ground beef, \emph{E. coli} O157:H7
 are randomly distributed with a mean concentration of $c=10$ bacteria
 (cfu) per gram of product;
-\item We assume that no bacterial growth occurs in storage, since the product
+\item no bacterial growth occurs in storage, since the product
 is kept frozen until it is cooked, just before consumption;
@@
 cooking practices: 
-
 \begin{itemize}
@@
 \end{itemize}
-\item The variability in steak serving sizes $s$ for <3 year children was
-estimated in a consumption survey. The best fit to the data
-was a gamma distribution with parameters: $shape=3.93,\: rate=0.0806$.
+\item The variability in steak serving sizes $s$ for <3 year children can 
+be described with a gamma distribution with parameters: $shape=3.93,\: rate=0.0806$.
 \item The dose-response relationship, describing the probability of illness,
@@
 \end{itemize}
-The question is: \emph{What is the distribution of the risk of
+The question is: \emph{What is the risk of
 illness in the population that consumed the contaminated lot?}
@@
 n & \sim & Poisson\left(c\times i\times s\right)\\
-P & = & 1-\left(1-0.001\right)^{n}
+r & = & 0.001\\
+P & = & 1-\left(1-r\right)^{n}
 \end{eqnarray*}
@@
 library(mc2d)
-ndvar(1000)
+ndvar(1001)
 conc <- 10
@@
 @
+
+The \texttt{print} output provides for each \texttt{node}: 
+the \texttt{mode} of the variable (numeric or logical), 
+\texttt{nsv}, the size of the variability dimension (here, 1000), 
+\texttt{nsu}, the size of the uncertainty dimension (here, 1 since no uncertainty is considered),
+\texttt{nva}, the number of variate (here, all the variables are univariate),
+some statistics such as the \texttt{min}imum, the \texttt{mean}, 
+the \texttt{median} and the \texttt{max}imum,
+evaluated after gathering the two dimensions,
+the number of missing data \texttt{Nas}, 
+the \texttt{type} of the variable (\texttt{0} for fix,
+\texttt{V} for variable, \texttt{U} for uncertain or \texttt{VU} for variable
+and uncertain (here, all nodes are variable parameters only)
+and, finally, the level of output \texttt{outm}, for multivariate parameters.
+The \texttt{summary} output proovides additionnal statistics.
 
@@
 while the distributions of $i$ and $s$ represent variability. $n$,
-which is a function of $c$, $i$ and $s,$ will be both variable
+which is a function of $c$, $i$ and $s,$ will become both variable
 \emph{and} uncertain.
@@
 argument is provided for each distribution, indicating whether the
-parameter distribution represents uncertainty (\texttt{type}=\texttt{U}),
-variability (\texttt{type}=V), or both
-(\texttt{type}=\texttt{VU}).
+parameter distribution represents variability (\texttt{type=V}, by default),
+uncertainty (\texttt{type=U}),
+ or both (\texttt{type=VU}).
 
@@
 distribution is chosen for each parent parameter. For developing
-an empirical distribution from data , the \texttt{fitdistrplus}
-package is recommended;
+a parametric distribution from data, the \texttt{fitdistrplus}
+package is a convenient way;
 \item For each parameter, an \texttt{mcnode} object is constructed (key
@@
 An \texttt{mcnode} object is the basic element of an \texttt{mc} object.
-It is an array of dimension \texttt{nsv} * \texttt{nsu} * \texttt{nvariates}
+An \texttt{mcnode}  is associated to one variable while an \texttt{mc} is 
+a set of associated variables.
+
+An \texttt{mcnode} is an array of dimension \texttt{nsv} * \texttt{nsu} * \texttt{nvariates}
 where $nsv$ is the dimension of variability, $nsu$ is the dimension
 of uncertainty and $nvariates$ is the number of variates of the \texttt{mcnode}%
-\footnote{In this section, we will only consider \texttt{mcnode}s with $nvariates=1$.%
+\footnote{In this section, we will only consider univariate \texttt{mcnode}s,
+that is \texttt{mcnode}s with $nvariates=1$.%
 }. Four types of \texttt{mcnode} exist: 
@@
 \noindent \begin{centering}
-\includegraphics[angle=270, width=.5\textwidth]{Illustration}
+\includegraphics[angle=270, width=.7\textwidth]{Illustration}
 \par\end{centering}
@@
 
-There are 5 ways to construct an \texttt{mcnode} object:
+There are several ways to construct an \texttt{mcnode} object:
 \begin{enumerate}
@@
 \item \texttt{mcprobtree} is a special function that constructs an \texttt{mcnode}
-from other \texttt{mcnodes} using a probability tree;
+from other \texttt{mcnodes} using a  mixture of distribution, called ''probability tree''
+in QRA;
 \item Some functions, such as \texttt{==} or \texttt{>}
@@
 
-The \texttt{mcstoc} function is written as%
-\footnote{as is standard in R, most arguments have logical default values and
-will be infrequently modified.%
-}:
+The \texttt{mcstoc} function \footnote{from \texttt{stoc}hastic)} is written as
+\footnote{as is standard in R, most arguments have default values and
+will be infrequently modified.}:
 
@@
 0), ..., nsv=ndvar(), nsu=ndunc(), nvariates=1,
-outm=each, nsample=n,
+outm=each, nsample='n',
 seed=NULL, rtrunc=FALSE, linf=-Inf, lsup=Inf, lhs=FALSE)}
@@
 This latter is calculated by the function according to \texttt{func},
-\texttt{type}, \texttt{nsv,} \texttt{nsu} and \texttt{nvariates}.
-If the name of the argument specifying the size of the sample is not
-\texttt{n} (\emph{e.g.} functions \texttt{rhyper} and \texttt{rwilcox},
-see table \ref{tab:Distributions}), the name of this parameter should
-be provided in the \texttt{nsample} argument. \emph{Note that all
-of the following arguments should be named};
+\texttt{type}, \texttt{nsv}, \texttt{nsu} and \texttt{nvariates}.
+\emph{Note that all of the following arguments should be named};
 \item \texttt{nsv} and \texttt{nsu} are the number of samples needed in
-the variability and uncertainty dimension, respectively. By default,
+the variability and uncertainty dimensions, respectively. By default,
 these values are the ones provided by \texttt{ndvar()} and \texttt{ndunc()},
 respectively;
-\item \texttt{nvariates} is the desired number of variates in the \texttt{mcnode};
-\item \texttt{outm} is the default output for multivariate nodes;
-\item \texttt{seed} optionally specifies a seed for the random number generator;
+\item \texttt{nvariates} is the desired number of variates in the \texttt{mcnode}. 
+see section \ref{sec:Multivar};
+\item \texttt{outm} is the default output for multivariate nodes. see section \ref{sec:Multivar};
+\item \texttt{nsample} is the name of the argument of \texttt{func} specifying 
+the size of the sample. It is usually \texttt{n}, with the notable exception of
+the \texttt{rhyper} and \texttt{rwilcox} where \texttt{nsample} should be 
+changed in \texttt{nn}.
+\item \texttt{seed} optionally specifies a seed for the random number generator.
+If \texttt{NULL}, the seed is unchanged;
 \item \texttt{rtrunc} allows truncation of a distribution between \texttt{linf}
-and \texttt{lsup}. This function is not valid for every distribution
+and \texttt{lsup}. This option is not valid for every distribution
 (see table \ref{tab:Distributions}). See the \texttt{rtrunc} function
@@
 sampling could be written%
-\footnote{Note that the mean and the standard deviation of the untruncatednormal
+\footnote{Note that the mean and the standard deviation of the untruncated normal
 distribution are not preserved in the truncated distribution.%
@@
 and the multivariate normal distribution (\texttt{rmultinormal}).
-The multinomial distribution has been adapted (vectorized): \texttt{rmultinomial}
+The multinomial distribution has been adapted (vectorized) and \texttt{rmultinomial}
 (library \texttt{mc2d}) should be used in place of \texttt{rmultinom}
@@
 
-Another way to construct an \texttt{mcnode} object is \emph{via} the
+Another way to construct a \texttt{mcnode} object is \emph{via} the
 \texttt{mcdata} function, when data are available.
@@
 and variability coherently within the model. Logically, the rules
-are as follows (illustrated here with a +)%
-\footnote{These rules are not the standard R rules for recycling.%
-}:
+are as follows (illustrated here with a \texttt{+}, 
+see table \ref{tab:Ops})\footnote{These rules are not the standard R rules for recycling.}:
 \begin{itemize}
-\item \texttt{0} + \texttt{0}
-= \texttt{0};
-\item \texttt{0} + \texttt{V}
-= \texttt{V}
-\item \texttt{0} + \texttt{U}
-= \texttt{U};
-\item \texttt{0} + \texttt{VU}
-= \texttt{VU};
-\item \texttt{V} + \texttt{V}
-= \texttt{V};
-\item \texttt{V} + \texttt{U}
-= \texttt{VU}: the \texttt{U}
+\item \texttt{0} + \texttt{0} = \texttt{0};
+\item \texttt{0} + \texttt{V} = \texttt{V}
+\item \texttt{0} + \texttt{U} = \texttt{U};
+\item \texttt{0} + \texttt{VU}= \texttt{VU};
+\item \texttt{V} + \texttt{V} = \texttt{V};
+\item \texttt{V} + \texttt{U} = \texttt{VU} \footnote{the \texttt{U}
 mcnode is recycled by row, the \texttt{V}
-mcnode is recycled in the standard manner by column;
-\item \texttt{V} + \texttt{VU}
-= \texttt{VU}: the \texttt{V}
-mcnode is recycled in the standard manner by column;
-\item \texttt{U} + \texttt{U}
-= \texttt{U};
-\item \texttt{U} + \texttt{VU}
-= \texttt{VU}: the \texttt{U}
-mcnode is recycled by row;
-\item \texttt{VU} + \texttt{VU}
-= \texttt{VU}
+mcnode is recycled in the standard manner by column.};
+\item \texttt{V} + \texttt{VU} = \texttt{VU} \footnote{the \texttt{V}
+mcnode is recycled in the standard manner by column.};
+\item \texttt{U} + \texttt{U} = \texttt{U};
+\item \texttt{U} + \texttt{VU} = \texttt{VU} \footnote{the \texttt{U}
+mcnode is recycled by row.};
+\item \texttt{VU} + \texttt{VU} = \texttt{VU}
 \end{itemize}
+
+\begin{table}
+\caption{\label{tab:Ops}\texttt{mcnode}s combinations}
+
+\centering{}\begin{tabular}{|c||c|c|c|c|}
+\hline 
+ & 0 & V & U & VU\tabularnewline
+\hline
+\hline 
+0 & 0 & V & U & VU\tabularnewline
+\hline
+V & V & V & VU & VU\tabularnewline
+\hline 
+U & U & VU & U & VU\tabularnewline
+\hline 
+VU & VU & VU & VU & VU\tabularnewline
+\hline
+\end{tabular}
+\end{table}
+
 Thus, in our example:
@@
 \texttt{mcnode}s: it is a \texttt{VU} \texttt{mcnode} with
-variability in the row dimension and uncertainty in the column dimension
-. \texttt{risk} is a function of a \texttt{U} and a \texttt{VU}
+variability in the row dimension and uncertainty in the column dimension. 
+\texttt{risk} is a function of a \texttt{U} and a \texttt{VU}
 \texttt{mcnode}: it is therefore a \texttt{VU} \texttt{mcnode}.
@@
 
-The \texttt{mcprobtree} function can be used if a probability
-tree is needed to construct an \texttt{mcnode}. Assume that the
+The \texttt{mcprobtree} function can be used if a mixture of distribution
+ is needed to construct an \texttt{mcnode}. Assume that the
 distribution representing the uncertainty on \texttt{conc} was not
@@
 \texttt{NA} (missing data), \texttt{NaN} (\emph{Not A Number}),
-or finite .
+finite or infinite .
 
@@
 samples \cite{IMAN-CONOVER-1982}. Assume that a study suggests that
-people who consume rare ground beef alsoconsume larger serving sizes.
+people who consume rare ground beef also consume larger serving sizes.
 We could specify this relation using:
@@
 
-Note that the resulting correlation (around 0.4) is obviously an approximation
+\footnote{Note that the resulting correlation (around 0.4) is obviously an approximation
 to the desired value of 0.5, because a discrete distribution (\texttt{cook}:
-3 categories) is correlated with a continuous distribution (\texttt{serving}).
+3 categories) is correlated with a continuous distribution (\texttt{serving}).}
 
@@
 
-One can re-run the model with various dimensions or random seeds in
+One can re-run the model with different dimensions or random seeds in
 one line:
@@
 
-If evaluating a high-dimensional model, R may exceed its memory limit.
-\texttt{evalmccut} evaluates a 2-dimensional Monte-Carlo model (written
+When evaluating a high-dimensional model, R may exceed its memory limit.
+To overcome this drawback, \texttt{evalmccut} evaluates a 2-dimensional 
+Monte-Carlo model (written
 with the \texttt{mcmodelcut} function) using a loop, and calculates
@@
 Readers should refer to the corresponding documentation for further
-details. Our example would be written as:
+details. Our example would be written as\footnote{Note that the use of 
+a \texttt{tornado} function in the model should
+be avoided as it slows the \texttt{evalmccut} function considerably.
+The \texttt{tornado} function will be rewritten in the near future
+to improve its performance.}:
 
@@
 
-Note that the use of a \texttt{tornado} function in the model should
-be avoided as it slows the \texttt{evalmccut} function considerably.
-The \texttt{tornado} function will be rewritten in the near future
-to improve its performance.
-
 \subsection{Analysing an \texttt{mc} Object}
@@
 Then, the median and the quantiles provided in the \texttt{lim} argument
-are evaluated on these statistics. Of course, these arguments should
-be changed if other quantiles are needed.
+are evaluated on these statistics.
 
@@
 
-The \texttt{plot} function provides a graph of the empirical distribution
+The \texttt{plot} function provides a graph of the cumulative empirical distribution
 function of the estimate (mean or median) of the quantiles.
@@
 \texttt{plot (x, prec = 0.01, stat = c(median,
-mean), lim = c(0.025,0.975), na.rm =
+mean), lim = c(0.025,0.25, 0.75, 0.975), na.rm =
 TRUE, griddim = NULL, xlab = NULL, ylab = Fn(x),
@@
 
-The output shows the impact of the uncertain nodes (type U
+The output shows the impact of the uncertain nodes (type \texttt{U}
 nodes) and some statistics (mean, median and, here, the $99^{th}$percentile)
-calculated on the variability dimension (type V
-and type VU nodes) of some output statistics
+calculated on the variability dimension (type \texttt{VU} nodes) of some output statistics
 .
@@
 
-Here is a (silly) example building a linear model (in fact \texttt{ndunc()}
+Here is a (silly) example building a linear model (in fact 100
 linear models) between the \texttt{risk} and the \texttt{dose} within
@@
 parameter $\alpha=(2,3,5)$. We want to derive the distribution (variability)
-of steak eaten by 500 babies sampled from these 500 families.
+of steak masses eaten by 500 babies sampled from such families.
 
@@
 <<sh20, echo=true, eval=true>>=
-(p <- mcstoc(rdirichlet, type="U", nsu=100, nvariates=3, alpha=c(2,3,5)))
-s <- mcstoc(rmultinomial,type="VU", nsv=500, nsu=100, nvariates=3,size=500, prob=p)
+(p <- mcstoc(rdirichlet, type="U", nvariates=3, alpha=c(2,3,5)))
+s <- mcstoc(rmultinomial,type="VU", nvariates=3, size=500, prob=p)
 summary(s)
@@
 
-Assume that each member of these families eats a normal distribution
+As a second example, assume that each member of these families eats a normal distribution
 (variability) of steak with mean 100, 150 and 250 g. There is a positive
@@
 <<sh21, echo=true, eval=true>>=
+sigma <- matrix(c(10,2,-5,2,10,-5,-5,-5,10), ncol=3)
 (x <- mcstoc(rmultinormal,type="V", nvariates=3, mean=c(100,150,250), 
-                              sigma=c(10,2,-5,2,10,-5,-5,-5,10))) 
+                              sigma=as.vector(sigma))) 
 cor(x[,1,])
@@
 In this example, \texttt{mean} could be variable or uncertain, as
-well as \texttt{sigma}%
-\footnote{Caution: the use of a varying \texttt{sigma} can make the analysis
-very slow.%
-}. You could have used, for an uncertain mean:
+well as \texttt{sigma}\footnote{\texttt{rmultinormal} is a vectorized
+version of \texttt{rmvnorm} (library \texttt{mvtnorm}).}. 
+You could have used, for an uncertain mean:
 
@@
 mun <- mcstoc(rnorm, type="U", nvariates=3, mean=m, sd=20)
-x <- mcstoc(rmultinormal, type="VU", nvariates=3, mean=mun, sigma=c(10,2,-5,2,10,-5,-5,-5,10))
+x <- mcstoc(rmultinormal, type="VU", nvariates=3, mean=mun, sigma=as.vector(sigma))
 cor(x[,1,])
@@
 
-Finally, multivariate nodes may be useful to derive a nonparametric
+As a third example, multivariate nodes may be useful to derive a nonparametric
 bootstrap. Assume that, based on a study, you obtained 6 individuals
@@
 <<sh22, echo=true, eval=true>>=
-(x <- mcstoc(rempiricalD, type="U", outm=c('min','mean','max'), nvariates=30, 
-              values=c(100,150,170,200),prob=c(6,12,6,6)))
+val <- c(100,150,170,200)
+pr <-  c(6,12,6,6)
+out <- c('min','mean','max')
+(x <- mcstoc(rempiricalD, type="U", outm=out, nvariates=30, 
+              values=val,prob=pr))
 mcstoc(rempiricalD,type="VU", values=x)
@@
 none, \texttt{each}, the default, for a series of statistics
-for each variate, or, as in the example, a vector of functions that
+for each variate, or, as in the example, a vector of function names that
 are applied over all the 30 variates).
@@
 The recycling rules in \texttt{mc2d} regarding the \texttt{nvariate}
-dimension are as follows: the recycling will be done from \texttt{nvariates=1}
+dimension are as follows: if needed, the recycling will be done from \texttt{nvariates=1}
 to \texttt{nvariates=n} with $n>1$. This allows you to use multivariates
@@
 We can \emph{i}) build a bivariate node reflecting these two
-options; ii) transfer these options into the final risk estimate.
+independent options; ii) transfer these options into the final risk estimate.
 We obtain a bivariate node for the risk, one using the first hypothesis,
@@
 risk <- 1-(1-r)^dose
-EC5 <- mc(conc,cook,serving,expo,dose,r,risk)
+EC5 <- mc(conc,risk)
 summary(EC5)
@@
 risk <- 1-(1-r)^dosetot
-EC6 <- mc(conc,cook,serving,expo,dose,dosetot,r,risk)
+EC6 <- mc(conc,risk)
 summary(EC6)
@@
 <<Crypto3, echo=true, eval=true>>=
-ndvar(1001);ndunc(1001)
+ndvar(1001);ndunc(101)
 mcstoc(rempiricalD,type="V",values=inca)
@@
 bottled water on those days). We could try a mixture of distributions,
-with \texttt{0}and non-\texttt{0}
-data.
+with \texttt{0} and non-\texttt{0} data.
 
@@
 We will use the \texttt{mcprobtree} function to construct a mixture
-of 0and non-0
+of 0 and non-0
 distributions:
@@
 estDR <- function(pos,ref){
-   -glm(cbind(ref$ni-pos,pos) ~ ref$dose + 0, binomial(link="log"))$coefficients}
+   -glm(cbind(ref$ni-pos,pos) ~ ref$dose + 0, 
+    binomial(link="log"))$coefficients}
 
@@
 
-This result can be compared (roughly since there is some differences
-in the modelsvariability) to the results shown in Table 2 in \cite{POUILLOT-2004}.
+This result can be compared (roughly since there are some differences
+in the model variability) to the results shown in Table 2 in \cite{POUILLOT-2004}.
 
@@
 to constuct and analyse their models, and that it may help in developing
-two-dimensional simulations. Nevertheless,
-\texttt{mc2d} is currently under development:
-
-\noindent \begin{center}
-\emph{CHECK YOUR MODEL CAREFULLY AND EXAMINE RESULTS TO DETECT BUGS}
-\par\end{center}
-
-and, if you would like to improve it, join us at
+two-dimensional simulations. Nevertheless, \texttt{mc2d} is still under 
+development: \emph{please report any bugs you get to rpouillot@yahoo.fr.}
+
+If you would like to improve \texttt{mc2d}, join us at
 
@@
 
-Please refer any comments or bugs to rpouillot@yahoo.fr.
-
 \bibliographystyle{plain}

### 2010-03-02 pkg/mc2d/R/extractvar.R

@@ -9,7 +9,7 @@ extractvar <- function(x, which = 1)
 #INPUTS
 #{x}<<a multivariates \samp{mcnode}.>>
 #[INPUTS]
-#{which}<<which variate should be extracted.>>
+#{which}<<a vector. which variate(s) should be extracted?>>
 #{\dots}<< \samp{mcnode}s to be gathered in a multivariate \samp{mcnode}.
 # These \samp{mcnode}s should be of same type and dimension.>>
 #VALUE
@@ -20,7 +20,7 @@ extractvar <- function(x, which = 1)
 # \code{\link{mcnode}} for \samp{mcnode} objects.
 #EXAMPLE
 #x <- mcdata(0:3,"0",nvariates = 4)
-#y <- extractvar(x, 2) 
+#y <- extractvar(x, c(1,3)) 
 #y
 #addvar(x,y)
 #AUTHOR Regis Pouillot
@@ -29,8 +29,8 @@ extractvar <- function(x, which = 1)
 {
   if(missing(x) || !inherits(x,"mcnode")) stop("extractvar need a mcnode object")
   dimm <- dim(x)
-  if((which > dimm[3]) || (which < 1)) stop("Incorrect value of which")
-  x <- mcdata(x[,,which],type=typemcnode(x),nsv=dimm[1],nsu=dimm[2],nvariates=1)
+  if(any(which > dimm[3]) || (which < 1)) stop("Incorrect value of which")
+  x <- mcdata(x[,,which],type=typemcnode(x),nsv=dimm[1],nsu=dimm[2],nvariates=length(which))
   return(x)
 }
 

### 2010-03-02 pkg/mc2d/R/plot.mc.R

@@ -1,5 +1,5 @@
 #<<BEGIN>>
-plot.mc <- function(x, prec=0.001, stat = c("median","mean"), lim = c(0.025,0.25, 0.75, 0.975), na.rm=TRUE, griddim = NULL, xlab = NULL, ylab = "Fn(x)", main = "", draw = TRUE, paint=TRUE,...)
+plot.mc <- function(x, prec=0.001, stat = c("median","mean"), lim = c(0.025, 0.25, 0.75, 0.975), na.rm=TRUE, griddim = NULL, xlab = NULL, ylab = "Fn(x)", main = "", draw = TRUE, paint=TRUE,...)
 #TITLE Plots Results of a Monte Carlo Simulation
 #DESCRIPTION
 # Plots the empirical cumulative distribution function of a \samp{mcnode} or a \samp{mc} object ("0" and "V" nodes) or the
@@ -24,8 +24,8 @@ plot.mc <- function(x, prec=0.001, stat = c("median","mean"), lim = c(0.025,0.25
 #\samp{plot.mcnode} is a user-friendly function that send the \samp{mcnode} to \samp{plot.mc}.</>
 #For \samp{"VU"} and \samp{"U"} \samp{mcnode}s, quantiles are calculated using \code{\link{quantile.mc}}
 #within each of the \samp{nsu} simulations (i.e. by columns of each \samp{mcnode}). The medians (but may be
-#the means using \samp{stat="mean"}) calculated from the \samp{nsu} values are plotted. The 0.025 and 0.975 quantiles (default values
-#of \samp{lim}) of these quantiles are used as the enveloppe.
+#the means using \samp{stat="mean"}) calculated from the \samp{nsu} values are plotted. The 0.025 and 0.975 quantiles, 
+#and the 0.25 and 0.75 quantiles (default values of \samp{lim}) of these quantiles are used as the enveloppe.
 #REFERENCE
 #Cullen AC and Frey HC (1999) Probabilistic techniques in exposure assessment. Plenum Press, USA, pp. 81-155. 
 #VALUE

### 2010-05-18 pkg/rs00/demo/rs00.demo.creating.R
plus de rs00
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@94 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,60 +1,0 @@
-chemin <- searchpaths()[grep("rs00",searchpaths())];
-chemin <- paste(chemin,"/files/",sep="");
-# creating rs00 S4 object
-# 
-# 10_02_11
-#
-#
-
-
-############################################################
-form3titre("Demonstrating the creation of 'des' objects");
-
-des1 <- new("des",name="To describe any object you want",
-                  orig="For instance, a reference",
-                  time=today("nor"),
-                  defi="Just a charcter(1) providing the definition",
-                  role="here, just an illustration",
-                  comm=c("The comment field",
-                         "It is the only ones which can comprise several paragraphs")
-            );
-print(des1,quoi="a");
-
-print(char2des("Just The name"));
-
-############################################################
-form3titre("Demonstrating the creation of 'daf' objects");
-
-# from an existing data frame
-ddd <- data.frame(A=11:36,B=letters,C=sin(1:26),row.names=NULL);
-dafd <- new("daf",des=char2des("From a data.frame"),
-                  what="d",
-                  valu="ddd"
-           );
-print(dafd);
-
-# from an existing text file
-if (!exists("chemin")) {
-    cat("A file into the private 'files' directory is needed\n");
-    erreur("The package must be installed to run this demo");
-}
-ttt <- paste(chemin,"rs00.demo.creating.txt",sep="");
-print(ttt);
-print(readLines(ttt));
-
-daft <- new("daf",des=char2des("From a text file"),
-                  what="t",
-                  valu=ttt
-           );
-print(daft);
-
-
-# from an existing function without argument
-fff <- function() {
-    data.frame(A=11:36,B=letters,C=sin(1:26),row.names=NULL);
-}
-daff <- new("daf",des=char2des("From a function"),
-                  what="f",
-                  valu="fff"
-           );
-print(daff);

### 2010-05-18 pkg/rs00/demo/rs00.demo.easyp3cut.R

@@ -1,27 +1,0 @@
-chemin <- searchpaths()[grep("rs00",searchpaths())];
-chemin <- paste(chemin,"/files/",sep="");
-# script test for easyp3cut et al.
-# 
-# 09_10_12
-#
-#
-form3titre("Testing the use of 'easyp3cut' and 'easyp3stickback'");
-
-# preparing the expressions to parse
-ex <- character(0);
-ex[1] <- "(A)+{B}+[C]";
-
-# parsing and recompacting
-for (ii in sjl(ex)) {
-    res0 <- ex[ii];
-    form3line(wid=1);
-    form3titre("Initial Expression",0);
-    print(res0);
-    res1 <- easyp3cut(ex[ii]);
-    form3titre("Decomposed Expression",0);
-    print(res1);
-    res2 <- easyp3stickback(res1);
-    form3titre("Recomposed Expression",0);
-    print(res2);
-    form3line(wid=1);
-}

### 2010-05-18 pkg/rs00/tests/rs00.demo.unex.R

@@ -1,28 +1,0 @@
-#
-# This was added by paquet.pl (version 0.04) (beginning)
-#
-library("rs00");
-#
-# This was added by paquet.pl (version 0.04) (end)
-#
-chemin <- searchpaths()[grep("rs00",searchpaths())];
-chemin <- paste(chemin,"/files/",sep="");
-# script test for unex
-# 
-# 09_09_22 09_09_28 09_09_29
-#
-#
-form3titre("Demonstrating the use of 'unex'");
-
-# preparing the data.frame
-nb <- 23; nf <- c(2,7,3);
-set.seed(1234);
-xv <-  as.data.frame(matrix(0,nb,0));
-for (ii in 1:length(nf)) {
-    xv <- cbind(xv,factor(sample(nf[ii],nb,TRUE),levels=1:nf[ii]));
-}
-names(xv) <- LETTERS[1:length(nf)];
-# computing the index
-Index <- unex(xv,nf);
-# printint the result
-print(cbind(xv,Index));

### 2010-05-18 pkg/rs00/R/f0.rs00.code.r
plus de rs00
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@94 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,4017 +1,0 @@
-
-#<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<
-rs003k <- function(whi)
-#TITLE (00) assigns the constants for the rs00 layer
-#DESCRIPTION
-# defines or returns the constants used within /rbsb00/. 
-# The performed action depends on the argument.
-#DETAILS
-# All constant names start with 'rbsb.'.
-# This solution was adopted to replace
-# a set of global constants that I had difficulty
-# to make acceptable with R packages standards.
-# It is recommended not to modify these constants
-# unless you are completely aware of the consequences.\cr
-# The constants can be any object type.
-#PKEYWORDS helpful
-#KEYWORDS misc
-#INPUTS
-#{whi}    <<a character(1) indicating either to reset or
-#           to return the names or the current values. The three
-#           values are \code{RESET}, \code{reset}, \code{names}, \code{definitions} or \code{values}.>>
-# >>
-#[INPUTS]
-#VALUE
-# When \code{whi=="RESET"} nothing (but the assignments are
-# performed for all layers: only \code{rs00} since it is the first one.).
-# When \code{whi=="reset"} nothing (but the assignments of 
-# the layer \code{rs00} are performed).
-# When \code{whi=="names"} the names as a character vector.
-# When \code{whi=="definitions"} the definitions as a named character vector.
-# When \code{whi=="values"} the values through a named list.
-#EXAMPLE
-## First assign the standard values
-# rs003k("reset");
-# print(rbsb.msi);
-## to get the short labels
-# rs003k("names");
-## to obtain the current values
-# rs003k("values");
-#REFERENCE
-#SEE ALSO
-#CALLING
-#COMMENT
-#FUTURE
-#AUTHOR J.-B. Denis
-#CREATED 09_09_16
-#REVISED 10_02_11
-#--------------------------------------------
-{
-# checking
-if (!expr3present(whi,c("RESET","reset","names","definitions","values"))) {
-    print.default(whi);
-    stop("rs003k does not accept this argument");
-}
-#
-#if (whi=="RESET") { }
-#
-# definition of the different constants
-sc <- character(0);
[…]

### 2010-05-18 pkg/rs00/demo/rs00.demo.unex.R

@@ -1,21 +1,0 @@
-chemin <- searchpaths()[grep("rs00",searchpaths())];
-chemin <- paste(chemin,"/files/",sep="");
-# script test for unex
-# 
-# 09_09_22 09_09_28 09_09_29
-#
-#
-form3titre("Demonstrating the use of 'unex'");
-
-# preparing the data.frame
-nb <- 23; nf <- c(2,7,3);
-set.seed(1234);
-xv <-  as.data.frame(matrix(0,nb,0));
-for (ii in 1:length(nf)) {
-    xv <- cbind(xv,factor(sample(nf[ii],nb,TRUE),levels=1:nf[ii]));
-}
-names(xv) <- LETTERS[1:length(nf)];
-# computing the index
-Index <- unex(xv,nf);
-# printint the result
-print(cbind(xv,Index));

### 2010-05-18 pkg/rs00/tests/rs00.demo.creating.R

@@ -1,67 +1,0 @@
-#
-# This was added by paquet.pl (version 0.04) (beginning)
-#
-library("rs00");
-#
-# This was added by paquet.pl (version 0.04) (end)
-#
-chemin <- searchpaths()[grep("rs00",searchpaths())];
-chemin <- paste(chemin,"/files/",sep="");
-# creating rs00 S4 object
-# 
-# 10_02_11
-#
-#
-
-
-############################################################
-form3titre("Demonstrating the creation of 'des' objects");
-
-des1 <- new("des",name="To describe any object you want",
-                  orig="For instance, a reference",
-                  time=today("nor"),
-                  defi="Just a charcter(1) providing the definition",
-                  role="here, just an illustration",
-                  comm=c("The comment field",
-                         "It is the only ones which can comprise several paragraphs")
-            );
-print(des1,quoi="a");
-
-print(char2des("Just The name"));
-
-############################################################
-form3titre("Demonstrating the creation of 'daf' objects");
-
-# from an existing data frame
-ddd <- data.frame(A=11:36,B=letters,C=sin(1:26),row.names=NULL);
-dafd <- new("daf",des=char2des("From a data.frame"),
-                  what="d",
-                  valu="ddd"
-           );
-print(dafd);
-
-# from an existing text file
-if (!exists("chemin")) {
-    cat("A file into the private 'files' directory is needed\n");
-    erreur("The package must be installed to run this demo");
-}
-ttt <- paste(chemin,"rs00.demo.creating.txt",sep="");
-print(ttt);
-print(readLines(ttt));
-
-daft <- new("daf",des=char2des("From a text file"),
-                  what="t",
-                  valu=ttt
-           );
-print(daft);
-
-
-# from an existing function without argument
-fff <- function() {
-    data.frame(A=11:36,B=letters,C=sin(1:26),row.names=NULL);
-}
-daff <- new("daf",des=char2des("From a function"),
-                  what="f",
-                  valu="fff"
-           );
-print(daff);

### 2010-05-18 pkg/mc2d/R/mc.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@95 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -72,8 +72,11 @@ mc <- function(..., name=NULL, devname=FALSE)
   rv <- vector(mode="list",length=0)
   nom <- character(0)
   
+  if(!all(sapply(args,inherits,"mcnode")|sapply(args,inherits,"mc"))) stop("arguments should be mc or mcnode objects")
+  
   for(i in 1:length(args)){   # Should find better                                                  # loop to help memory
-    if(is.list(args[[i]])){
+    
+	if(is.list(args[[i]])){
       rv <- c(rv,args[[i]])
       if(devname) nom <- c(nom,paste(nameobj[i],names(args[[i]]),sep="."))
       else nom <- c(nom,names(args[[i]]))}

### 2010-05-18 pkg/mc2d/R/plot.mc.R

@@ -34,8 +34,17 @@ plot.mc <- function(x, prec=0.001, stat = c("median","mean"), lim = c(0.025, 0.2
 #\code{\link{ecdf}}, \code{\link{plot}}, \code{\link{quantile.mc}}
 #EXAMPLE
 #data(total)
+
 #plot(xVUM3)
+### only one enveloppe corresponding to quantiles 0.025 and 0.975
+#plot(xVUM3,lim=c(0.025,0.975)) 
+### only one enveloppe not painted
+#plot(xVUM3,lim=c(0.025,0.975),paint=FALSE) 
+
+#def.par <- par(no.readonly = TRUE)
+#par(mar=c(4,4,1,1))
 #plot(total)
+#par(def.par)
 #AUTHOR Regis Pouillot
 #CREATED 07-08-01
 #REVISED 10-02-10

### 2010-05-18 pkg/mc2d/R/pmin.R

@@ -1,0 +1,217 @@
+#<<BEGIN>>
+pmin <- function (..., na.rm = FALSE)
+#TITLE Maxima and Minima for \samp{mcnode}s
+#DESCRIPTION
+# Returns the parallel maxima and minima of the input values.
+#KEYWORDS utilities
+#INPUTS
+#{\dots}<<One or more \samp{mcnodes}s or one or more \samp{mcnode}s and vector(s) of compatible size. Note that one \samp{mcnode} must be at the first place.>> 
+#{na.rm}<<a logical indicating whether missing values should be removed.>>
+#VALUE
+#an \samp{mcnode} of adequate type and dimension.
+#DETAILS
+#\samp{pmax} and \samp{pmin} take one or more \samp{mcnode} and/or vectors as arguments and return a \samp{mcnode} of adequate type and size 
+#giving the "parallel" maxima (or minima) of the \samp{mcnode} and\or vectors. 
+#Note that the first element of ... should be an \samp{mcnode}. 
+#The resulting type of \samp{mcnode} is variable according to the elements that are passed. The same rules as in \code{\link{Ops.mcnode}} are applied. 
+#SEE ALSO
+#\code{\link{min}}, \code{\link{Ops.mcnode}}
+#EXAMPLE
+#ndvar(10);ndunc(21)
+#x <- mcstoc(rnorm,"V")
+#pmin(x,0)
+#y <- mcdata(rep(c(-1,1),length=ndunc()),"U")
+#unclass(pmin(x,y))
+#--------------------------------------------
+ UseMethod("pmin")
+
+#<<BEGIN>>
+pmax <- function (..., na.rm = FALSE) 
+#ISALIAS pmin
+#--------------------------------------------
+ UseMethod("pmax")
+
+
+#<<BEGIN>>
+pmin.default <- base::pmin
+#ISALIAS pmin
+#--------------------------------------------
+
+#<<BEGIN>>
+pmax.default <- base::pmax
+#ISALIAS pmin
+#--------------------------------------------
+
+#<<BEGIN>>
+pmin.mcnode <- function(..., na.rm = FALSE)
+#ISALIAS pmin
+#--------------------------------------------
+{
+    argsd <- list(...)
+
+    largsd <- length(argsd)
+	if(largsd == 1) return(argsd)
+	
+	outm <- attr(argsd[[1]],"outm")
+	
[…]

### 2010-05-18 pkg/mc2d/inst/doc/docmcEnglish.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@95 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 %\VignettePackage{mc2d}
-%\VignetteDepends{mvtnorm}
+%\VignetteDepends{mvtnorm,fitdistrplus}
 
@@
 to specify a relationship such as
-\[ R\left|a,b\right.\sim N\left(a,b\right) \]
+\[ r\left|a,b\right.\sim N\left(a,b\right) \]
 where the specified normal distribution represents variability in
@@
 
+\emph{Warning}: the data are fictitous and the model is over-simplified 
+to better illustrate the use of the package: 
+the results will not and should not be interpreted.
+
 The model is built assuming that:
@@
 \end{eqnarray*}
-
 where $emp\left(X,P\right)$ is an empirical
@@
 $P_{i}$.We will use a classical one
-dimensional Monte-Carlo simulation, with 1,000 iterations. Using the
+dimensional Monte-Carlo simulation, with 1,001 iterations. Using the
 \texttt{mc2d} package, the model may be written as:
@@
 the \texttt{mode} of the variable (numeric or logical), 
-\texttt{nsv}, the size of the variability dimension (here, 1000), 
+\texttt{nsv}, the size of the variability dimension (here, 1001), 
 \texttt{nsu}, the size of the uncertainty dimension (here, 1 since no uncertainty is considered),
@@
 and uncertain (here, all nodes are variable parameters only)
-and, finally, the level of output \texttt{outm}, for multivariate parameters.
+and, finally, the level of output \texttt{outm} (used for multivariate parameters).
 The \texttt{summary} output proovides additionnal statistics.
@@
 
-We will use a two-dimensional Monte-Carlo simulation, with 1,000 iterations
-in the variability dimension and 100 iterations in the uncertainty
+We will use a two-dimensional Monte-Carlo simulation, with 1,001 iterations
+in the variability dimension and 101 iterations in the uncertainty
 dimension. Using the \texttt{mc2d} package, the model may
@@
 <<sh2, eval=true, echo=true>>=
-ndunc(100)
+ndunc(101)
 conc <- mcstoc(rnorm,type="U",mean=10,sd=2)
@@
 around \texttt{conc} and \texttt{r}. The estimate of the mean risk
-is now uncertain. The median of the 100 simulations leads to a best
-estimateof 0.0445, with a 95\% credible
+is now uncertain. The median of the 101 simulations leads to a best
+estimate of 0.0445, with a 95\% credible
 interval of {[}0.191, 0.0730{]}.
@@
 
-An \texttt{mcnode} is an array of dimension \texttt{nsv} * \texttt{nsu} * \texttt{nvariates}
+An \texttt{mcnode} is an array of dimension \texttt{nsv x nsu x nvariates}
 where $nsv$ is the dimension of variability, $nsu$ is the dimension
@@
 people who consume rare ground beef also consume larger serving sizes.
-We could specify this relation using:
+We could specify this relation using\footnote{Note that the resulting correlation (around 0.4) is obviously an approximation
+to the desired value of 0.5, because a discrete distribution (\texttt{cook}:
+3 categories) is correlated with a continuous distribution (\texttt{serving}).}:
 
@@
 @
-
-\footnote{Note that the resulting correlation (around 0.4) is obviously an approximation
-to the desired value of 0.5, because a discrete distribution (\texttt{cook}:
-3 categories) is correlated with a continuous distribution (\texttt{serving}).}
 
@@
 \noindent \begin{flushleft}
-\texttt{plot (x, prec = 0.01, stat = c(median,
-mean), lim = c(0.025,0.25, 0.75, 0.975), na.rm =
-TRUE, griddim = NULL, xlab = NULL, ylab = Fn(x),
-main = , draw = TRUE, ...)}
+\texttt{plot(x, prec = 0.001, stat = c("median", "mean"), lim = c(0.025, 
+    0.25, 0.75, 0.975), na.rm = TRUE, griddim = NULL, xlab = NULL, 
+    ylab = "Fn(x)", main = "", draw = TRUE, paint = TRUE, ...)}
 \par\end{flushleft}
@@
 For our example, see Figure \ref{fig:Function-plot}, a default graph.
+The 0.25 and 0.75 quantiles (default values of \texttt{lim}) in the uncertainty dimension
+of the quantiles (variability dimension) are used as the enveloppe.
 
@@
 .
+
+\subsubsection{The \texttt{mcratio} function}
+
+The \texttt{mcratio} function provides measures of variability, uncertainty, 
+and both combined propose by \cite{OZKAYNAK-2009} for an \texttt{mc} or an \texttt{mcnode} object.
+Given: 
+\begin{description}
+\item[A] the median of uncertainty for the median of variability; 
+\item[B] the median of uncertainty for the 97.5th percentile of variability; 
+\item[C] the 97.5th percentile of uncertainty for the median percentile of variability; 
+\item[D] the 97.5th percentile of uncertainty for the 97.5th percentile of variability. 
+\end{description}
+The following ratio are estimated: 
+\begin{itemize}
+\item Variability Ratio: B / A 
+\item Uncertainty Ratio: C / A 
+\item Overall Uncertainty Ratio: D / A
+\end{itemize}
+
+%true
+<<sh19d, echo=true, eval=true, fig=false>>=
+mcratio(risk)
+@
 
@@
 \end{document}
+<- mcstoc(rbeta, type="U", shape1=2.65, shape2=3.64)
+w <- mcstoc(rbeta, type="V", shape1=2.6, shape2=3.4) 
+@
+
+Given that $O_{o}=2$ oocysts are observed in 100 l of water, the
+expected number of oocysts in the sample is \texttt{l}:
+
+<<Crypto11, echo=true, figure=false>>=
+Oo <- 2
+l <- (Oo + mcstoc(rnbinom, type="U", size=Oo+1, prob=Rr))/100
+@
+
+The expected number of oocysts drunk by the individuals is \texttt{Or}
+and the risk ($\times10000$) is estimated by:
+
+<<Crypto12, echo=true, figure=false>>=
+Or <- l * v * w
+P <- 10000 * (1-exp(-r*Or))
+summary(P)
+@
+
+This result can be compared (roughly since there is some differences
+in the modelsvariability) to the results shown in Table 2 in \cite{POUILLOT-2004}.
+
+Improvement: the results for $O_{o}=\left\{ 0,1,2,5,10,20,50,100,1000\right\} $
+can be obtained in one step using:
+
+<<Crypto13, echo=true, eval=false, figure=false>>=
+Oo <- mcdata(c(0,1,2,5,10,20,50,100,1000),type="0",nvariates=9) 
+@
+
+\section*{As a Conclusion}
+
+We think and hope that {}``\texttt{mc2d}'' could help risk assessors
+to constuct and analyse their models, and that it may help in developing
+\textquotedbl{}two-dimensional\textquotedbl{} simulations. Nevertheless,
+\textquotedbl{}\texttt{mc2d}\textquotedbl{} is currently under development:
+
+\noindent \begin{center}
+\emph{CHECK YOUR MODEL CAREFULLY AND EXAMINE RESULTS TO DETECT BUGS}
+\par\end{center}
+
+and, if you would like to improve it, join us at
+
+\noindent \begin{center}
+\emph{\url{http://riskassessment.r-forge.r-project.org/}}
+\par\end{center}
+
+Please refer any comments or bugs to \url{rpouillot@yahoo.fr}.
+
+\bibliographystyle{plain}
+\bibliography{Bibtex}
+
+\end{document}

### 2010-05-18 pkg/mc2d/R/empiricalC.R

@@ -57,12 +57,13 @@ dempiricalC <- function(x, min, max, values, prob=rep(1,length(values)), log=FAL
   d <- rep(NA,length(x))
   d[x > max | x < min] <- 0
   lesquel <- which(!is.na(x) & x <= max & x >= min)
-  quel <- sapply(x[lesquel],function(y) min(which(y < val)))
+  quel <- findInterval(x[lesquel],val)+1
   d[lesquel] <- prob[quel-1]+(x[lesquel]-val[quel-1])/(val[quel]-val[quel-1])*(prob[quel]-prob[quel-1])
 
+  d <- d/Integ
   if(log) d <- log(d)
 	if(any(is.na(d))) warning("NaN in dempiricalC")
-  return(d/Integ)}
+  return(d)}
 
 #<<BEGIN>>
 pempiricalC <- function(q, min, max, values, prob=rep(1,length(values)),lower.tail = TRUE, log.p = FALSE)                 
@@ -72,7 +73,7 @@ pempiricalC <- function(q, min, max, values, prob=rep(1,length(values)),lower.ta
   if(min > max | min > min(values) | max < max(values) | !is.finite(min) | !is.finite(max)) stop("Error in min or max")
 
   val2 <- sort(unique(values))
-  probi   <- dempiricalC(val2,min=min,max=max,values=values,prob=prob,log=FALSE)
+  probi   <- dempiricalC(val2, min=min, max=max, values=values, prob=prob, log=FALSE)
 
   h <- c(val2,max)-c(min,val2)
   a <- c(0,probi)
@@ -87,7 +88,7 @@ pempiricalC <- function(q, min, max, values, prob=rep(1,length(values)),lower.ta
   p[q >= max] <- 1
   p[q <= min] <- 0
   lesquel <- which(!is.na(q) & q < max & q > min)
-  quel <- sapply(q[lesquel],function(y) min(which(y < val)))
+  quel <- findInterval(q[lesquel], val) + 1
 
   p[lesquel] <- probcum[quel-1]+(q[lesquel]-val[quel-1])*
                 (probi[quel-1]+((probi[quel]-probi[quel-1])*(q[lesquel]-val[quel-1])/(2*(val[quel]-val[quel-1]))))
[…]

### 2010-05-18 pkg/rs00/demo/rs00.demo.creating.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@98 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,60 @@
+chemin <- searchpaths()[grep("rs00",searchpaths())];
+chemin <- paste(chemin,"/files/",sep="");
+# creating rs00 S4 object
+# 
+# 10_02_11 10_03_10
+#
+#
+
+
+############################################################
+form3titre("Demonstrating the creation of 'des' objects");
+
+des1 <- new("des",name="To describe any object you want",
+                  orig="For instance, a reference",
+                  time=today("nor"),
+                  defi="Just a charcter(1) providing the definition",
+                  role="here, just an illustration",
+                  comm=c("The comment field",
+                         "It is the only ones which can comprise several paragraphs")
+            );
+print(des1,quoi="a");
+
+print(char2des("Just The name"));
+
+############################################################
+form3titre("Demonstrating the creation of 'daf' objects");
+
+# from an existing data frame
+ddd <- data.frame(A=11:36,B=letters,C=sin(1:26),row.names=NULL);
+dafd <- new("daf",des=char2des("From a data.frame"),
+                  what="d",
+                  valu="ddd"
+           );
+print(dafd);
+
+# from an existing text file
+if (!exists("chemin")) {
+    cat("A file into the private 'files' directory is needed\n");
+    erreur("The package must be installed to run this demo");
+}
+ttt <- paste(chemin,"rs00.demo.data.table.txt",sep="");
+print(ttt);
+print(readLines(ttt));
+
+daft <- new("daf",des=char2des("From a text file"),
+                  what="t",
+                  valu=ttt
+           );
+print(daft);
+
+
+# from an existing function without argument
+fff <- function() {
+    data.frame(A=11:36,B=letters,C=sin(1:26),row.names=NULL);
+}
+daff <- new("daf",des=char2des("From a function"),
+                  what="f",
+                  valu="fff"
+           );
+print(daff);

### 2010-05-18 pkg/rs00/demo/rs00.demo.easyp3cut.R

@@ -1,0 +1,27 @@
+chemin <- searchpaths()[grep("rs00",searchpaths())];
+chemin <- paste(chemin,"/files/",sep="");
+# script test for easyp3cut et al.
+# 
+# 09_10_12
+#
+#
+form3titre("Testing the use of 'easyp3cut' and 'easyp3stickback'");
+
+# preparing the expressions to parse
+ex <- character(0);
+ex[1] <- "(A)+{B}+[C]";
+
+# parsing and recompacting
+for (ii in sjl(ex)) {
+    res0 <- ex[ii];
+    form3line(wid=1);
+    form3titre("Initial Expression",0);
+    print(res0);
+    res1 <- easyp3cut(ex[ii]);
+    form3titre("Decomposed Expression",0);
+    print(res1);
+    res2 <- easyp3stickback(res1);
+    form3titre("Recomposed Expression",0);
+    print(res2);
+    form3line(wid=1);
+}

### 2010-05-18 pkg/rs00/tests/rs00.demo.creating.R

@@ -1,0 +1,67 @@
+#
+# This was added by paquet.pl (version 0.04) (beginning)
+#
+library("rs00");
+#
+# This was added by paquet.pl (version 0.04) (end)
+#
+chemin <- searchpaths()[grep("rs00",searchpaths())];
+chemin <- paste(chemin,"/files/",sep="");
+# creating rs00 S4 object
+# 
+# 10_02_11 10_03_10
+#
+#
+
+
+############################################################
+form3titre("Demonstrating the creation of 'des' objects");
+
+des1 <- new("des",name="To describe any object you want",
+                  orig="For instance, a reference",
+                  time=today("nor"),
+                  defi="Just a charcter(1) providing the definition",
+                  role="here, just an illustration",
+                  comm=c("The comment field",
+                         "It is the only ones which can comprise several paragraphs")
+            );
+print(des1,quoi="a");
+
+print(char2des("Just The name"));
+
+############################################################
+form3titre("Demonstrating the creation of 'daf' objects");
+
+# from an existing data frame
+ddd <- data.frame(A=11:36,B=letters,C=sin(1:26),row.names=NULL);
+dafd <- new("daf",des=char2des("From a data.frame"),
+                  what="d",
+                  valu="ddd"
+           );
+print(dafd);
+
+# from an existing text file
+if (!exists("chemin")) {
+    cat("A file into the private 'files' directory is needed\n");
+    erreur("The package must be installed to run this demo");
+}
+ttt <- paste(chemin,"rs00.demo.data.table.txt",sep="");
+print(ttt);
+print(readLines(ttt));
+
+daft <- new("daf",des=char2des("From a text file"),
+                  what="t",
+                  valu=ttt
+           );
+print(daft);
+
+
+# from an existing function without argument
+fff <- function() {
+    data.frame(A=11:36,B=letters,C=sin(1:26),row.names=NULL);
+}
+daff <- new("daf",des=char2des("From a function"),
+                  what="f",
+                  valu="fff"
+           );
+print(daff);

### 2010-05-18 pkg/rs00/R/zzz.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@98 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,4 @@
+.First.lib <- function(libname, pkgname) {
+    # creating the constant for the rs00 package
+    rs003k("reset");
+}

### 2010-05-18 pkg/rs00/demo/rs00.demo.unex.R

@@ -1,0 +1,21 @@
+chemin <- searchpaths()[grep("rs00",searchpaths())];
+chemin <- paste(chemin,"/files/",sep="");
+# script test for unex
+# 
+# 09_09_22 09_09_28 09_09_29
+#
+#
+form3titre("Demonstrating the use of 'unex'");
+
+# preparing the data.frame
+nb <- 23; nf <- c(2,7,3);
+set.seed(1234);
+xv <-  as.data.frame(matrix(0,nb,0));
+for (ii in 1:length(nf)) {
+    xv <- cbind(xv,factor(sample(nf[ii],nb,TRUE),levels=1:nf[ii]));
+}
+names(xv) <- LETTERS[1:length(nf)];
+# computing the index
+Index <- unex(xv,nf);
+# printint the result
+print(cbind(xv,Index));

### 2010-05-18 pkg/rs00/tests/rs00.demo.unex.R

@@ -1,0 +1,28 @@
+#
+# This was added by paquet.pl (version 0.04) (beginning)
+#
+library("rs00");
+#
+# This was added by paquet.pl (version 0.04) (end)
+#
+chemin <- searchpaths()[grep("rs00",searchpaths())];
+chemin <- paste(chemin,"/files/",sep="");
+# script test for unex
+# 
+# 09_09_22 09_09_28 09_09_29
+#
+#
+form3titre("Demonstrating the use of 'unex'");
+
+# preparing the data.frame
+nb <- 23; nf <- c(2,7,3);
+set.seed(1234);
+xv <-  as.data.frame(matrix(0,nb,0));
+for (ii in 1:length(nf)) {
+    xv <- cbind(xv,factor(sample(nf[ii],nb,TRUE),levels=1:nf[ii]));
+}
+names(xv) <- LETTERS[1:length(nf)];
+# computing the index
+Index <- unex(xv,nf);
+# printint the result
+print(cbind(xv,Index));

### 2010-05-28 pkg/mc2d/R/empiricalC.R
Version with 2 vignettes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@101 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -34,7 +34,7 @@ dempiricalC <- function(x, min, max, values, prob=rep(1,length(values)), log=FAL
 #par(mfrow=c(1,2))
 #curve(dempiricalC(x, min=0, max=6, values, prob), from=-1, to=7, n=1001)
 #curve(pempiricalC(x, min=0, max=6, values, prob), from=-1, to=7, n=1001)
-#AUTHOR Regis Pouillot
+
 #CREATED 08-02-20
 #--------------------------------------------
 {

### 2010-05-28 pkg/mc2d/R/mc.R

@@ -37,7 +37,7 @@ mc <- function(..., name=NULL, devname=FALSE)
 #z <- x * y
 #(m <- mc(x,y,z,name=c('n1','n2','n3')))
 #mc(m,x,devname=TRUE)
-#AUTHOR Regis Pouillot
+
 #CREATED 07-08-01
 #REVISED 07-08-01
 #--------------------------------------------

### 2010-05-28 pkg/mc2d/R/plot.mc.R

@@ -45,7 +45,7 @@ plot.mc <- function(x, prec=0.001, stat = c("median","mean"), lim = c(0.025, 0.2
 #par(mar=c(4,4,1,1))
 #plot(total)
 #par(def.par)
-#AUTHOR Regis Pouillot
+
 #CREATED 07-08-01
 #REVISED 10-02-10
 #--------------------------------------------
@@ -73,8 +73,8 @@ plot.mc <- function(x, prec=0.001, stat = c("median","mean"), lim = c(0.025, 0.2
 	 ylab <- rep(ylab,n)
 
   if(is.null(griddim)) griddim <- beau(n)
-  if(prod(griddim) < n) op <- par(mfrow=griddim,ask=TRUE)
-     else op <- par(mfrow=griddim )
+  if(prod(griddim) < n) op <- par(mfrow=griddim,ask=TRUE,mar=c(5,4,.2,.2))
+     else op <- par(mfrow=griddim, mar=c(5,4,.2,.2))
 
   try({   #to restore par in case of error
 

### 2010-05-28 pkg/mc2d/inst/doc/mc2dLmEnglish.rnw
Version with 2 vignettes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@101 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% mc2d: Tools for Two-Dimensional Monte-Carlo Simulations
%%%
%%%         New TP V.2
%%%

\documentclass[10pt, english]{article}

%\usepackage[letterpaper]{geometry}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[scale=0.8, centering]{geometry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
% sweave commands for vignette
%\VignetteIndexEntry{Case study: L. monocytogenes in cold-smoked salmon}
%\VignettePackage{mc2d}
%\VignetteDepends{mvtnorm,fitdistrplus}

%\makeatother

\begin{document}

% Command for Sweave: figure = .5 the text width
\setkeys{Gin}{width=0.5\textwidth}

\title{Case study: \emph{L.~monocytogenes} in cold-smoked salmon}
\author{R. POUILLOT, M.-L. DELIGNETTE-MULLER, M. CORNU}

\maketitle

The objective of this case study is to assess the risk of invasive listeriosis from consumption
of cold-smoked salmon in France. 
The process of interest lays from the end of the production line in the factory, when the cold-smoked 
salmon is vacuum-packed, to the consumption.

The data and the model are adapted to illustrate the use 
of \texttt{mc2d}: the results will not and \emph{should not} be interpreted as an assessment
of the actual risk of listeriosis from consumption of cold-smoked salmon. 
Interested readers could refer to \cite{POUILLOT-2007} and \cite{POUILLOT-2009} for a complete
risk assessment on that issue.

The model will be developed in a first section, without considering variability or uncertainty (deterministic model).
Variability will then be introduced in a second section, and a last section will consider variability 
and a part of the data uncertainty.

\section{The Model}

In this section, no variability nor uncertainty is considered.
We assess the final level of \emph{L.~monocytogenes} in the product,
the exposure and the risk of invasive listeriosis for an ``average'' individual
of the ``healthy'' French population\footnote{Yes, it makes no sense, but it will help us introducing
smoothly the model.}.

During the logistic, the retail and the home step,
a bacterial growth is modeled considering \emph{i}) the fluctuating temperature
during the various stages and; \emph{ii}) the bacterial competition with the food flora.
We use the models developed and/or used in \cite{POUILLOT-2007}. The data are adapted from
\cite{POUILLOT-2007} and \cite{DELIGNETTE-2006}:

\begin{itemize}
  \item The DMS model predicts the bacterial growth during a stage of duration $d$,
  when the temperature is fluctuating, with an intra-stage average temperature $m_{T}$
  and an intra-stage standard deviation of the temperature $s_{T}$. It is written:
  %
  \begin{equation}
  N_{1}=
  \min\left(
    N_{0} +
    \frac{\mu_{ref}}{\ln(10)} \times d \times 
    \frac{\left(s_{T}^{2}+\left(m_{T}-T_{min}\right)^{2}\right)}{\left(T_{ref}-T_{min}\right)^{2}},
    \: N_{max}\right) \label{DMS}
  \end{equation}
  %
  if $m_{T}>T_{min}$, with $N_{1}$ the $\log_{10}$ concentration of bacteria
  ($\log_{10}$ (CFU/g)) in the product at the end of the stage,
  $N_{0}$ the $\log_{10}$ concentration of bacteria ($\log_{10}$ (CFU/g))
  in the product at the beginning of the stage, $\mu_{ref}$ the specific growth
  rate ($\text{day}^{-1}$) at a reference temperature $T_{ref}$ (\textdegree{}C),
   $T_{min}$ the minimal temperature (\textdegree{}C) of growth
  and $N_{max}$   the maximum achievable concentration in the product ($\log_{10}$ (CFU/g)).
  If $m_{T} \leq T_{min}$, $N_{1}=N_{0}$.
  \item We will use $T_{ref}=25$\textdegree{}C.
  We have in this section $N_{max}=7.27 \log_{10}(\text{CFU/g})$;
  \item The model for \emph{L.~monocytogenes} uses $\mu_{ref,Lm}=6.2\: \text{day}^{-1}$ and
  $T_{min,Lm}=-2.9$\textdegree{}C;
  \item  The same model is used for the food flora, with $\mu_{ref,ff}=4.1\: \text{day}^{-1}$ and
  $T_{min,ff}=-4.5$\textdegree{}C;
  \item The growth model for the bacterial competition consider the Jameson effect,
  i.e.~consider that the bacterial growth of \emph{L.~monocytogenes and} the growth of the food flora
  are stopped as soon as one population reaches $N_{max}$.
\end{itemize}
  In practice, one will evaluate $d_{Lm}$ and $d_{ff}$, the time needed for \emph{L.~monocytogenes}
  or the food flora to reach $N_{max}$, respectively, and model a growth for the
  given stage during an effective duration of $\min(d,d_{Lm},d_{ff})$.
  The time needed to reach $N_{max}$ is evaluated by inverting \eqref{DMS}:
  %
  \[
  d_{\left(N_{1}=N_{max}\right)}=
  \left(N_{max}-N_{0}\right)\times
    \frac{\ln(10)}{\mu_{ref}} \times
    \frac{\left(T_{ref}-T_{min}\right)^{2}}{\left(s_{T}^{2}+\left(m_{T}-T_{min}\right)^{2}\right)}
  \]
  %

The other assumptions are:
\begin{itemize}
  \item  A cold-smoked salmon package is homogeneously contaminated with \emph{L.~monocytogenes}
  at the end of the production at a level of $0.1$ CFU/g;
  \item  The food flora level at the end of the production is $10^{2.78}$ CFU/g;
  \item  The time-temperature profile is:

    \begin{itemize}
    \item  1.1 days at an average temperature of 3.2\textdegree{}C from the factory to the retail (logistic step),
    with an intra-stage standard deviation of the temperature of 2.1 \textdegree{}C;
    \item  4.7 days at an average temperature of 5.5\textdegree{}C at retail
    with an intra-stage standard deviation of the temperature of 1.0 \textdegree{}C;
    \item  4.3 days at an average temperature of 8.2\textdegree{}C in the consumer's home
    with an intra-stage standard deviation of the temperature of 2.0 \textdegree{}C;
    \end{itemize}
  \item  An healthy, non elderly, non pregnant individual eats 35g of this product;
  \item  The individual dose-response model for this population is a one hit model
  \[
  \Pr(\text{Illness} \mid D)=1-(1-r)^{D}
  \]
  with $r=4.7\times10^{-14}$ for an individual from this healthy sub-population.
  The populational dose-response that evaluates the mean risk for
  a population exposed to food where the number of bacteria follows a Poisson distribution of
  mean paramer $D$ is the exponential dose-response
  \[
  \Pr(\text{Illness} \mid D)=1-\exp(r\times D)
  \]
\end{itemize}

The question is ``What is the risk for this `average' individual?''. 
One way to write this model is as following:

%Specify the width opf the output
<<sh0, echo=false, eval=true>>=
options("width"=100,"digits"=3)
set.seed(666)
@

<<sh1, echo=true, eval=true>>=
Nmax <- 7.3
murefLm <- 6.2; TminLm <- -2.9

murefFF <- 4.1; TminFF <- -4.5

Lm0 <- log10(1); FF0 <- 2.78

d1 <- 1.1; mT1 <- 3.2; sdT1 <- 2.1

d2 <- 4.7; mT2 <- 5.5; sdT2 <- 1.0

d3 <- 4.3; mT3 <- 8.2; sdT3 <- 2.0

conso <- 35
r <- 4.7e-14

modGrowth <- function(duration, mTemp, sdTemp,
                      N0Lm, murefLm, TminLm,
                      N0FF, murefFF, TminFF,
                      Nmax, Tref=25) {
  N0Lm <- pmin(N0Lm, Nmax)
  N0FF <- pmin(N0FF, Nmax)
  dLm <- (Nmax-N0Lm) * log(10)/murefLm * (Tref-TminLm)^2 / (sdTemp^2 + (mTemp-TminLm)^2)
  dLm <- ifelse(mTemp < TminLm & N0Lm!=Nmax, Inf, dLm)
  dFF <- (Nmax-N0FF) * log(10)/murefFF * (Tref-TminFF)^2 / (sdTemp^2 + (mTemp-TminFF)^2)
  dFF <- ifelse(mTemp < TminFF & N0FF!=Nmax, Inf, dFF)
  realDuration <- pmin(duration, dLm , dFF)
  xLm <- N0Lm + (mTemp > TminLm) * murefLm/log(10) * (sdTemp^2 + (mTemp - TminLm)^2) /  ((Tref - TminLm)^2) * realDuration
  xFF <- N0FF + (mTemp > TminFF) * murefFF/log(10) * (sdTemp^2 + (mTemp - TminFF)^2) /  ((Tref - TminFF)^2) * realDuration
  return(list(xLm = xLm, xFF=xFF))}

x1 <- modGrowth(d1, mT1, sdT1,
                Lm0, murefLm, TminLm,
                FF0, murefFF, TminFF,
                Nmax)
x2 <- modGrowth(d2, mT2, sdT2,
                x1$xLm, murefLm, TminLm,
                x1$xFF, murefFF, TminFF,
                Nmax)
x3 <- modGrowth(d3, mT3, sdT3,
                x2$xLm, murefLm, TminLm,
                x2$xFF, murefFF, TminFF,
                Nmax)
x3
conta <-10^x3$xLm
conta
expo <- conso * conta
expo
risk <- 1 - (1 - r)^expo
risk
@

\texttt{modGrowth} is a convenient function for the growth model. Within this function
\texttt{dLm} is the time needed for \emph{L.~monocytogenes} to reach \texttt{Nmax},
\texttt{dFF} is the time needed for the food flora to reach \texttt{Nmax} and,
\texttt{realDuration} is the effective time of growth during the stage. Note that: 
\begin{itemize}
\item this function is ``vectorized'', meaning that it can deal with a vector for any of its parameters, returning consequently 
a vector. This is a strength of R, notably for Monte-Carlo simulations, but it requests a bit of knowledge on
the way to code the functions. As an example: \texttt{pmin}, a function that takes one or more vectors 
as arguments and return a single vector giving the ``parallel'' minima of the vectors  is used instead
of the more classical function \texttt{min} function, that would return the maximum or minimum of \emph{all} the values.
Another example is the use of the \texttt{ifelse} instead of \texttt{if};
\item it is also written to handle \emph{all} specific cases that could occur in the Monte-Carlo simulation, such as 
$N_{0} \geq N_{max}$ or $m_{T} \leq T_{min}$ or both, for any or both bacterial populations.    
\end{itemize}
\texttt{x1}, \texttt{x2} and \texttt{x3} are the bacterial concentrations at the end of the logistic, the retail and the home step, respectively.

\section{Including Variability}

We now specify now some variability distributions for some inputs, following \cite{DELIGNETTE-2006} and
\cite{POUILLOT-2007}.
We first have to call the needed libraries, and define the desired number of iterations:

<<shCallLibrary, echo=true, eval=true>>=
library(fitdistrplus)
library(mc2d)
ndvar(10001)
@

\subsection{Specifying Variability Distribution}
\subsubsection{Initial Contamination}
For the initial contamination levels in \emph{L.~monocytogenes}, 
we have a set of 62 enumeration data from a representative sample of
packages of cold smoked salmon positive in detection: 43 samples have less than $0.2$ CFU/g,
7 samples have $0.2$ CFU/g, 4 samples have $0.4$ CFU/g, 2 samples have $0.6$ CFU/g,
and the other values are $0.3$, $1.0$, $1.6$, $2.4$, $5.4$ and $7.0$ CFU/g \cite{POUILLOT-2007}. 
We will use
the \texttt{fitdistrplus} package to fit a normal distribution on the $\log_{10}$ of
these values, taking into account the censored values.
Using the fitted parameters, we model thereafter these initial concentrations
in contaminated packages through a normal distribution truncated\footnote{so that at least one CFU is included in one 100g package}
on $[-2,\infty)\: \log_{10}$ (CFU/g).

For the food flora, we use the distribution proposed by \cite{DELIGNETTE-2006}, 
$N_{0ff} \sim N(2.78,1.14)$.

<<shLm0FF0V, echo=true, eval=true>>=
dataC <- data.frame(
  left =c(rep(NA,43), rep(.2,7),.3,rep(.4,4),1,1.6,.6,.6,2.4,5.4,7),
  right=c(rep(0.2,43),rep(.2,7),.3,rep(.4,4),1,1.6,.6,.6,2.4,5.4,7)
  )
fit <- fitdistcens(log10(dataC), "norm")
fit
Lm0V <- mcstoc(rnorm, mean = fit$est["mean"], sd = fit$est["sd"], rtrunc=TRUE, linf=-2)
FF0V <- mcstoc(rnorm, mean=2.78, sd=1.14)
@
Note that, by default, the type of alea that is modeled is ``variability'' (\texttt{type="V"}).

\subsubsection{Growth Parameters}
Distributions are derived from \cite{DELIGNETTE-2006}: 
\begin{itemize}
\item $N_{max}$ follows a normal distribution with mean 7.27 $\log_{10}$ CFU/g
and standard deviation 0.86 $\log_{10}$ CFU/g;
\item The specific growth rate at the reference temperature of 25\textdegree{}C
for \emph{L.~monocytogenes} follows a normal distribution with mean 6.24 $\text{day}^{-1}$ and standard
deviation 0.75 $\text{day}^{-1}$ truncated on $[0,\infty)$. The minimal growth temperature follows a normal distribution with
mean -2.86\textdegree{}C and standard deviation 1.93\textdegree{}C;
\item The specific growth rate at the reference temperature of 25\textdegree{}C
for the food flora follows a normal distribution with mean 4.12 $\text{day}^{-1}$ and standard
deviation 1.97 $\text{day}^{-1}$ truncated on $[0,\infty)$. The minimal growth temperature follows a normal distribution with
mean -4.52\textdegree{}C and standard deviation 7.6\textdegree{}C.
\end{itemize}

<<shGrowtV, echo=true, eval=true>>=

NmaxV    <- mcstoc(rnorm, mean=7.27, sd = 0.86)

murefLmV <- mcstoc(rnorm, mean = 6.24, sd = 0.75, rtrunc=TRUE, linf=0)
TminLmV  <- mcstoc(rnorm, mean = -2.86, sd = 1.93)

murefFFV <- mcstoc(rnorm, mean = 4.12, sd = 1.97, rtrunc=TRUE, linf=0)
TminFFV <-  mcstoc(rnorm, mean = -4.52, sd = 7.66)
@

\subsubsection{Time-Temperature Profiles}
The time temperature profiles in the three steps are modelled using
the distribution provided in the table \ref{tab:4} (adapted from \cite{POUILLOT-2007} from representative
data from France)\footnote{$\Gamma$ is the Gamma distribution parameterized as $\Gamma(shape,\: scale)$.
The Exponential($x$) ditribution is the exponential distribution with mean $x$.}. We
assume a shelf life of 28 days. A simple way to model this shelf life will
be to have $d_{1}+d_{2}+d_{3}\leq28$ days, with $d_{1}$ the duration of the logistic stage,
$d_{2}$ the duration of the retail stage and $d_{3}$ the duration of the consumer stage\footnote{See
the code for a way to model this shelf life using truncated distributions.};

\begin{table}
 \caption{ \label{tab:4}Time Temperature Profiles}

 \centering{}\begin{tabular}{|c|c|c|c|}
\hline 
 Stage &  Mean Temperature (\textdegree{}C) & Intra-Stage Variance of T (\textdegree{}C) &  time (days)%
\tabularnewline
\hline
\hline 
 logistic &  normal(3.2, 2.2) truncated on {[}-3;25{]} & $\Gamma(1.16, 4.61)$ &  Exponential(1.1)\tabularnewline
\hline 
 retail &  normal(5.5, 2.2) truncated on {[}-3;25{]} & $\Gamma(0.65, 2.09)$ &  Exponential(4.7)\tabularnewline
\hline 
 consumer &  normal(8.2, 3.8) truncated on {[}-3; 25{]} & $\Gamma(0.35, 19.7)$  &  Exponential(4.3)\tabularnewline
\hline
\end{tabular}
\end{table}
%

<<shTtV, echo=true, eval=true>>=
d1V <- mcstoc(rexp, rate = 1/1.1)
mT1V <- mcstoc(rnorm, mean = 3.2, sd = 2.2, rtrunc = TRUE, linf = -3, lsup = 25)
sdT1V <- sqrt(mcstoc(rgamma, shape = 1.16, scale=4.61))

d2V <- mcstoc(rexp, rate = 1/4.7, rtrunc=TRUE, lsup=28-d1)
mT2V <- mcstoc(rnorm, mean = 5.5, sd = 2.2, rtrunc = TRUE, linf = -3, lsup = 25)
sdT2V <- sqrt(mcstoc(rgamma, shape = 0.65, scale=2.09))

d3V <- mcstoc(rexp, rate = 1/4.3, rtrunc=TRUE, lsup=28-(d1+d2))
mT3V <- mcstoc(rnorm, mean = 8.2, sd = 3.8, rtrunc = TRUE, linf = -3, lsup = 25)
sdT3V <- sqrt(mcstoc(rgamma, shape = 0.35, scale=19.7))
@

\subsubsection{Serving Size}
As for the serving size, we consider, from observed data, a discrete empirical distribution with values \cite{POUILLOT-2007}:
$V$=\{10, 12, 19, 20, 30, 34, 40, 50, 60, 67.5, 80, 100, 250\} grams, observed
$F$=\{11, 1, 1 ,29, 12, 1, 41, 4, 4, 1, 4, 1, 1\} time, respectively.
%
<<shConsoV, echo=true, eval=true>>=
consoV <- mcstoc(rempiricalD, values = c(10, 12, 19, 20, 30, 34, 40, 50, 60, 67.5, 80, 100, 250), prob = c(11, 1, 1, 29, 12, 1, 41, 4, 4, 1, 4, 1, 1))
@

\subsection{Applying the Model}
The model may then be evaluated straightforwardly:

<<shModelV, echo=true, eval=true>>=
r <- mcdata(4.7e-14, type = "0")
x1V <- modGrowth(d1V, mT1V, sdT1V,
                Lm0V, murefLmV, TminLmV,
                FF0V, murefFFV, TminFFV,
                NmaxV)
x2V <- modGrowth(d2V, mT2V, sdT2V,
                x1V$xLm, murefLmV, TminLmV,
                x1V$xFF, murefFFV, TminFFV,
                NmaxV)
x3V <- modGrowth(d3V, mT3V, sdT3V,
                x2V$xLm, murefLmV, TminLmV,
                x2V$xFF, murefFFV, TminFFV,
                NmaxV)

contaV <-10^x3V$xLm
expoV <- consoV * contaV
riskV <- 1 - exp(-r * expoV )
Lm1 <- mc(Lm0V, FF0V, NmaxV, murefLmV, TminLmV, murefFFV, TminFFV,
          d1V, mT1V, sdT1V, d2V, mT2V, sdT2V, d3V, mT3V, sdT3V,
          consoV, r, contaV, expoV, riskV)
Lm1
sLm1 <- mc(contaV=Lm1$contaV, expoV=Lm1$expoV, riskV=Lm1$riskV)
summary(sLm1, probs = c(0, 0.5, 0.75, 0.95, 1))
@
\texttt{Lm1} is a \texttt{mc} object that contains all the parameters and outputs.
We extract some of these outputs in \texttt{sLm1} to provide a short summary.

\subsection{Final Estimate}
If 6.5\% of cold-smoked salmon package are contaminated,
if 49,090,000 Frenchs are part of the ``non susceptible''
population and if, on average, those people consume some smoked salmon
$6.4$ times per year, the expected number of cases of listeriosis
from consumption of cold smoked salmon in this population is estimated through:

<<sh4, echo=true, eval=true>>=
meanRisk <-  mcapply(riskV,"var",mean)
expectedN <- round(0.065 * unmc(meanRisk) * 6.4 * 49090000)
expectedN
@

\section{Including (a Part of the) Uncertainty}
We eventually include both variability and uncertainty in the model.
For this example, we will only consider the uncertainty linked to the initial contamination,
the growth parameters and the prevalence.

\subsection{Specifying Uncertainty}
\subsubsection{Initial Contamination}

The uncertainty surrounding the initial contamination levels of the \emph{L.~monocytogenes}
will be modeled using a bootstrap procedure, obtained straightforwardly with the help of the \texttt{fitdistrplus}
package and its \texttt{bootdistcens} function. Before this, we define the number of iterations needed in 
the uncertainty dimension.
%
<<sh8, echo=true, eval=true>>=
ndunc(101)
bootLm0 <- bootdistcens(fit, niter=ndunc())
MLm0 <- mcdata(bootLm0$est$mean,type="U")
SLm0 <- mcdata(bootLm0$est$sd,type="U")
Lm0VU <- mcstoc(rnorm, type="VU", mean=MLm0, sd=SLm0, rtrunc=TRUE, linf=-2)
@

In order to consider uncertainty for the food flora initial contamination, we have, from \cite{DELIGNETTE-2006},
a set of uncertain hyperparameters, $M_{N0ff}$ and $\sigma_{N0ff}$, that are used
as parameters for the uncertain \emph{and} variable parameter $N_{0ff}$:
\begin{eqnarray*}
  N_{0ff} & \sim & N(M_{N0ff},\sigma_{N0ff})\\
  M_{N0ff} & \sim & N(2.78,0.265)\\
  \ln(\sigma_{N0ff}) & \sim & N(0.114,0.172)
\end{eqnarray*}

This hierarchical simulation is written with \texttt{mc2d}:

<<shFF0VU, echo=true, eval=true>>=
MLm0FF <- mcstoc(rnorm, type="U", mean=2.78, sd=0.265)
SLm0FF <- mcstoc(rlnorm, type="U", meanlog=0.114, sdlog=0.172)
FF0VU <- mcstoc(rnorm, type="VU", mean=MLm0FF, sd=SLm0FF)
@
\subsubsection{Growth Parameters}
The uncertainty around $\mu_{ref,Lm}$, $T_{min,Lm}$, $\mu_{ref,ff}$, $T_{min,ff}$
and $N_{max}$ are modeled similarly through the
specification of hyperparameters \cite{DELIGNETTE-2006}\footnote{Note
that there was a typo in \cite{DELIGNETTE-2006} that lead to an error in
\cite{POUILLOT-2007}: the standard-error for $\ln(\sigma_{\mu ref,Lm})$ is $1.03$
and not $-1.03$ as written in \cite{DELIGNETTE-2006}. We will use here the correct value.}:
\begin{eqnarray*}
  \mu_{ref,Lm} & \sim & N(M_{\mu ref,Lm},\sigma_{\mu ref,Lm})\\
  M_{\mu ref,Lm} & \sim & \Gamma(shape: 69.7, scale: 0.0896)\\
  \ln(\sigma_{\mu ref,Lm}) & \sim & N(1.03,0.191)\\
\\
  T_{min,Lm} & \sim & N(M_{Tmin,Lm},\sigma_{Tmin,Lm})\\
  M_{Tmin,Lm} & \sim & N(-2.86,0.459)\\
  \ln(\sigma_{Tmin,Lm}) & \sim & N(0.638,0.208) \\
\\
  \mu_{ref,ff} & \sim & N(M_{\mu ref,ff},\sigma_{\mu ref,ff})\\
  M_{\mu ref,ff} & \sim & \Gamma(shape: 32.5, scale: 0.127)\\
  \ln(\sigma_{\mu ref,ff}) & \sim & N(-0.656,0.221)\\
\\
  T_{min,ff} & \sim & N(M_{Tmin,ff},\sigma_{Tmin,ff})\\
  M_{Tmin,ff} & \sim & N(-4.52,1.23)\\
  \ln(\sigma_{Tmin,ff}) & \sim & N(2.00,0.257) \\
\\
  N_{max} & \sim & N(M_{Nmax},\sigma_{Nmax})\\
  M_{Nmax} & \sim & N(7.27,0.276)\\
  \ln(\sigma_{Nmax}) & \sim & N(-0.172,0.218)
\end{eqnarray*}

with $\mu_{ref}>0$ and $T_{min}<25$. We simply translated the preceeding distributions:

<<sh5, echo=true, eval=true>>=
MmurefLm <- mcstoc(rgamma, type="U", shape=69.7, scale=0.0896)
SmurefLm <- mcstoc(rlnorm, type="U", meanlog = 1.03, sdlog = 0.191)
murefLmVU <- mcstoc(rnorm, type="VU", mean=MmurefLm, sd=SmurefLm, rtrunc=TRUE, linf=0)

MTminLm <- mcstoc(rnorm, type="U", mean=-2.86, sd=0.459)
STminLm <- mcstoc(rlnorm, type="U", meanlog = 0.638, sdlog = 0.208)
TminLmVU <- mcstoc(rnorm, type="VU", mean = MTminLm, sd = STminLm, rtrunc=TRUE, lsup=25)

MmurefFF <- mcstoc(rgamma, type="U", shape=32.5, scale=.127)
SmurefFF <- mcstoc(rlnorm, type="U", meanlog = -.656, sdlog = 0.221)
murefFFVU <- mcstoc(rnorm, type="VU", mean=MmurefFF, sd=SmurefFF, rtrunc=TRUE, linf=0)

MTminFF <- mcstoc(rnorm, type="U", mean=-4.52, sd=1.23)
STminFF <- mcstoc(rlnorm, type="U", meanlog = 2.00, sdlog = 0.257)
TminFFVU <- mcstoc(rnorm, type="VU", mean = MTminFF, sd = STminFF, rtrunc=TRUE, lsup=25)

MNmax <- mcstoc(rnorm, type="U", mean=7.27, sd=0.276)
SNmax <- mcstoc(rlnorm, type="U", meanlog = -0.172, sdlog = 0.218)
NmaxVU <- mcstoc(rnorm, type="VU", mean = MNmax, sd = SNmax)
@
%
\subsubsection{Prevalence}
The prevalence level of contaminated cold-smoked salmon packages
(6.5\%) was estimated from 41 positive packages out of 626 tested \cite{POUILLOT-2007}.
We assume a sensitivity and a specificity of the method of 100\%.
We model the data uncertainty around the true prevalence of contaminated
package using a bayesian reasonning, with a Beta(1, 1) distribution
as a prior. The number of expected cases may be estimated using:
<<shPrevU, echo=true, eval=true>>=
prevU <- mcstoc(rbeta,type="U", shape1=41+1, shape2=626-41+1)
@
\subsection{Applying the Model}
Applying the model is just a copy-paste from the previous version (+ we change the name of the parameters).

<<sh9, echo=true, eval=true>>=
x1VU <- modGrowth(d1V, mT1V, sdT1V,
                Lm0VU, murefLmVU, TminLmVU,
                FF0VU, murefFFVU, TminFFVU,
                NmaxVU)
x2VU <- modGrowth(d2V, mT2V, sdT2V,
                x1VU$xLm, murefLmVU, TminLmVU,
                x1VU$xFF, murefFFVU, TminFFVU,
                NmaxVU)
x3VU <- modGrowth(d3V, mT3V, sdT3V,
                x2VU$xLm, murefLmVU, TminLmVU,
                x2VU$xFF, murefFFVU, TminFFVU,
                NmaxVU)

contaVU <-10^x3VU$xLm
expoVU <- consoV * contaVU
riskVU <- 1 - exp(-r * expoVU)
Lm2 <- mc(Lm0VU, FF0VU, NmaxVU, murefLmVU, TminLmVU, murefFFVU, TminFFVU,
          d1V, mT1V, sdT1V, d2V, mT2V, sdT2V, d3V, mT3V, sdT3V,
          consoV, r, contaVU, expoVU, riskVU)
Lm2
sLm2 <- mc(contaVU=Lm2$contaVU, expoVU=Lm2$expoVU, riskVU=Lm2$riskVU)
summary(sLm2, probs = c(0, 0.5, 0.75, 0.95, 1))
@
The summary provides the estimate of the mean, the standard deviation, the minimum, the median \ldots and a 95\% credible interval.
The estimate is the median of the 101 values obtained in the uncertainty dimension. The credible interval lays
between the $2.5^{\text{th}}$ and the $97.5^{\text{th}}$ percentiles obtained in the uncertainty dimension.

\subsection{Final Estimate}
The uncertainty around the number of expected cases is estimated using:
<<sh7, echo=true, eval=true>>=
meanRiskU <-  mcapply(riskVU,"var",mean)
expectedNU <- round(prevU * meanRiskU * 6.4 * 49090000)
summary(expectedNU)
@

This is an estimate of the uncertainty around the number of cases linked to the uncertainty 
around the initial contamination, the bacterial growth parameter and
the sampling uncertainty for positive packages.
A lot of other uncertainties exist but are not considered here, notably 
the uncertainty around the dose-response model
and parameters. See \cite{POUILLOT-2007,POUILLOT-2009} for a complete analysis.
The study of the model through a Tornado chart in the variability dimension
leads to the Figure \ref{fig:TorLm}. 
It suggests a big impact of the growth rate of \emph{L.~monocytogenes}, of the storage duration
during the consumer step, and of the initial level of \emph{L.~monocytogenes}. 
The Tornado chart in the uncertainty dimension
leads to the Figure \ref{fig:TorLmU} and suggests the impact of the uncertainty around
$N_{max}$ on the mean risk, and thus the expected number of cases.

<<sh3, echo=true, eval=true, figure=false>>=
torn <- tornado(Lm2)
torn
tornunc <- tornadounc.mc(Lm2, quant=NULL)
tornunc
@
\texttt{quant=NULL} request an analysis only on the mean risk, and not on some qunatiles.
<<sh3b, echo=true, eval=false, figure=false>>=
plot(torn)
plot(tornunc, stat="mean risk")
@

\begin{figure}
\caption{\label{fig:TorLm}Tornado chart for the \emph{L.~monocytogenes} example (Variability).}
\begin{center}
<<sh3c, echo=false, eval=true, fig=true>>=
plot(torn)
@
\end{center}
\end{figure}

\begin{figure}
\caption{\label{fig:TorLmU}Tornado chart for the \emph{L.~monocytogenes} example (Uncertainty).}
\begin{center}
<<sh3d, echo=false, eval=true, fig=true>>=
plot(tornunc, stat="mean risk")
@
\end{center}
\end{figure}

As a conclusion, this example illustrates how predictive growth models may 
be implemented within \texttt{mc2d}\ldots

\bibliographystyle{plain}
\bibliography{docmcEnglish}

\end{document}

### 2011-08-11 share/useR11/code/4 useR 2011 v2.R
add slide for useR 2011.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@139 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,242 @@
+library(fitdistrplus)
+
+?mledist
+par(mar=c(4,4,2,1))
+
+
+#------------------------------------------------------------------
+x1 <- c(6.4,13.3,4.1,1.3,14.1,10.6,9.9,9.6,15.3,22.1,13.4,
+13.2,8.4,6.3,8.9,5.2,10.9,14.4)
+x <- seq(0, 1.1*max(x1), length=100)
+
+(f1 <- mledist(x1,"norm"))
+
+dgumbel<-function(x,a,b) 1/b*exp((a-x)/b)*exp(-exp((a-x)/b))
+
+(f2 <- mledist(x1,"gumbel",start=list(a=10,b=5)))
+
+(f3 <- mledist(x1, "exp"))
+
+hist(x1, 10, prob=TRUE)
+lines(x, dnorm(x, f1$estimate[1], f1$estimate[1]), col="red")
+lines(x, dgumbel(x, f2$estimate[1], f2$estimate[1]), col="green")
+lines(x, dexp(x, f3$estimate[1]), col="blue")
+
+legend("topright", lty=1, leg=c("Normal","Gumbel","Exp"), col=c("red","green","blue"))
+
+
+plotdist(x1)
+
+pgumbel<-function(q,a,b) exp(-exp((a-q)/b))
+
+
+plot(ecdf(x1))
+lines(x, pnorm(x, f1$estimate[1], f1$estimate[2]), col="red")
+lines(x, pgumbel(x, f2$estimate[1], f2$estimate[2]), col="green")
+lines(x, pexp(x, f3$estimate[1]), col="blue")
+
+legend("bottomright", lty=1, leg=c("Normal","Gumbel","Exp"), col=c("red","green","blue"))
+
+
+descdist(x1, boot=10, boot.col="blue")
+
+
+(f4 <- mledist(x1, "gumbel", start=list(b=5), fix.arg=list(a=7) ))
+
+plot(ecdf(x1))
+lines(x, pgumbel(x, f2$estimate[1], f2$estimate[2]), col="green")
+lines(x, pgumbel(x, 8, f4$estimate[1]), col="red")
+
+legend("bottomright", lty=1, leg=c("Gumbel(a,b)","Gumbel(a,b), a fixed"), col=c("green","red"))
+
+
+#------------------------------------------------------------------
+
+fit1 <- mledist(x1, "gamma")
+
+fit1bis <- mledist(x1, "gamma", optim.method="BFGS")
+
+#wrap genoud function 
[…]

### 2012-11-15 JSS/fitdistrplus/paper2JSS.Rnw
start jss file.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@183 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[article]{jss}

%%%
%%% TITLE:       fitdistrplus: an R Package for Distribution Fitting Methods
%%% AUTHORS:     Marie Laure Delignette Muller*, Christophe Dutang
%%%              * Corresponding author
%%% AFFILIATION: LBBE, Université Claude Bernard Lyon 1, Lyon, France
%%% ADDRESS:     VetAgro Sup Campus Vétérinaire de Lyon
%%%              1, avenue Bourgelat
%%%              69820 MARCY L'ETOILE, France
%%% PHONE:       +33 4 78 87 27 40
%%% FAX:         +33 4 78 87 27 12
%%% EMAIL:       marielaure.delignettemuller@vetagro-sup.fr
%%%

\usepackage{amsmath,amsthm}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
%% need no \usepackage{Sweave.sty}

\usepackage{wrapfig, subfig}
%\usepackage[a4paper, textwidth=18cm, textheight=27cm]{geometry}

\author{ Marie Laure Delignette-Muller \\ Université Claude Bernard Lyon 1 \And
    Christophe Dutang \\ Université de Strasbourg}
\Plainauthor{Marie Laure Delignette-Muller}

\title{\pkg{fitdistrplus}: An \proglang{R} Package for Distribution Fitting Methods}
\Plaintitle{fitdistrplus: an R Package for Distribution Fitting Methods}

\Abstract{%
    blabla.}

\Keywords{distributions, censored data, \proglang{R}}
\Plainkeywords{distributions, censored data, R}

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-09-04}
%% \Acceptdate{2012-09-04}

\Address{
    Marie Laure Delignette-Muller \\
    VetAgro Sup Campus Vétérinaire de Lyon \\
    1, avenue Bourgelat \\
    69820 MARCY L'ETOILE\\
    France \\
    E-mail: \email{marielaure.delignettemuller@vetagro-sup.fr} \\
    URL: \url{http://lbbe.univ-lyon1.fr/-Delignette-Muller-Marie-Laure-.html}
}

\newcommand{\Sconcordance}[1]{%
  \ifx\pdfoutput\undefined%
  \csname newcount\endcsname\pdfoutput\fi%
  \ifcase\pdfoutput\special{#1}%
  \else%
   \begingroup%
     \pdfcompresslevel=0%
     \immediate\pdfobj stream{#1}%
     \pdfcatalog{/SweaveConcordance \the\pdflastobj\space 0 R}%
   \endgroup%
  \fi}

\begin{document}
\SweaveOpts{concordance=TRUE, prefix.string=figs/jss}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

This manuscript reviews the various features of version 1.0-0 of \pkg{fitdistrplus}. The package
is available from the Comprehensive \proglang{R} Archive Network at \url{http://cran.r-project.org/package=fitdistrplus}.
The development version of the package is located at \proglang{R}-forge as one the packages of the 
project ``Risk Assessment with R''
(\url{http://r-forge.r-project.org/projects/riskassessment/})
The following command will load the package.
%%% R code set default options for all R schunks
<<par4vignette, echo=FALSE, results=hide>>=
options(digits=4, SweaveHooks=list(fig=function() par(mar=c(5.1, 4.1, 1.1, 2.1))) )
@
%%% R code
<<pkgload1, results=hide>>=
library(fitdistrplus)
@

%\section*{Acknowledgments}

%\bibliography{fitdistrplus}

\end{document}

### 2013-02-05 JSS/script_ex_et_plan_JSS2.R
example with mixing.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@208 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,426 @@
+# Exemples pour le papier JSS
+require(fitdistrplus)
+require(actuar)
+options(digits=3)
+
+#### 1/ intro
+
+#### 2/ Fitting distributions to continuous non-censored data
+
+#### 2.1/ Choice of candidate distributions
+
+data(endosulfan)
+ATV <-endosulfan$ATV
+plotdist(ATV) # pas sûr que celui-là soit très intéressant sur cet exemple
+# mais on n'est pas obligé de mettre toutes les figures
+# ---
+# oui c'est pas hyper intéressant sur cet exemple
+
+descdist(ATV,boot=1000)
+# sur celle-là je commenterais le fait qu'en général les écotoxicologues
+# ajustent soit une lognormale soit une loglogistique sur ce type de données
+# mais qu'il semble qu'ici la skewness soit plus forte qu'attendue 
+# pour une loi [log?]normale, par rapport à la kurtosis
+# ---
+# oui, en effet pour une loi lognormale la skewness serait plus petite. 
+# pour la log-logistique on ne peut pas comparer? il me semblait qu'il y avait
+# un graph de C-F avec la loglogistique?
+
+
+# cet exemple n'est pas forcément très illustratif pour cette partie mais intéressant
+# pour la suite
+# ???? FAUT-IL démarrer par un exemple plus simple (l'actuel de la vignette)
+# ???? et présenter celui-là ensuite uniquement?
+# ???? j'attends ton avis là-dessus
+# ---
+# pour les fonctions utilisés ci dessus, je pense groundbeef est mieux que ATV. 
+# car les graphes sont plus démonstratifs et facilement commentable.
+
+
+#### 2.2/ Fit of distributions by MLE and comparison of fits
+
+fln <- fitdist(ATV, "lnorm")
+summary(fln)
+gofstat(fln)
+
+# use of plotdist to find good reasonable initial values for parameters
+plotdist(ATV, "llogis", para=list(shape=1,scale=500))
+fll <- fitdist(ATV, "llogis", start=list(shape=1,scale=500))
+
+# use of plotdist to find good reasonable initial values for parameters
+plotdist(ATV, "pareto", para=list(shape=1,scale=500))
+fP <- fitdist(ATV, "pareto", start=list(shape=1,scale=500))
+
[…]

### 2013-02-06 JSS/script_ex_et_plan_JSS.R
work on the vignette (to continue)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@209 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -9,14 +9,39 @@ options(digits=3)
 
 #### 2.1/ Choice of candidate distributions
 
+data(groundbeef)
+str(groundbeef)
+plotdist(groundbeef$serving)
+descdist(groundbeef$serving, boot=1000)
+
+
+
+
+#### 2.2/ Fit of distributions by MLE and comparison of fits
+
+fw <- fitdist(groundbeef$serving, "weibull")
+summary(fw)
+fg <- fitdist(groundbeef$serving,"gamma")
+fln <- fitdist(groundbeef$serving,"lnorm")
+x11()
+par(mfrow=c(2, 2))
+denscomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"),
+         xlab="serving sizes (g)")
+qqcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+cdfcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+ppcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+
+
 data(endosulfan)
+str(endosulfan)
 ATV <-endosulfan$ATV
-plotdist(ATV) # pas sûr que celui-là soit très intéressant sur cet exemple
+# plotdist(ATV) 
+# pas sûr que celui-là soit très intéressant sur cet exemple
 # mais on n'est pas obligé de mettre toutes les figures
 # ---
 # oui c'est pas hyper intéressant sur cet exemple
 
-descdist(ATV,boot=1000)
+# descdist(ATV,boot=1000)
 # sur celle-là je commenterais le fait qu'en général les écotoxicologues
 # ajustent soit une lognormale soit une loglogistique sur ce type de données
 # mais qu'il semble qu'ici la skewness soit plus forte qu'attendue 
@@ -36,32 +61,29 @@ descdist(ATV,boot=1000)
 # pour les fonctions utilisés ci dessus, je pense groundbeef est mieux que ATV. 
 # car les graphes sont plus démonstratifs et facilement commentable.
 
-
-#### 2.2/ Fit of distributions by MLE and comparison of fits
-
-fln <- fitdist(ATV, "lnorm")
-summary(fln)
-gofstat(fln)
+fendo.ln <- fitdist(ATV, "lnorm")
+# summary(fendo.ln)
+# gofstat(fendo.ln)
 
[…]

### 2013-02-10 JSS/script_ex_et_plan_JSS.R
pdf for collab, dir to remove
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@212 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -120,6 +120,9 @@ qqcomp(list(fendo.ln,fendo.ll,fendo.P,fendo.B),xlogscale=TRUE,ylogscale=TRUE,
 
 # A remplacer par le print de gofstat
 # avec la nouvelle version de fitdistrplus
+
+gofstat(list(fendo.ln, fendo.ll, fendo.P, fendo.B))
+
 cbind(
   ln = c(aic=summary(fendo.ln)$aic,bic=summary(fendo.ln)$bic, unlist(gofstat(fendo.ln)[c("cvm","ks","ad")])),
   ll = c(aic=summary(fendo.ll)$aic, bic=summary(fendo.ll)$bic, unlist(gofstat(fendo.ll)[c("cvm","ks","ad")])),

### 2013-03-27 JSS/fitdistrplus/paper2JSS.Rnw
first working version of JSS template.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@236 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\usepackage{amsmath,amsthm}
+\usepackage{amsmath,amsthm,amssymb}
 \usepackage[utf8]{inputenc}
@@
 
-\usepackage{wrapfig, subfig}
+%\usepackage{wrapfig, subfig}
 %\usepackage[a4paper, textwidth=18cm, textheight=27cm]{geometry}
+
+\newcommand{\sigle}{\textsc}
+%\newcommand{\pkg}{\textbf} defined in jss.cls
+%\newcommand{\code}{\texttt} defined in jss.cls
+%\newcommand{\proglang}{\textsf} defined in jss.cls
 
@@
 \Abstract{%
-    blabla.}
+    The package \pkg{fitdistrplus} provides functions for fitting univariate distributions
+to different types of data (continuous censored or non-censored data and discrete data) and allowing
+different estimation methods (maximum likelihood, moment matching, quantile matching
+and maximum goodness-of-fit estimation).
+Outputs of \code{fitdist} and \code{fitdistcens} functions are S3 objects, for
+which kind generic methods are provided, including \code{summary}, \code{plot} and
+\code{quantile}. 
+This package also provides various functions to compare the fit of several distributions to a same data set 
+and can handle bootstrap of parameter estimates.
+Detailed examples are given in food risk assessment, ecotoxicology and insurance contexts.}
 
-\Keywords{distributions, censored data, \proglang{R}}
-\Plainkeywords{distributions, censored data, R}
+\Keywords{probability distribution fitting, bootstrap, censored data, maximum likelihood, moment matching, 
+quantile matching, maximum goodness-of-fit, distributions, \proglang{R}}
+\Plainkeywords{probability distribution fitting, bootstrap, censored data, maximum likelihood, moment matching, 
+quantile matching, maximum goodness-of-fit, distributions, R}
 
@@
 \begin{document}
-\SweaveOpts{concordance=TRUE, prefix.string=figs/jss}
+\SweaveOpts{concordance=TRUE}
+%\SweaveOpts{concordance=TRUE}
+
+%, prefix.string=figs/jss
 
@@
 
-This manuscript reviews the various features of version 1.0-0 of \pkg{fitdistrplus}. The package
+\section{Introduction}
+\label{Introduction}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Fitting distributions to data is a very common task in statistics and consists in choosing
+a probability distribution modelling the random
+variable, as well as finding parameter estimates for that distribution. 
+This requires judgment and expertise and generally needs an iterative process of
+distribution choice, parameter estimation, and quality of fit assessment.
+In the \proglang{R} (\cite{R13}) package \pkg{MASS} (\cite{MASS}), maximum likelihood estimation 
+is available via the \code{fitdistr} function; other steps of the
+fitting process can be done using other \proglang{R} functions, e.g. \cite{Ricci05}. 
+In this paper, we present the \proglang{R} package \pkg{fitdistrplus} (\cite{fitdistrplus})
+implementing several methods for fitting univariate parametric distribution.
+A first objective in developing this package was to provide \proglang{R} 
+users a set of functions dedicated to help this overall process.
+
+The \code{fitdistr} function estimates distribution parameters by maximizing the likelihood
+function using the \code{optim} function. 
+In some cases, other estimation methods could be prefered, 
+such as maximum goodness-of-fit estimation (also called minimum distance estimation), as
+proposed in the \proglang{R} package \pkg{actuar} with three different goodness-of-fit distances
+(see \cite{actuarJSS}). 
+While developping  the \pkg{fitdistrplus} package, a second objective  was to consider 
+various estimation methods in addition to maximum likelihood estimation (MLE). Functions
+were developped to enable moment matching estimation (MME), quantile matching estimation (QME), and 
+maximum goodness-of-fit estimation (MGE) using eight different distances.
+Moreover, the \pkg{fitdistrplus} package offers the possibility to specify a user-supplied function
+for optimization, 
+useful in cases where classical optimization techniques not included in \code{optim} are more adequate.
+
+In applied statistics, it is frequent to have to fit distributions to censored data \cite{kleinmoeschberger03,
+helsel05,busschaertetal10,lehaetal11,commeauetal12}. 
+The \pkg{MASS} \code{fitdistr} function 
+does not enable maximum likelihood estimation with this type of data. Some packages 
+deal with censored data, especially survival data \cite{survival} (see \cite{hiranoetal94,jordan05}
+for examples),
+but those packages generally focused on specific models, enabling the fit of  a 
+restricted set of distributions. 
+A third objective was thus to provide \proglang{R} users a function 
+to estimate univariate distribution parameters from right-, left- and interval-censored data.
+
+Few packages on \sigle{CRAN} provide estimation procedures for any user-supplied parametric distribution 
+and support different types of data. 
+The \pkg{distrMod} package of \cite{distrModJSS} provides an object-oriented (S4) 
+implementation of probability models for non-censored data
+and includes distribution fitting procedures for a given minimization criterion.
+This package is based on the \proglang{R}-shipped function \code{mle} of the \pkg{stats4} package
+producing S4 objects of class \code{"mle"} for which generic methods are 
+implemented, e.g. \code{confint} and \code{logLik}.
+In \pkg{fitdistrplus}, we choose to use the standard S3 class system for its understanding by most \proglang{R} users.
+When designing the \pkg{fitdistrplus} package, we do not forget to implement
+generic functions also available for S3 classes.
+Finally, various other packages provide functions to estimate the mode, the moments or the L-moments of a distribution, see the reference manuals of  \pkg{modeest}, \pkg{lmomco} and \pkg{Lmoments} packages.
+
+This manuscript reviews the various features of version 1.0-1 of \pkg{fitdistrplus}. The package
 is available from the Comprehensive \proglang{R} Archive Network at \url{http://cran.r-project.org/package=fitdistrplus}.
-The development version of the package is located at \proglang{R}-forge as one the packages of the 
+The development version of the package is located at \proglang{R}-forge as one package of the 
 project ``Risk Assessment with R''
-(\url{http://r-forge.r-project.org/projects/riskassessment/})
-The following command will load the package.
+(\url{http://r-forge.r-project.org/projects/riskassessment/}).
+The paper is organized as follows: Section \ref{fitnoncenscont} presents tools for fitting continuous
+distributions to classic non-censored data. Section \ref{advtopic} deals with other
+estimation methods and other types of data, before Section \ref{ccl} concludes.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section[Fitting distributions]{Fitting distributions to continuous non-censored data}\label{fitnoncenscont}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Choice of candidate distributions}
+\label{Choice}
+For illustrating the use of various functions of the \pkg{fitdistrplus} package 
+with continuous non-censored data, we   first use
+a data set named \code{groundbeef} which is included in our package.
+This data set contains pointwise values of
+serving sizes in grams, collected in a French survey, for ground beef patties consumed by 
+children under 5 years old. It was used in a quantitative risk assessment 
+published in the International Journal of Food Microbiology \cite{Delignette08}.
 %%% R code set default options for all R schunks
@@
 options(digits=4, SweaveHooks=list(fig=function() par(mar=c(5.1, 4.1, 1.1, 2.1))) )
+set.seed(1234)
 @
 %%% R code
-<<pkgload1, results=hide>>=
+<<datgroundbeef, echo=TRUE>>=
 library(fitdistrplus)
+data(groundbeef)
+str(groundbeef)
 @
 
-%\section*{Acknowledgments}
+Before fitting one or more distributions to a data set, it is generally
+necessary to choose good candidates among a predefined set of distributions.
+This choice may be guided by the knowledge of stochastic processes governing
+the modelled variable, or, in the absence of knowledge regarding the underlying process, 
+by the observation of its empirical distribution.
+To help the user in this choice, we developed functions to plot and
+characterise the empirical distribution.
 
-%\bibliography{fitdistrplus}
+First of all, it is common to start with plots of the empirical distribution 
+function and the histogram, which can be obtained with the \code{plotdist} function of the \pkg{fitdistrplus} package.
+This function provides two plots (see Figure 1): the left-hand plot is the histogram (on a density scale) 
+and the right-hand  plot the empirical cumulative distribution function (CDF).
+%%% R code
+<<figgroundbeef.echo, echo=TRUE, fig=FALSE, eval=FALSE>>=
+plotdist(groundbeef$serving)
+@
+
+\setkeys{Gin}{width=0.8\textwidth}
+
+\begin{figure}[htb!]
+  \centering
+  %%% R code
+<<figgroundbeefplot, echo=FALSE, fig=TRUE, width=10, height=5, eps=FALSE>>=
+plotdist(groundbeef$serving)
+@
+  \caption{Histogram and CDF plots of an empirical distribution for a continuous variable 
+  (serving size from the \code{groundbeef} data set) as provided by the \code{plotdist} function}
+  \label{plotdistcont}
+\end{figure}
+
+In addition to empirical plots, descriptive statistics may help to choose candidates to describe a distribution
+among a set of parametric distributions. Especially the skewness and kurtosis, linked to the third and fourth moments, 
+are useful for this purpose. A non-zero skewness reveals a lack of symmetry of the empirical distribution,
+while the kurtosis value quantifies the weight of tails in comparison to the normal distribution for which
+the kurtosis equals 3.
+The skewness and kurtosis and their corresponding unbiased estimator (e.g. \cite{casellaberger02}) from a sample 
+$(X_i)_i \stackrel{\text{i.i.d.}}{\sim} X$ with observations $(x_i)_i$ are given by
+  \begin{equation}
+    \label{skewness}
+    sk(X) = \frac{E[(X-E(X))^3]}{Var(X)^{\frac{3}{2}}}~,~
+    \widehat{sk}=\frac{\sqrt{n(n-1)}}{n-2}\times\frac{m_{3}}{m_{2}^{\frac{3}{2}}},
+  \end{equation}
+
+  \begin{equation}
+    \label{kurtosis}
+    kr(X) = \frac{E[(X-E(X))^4]}{Var(X)^{2}}~,~
+    \widehat{kr}=\frac{n-1}{(n-2)(n-3)}((n+1) \times \frac{m_{4}}{m_{2}^{2}}-3(n-1)) + 3,
+  \end{equation}
+where $m_{2}$, $m_{3}$, $m_{4}$ denote empirical moments defined by
+$m_{r}=\frac{1}{n}\sum_{i=1}^n(x_{i}-\overline{x})^{r}$, with
+$x_{i}$ the $n$ observations of variable $x$ and $\overline{x}$ their mean value.
+
+The  \code{descdist} function provides classical descriptive statistics
+(minimum, maximum, median, mean, standard deviation), skewness and kurtosis. By default, 
+unbiased estimations of the three last statistics are provided. Nevertheless,
+the argument \code{method}
+can be changed from \code{"unbiased"} (default) to \code{"sample"} to obtain  
+them without correction for bias. 
+A skewness-kurtosis plot such as the one proposed by \cite{Cullen99} is  provided by 
+the function \code{descdist} for the empirical distribution 
+(see Figure~\ref{Cullenplotcont} for the \code{groundbeef} data set). 
+On this plot, values for common distributions are displayed in order 
+to help the choice of distributions to fit to data. For some distributions (normal, uniform,
+logistic, exponential), there is only one possible value for the skewness and the kurtosis.
+ Thus, the distribution is represented by a single point on the plot. For other distributions, 
+areas of possible values are represented, consisting in lines (as for gamma and lognormal distributions), 
+or larger areas (as for beta distribution).
+
+Skewness and kurtosis are known not to be robust. In order to take into account the uncertainty 
+of the estimated values of kurtosis and skewness from data, a nonparametric bootstrap 
+procedure (\cite{efrontibshirani94}) can be performed by using the argument \code{boot}. %to an integer above 10.
+Values of skewness and kurtosis are computed on bootstrap samples (constructed by 
+random sampling with replacement from the original data set) and reported on the 
+skewness-kurtosis plot. 
+The properties of the random variable should be considered, notably its expected value, its range,
+as a complement to the use of the \code{plotdist} and \code{descdist}. 
+Below is a call to the \code{descdist} function to describe the distribution 
+of the serving size from the \code{groundbeef} data set and to draw the corresponding skewness-kurtosis
+plot (see Figure~\ref{Cullenplotcont}). Looking at the results on this example
+with a positive skewness and a kurtosis not far from 3,
+the fit of three common right-skewed distributions could be considered, Weibull, gamma and 
+lognormal distributions.
+%%% R code
+<<descgroundbeef.echo, echo=TRUE, fig=FALSE, eval=FALSE>>=
+descdist(groundbeef$serving, boot=1000)
+@
+
+\setkeys{Gin}{width=0.5\textwidth}
+
+\begin{figure}[htb]
+  \centering
+ %%% R code
+<<descgroundbeefplot, echo=FALSE, fig=TRUE, width=6, height=6, eps=FALSE>>=
+descdist(groundbeef$serving, boot=1000)
+@
+  \caption{Skewness-kurtosis plot for a continuous variable 
+  (serving size from the \code{groundbeef} data set) as provided by the \code{descdist} function}
+  \label{Cullenplotcont}
+\end{figure}
+
+\subsection[Maximum likelihood]{Fit of distributions by maximum likelihood estimation}
+\label{FIT}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Once selected, one or more parametric distributions $f(.\vert \theta)$ 
+(with parameter $\theta\in\mathbb{R}^d$) may be fitted to the data set, 
+one at a time, using the \code{fitdist} function. 
+Under the i.i.d. sample assumption, distribution parameters $\theta$ are by default estimated by maximizing the
+likelihood function defined as:
+  \begin{equation}
+    \label{likelihood}
+    L(\theta)=\prod_{i=1}^n f(x_{i}\vert \theta)
+  \end{equation}
+with $x_{i}$ the $n$ observations of variable $X$ and $f(.\vert \theta)$ the density function of the 
+parametric distribution.
+The other proposed estimation methods are described in Section~\ref{Alternatives}.
+
+The \code{fitdist} function returns an S3 object of class \code{"fitdist"} 
+for which \code{print}, \code{summary} and \code{plot} functions are provided.
+The fit of a distribution using \code{fitdist} assumes that the corresponding 
+\code{d}, \code{p}, \code{q} functions (standing respectively for the density, 
+the distribution and the quantile functions) are defined.
+Classical distributions are already defined in that way in the \pkg{stats} package,
+e.g. \code{dnorm}, \code{pnorm} and \code{qnorm} for the normal distribution
+(see \code{?Distributions}). Others
+may be found in various packages (see the CRAN task view: Probability Distributions). 
+%%%%%%%%%%% PEUT-ON METTRE UNE REF POUR LA CRAN TASK VIEW?
+Distributions not found in any package must be implemented by the user as
+\code{d}, \code{p}, \code{q} functions.
+In the call to \code{fitdist}, a distribution  has to be specified via
+the argument \code{dist} either by the character string corresponding to its common root name
+used in the names of \code{d}, \code{p}, \code{q} functions
+(e.g. \code{"norm"} for the normal distribution) 
+or by the density function itself from which
+the root name is extracted (e.g. \code{dnorm} for the normal distribution).
+Numerical results returned by the \code{fitdist} function are 
+(1) the parameter estimates, (2) the estimated standard errors (computed from the estimate
+of the Hessian matrix at the maximum likelihood solution), (3) the loglikelihood, 
+(4) Akaike and Bayesian information criteria (the so-called AIC and BIC), and (5) the correlation matrix
+between parameter estimates.
+Below is a call to the \code{fitdist} function to fit a Weibull distribution 
+to the serving size from the \code{groundbeef} data set.
+%%% R code
+<<fitgroundbeef.weibull, echo = TRUE, fig = FALSE>>=
+fw <- fitdist(groundbeef$serving, "weibull")
+summary(fw)
+@
+
+The plot of an object of class \code{"fitdist"} provides 
+four classical goodness-of-fit plots (\cite{Cullen99}) 
+presented on Figures \ref{groundbeef:comp} and \ref{endo:comp}: 
+\begin{itemize}
+\item a density plot representing the density function of the fitted distribution 
+along with the histogram of the empirical distribution, 
+\item a CDF plot of both the empirical distribution and the fitted distribution, 
+\item a Q-Q plot
+representing the empirical quantiles (y-axis) against the theoretical quantiles (x-axis)
+\item a P-P plot representing the empirical distribution function evaluated at each data point (y-axis)
+against the fitted distribution function (x-axis).
+\end{itemize}
+
+For CDF, Q-Q and P-P plots, the probability plotting position is defined by default using the Hazen's rule, with probability points of the empirical distribution defined as
+ \code{(1:n - 0.5)/n}, as recommended by Blom \cite{Blom}.
+This plotting position can be easily changed (see the reference 
+manual~\cite{fitdistrplus} for details).
+% using the arguments 
+%\code{use.ppoints} and \code{a.ppoints}. When \code{use.ppoints = TRUE}, 
+%the argument \code{a.ppoints} is passed to the \code{ppoints} function from the \pkg{stats} package to %define the 
+%probability points of the empirical distribution as \code{(1:n - a.ppoints)/(n - 2a.ppoints + 1)}. 
+%When \code{use.ppoints = FALSE}, the probability points are simply defined as \code{1:n / n}.
+
+Unlike the generic \code{plot} function, the \code{denscomp}, \code{cdfcomp},
+\code{qqcomp} and \code{ppcomp} functions enable to draw separately each
+of these four plots, in order to compare the empirical distribution 
+and multiple parametric distributions fitted
+on a same data set. These functions must be called with a first argument 
+corresponding to a list of objects of class \code{fitdist}, and optionally
+further arguments to customize the plot (see the reference 
+manual~\cite{fitdistrplus} for lists of arguments that may be specific to each plot). 
+In the following example, we compare the fit of
+a Weibull, a lognormal and a gamma distributions to the \code{groundbeef} data set (Figure~\ref{groundbeef:comp}).
+%%% R code
+<<fitgroundbeef.echo, echo=TRUE, fig=FALSE>>=
+fg <- fitdist(groundbeef$serving,"gamma")
+fln <- fitdist(groundbeef$serving,"lnorm")
+par(mfrow=c(2, 2))
+denscomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+qqcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+cdfcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+ppcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+@
+
+\setkeys{Gin}{width=0.8\textwidth}
+
+\begin{figure}[htb!]
+  \centering
+<<fitgroundbeef, echo=FALSE, fig=TRUE, width=7, height=7, eps=FALSE>>=
+par(mfrow=c(2, 2))
+denscomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+qqcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+cdfcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+ppcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+@ 
+  \caption{Four Goodness-of-fit plots for various distributions fitted to continuous data
+  (Weibull, gamma and lognormal
+  distributions fitted to serving sizes from the \code{groundbeef} data set)  
+  as provided by functions \code{denscomp}, \code{qqcomp}, \code{cdfcomp} and \code{ppcomp}}
+\label{groundbeef:comp}  
+\end{figure}
+
+The density plot and the CDF plot may be considered as the basic classical goodness-of-fits plots.
+% ca te va mieux reformule comme ca?
+The two other plots are complementary and can be very informative in some cases.
+The Q-Q plot emphasizes the lack-of-fit at the distribution tails while the P-P plot emphasizes
+the lack-of-fit at the distribution center.
+As an example in Figure~\ref{groundbeef:comp}, none of the three fitted distributions correctly describes
+the center of the distribution, but the Weibull and gamma distributions 
+could be prefered for their better description of the right tail of the empirical distribution, especially
+if this tail is important in the use of the fitted distribution, as it is in the context of
+food risk assessment.
+
+The data set named  \code{endosulfan} is now used to illustrate other features of 
+the \pkg{fitdistrplus} package. This data set contains
+acute toxicity values for the organochlorine pesticide endosulfan
+(geometric mean of LC50 ou EC50 values in $\mu g.L^{-1}$),
+tested on Australian and non-Australian laboratory-species 
+(arthropods, fish or nonarthropod invertebrates, see \cite{Hose04}). 
+<<datendosulfan, echo=TRUE>>=
+data(endosulfan)
+str(endosulfan)
+@
+
+In ecotoxicology, a
+lognormal or a loglogistic distribution is often fitted to such a data set in order
+to characterize the species sensitivity distribution (SSD) for a pollutant. A low percentile
+of the fitted distribution, generally the 5$\%$ percentile, is calculated and named 
+the hazardous concentration 5$\%$ (HC5). It is interpreted as the value of the
+pollutant concentration protecting 95$\%$ of the species (\cite{Posthuma2010}).
+But the fit of a lognormal or a loglogistic distribution to the whole \code{endosulfan} data set is rather bad,
+especially due to a minority of very high values. 
+The two-parameter Pareto distribution and the three-parameter Burr distribution (which is an extension
+of both the loglogistic and the Pareto distributions) have been fitted.
+Pareto and Burr distributions are provided in the package \pkg{actuar}. 
+Until here, we did not have to  define starting values (in the optimization process)
+as reasonable starting values are implicity defined within the \code{fitdist} function 
+for most of the distributions defined in \proglang{R} (see \code{?fitdist} for details).
+For other distributions like the Pareto and the Burr distribution, initial 
+values for the distribution parameters have to be supplied
+in the argument \code{start} as a named list  with initial values for each parameter 
+(as they appear in the \code{d}, \code{p}, \code{q} functions). 
+Having defined reasonable starting values\footnote{%----
+The \code{plotdist} function can plot any 
+parametric distribution with specified parameter values in argument \code{para}.
+It can thus help to find correct initial values for the distribution parameters in non trivial cases, by iterative calls if necessary (see the reference
+manual~\cite{fitdistrplus} for examples).
+},%----
+various distributions can be fitted and graphically
+compared. On this example, the function \code{cdfcomp} can be used to report CDF values in a logscale so as to emphasize discrepancies on the tail of
+interest while defining an HC5 value.
+
+%%% R code
+<<fitendo.echo, echo=TRUE, fig=FALSE>>=
+ATV <-endosulfan$ATV
+fendo.ln <- fitdist(ATV, "lnorm")
+library(actuar)
+fendo.ll <- fitdist(ATV, "llogis", start=list(shape=1, scale=500))
+fendo.P <- fitdist(ATV, "pareto", start=list(shape=1, scale=500))
+fendo.B <- fitdist(ATV, "burr", start=list(shape1=0.3, shape2=1, rate=1))
+cdfcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), xlogscale = TRUE,
+        ylogscale = TRUE,legendtext = c("lognormal","loglogistic","Pareto","Burr"))
+@
+% qqcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), xlogscale=TRUE, ylogscale=TRUE,
+%       legendtext = c("lognormal","loglogistic","Pareto","Burr"))
+
+\setkeys{Gin}{width=0.5\textwidth}
+
+\begin{figure}[htb!]
+  \centering
+<<fitendo, echo=FALSE, fig=TRUE, width=6, height=4, eps=FALSE>>=
+cdfcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), xlogscale = TRUE,
+        ylogscale = TRUE,legendtext = c("lognormal","loglogistic","Pareto","Burr"))
+@
+%qqcomp(list(fendo.ln,fendo.ll,fendo.P,fendo.B),xlogscale=TRUE,ylogscale=TRUE,
+%       legendtext = c("lognormal","loglogistic","Pareto","Burr"))
+
+  \caption{CDF plot to compare the fit of four distributions to
+  acute toxicity values of various organisms for the organochlorine pesticide endosulfan
+  (\code{endosulfan} data set) as provided by the \code{cdfcomp} function, with CDF values
+  in a logscale to emphasize discrepancies on the left tail}
+\label{endo:comp}  
+\end{figure}
+
+None of the fitted distribution correctly
+describes the right tail observed in the data set, but as shown in Figure~\ref{endo:comp},
+the left-tail seems to be better described by the Burr distribution. Its use could then be considered
+to estimate the HC5 value as the $5\%$ quantile of the distribution. This can be 
+easily done using the \code{quantile} generic function defined for an object 
+of class \code{"fitdist"}. Below is this calculation together with the calculation of the 
+empirical quantile for comparison.
+%%% R code
+<<quantilefitdist, echo=TRUE, fig=FALSE>>=
+quantile(fendo.B, probs = 0.05)
+quantile(ATV, probs = 0.05)
+@
+
+The computation of different goodness-of-fit statistics is proposed in the \pkg{fitdistrplus} package
+in order to further compare fitted distributions.
+The purpose of goodness-of-fit statistics aims to measure the distance between the
+fitted parametric distribution and the empirical distribution: e.g. 
+the distance between the fitted cumulative distribution function $F$ 
+and the empirical distribution function $F_{n}$.
+When fitting continuous distributions, three  goodness-of-fit statistics are classicaly considered: 
+Cramer-von Mises, Kolmogorov-Smirnov and Anderson-Darling statistics \cite{Stephens86}.
+Naming $x_{i}$ the $n$ observations of a continuous variable $X$ arranged in an ascending order,
+Table \ref{tabKSCvMAD} gives the definition and the empirical estimate of the three considered 
+goodness-of-fit statistics.
+
+\begin{table}[htb!]
+  \begin{center}
+    \caption{Goodness-of-fit statistics as defined by Stephens \cite{Stephens86}.}
+    \begin{tabular}{lll}
+    \hline
+      Statistic  & General formula & Computational formula\\
+    \hline
+      Kolmogorov-Smirnov   & $\sup|F_{n}(x) - F(x)|$ &
+      $\max(D^{+},D^{-})$ with\\
+      (KS)&& $D^{+}=\max\limits_{i=1,\dots,n}\left(\frac{i}{n} - F(x_{i})\right)
+      ;D^{-}=\max\limits_{i=1,\dots,n}\left(F(x_{i})-\frac{i-1}{n}\right)$ \\
+    \hline
+      Cramer-von Mises  & n $\int_{-\infty}^{\infty}(F_{n}(x) - F(x))^2 dx$ &
+        $\frac{1}{12n} + \sum\limits_{i=1}^n \left(F(x_{i})-\frac{2i-1}{2n} \right)^{2}$\\
+      (CvM)&&\\
+% cvm <- 1/(12*n) + sum( ( theop - (2 * seq(1:n) - 1)/(2 * n) )^2 )
+    \hline
+      Anderson-Darling  & n $\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2}{F(x) (1 - F(x))} dx$
+      & $-n -\frac{1}{n}\sum\limits_{i=1}^n \left((2i-1)(\log(F(x_{i}))+\log(1-F(x_{n+1-i})) )\right)$ \\
+      (AD) &    & \\
+% ad <- - n - mean( (2 * seq(1:n) - 1) * (log(theop) + log(1 - rev(theop))) )
+    \hline
+    \end{tabular}
+    \label{tabKSCvMAD}
+  \end{center}
+\end{table}
+
+They can be computed using the function \code{gofstat}
+as defined by Stephens \cite{Stephens86}.
+
+<<fendo.gof.print, echo=TRUE, fig=FALSE>>=
+gofstat(list(fendo.ln, fendo.ll, fendo.P, fendo.B))
+@
+As giving  more weight to distribution tails, 
+Anderson-Darling statistics is of special interest when it matters to equally 
+emphasize the tails as well as the main body of a distribution.
+This is often the case in risk assessment, e.g. \cite{Cullen99,Vose10}. For this reason, 
+this statistics is often used to select the best distribution among those fitted.
+Nevertheless, this statistics should be used cautiously when comparing fits of various distributions. 
+Keeping in mind that the weighting of each CDF quadratic difference 
+depends on the parametric distribution in its definition (see Table \ref{tabKSCvMAD}), 
+Anderson-Darling statistics computed for several distributions fitted on a same data set
+are theoretically difficult to compare. Moreover, such a statistic, as Cramer-von Mises and
+Kolmogorov-Smirnov ones, does not take into account the complexity of 
+the model (i.e. parameter number). It is not
+a problem when compared distributions are characterized by the same number of parameters, but
+it could systematically promote the selection of the more complex distributions in the other case. 
+Looking at classical penalized criteria based on the loglikehood (AIC, BIC) seems thus also interesting, especially
+to discourage overfitting.
+
+In the previous example, all the goodness-of-fit statistics based on the CDF distance are in favor of 
+the Burr distribution, the only one characterized by three parameters, while AIC and BIC values
+respectively give the preference to the Burr distribution or the Pareto distribution.
+The choice between these two distributions seems thus less obvious and could be discussed. 
+Even if specifically recommended for discrete distributions, the Chi-squared statistic may also be used for
+continuous distributions (see Section~\ref{otherdata} and the reference 
+manual~\cite{fitdistrplus} for examples).
+
+\subsection[Uncertainty]{Uncertainty in parameter estimates}
+\label{Uncertainty}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The uncertainty in the parameters of the fitted distribution may be estimated by parametric or nonparametric
+bootstraps using the \code{boodist} function for non-censored data, see \cite{efrontibshirani94}.
+This function returns the bootstrapped values of parameters in a 
+S3 class object which may be plotted to visualize the bootstrap region. 
+The medians and the 95 percent confidence intervals of parameters (2.5 and 97.5 percentiles) are printed 
+in the summary.
+When inferior to the whole number of iterations (due to lack of convergence of the optimization
+algorithm for some bootstrapped data sets), the number of iterations for which the estimation converges 
+is also printed in the summary.
+
+The plot of an object of class \code{"bootdist"}  
+consists in a scatterplot or a matrix of scatterplots
+of the bootstrapped values of parameters providing
+a representation of the joint uncertainty distribution of the fitted parameters (see Figure~\ref{fig:bootstrap}).
+Below is an example of the use of the \code{bootdist} function with the previous fit of the Burr distribution to the
+\code{endosulfan} data set. 
+%%% R code
+<<fitBurr.boot.echo, echo=TRUE, results=hide, fig=FALSE>>=
+bendo.B <- bootdist(fendo.B, niter=1001)
+@
+<<fitBurr.boot.results, echo=TRUE, fig=FALSE>>=
+summary(bendo.B)
+plot(bendo.B)
+@
+
+\setkeys{Gin}{width=0.5\textwidth} 
+\begin{figure}[htb!]
+  \centering
+<<fitBurrbootplot, echo=FALSE, fig=TRUE, width=6, height=6, eps=FALSE>>=
+plot(bendo.B)
+@
+  \caption{Bootstrappped values of parameters for a fit of the Burr distribution
+characterized by three parameters (example on the \code{endosulfan} data set) as provided 
+by the plot of an object of class \code{"bootdist"}}
+  \label{fig:bootstrap}
+\end{figure}
+
+Bootstrap samples of parameter estimates are useful especially to calculate confidence intervals on each parameter
+of the fitted distribution from the marginal distribution of the bootstraped values. It is also interesting to look at the joint distribution of the bootstraped
+values in a scatterplot (or a matrix of scatterplots if the number of parameters exceeds two) in order
+to understand the potential structural correlation between parameters (see Figure~\ref{fig:bootstrap}).
+
+The use of the whole bootstrap sample is also of interest in the risk assessment field. Its use enables
+the characterization of uncertainty in distribution parameters. 
+It can be directly used within a second-order Monte Carlo
+simulation framework, especially within the package \pkg{mc2d} (\cite{mc2d}). One could refer to Pouillot and Delignette-Muller
+(\cite{Pouillot10}) for an introduction to the use of \pkg{mc2d} 
+and \pkg{fitdistrplus} packages in the context of quantitative 
+risk assessment.
+
+The bootstrap method can also be used to calculate confidence intervals on quantiles 
+of the fitted distribution.
+For this purpose, a generic \code{quantile} function is provided for class \code{bootdist}. 
+By default, $95\%$ percentiles bootstrap confidence intervals of quantiles are provided.
+Going back to the previous example from ecotoxicolgy, this function can be used to 
+estimate the uncertainty associated to the HC5 estimation, for example from the previously fitted
+Burr distribution to the \code{endosulfan} data set.
+ %%% R code
+<<fitATV.lnorm.quantile, echo=TRUE>>=
+quantile(bendo.B, probs = 0.05)
+@
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Advanced topics}\label{advtopic}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection[Alternative estimation]{Alternative methods for parameter estimation}
+\label{Alternatives}
+
+This subsection focuses on alternative estimation methods.
+One of the alternative for continuous distributions
+is the maximum goodness-of-fit estimation method also called
+minimum distance estimation method \cite{Stephens86,actuarJSS}. 
+In this package this method is proposed with
+eight different distances: the three classical distances defined
+in Table~\ref{tabKSCvMAD},
+or one of the variants of the Anderson-Darling distance proposed by \cite{Luceno06} and defined in 
+Table~\ref{modifiedAD}.
+The right-tail AD gives more weight to the right-tail, the left-tail AD
+gives more weight only to the left tail. Either of the tails, or both of them,
+can receive even larger
+weights by using second order Anderson-Darling Statistics.
+
+\begin{table}[htb!]
+  \begin{center}
+    \caption{Modified Anderson-Darling statistics as defined by Luceno \cite{Luceno06}.}
+    \begin{tabular}{lll}
+    \hline
+      Statistic  & General formula & Computational formula\\
+    \hline
+      Right-tail AD  & $\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{1 - F(x)}  dx$
+      & $\frac{n}{2} -2\sum_{i}F(x_{i}) -\frac{1}{n}\sum_{i}((2i-1)ln(1-F(x_{n+1-i})))$ \\
+      (ADR) &  &\\
+    \hline
+      Left-tail AD  & $\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{(F(x))}  dx$
+      & $-\frac{3n}{2} +2\sum_{i}F(x_{i}) -\frac{1}{n}\sum_{i}((2i-1)ln(F(x_{i})))$ \\
+      (ADL) &  &\\
+    \hline
+      Right-tail AD  & $ad2r=\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{(1 - F(x))^{2}}  dx$
+      & $ad2r=2\sum_{i}ln(1-F(x_{i})) +\frac{1}{n}\sum_{i} \frac{2i-1}{1-F(x_{n+1-i})}$ \\
+      2nd order (AD2R) &  &\\
+    \hline
+      Left-tail AD  & $ad2l=\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{(F(x))^{2}}  dx$
+      & $ad2l=2\sum_{i}ln(F(x_{i})) +\frac{1}{n}\sum_{i}\frac{2i-1}{F(x_{i})}$ \\
+      2nd order (AD2L) &  &\\
+    \hline
+    AD 2nd order & $ad2r+ad2l$
+      & $ad2r+ad2l$ \\
+    (AD2) &  &\\
+    \hline
+    \end{tabular}
+    \label{modifiedAD}
+  \end{center}
+\end{table}
+
+To fit a distribution by maximum goodness-of-fit estimation, one
+needs to fix the argument \code{method}
+to \code{"mge"} in the call to \code{fitdist} and to specify the argument \code{gof}
+coding for the chosen goodness-of-fit distance. 
+This function is intended to be used only with continuous non-censored data.
+
+Maximum goodness-of-fit estimation may be useful to give
+more weight to data at one tail of the distribution.
+Going back to the previous example from ecotoxicology, 
+instead of trying to find a less classical distribution to correctly
+fit the empirical distribution especially on its left tail, one could
+consider the fit of the classical lognormal distribution, but minimizing a
+goodness-of-fit distance giving more weight to the left tail of the empirical distribution,
+so as to correctly estimate the 5$\%$ percentile. 
+In the following example of \code{endosulfan} data set, we use left tail 
+Anderson-Darling distances of first or second order
+(see Figure~\ref{plotfitMGE}).
+
+%%% R code
+<<mge.gofcomp.echo, echo=TRUE, fig=FALSE>>=
+fendo.ln.ADL <- fitdist(ATV,"lnorm",method="mge",gof="ADL")
+fendo.ln.AD2L <- fitdist(ATV,"lnorm",method="mge",gof="AD2L")
+cdfcomp(list(fendo.ln, fendo.ln.ADL, fendo.ln.AD2L),
+xlogscale = TRUE, ylogscale = TRUE,
+main = "Fitting a lognormal distribution",xlegend = "bottomright",
+legendtext = c("MLE","Left-tail AD", "Left-tail AD 2nd order"))
+@
+
+\setkeys{Gin}{width=0.5\textwidth}
+
+\begin{figure}[htb!]
+  \centering
+  %%% R code
+<<mgegofcompplot, echo=FALSE, fig=TRUE, width=6, height=4, eps=FALSE>>=
+cdfcomp(list(fendo.ln, fendo.ln.ADL, fendo.ln.AD2L),
+xlogscale = TRUE, ylogscale = TRUE,
+main = "Fitting a lognormal distribution",
+legendtext = c("MLE","Left-tail AD", "Left-tail AD 2nd order"),
+xlegend = "bottomright")
+@
+  \caption{Comparison of a lognormal distribution fitted by MLE
+  and by MGE using  two different goodness-of-fit distances : left-tail Anderson-Darling and left-tail Anderson Darling of seond order (example on the \code{endosulfan} data set)
+  as provided by the \code{cdfcomp} function, with CDF values
+  in a logscale to emphasize discrepancies on the left tail}
+  \label{plotfitMGE}
+\end{figure}
+
+Comparing the $5\%$ percentiles (HC5) calculated using these three fits to the one calculated
+from the MLE fit of the Burr distribution, we can observe, on this example,
+that fitting the lognormal distribution by maximizing left tail 
+Anderson-Darling distances of first or second order
+enables to approach the value obtained by fitting the Burr distribution by MLE. 
+%%% R code
+<<quantilefitdist, echo=TRUE, fig=FALSE>>=
+( HC5.estimates <- cbind(empirical = as.numeric(quantile(ATV, probs = 0.05)),
+      Burr = as.numeric(quantile(fendo.B,probs = 0.05)$quantiles),
+      lognormal_MLE = as.numeric(quantile(fendo.ln,probs = 0.05)$quantiles),
+      lognormal_AD2 = as.numeric(quantile(fendo.ln.ADL,probs = 0.05)$quantiles),
+      lognormal_AD2L = as.numeric(quantile(fendo.ln.AD2L,probs = 0.05)$quantiles)) )
+@
+
+%\subsubsection{Moment matching estimation}
+%\label{MME}
+%%%%%%%%%%
+Another method commonly used to fit parametric distribution is the moment matching estimation (MME). 
+It consists in finding the value of the parameter $\theta$  that equalizes the first theoretical 
+raw moments of the parametric distribution  to the empirical moments (Equation~\ref{moments})
+  \begin{equation}
+    \label{moments}
+    E(X^{k}|\theta)=\frac{1}{n}\sum_{i=1}^{n}x_{i}^{k} ,
+  \end{equation}
+for $k=1,\ldots,d$, with $d$ the number of parameters to estimate and
+ $x_{i}$ the $n$ observations of variable $X$.
+For moments of order greater or equal than 2, it may  also be relevant to match centered moments as
+given by Equation (\ref{centmoments}).
+  \begin{equation}
+    \label{centmoments}
+    E\left((X-E(X))^{k}|\theta\right)=\frac{1}{n}\sum_{i=1}^{n}(x_{i}-\bar x_n)^{k} 
+  \end{equation}
+This method can be performed by setting the argument \code{method} to \code{"mme"} in the call to fitdist.
+The estimate is computed by a closed-form formula for the following distributions:
+normal, lognormal, exponential, Poisson, gamma, logistic, negative binomial,
+geometric, beta and uniform distributions.
+In this case, for distributions characterized by one parameter (geometric, Poisson and exponential),
+this parameter is simply estimated by matching theoretical and observed means, and for distributions 
+characterized by two parameters, these parameters are estimated by matching theoretical and observed 
+means and variances (see e.g. \cite{Vose10}).
+For other distributions, the equation of moments is solved numerically using 
+the \code{optim} function  by minimizing the
+sum of squared differences between observed and theoretical moments (see the \pkg{fitdistrplus} reference
+manual~\cite{fitdistrplus}
+for technical details).
+
+A classical data set from the Danish insurance industry 
+published in \cite{mcneil97} is used to illustrate this method.
+In \pkg{fitdistrplus}, the data set is stored in \code{danishuni} for the univariate version and contains
+the losse amounts collected at Copenhagen Reinsurance between 1980 and 1990.
+In actuarial science, it is standard to consider positive heavy-tailed distribution and have 
+a special focus on the right-tail of the distribution.
+In this numerical experiment,  the lognormal distribution is first considered and then the Pareto type II distribution (\cite{Klugmanetal09}).
+
+The lognormal distribution on \code{danishuni} data set is 
+fitted by matching moments implemented as a closed-form formula.
+On the left-hand graph of Figure \ref{fig:danish:mme}, we compare the fitted distribution function between
+ MME and MLE method.
+The moment matching estimation provides a more cautious estimation of the insurance risk 
+as the MME-fitted distribution function (resp. MLE-fitted) underestimates (overestimates) 
+the empirical distribution function for large values of claim amounts.
+
+%%% R code
+<<danish.mme, echo=TRUE, fig=FALSE>>=
+data(danishuni)
+str(danishuni)
+fdanish.ln.MLE <- fitdist(danishuni$Loss, "lnorm")
+fdanish.ln.MME <- fitdist(danishuni$Loss, "lnorm", method="mme", order=1:2)
+cdfcomp(list(fdanish.ln.MLE, fdanish.ln.MME), 
+        legend=c("lognormal MLE", "lognormal MME"), main="Fitting a lognormal distribution",
+        xlogscale=TRUE, datapch=".")
+@
+
+In a second time, a Pareto type-II distribution, which gives more weight to 
+the right-tail of the distribution, is fitted.
+As the lognormal distribution, the Pareto type-II has two parameters, which
+allows a fair comparison.
+%The Burr distribution (with its three parameters) would lead to a better fit.
+
+We use the implementation of the \pkg{actuar} package providing raw and centered moments
+for that distribution (in addition to \code{d}, \code{p}, \code{q} and \code{r} functions,
+see \cite{actuar12}).
+Fitting a heavy-tailed distribution for which the first and the second moments do not exist for
+certain values of the shape parameter requires some cautiousness.
+This is carried out by providing for the optimization process 
+a lower and an upper bound for each parameter.
+Our call below immediately calls the L-BFGS-B optimization method in \code{optim}, since this quasi-Newton
+allows box constraints\footnote{That is what the B stands for.}.
+Note that a function (called \code{memp} in our example) 
+for computing the empirical raw moment to \code{fitdist} has to be passed, 
+since the user may choose either Equation (\ref{moments}) or (\ref{centmoments}).
+
+<<danish.mme.pareto, echo=TRUE, fig=FALSE>>=
+library(actuar)
+fdanish.P.MLE <- fitdist(danishuni$Loss, "pareto",start=c(shape=10, scale=10), 
+      lower=2+1e-6, upper=Inf)
+memp <- function(x, order) ifelse(order == 1, mean(x), sum(x^order)/length(x))
+fdanish.P.MME <- fitdist(danishuni$Loss, "pareto", method="mme", order=1:2, 
+      memp="memp", start=c(shape=10, scale=10), lower=c(2+1e-6,2+1e-6), upper=c(Inf,Inf))
+cdfcomp(list(fdanish.P.MLE, fdanish.P.MME), 
+        legend=c("Pareto MLE", "Pareto MME"), main="Fitting a Pareto distribution",
+        xlogscale=TRUE, datapch=".")
+gofstat(list(fdanish.ln.MLE, fdanish.P.MLE, fdanish.ln.MME, fdanish.P.MME))
+@
+
+\setkeys{Gin}{width=0.8\textwidth} %default
+
+\begin{figure}[htb!]
+  \centering
+  %%% R code
+<<danishmmeplot, echo=FALSE, fig=TRUE, width=7, height=3.5, eps=FALSE>>=
+par(mfrow=c(1, 2))
+cdfcomp(list(fdanish.ln.MLE, fdanish.ln.MME), 
+        legend=c("lognormal MLE", "lognormal MME"), main="Fitting a lognormal distribution",
+        xlogscale=TRUE, datapch=".")
+cdfcomp(list(fdanish.P.MLE, fdanish.P.MME), 
+        legend=c("Pareto MLE", "Pareto MME"), main="Fitting a Pareto distribution",
+        xlogscale=TRUE, datapch=".")
+@
+\caption{Comparison between MME and MLE when fitting a lognormal or a Pareto distribution 
+to loss data from the \code{danishuni} data set}
+\label{fig:danish:mme}
+\end{figure}
+
+% je trouve nos legendes de graph trop longues?
+%
+
+As shown on Figure \ref{fig:danish:mme}, matching moments and maximum likelihood fits
+are far less distant (when looking at the right-tail) for the Pareto distribution 
+than for the lognormal distribution on this data set. 
+Furthermore for these two distributions, the MME method better fits the
+right-tail of the distribution, which seems logical since empirical moments 
+are influenced by large observed values.
+In the previous traces, we give the values of goodness-of-fit statistics.
+Whatever the statistic considered, the MLE-fitted lognormal always provides the best fit 
+to the observed data.
+
+Maximum likelihood and moment matching 
+%(the default on many softwares)
+% si on ne trouve pas de references on l'enleve?
+estimations are 
+certainly the most commonly used method for fitting distributions. 
+Keeping in mind that these two methods may produce very different results,
+the user should be aware of its great sensitivity to outliers
+when choosing the matching moment estimation (MME). This may be seen as an advantage in our example if the
+objective is to better describe the right tail of the distribution, but it may be seen as a
+drawback if the objective is different.
+% As-tu des references indiquant que le MLE est meilleurs dans certains cas, cf. commentaires de Regis ?
+
+%\subsubsection{Quantile matching estimation}
+%\label{QME}
+%%%%%%%%%%
+Fitting of a parametric distribution may also be done by matching theoretical quantiles
+of the parametric distributions (for specified probabilities) against the empirical quantiles.
+Equation (\ref{quantiles}) below is thus very similar to Equations (\ref{moments}) and (\ref{centmoments})
+  \begin{equation}
+    \label{quantiles}
+    F^{-1}(p_{k}|\theta)=Q_{n,p_{k}} 
+  \end{equation}
+for $k=1,\ldots,d$, with $d$ the number of parameters to estimate (dimension of $\theta$ if there is
+no fixed parameters) and
+ $Q_{n,p_{k}}$ the empirical quantiles calculated from data for specified probabilities $p_{k}$.
+
+Quantile matching estimation is performed by setting the argument \code{method}
+to \code{"qme"} in the call to \code{fitdist} and adding an argument \code{probs}
+defining the probabilities for which the quantile matching is performed.
+The length of this vector must be equal
+to the number of parameters to estimate (as the vector of moment orders for MME).
+Empirical quantiles are computed using the
+\code{quantile} function of the \pkg{stats} package using  \code{type=7} 
+by default (see \code{?quantile} and \cite{hyndmanfan96}).
+But the type of quantile can be easily changed by using the \code{qty} 
+argument in the call to the \code{qme} function.   
+The quantile matching is carried out numerically, 
+by minimizing the sum of squared differences between observed and theoretical quantiles.
+%%% R code
+<<danish.qme.echo, echo=TRUE, fig=FALSE>>=
+fdanish.ln.QME1 <- fitdist(danishuni$Loss, "lnorm", method="qme", probs=c(1/3, 2/3))
+fdanish.ln.QME2 <- fitdist(danishuni$Loss, "lnorm", method="qme", probs=c(8/10, 9/10))
+cdfcomp(list(fdanish.ln.MLE, fdanish.ln.QME1, fdanish.ln.QME2), 
+        legend=c("MLE", "QME(1/3, 2/3)", "QME(8/10, 9/10)"), 
+        main="Fitting a lognormal distribution", xlogscale=TRUE, datapch=".")
+@
+Above is an example of fitting of a lognormal distribution to \code{danishuni} data set 
+by matching probabilities $(p_1= 1/3, p_2=2/3)$ and $(p_1= 8/10, p_2=9/10)$.
+As expected, the second QME fit gives more weight to the right-tail of the distribution, despite
+we do not choose the Pareto type-II distribution.
+Compared to the maximum likelihood estimation, the second QME fit is also more conservative, whereas
+the first QME fit is less conservative.
+The quantile matching estimation is of particular interest when we need a focus 
+around particular quantiles, e.g. $p=99.5\%$ in the Solvency II insurance context or
+$p=5\%$ for the HC5 estimation in the ecotoxicology context.
+
+\setkeys{Gin}{width=0.5\textwidth} %default
+
+\begin{figure}[htb!]
+  \centering
+  %%% R code
+<<danishqmeplot, echo=FALSE, fig=TRUE, width=4, height=4, eps=FALSE>>=
+cdfcomp(list(fdanish.ln.MLE, fdanish.ln.QME1, fdanish.ln.QME2), 
+        legend=c("MLE", "QME(1/3, 2/3)", "QME(8/10, 9/10)"), main="Fitting a lognormal distribution",
+        xlogscale=TRUE, datapch=".")
+@
+\caption{Comparison between QME and MLE when fitting a lognormal distribution to loss data from the \code{danishuni} data set}
+\label{fig:danish:qme}
+\end{figure}
+
+\subsection[Customizing optimization]{Customization of the optimization algorithm}
+\label{Customization}
+%%%%%%%%%%
+Each time a numerical minimization (or maximization) is carried out in the \code{fitdistrplus} package, 
+the \code{optim} function of the \pkg{stats} package  is used by default with the 
+\code{"Nelder-Mead"} method for distributions characterized by more than one parameter and
+the \code{"BFGS"} method for distributions characterized by only one parameter.
+Sometimes the default algorithm fails to converge. It is then
+interesting to change some options of the \code{optim} function or to use another optimization
+function than \code{optim} to maximize the likelihood or to minimize a squared difference.
+The argument \code{optim.method} can be used in the call to \code{fitdist} or
+\code{fitdistcens}. It will internally be passed to \code{mledist} and to \code{optim}
+(see \code{?optim} for details about the different algorithms available).
+
+Below are examples of fits of a gamma distribution $\mathcal{G}(\alpha, \lambda)$ to 
+the \code{groundbeef} data set with various algorithms. Note that the conjugate gradient algorithm
+(\code{"CG"}) needs far more 
+iterations to converge (around 2500 iterations)
+compared to other algorithms (converging in less than 100 iterations).
+%%% R code
+<<optimmethod.gamma, echo=TRUE>>=
+data(groundbeef)
+fNM <- fitdist(groundbeef$serving, "gamma", optim.method="Nelder-Mead")
+fBFGS <- fitdist(groundbeef$serving, "gamma", optim.method="BFGS") 
+fSANN <- fitdist(groundbeef$serving, "gamma", optim.method="SANN")
+fCG <- try(fitdist(groundbeef$serving, "gamma", optim.method="CG", control=list(maxit=10000)))
+if(class(fCG) == "try-error")
+  fCG <- list(estimate=NA)
+@
+
+It is also possible to use another function than \code{optim} to maximize the likelihood by
+specifying by the argument \code{custom.optim} in the call to \code{fitdist}. 
+It may be necessary to customize this optimization function to meet
+the following requirements. (1) \code{custom.optim} function must have the following arguments:
+\code{fn} for the function to be optimized and \code{par} for the initialized parameters. 
+(2) \code{custom.optim} should carry out a MINIMIZATION and must return
+ the following components: \code{par} for the estimate, \code{convergence} for the convergence
+code, \code{value=fn(par)} and \code{hessian}.
+Below is an example of code written to wrap the \code{genoud} function from the \pkg{rgenoud} package in
+order to respect our optimization ``template''.
+The \pkg{rgenoud} package implements the genetic (stochastic) algorithm.
+%%% R code
+<<optimmethod.customgenoud, echo=TRUE>>=
+mygenoud <- function(fn, par, ...) 
+{
+   require(rgenoud)
+   res <- genoud(fn, starting.values=par, ...)        
+   standardres <- c(res, convergence=0)
+   return(standardres)
+}
+@
+
+The customized optimization function can then be passed as the argument \code{custom.optim} 
+in the call to \code{fitdist}
+or \code{fitdistcens}. The following code can for example be used to fit a gamma distribution to the
+\code{groundbeef} data set. Note that in this example various arguments are also passed from \code{fitdist} 
+to \code{genoud} :
+\code{nvars}, \code{Domains}, \code{boundary.enforcement}, 
+\code{print.level} and \code{hessian}.
+The code below compares all the parameter estimates ($\hat\alpha$, $\hat\lambda$) by the different algorithms:
+shape $\alpha$ and rate $\lambda$ parameters are relatively similar on this 
+example, roughly 4.00 and 0.05, respectively.
+%%% R code
+<<optimmethod.customgenoud.fitdist, echo=TRUE, eval=TRUE>>=
+fgenoud <- mledist(groundbeef$serving, "gamma", custom.optim= mygenoud, nvars=2, 
+    max.generations=10, Domains=cbind(c(0,0), c(10,10)), boundary.enforcement=1, 
+    hessian=TRUE, print.level=0, P9=10)
+cbind(NM = fNM$estimate,
+BFGS = fBFGS$estimate,
+SANN = fSANN$estimate,
+CG = fCG$estimate,
+fgenoud = fgenoud$estimate)
+@
+
+%%%%%%%%%%%%% J'EN SUIS LA !!!!!!!!!!!!!!
+
+\subsection[Other types of data]{Fitting distributions to other types of data}
+\label{otherdata}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Analytical methods often lead to semi-quantitative results which are referred to as
+censored data. Observations only known to be under a limit of detection are 
+left-censored data. Observations only known to be above a limit of quantification
+are  right-censored data. Results known to lie between two bounds are 
+interval-censored data. These two bounds may correspond to a limit of detection and 
+a limit of quantification, or more generally to uncertainty bounds around the observation.
+Right-censored data are also commonly encountered with survival data \cite{kleinmoeschberger03}.
+A data set may thus contain right-, left-, or interval-censored data, or may be a mixture
+of these categories, possibly with different upper and lower bounds.
+Censored data are sometimes excluded from the data analysis or replaced by a fixed value, which
+in both cases may lead to biased results. A more recommended approach to correctly model
+such data is based upon maximum likelihood \cite{kleinmoeschberger03,helsel05}.
+
+Censored data can thus contain left-censored, right-censored and interval-censored values, 
+with several lower and upper bounds. 
+Such data must be coded into a dataframe with two columns, respectively named \code{left} 
+and \code{right}, describing each observed value as an interval.
+The \code{left} column contains either \code{NA} for left censored observations,
+the left bound of the interval for interval censored observations,
+or the observed value for non-censored observations.
+The \code{right} column contains either \code{NA} for right censored observations,
+the right bound of the interval for interval censored observations,
+or the observed value for non-censored observations.
+To illustrate the use of package \pkg{fitdistrplus} to fit distributions
+to censored continous data, we  use another
+data set from ecotoxicology, included in our package and named  \code{salinity}. 
+This data set contains acute salinity tolerance (LC50 values in electrical conductivity, $mS$.$cm^{-1}$)
+of riverine macro-invertebrates taxa from the southern Murray-Darling Basin in 
+Central Victoria, Australia
+(see \cite{kefford07}).
+
+%%% R code
+<<datsalinity, echo=TRUE>>=
+data(salinity)
+str(salinity)
+@
+
+%\subsubsection{Graphical display of the observed distribution}
+%\label{censored:graph}
+%%%%%%%%%%
+
+Using censored data such as those coded in the \code{salinity} data set,
+the empirical distribution can be plotted using the \code{plotdistcens} function.
+By default, this function uses the Expectation-Maximization approach of Turnbull \cite{Turnbull74} to compute the overall
+empirical cdf curve with optional confidence intervals, by calls to \code{survfit}
+and \code{plot.survfit} functions from the \pkg{survival} package 
+(Figure~\ref{cdfcompcens} shows the Turnbull plot of data together with two fitted distributions). 
+A less rigorous
+but sometimes more illustrative plot to see the real nature of censored data can be 
+obtained by fixing the argument \code{Turnbull} to \code{FALSE} in the call to \code{plotdistcens}
+(see Figure~\ref{plotdistcens} for an example and the help page of Function
+\code{plotdistcens} for details).
+
+%%% R code
+%<<plotsalinity.echo, echo=TRUE, fig=FALSE>>=
+%plotdistcens(salinity)
+%@
+
+%%% R code
+<<plotsalinity2.echo, echo=TRUE, fig=FALSE>>=
+plotdistcens(salinity,Turnbull = FALSE)
+@
+
+% pourquoi ne pas travailler avec un exemple simulé?
+%
+
+\setkeys{Gin}{width=0.5\textwidth}
+
+\begin{figure}[htb!]
+  \centering
+  %%% R code
+<<plotsalinity, echo=FALSE, fig=TRUE, width=6, height=4, eps=FALSE, result=hide>>=
+plotdistcens(salinity,Turnbull = FALSE)
+@
+  \caption{Simple plot of censored raw data (72-hour acute salinity tolerance of riverine macro-invertebrates from 
+   the \code{salinity} data set) as ordered points 
+   and intervals}
+  \label{plotdistcens}
+\end{figure}
+
+%\subsubsection{Maximum likelihood estimation}
+%\label{censored:MLE}
+%%%%%%%%%%
+
+As for non censored data, one or more parametric distributions can
+be fitted to the censored data set, one at a time, but using in this case
+the \code{fitdistcens} function. This function estimates the vector of distribution parameters $\theta$
+by maximizing the likelihood for censored data defined as:
+  \begin{equation}
+    \label{likelihoodC}
+    L(\theta)=\prod_{i=1}^{N_{nonC}} f(x_{i}|\theta)
+    \times \prod_{j=1}^{N_{leftC}} F(x^{upper}_{j}|\theta)
+    \times \prod_{k=1}^{N_{rightC}} (1- F(x^{lower}_{k}|\theta))
+    \times \prod_{m=1}^{N_{intC}} (F(x^{upper}_{m}|\theta)- F(x^{lower}_{j}|\theta))
+  \end{equation}
+with $x_{i}$ the $N_{nonC}$ non-censored observations,
+$x^{upper}_{j}$ upper values defining the $N_{leftC}$ left-censored observations,
+$x^{lower}_{k}$ lower values defining the $N_{rightC}$ right-censored observations,
+$[x^{lower}_{m} ; x^{upper}_{m}]$ the intervals defining the $N_{intC}$ interval-censored observations,
+and F the cumulative distribution function of the parametric distribution 
+(\cite{kleinmoeschberger03,helsel05}).
+
+As \code{fitdist}, \code{fitdistcens} returns the results of the fit of any parametric distribution
+to a data set as a S3 class object that can be easily printed,
+summarized or plotted.
+For the \code{salinity} data set, a lognormal distribution or a loglogistic can be fitted 
+as commonly done in ecotoxicology for such data. 
+As with \code{fitdist}, for some distributions (see \cite{fitdistrplus} for details), 
+it is necessary to specify initial 
+values for the distribution parameters in the argument \code{start}.
+The \code{plotdistcens} function can help to find correct initial values for 
+the distribution parameters in non trivial cases, by a manual iterative use if necessary.
+%%% R code
+<<fitsalinity.echo, echo=TRUE, fig=FALSE>>=
+fsal.ln <- fitdistcens(salinity, "lnorm")
+fsal.ll <- fitdistcens(salinity, "llogis", start=list(shape=5, scale=40))
+summary(fsal.ln)
+summary(fsal.ll)
+@
+
+To our knowledge, computations of goodness-of-fit statistics have not yet been 
+developed for fits using censored data but the quality of fit 
+can be judged using Akaike and Schwarz's Bayesian 
+information criteria (AIC and BIC) and the goodness-of-fit CDF plot, 
+respectively provided when summarizing or plotting an object of class \code{"fitdistcens"}. 
+Function \code{cdfcompcens} can also be used to compare the fit of various
+distributions to the same censored data set. Its call is similar to the one of \code{cdfcomp}. Below is
+an example of comparison of the two fitted distributions to the \code{salinity} data set (see Figure~\ref{cdfcompcens}).
+
+%%% R code
+<<fitsalinity.cdfcomp.echo, echo=TRUE, fig=FALSE, eval=FALSE>>=
+cdfcompcens(list(fsal.ln, fsal.ll),
+    legendtext=c("lognormal", "loglogistic "))
+@
+
+\setkeys{Gin}{width=0.5\textwidth} %default
+\begin{figure}[htb!]
+  \centering
+  %%% R code
+<<fitsalinitycdfcompplot, echo=FALSE, fig=TRUE, width=6, height=4, eps=FALSE>>=
+cdfcompcens(list(fsal.ln, fsal.ll),
+    legendtext=c("lognormal", "loglogistic "))
+@
+  \caption{Goodness-of-fit CDF plot for fits of a lognormal and a loglogistic distribution
+  to censored data: LC50 values from 
+   the \code{salinity} data set}
+  \label{cdfcompcens}
+\end{figure}
+
+Function \code{bootdistcens} is the equivalent of \code{bootdist} for censored data, except that it 
+only proposes nonparametric bootstrap, as it is not obvious to simulate censoring within a parametric 
+bootstrap resampling procedure. The generic function \code{quantile} can also be applied, as for 
+continuous non-censored data,
+to an object of class \code{"fitdistcens"} or \code{"bootdistcens"}.
+
+In addition to the fit of distributions to censored or non censored continuous data, our package
+can also accomodate discrete variables, such as count numbers, using the functions 
+developped for continuous non-censored data. 
+These functions will provide somewhat different graphs
+and statistics, taking into account the discrete type of the modeled variable.
+The discrete nature of the variable is automatically recognized when a classical distribution is fitted to data
+(binomial, negative binomial, geometric, hypergeometric and Poisson distributions) but must 
+be indicated by fixing argument \code{discrete} to \code{TRUE} in the call to functions in other cases.
+The \code{toxocara} data set included in the package corresponds to the observation of such a discrete variable. 
+It reports numbers of \emph{Toxocara cati} parasites present in digestive tract,
+from a random sampling of feral cats living on Kerguelen island (\cite{Fromont01}). We  use it 
+to illustrate the case of discrete data.
+%%% R code
+<<dattoxocara, echo=TRUE>>=
+data(toxocara)
+str(toxocara)
+@
+
+The fit of a discrete distribution to discrete data by maximum
+likelihood estimation requires the same procedure as for continuous non-censored data. 
+As an example, 
+using the \code{toxocara} data set, Poisson and
+negative binomial distributions can be easily fitted.
+%%% R code
+<<fittoxocara.poisnbinom, echo = TRUE, fig = FALSE>>=
+(ftoxo.P <- fitdist(toxocara$number, "pois"))
+(ftoxo.nb <- fitdist(toxocara$number, "nbinom"))
+@
+
+For discrete distributions, the plot of an object of class \code{"fitdist"} simply provides
+two goodness-of-fit plots comparing empirical and theoretical distributions
+in pdf and in cd f(Figure~\ref{plotdiscfitP}). Function \code{cdfcomp} can also be used to compare several plots
+to the same data set, as follows for the previous fits (Figure~\ref{plotdiscfit}). 
+<<fittoxocara.pois.echo, echo=TRUE, fig=FALSE>>=
+plot(ftoxo.P)
+@
+
+\setkeys{Gin}{width=0.8\textwidth} %default
+
+\begin{figure}[htb]
+  \centering
+<<fittoxocarapois, echo=FALSE, results=hide, fig=TRUE, width=7, height=3.5, eps=FALSE>>=
+plot(ftoxo.P)
+@
+  \caption{Plot of the fit of a discrete distribution (Poisson distribution
+  fitted to numbers of \emph{Toxocara cati} parasites from the \code{toxocara} data set)}
+  \label{plotdiscfitP}
+\end{figure}
+
+<<fittoxocara.poisnbinom.echo, echo=TRUE, fig=FALSE>>=
+cdfcomp(list(ftoxo.P,ftoxo.nb),legendtext=c("Poisson", "negative binomial"))
+@
+
+\setkeys{Gin}{width=0.5\textwidth} %default
+
+\begin{figure}[htb]
+  \centering
+<<fittoxocarapoisnbinomplot, echo=FALSE, results=hide, fig=TRUE, width=6, height=4, eps=FALSE>>=
+cdfcomp(list(ftoxo.P,ftoxo.nb),legendtext=c("Poisson", "negative binomial"))
+@
+  \caption{Comparison of the fits of a negative binomial and a Poisson distribution
+   to numbers of \emph{Toxocara cati} parasites from the \code{toxocara} data set)}
+  \label{plotdiscfit}
+\end{figure}
+
+When fitting discrete distributions, the Chi-squared statistic is computed by 
+the \code{gofstat} function using cells defined 
+by the argument \code{chisqbreaks} or cells automatically defined from the data in order 
+to reach roughly the same number of observations per cell, roughly equal to the argument
+\code{meancount}, or sligthly more if there are some ties. 
+The choice to define cells from the empirical distribution (data), and not from the
+theoretical distribution, was done to enable the comparison of Chi-squared values obtained
+with different distributions fitted on a same data set.
+If arguments \code{chisqbreaks} and \code{meancount} 
+are both omitted, \code{meancount} is fixed in order to obtain roughly $(4n)^{2/5}$ cells, 
+with $n$ the length of the data set~\cite{Vose10}. Using this default option with 
+the fits of the Poisson and the negative binomial distribution to \code{toxocara} data set 
+are compared as follows, 
+giving the preference
+to the negative binomial distribution, from both Chi-squared statistics and information criteria:
+%%% R code
+<<fittoxocara.poisnbinom.gof, echo=TRUE, fig=FALSE>>=
+gofstat(list(ftoxo.P,ftoxo.nb))
+@
+
+\section{Conclusion}
+\label{ccl}
+%%%%%%%%%%
+
+The \proglang{R}  package \pkg{fitdistrplus} allows to easily fit distributions. 
+Our main objective while developing this package was to provide tools for helping 
+\proglang{R} users to fit distributions to data. 
+We have been encouraged to pursue our work by feedbacks 
+from users of our package in various areas as food or environmental risk assessment, epidemiology, ecology, 
+molecular biology, genomics, bioinformatics, hydraulics, mechanics, financial and actuarial mathematics 
+or operations research. Indeed, this package is already used a lot by practionners and academics 
+for simple MLE fits in \cite{distrModJSS,jaloustreetal11,saketal11,wilson12,kochetal12,marquetouxetal12,nordan12,
+scholletal12,suuronenetal12,varoetal12,mandletal13}, for MLE fits and goodness-of-fit statistics in 
+\cite{garcia12,tarnczi11,bagariaetal12,benavidesetal12,breitbach12,poduje12,pouillotetal10}, for MLE fits 
+and bootstrap in  \cite{croucheretal12,meheustetal12,orellanoetal12,telloetal12,hoelzeretal12}, for MLE fits, 
+bootstrap and goodness-of-fit statistics in \cite{rebuge12}, for MME fit in \cite{luangkesornetal12}, for 
+censored MLE and bootstrap in \cite{lehaetal11,poulliotetal12,jongenburgeretal12,commeauetal12,pouillotetal11}, 
+for graphic analysing in \cite{anandetal12,rosa12}, for grouped-data fitting methods in  \cite{fusteinercostafreda12} 
+and more generally in  \cite{eling12,busschaertetal10,brooksetal11,aktassjostrand11,eikhermann12,bakos11}.
+Many extensions of this package are planned in the future: 
+we target to extend to censored data some methods for the moment only available for non-censored data, 
+especially concerning goodness-of-fit evaluation and fitting methods. We will also enlarge the choice of 
+fitting methods for non-censored data, by proposing new goodness-of-fit distances (e.g. distances 
+based on quantiles) for maximum goodness-of-fit estimation 
+and new types of moments (e.g. limited expected values) for moment matching estimation. At last, we will
+consider the case of multivariate distribution fitting.
+
+\section{Acknowledgments}
+\label{merci}
+%%%%%%%%%%
+The package would not have been at this stage without the stimulating contribution of R\'egis Pouillot  and 
+Jean-Baptiste Denis, especially for its conceptualization. We also want to thank R\'egis Pouillot for 
+his very valuable comments on the first version of this paper.
+
+%\bibliographystyle{plain}
+\bibliography{../vignettes/fitdistrplus,../vignettes/paperscitingfitdistrplus}
 

### 2013-03-28 JSS/fitdistrplus/paper2JSS.Rnw
new version of the paper for an easier compilation and with my affiliation modified
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@237 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\author{ Marie Laure Delignette-Muller \\ Université Claude Bernard Lyon 1 \And
-    Christophe Dutang \\ Université de Strasbourg}
+\author{ Marie Laure Delignette-Muller \\ Universit\'e de Lyon \And
+    Christophe Dutang \\ Universit\'e de Strasbourg}
 \Plainauthor{Marie Laure Delignette-Muller}
@@
     Marie Laure Delignette-Muller \\
-    VetAgro Sup Campus Vétérinaire de Lyon \\
+    Universit\'e de Lyon\\
+    Universit\'e Lyon 1, CNRS, UMR5558, Laboratoire de Biom\'etrie et Biologie \'evolutive\\
+    VetAgro Sup, Campus V\'et\'erinaire de Lyon \\
     1, avenue Bourgelat \\
@@
 %\bibliographystyle{plain}
-\bibliography{../vignettes/fitdistrplus,../vignettes/paperscitingfitdistrplus}
+\bibliography{fitdistrplus,paperscitingfitdistrplus}
 

### 2013-03-28 JSS/fitdistrplus/paper2JSS.Rnw
last version before submit?
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@240 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 <<par4vignette, echo=FALSE, results=hide>>=
-options(digits=4, SweaveHooks=list(fig=function() par(mar=c(5.1, 4.1, 1.1, 2.1))) )
+options(digits=4, SweaveHooks=list(fig=function() par(mar=c(5.1, 4.1, 1.1, 2.1))))
 set.seed(1234)
@@
 (arthropods, fish or nonarthropod invertebrates, see \cite{Hose04}). 
-<<datendosulfan, echo=TRUE>>=
+<<datendosulfan, echo=TRUE, eval=FALSE>>=
 data(endosulfan)
 str(endosulfan)
+@
+<<datendosulfan2, echo=FALSE, eval=TRUE>>=
+data(endosulfan)
+str(endosulfan, width = 70, strict.width = "cut")
 @
@@
 cdfcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), xlogscale = TRUE,
-        ylogscale = TRUE,legendtext = c("lognormal","loglogistic","Pareto","Burr"))
+      ylogscale = TRUE, legendtext = 
+          c("lognormal", "loglogistic", "Pareto", "Burr"))
 @
@@
   \centering
-<<fitendo, echo=FALSE, fig=TRUE, width=6, height=4, eps=FALSE>>=
+<<fitendo, echo=FALSE, fig=TRUE, width=4, height=4, eps=FALSE>>=
 cdfcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), xlogscale = TRUE,
@@
 goodness-of-fit statistics.
+They can be computed using the function \code{gofstat}
+as defined by Stephens \cite{Stephens86}.
 
@@
       $\max(D^{+},D^{-})$ with\\
-      (KS)&& $D^{+}=\max\limits_{i=1,\dots,n}\left(\frac{i}{n} - F(x_{i})\right)
-      ;D^{-}=\max\limits_{i=1,\dots,n}\left(F(x_{i})-\frac{i-1}{n}\right)$ \\
+      (KS) & &  $D^{+}=\max\limits_{i=1,\dots,n}\left(\frac{i}{n} - F_i\right)$ \\
+           & &  $D^{-}=\max\limits_{i=1,\dots,n}\left(F_{i}-\frac{i-1}{n}\right)$ \\
     \hline
       Cramer-von Mises  & n $\int_{-\infty}^{\infty}(F_{n}(x) - F(x))^2 dx$ &
-        $\frac{1}{12n} + \sum\limits_{i=1}^n \left(F(x_{i})-\frac{2i-1}{2n} \right)^{2}$\\
+        $\frac{1}{12n} + \sum\limits_{i=1}^n \left(F_i-\frac{2i-1}{2n} \right)^{2}$\\
       (CvM)&&\\
@@
       Anderson-Darling  & n $\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2}{F(x) (1 - F(x))} dx$
-      & $-n -\frac{1}{n}\sum\limits_{i=1}^n \left((2i-1)(\log(F(x_{i}))+\log(1-F(x_{n+1-i})) )\right)$ \\
+      & $-n -\frac{1}{n}\sum\limits_{i=1}^n (2i-1)\log(F_i(1-F_{n+1-i}))$ \\
       (AD) &    & \\
@@
     \hline
+     where $F_i\stackrel{\triangle}{=} F(x_i)$
     \end{tabular}
@@
 
-They can be computed using the function \code{gofstat}
-as defined by Stephens \cite{Stephens86}.
-
 <<fendo.gof.print, echo=TRUE, fig=FALSE>>=
@@
 @
+
+\newpage
 As giving  more weight to distribution tails, 
@@
       Right-tail AD  & $\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{1 - F(x)}  dx$
-      & $\frac{n}{2} -2\sum_{i}F(x_{i}) -\frac{1}{n}\sum_{i}((2i-1)ln(1-F(x_{n+1-i})))$ \\
+      & $\frac{n}{2} -2\sum\limits_{i=1}^nF_i -\frac{1}{n}\sum\limits_{i=1}^n(2i-1)ln(\overline F_{n+1-i})$ \\
       (ADR) &  &\\
@@
       Left-tail AD  & $\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{(F(x))}  dx$
-      & $-\frac{3n}{2} +2\sum_{i}F(x_{i}) -\frac{1}{n}\sum_{i}((2i-1)ln(F(x_{i})))$ \\
+      & $-\frac{3n}{2} +2\sum\limits_{i=1}^nF_i -\frac{1}{n}\sum\limits_{i=1}^n(2i-1)ln(F_i)$ \\
       (ADL) &  &\\
@@
       Right-tail AD  & $ad2r=\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{(1 - F(x))^{2}}  dx$
-      & $ad2r=2\sum_{i}ln(1-F(x_{i})) +\frac{1}{n}\sum_{i} \frac{2i-1}{1-F(x_{n+1-i})}$ \\
+      & $ad2r=2\sum\limits_{i=1}^nln(\overline F_i) +\frac{1}{n}\sum\limits_{i=1}^n \frac{2i-1}{\overline F_{n+1-i}}$ \\
       2nd order (AD2R) &  &\\
@@
       Left-tail AD  & $ad2l=\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{(F(x))^{2}}  dx$
-      & $ad2l=2\sum_{i}ln(F(x_{i})) +\frac{1}{n}\sum_{i}\frac{2i-1}{F(x_{i})}$ \\
+      & $ad2l=2\sum\limits_{i=1}^nln(F_i) +\frac{1}{n}\sum\limits_{i=1}^n\frac{2i-1}{F_i}$ \\
       2nd order (AD2L) &  &\\
@@
     \hline
+    where $F_i\stackrel{\triangle}{=} F(x_{i})$; & $\overline F_i\stackrel{\triangle}{=}1-F(x_{i})$
     \end{tabular}
@@
   %%% R code
-<<mgegofcompplot, echo=FALSE, fig=TRUE, width=6, height=4, eps=FALSE>>=
+<<mgegofcompplot, echo=FALSE, fig=TRUE, width=4, height=4, eps=FALSE>>=
 cdfcomp(list(fendo.ln, fendo.ln.ADL, fendo.ln.AD2L),
@@
 <<quantilefitdist, echo=TRUE, fig=FALSE>>=
-( HC5.estimates <- cbind(empirical = as.numeric(quantile(ATV, probs = 0.05)),
-      Burr = as.numeric(quantile(fendo.B,probs = 0.05)$quantiles),
-      lognormal_MLE = as.numeric(quantile(fendo.ln,probs = 0.05)$quantiles),
-      lognormal_AD2 = as.numeric(quantile(fendo.ln.ADL,probs = 0.05)$quantiles),
-      lognormal_AD2L = as.numeric(quantile(fendo.ln.AD2L,probs = 0.05)$quantiles)) )
+(HC5.estimates <- c(
+empirical = as.numeric(quantile(ATV, probs=0.05)),
+Burr = as.numeric(quantile(fendo.B, probs=0.05)$quantiles),
+lognormal_MLE = as.numeric(quantile(fendo.ln, probs=0.05)$quantiles),
+lognormal_AD2 = as.numeric(quantile(fendo.ln.ADL, probs=0.05)$quantiles),
+lognormal_AD2L = as.numeric(quantile(fendo.ln.AD2L,probs=0.05)$quantiles)))
 @
@@
 cdfcomp(list(fdanish.ln.MLE, fdanish.ln.MME), 
-        legend=c("lognormal MLE", "lognormal MME"), main="Fitting a lognormal distribution",
+        legend=c("lognormal MLE", "lognormal MME"), 
+        main="Fitting a lognormal distribution",
         xlogscale=TRUE, datapch=".")
@@
 library(actuar)
-fdanish.P.MLE <- fitdist(danishuni$Loss, "pareto",start=c(shape=10, scale=10), 
-      lower=2+1e-6, upper=Inf)
-memp <- function(x, order) ifelse(order == 1, mean(x), sum(x^order)/length(x))
+fdanish.P.MLE <- fitdist(danishuni$Loss, "pareto", 
+start=c(shape=10, scale=10), lower=2+1e-6, upper=Inf)
+memp <- function(x, order) sum(x^order)/length(x)
 fdanish.P.MME <- fitdist(danishuni$Loss, "pareto", method="mme", order=1:2, 
-      memp="memp", start=c(shape=10, scale=10), lower=c(2+1e-6,2+1e-6), upper=c(Inf,Inf))
+memp="memp", start=c(shape=10, scale=10), lower=c(2+1e-6,2+1e-6), 
+upper=c(Inf,Inf))
 cdfcomp(list(fdanish.P.MLE, fdanish.P.MME), 
@@
 
-\setkeys{Gin}{width=0.8\textwidth} %default
+\setkeys{Gin}{width=0.9\textwidth} %default
 
@@
 <<danish.qme.echo, echo=TRUE, fig=FALSE>>=
-fdanish.ln.QME1 <- fitdist(danishuni$Loss, "lnorm", method="qme", probs=c(1/3, 2/3))
-fdanish.ln.QME2 <- fitdist(danishuni$Loss, "lnorm", method="qme", probs=c(8/10, 9/10))
+fdanish.ln.QME1 <- fitdist(danishuni$Loss, "lnorm", method="qme", 
+probs=c(1/3, 2/3))
+fdanish.ln.QME2 <- fitdist(danishuni$Loss, "lnorm", method="qme", 
+probs=c(8/10, 9/10))
 cdfcomp(list(fdanish.ln.MLE, fdanish.ln.QME1, fdanish.ln.QME2), 
@@
 fSANN <- fitdist(groundbeef$serving, "gamma", optim.method="SANN")
-fCG <- try(fitdist(groundbeef$serving, "gamma", optim.method="CG", control=list(maxit=10000)))
+fCG <- try(fitdist(groundbeef$serving, "gamma", optim.method="CG", 
+    control=list(maxit=10000)))
 if(class(fCG) == "try-error")
@@
 <<optimmethod.customgenoud.fitdist, echo=TRUE, eval=TRUE>>=
-fgenoud <- mledist(groundbeef$serving, "gamma", custom.optim= mygenoud, nvars=2, 
-    max.generations=10, Domains=cbind(c(0,0), c(10,10)), boundary.enforcement=1, 
-    hessian=TRUE, print.level=0, P9=10)
+fgenoud <- mledist(groundbeef$serving, "gamma", custom.optim= mygenoud, 
+    nvars=2, max.generations=10, Domains=cbind(c(0,0), c(10,10)), 
+    boundary.enforcement=1, hessian=TRUE, print.level=0, P9=10)
 cbind(NM = fNM$estimate,
@@
   %%% R code
-<<plotsalinity, echo=FALSE, fig=TRUE, width=6, height=4, eps=FALSE, result=hide>>=
+<<plotsalinity, echo=FALSE, fig=TRUE, width=6, height=6, eps=FALSE, result=hide>>=
 plotdistcens(salinity,Turnbull = FALSE)
@@
   %%% R code
-<<fitsalinitycdfcompplot, echo=FALSE, fig=TRUE, width=6, height=4, eps=FALSE>>=
+<<fitsalinitycdfcompplot, echo=FALSE, fig=TRUE, width=6, height=6, eps=FALSE>>=
 cdfcompcens(list(fsal.ln, fsal.ll),
@@
   \centering
-<<fittoxocarapoisnbinomplot, echo=FALSE, results=hide, fig=TRUE, width=6, height=4, eps=FALSE>>=
+<<fittoxocarapoisnbinomplot, echo=FALSE, results=hide, fig=TRUE, width=4, height=4, eps=FALSE>>=
 cdfcomp(list(ftoxo.P,ftoxo.nb),legendtext=c("Poisson", "negative binomial"))

### 2013-04-03 JSS/fitdistrplus/paper2JSS.Rnw
work on the bibtex files
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@241 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 estimations are 
-certainly the most commonly used method for fitting distributions. 
+certainly the most commonly used method for fitting distributions (\cite{Cullen99}). 
 Keeping in mind that these two methods may produce very different results,
@@
 Fitting of a parametric distribution may also be done by matching theoretical quantiles
-of the parametric distributions (for specified probabilities) against the empirical quantiles.
+of the parametric distributions (for specified probabilities) against the empirical quantiles (\cite{Tse2009}).
 Equation (\ref{quantiles}) below is thus very similar to Equations (\ref{moments}) and (\ref{centmoments})
@@
 or operations research. Indeed, this package is already used a lot by practionners and academics 
-for simple MLE fits in \cite{distrModJSS,jaloustreetal11,saketal11,wilson12,kochetal12,marquetouxetal12,nordan12,
+for simple MLE fits in \cite{distrModJSS,jaloustreetal11,saketal11,kochetal12,marquetouxetal12,nordan12,
 scholletal12,suuronenetal12,varoetal12,mandletal13}, for MLE fits and goodness-of-fit statistics in 
-\cite{garcia12,tarnczi11,bagariaetal12,benavidesetal12,breitbach12,poduje12,pouillotetal10}, for MLE fits 
+\cite{garcia12,tarnczi11,bagariaetal12,benavidesetal12,breitbach12,poduje12,Pouillot10}, for MLE fits 
 and bootstrap in  \cite{croucheretal12,meheustetal12,orellanoetal12,telloetal12,hoelzeretal12}, for MLE fits, 

### 2013-04-10 JSS/fitdistrplus/paper2JSS.Rnw
improvement of the form of the paper (espacially citation of references) and minor corrections
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@244 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 %%%
-%%% TITLE:       fitdistrplus: an R Package for Distribution Fitting Methods
+%%% TITLE:       fitdistrplus: an R Package for Fitting Distributions 
 %%% AUTHORS:     Marie Laure Delignette Muller*, Christophe Dutang
@@
 
-\title{\pkg{fitdistrplus}: An \proglang{R} Package for Distribution Fitting Methods}
-\Plaintitle{fitdistrplus: an R Package for Distribution Fitting Methods}
+\title{\pkg{fitdistrplus}: An \proglang{R} Package for Fitting Distributions}
+\Plaintitle{fitdistrplus: an R Package for Fitting Distributions}
 
@@
 distribution choice, parameter estimation, and quality of fit assessment.
-In the \proglang{R} (\cite{R13}) package \pkg{MASS} (\cite{MASS}), maximum likelihood estimation 
+In the \proglang{R} \citep{R13} package \pkg{MASS} \citep{MASS}, maximum likelihood estimation 
 is available via the \code{fitdistr} function; other steps of the
-fitting process can be done using other \proglang{R} functions, e.g. \cite{Ricci05}. 
-In this paper, we present the \proglang{R} package \pkg{fitdistrplus} (\cite{fitdistrplus})
+fitting process can be done using other \proglang{R} functions \citep{Ricci05}. 
+In this paper, we present the \proglang{R} package \pkg{fitdistrplus} \citep{fitdistrplus}
 implementing several methods for fitting univariate parametric distribution.
@@
 proposed in the \proglang{R} package \pkg{actuar} with three different goodness-of-fit distances
-(see \cite{actuarJSS}). 
+\citep{actuarJSS}. 
 While developping  the \pkg{fitdistrplus} package, a second objective  was to consider 
@@
 for optimization, 
-useful in cases where classical optimization techniques not included in \code{optim} are more adequate.
-
-In applied statistics, it is frequent to have to fit distributions to censored data \cite{kleinmoeschberger03,
+useful in cases where classical optimization techniques, not included in \code{optim}, are more adequate.
+
+In applied statistics, it is frequent to have to fit distributions to censored data \citep{kleinmoeschberger03,
 helsel05,busschaertetal10,lehaetal11,commeauetal12}. 
@@
 does not enable maximum likelihood estimation with this type of data. Some packages 
-deal with censored data, especially survival data \cite{survival} (see \cite{hiranoetal94,jordan05}
-for examples),
+deal with censored data, especially survival data \citep{survival,hiranoetal94,jordan05},
 but those packages generally focused on specific models, enabling the fit of  a 
@@
 and support different types of data. 
-The \pkg{distrMod} package of \cite{distrModJSS} provides an object-oriented (S4) 
+The \pkg{distrMod} package \citep{distrModJSS} provides an object-oriented (S4) 
 implementation of probability models for non-censored data
@@
 implemented, e.g. \code{confint} and \code{logLik}.
-In \pkg{fitdistrplus}, we choose to use the standard S3 class system for its understanding by most \proglang{R} users.
-When designing the \pkg{fitdistrplus} package, we do not forget to implement
+In \pkg{fitdistrplus}, we chose to use the standard S3 class system for its understanding by most \proglang{R} users.
+When designing the \pkg{fitdistrplus} package, we did not forget to implement
 generic functions also available for S3 classes.
@@
 For illustrating the use of various functions of the \pkg{fitdistrplus} package 
-with continuous non-censored data, we   first use
+with continuous non-censored data, we will  first use
 a data set named \code{groundbeef} which is included in our package.
@@
 children under 5 years old. It was used in a quantitative risk assessment 
-published in the International Journal of Food Microbiology \cite{Delignette08}.
+published by \cite{Delignette08}.
 %%% R code set default options for all R schunks
@@
 the kurtosis equals 3.
-The skewness and kurtosis and their corresponding unbiased estimator (e.g. \cite{casellaberger02}) from a sample 
+The skewness and kurtosis and their corresponding unbiased estimator \citep{casellaberger02} from a sample 
 $(X_i)_i \stackrel{\text{i.i.d.}}{\sim} X$ with observations $(x_i)_i$ are given by
@@
 A skewness-kurtosis plot such as the one proposed by \cite{Cullen99} is  provided by 
-the function \code{descdist} for the empirical distribution 
+the \code{descdist} function for the empirical distribution 
 (see Figure~\ref{Cullenplotcont} for the \code{groundbeef} data set). 
@@
 of the estimated values of kurtosis and skewness from data, a nonparametric bootstrap 
-procedure (\cite{efrontibshirani94}) can be performed by using the argument \code{boot}. %to an integer above 10.
+procedure \citep{efrontibshirani94} can be performed by using the argument \code{boot}. %to an integer above 10.
 Values of skewness and kurtosis are computed on bootstrap samples (constructed by 
@@
 skewness-kurtosis plot. 
-The properties of the random variable should be considered, notably its expected value, its range,
-as a complement to the use of the \code{plotdist} and \code{descdist}. 
+The properties of the random variable should be considered, notably its expected value and its range,
+as a complement to the use of the \code{plotdist} and \code{descdist} functions. 
 Below is a call to the \code{descdist} function to describe the distribution 
@@
 (e.g. \code{"norm"} for the normal distribution) 
-or by the density function itself from which
+or by the density function itself, from which
 the root name is extracted (e.g. \code{dnorm} for the normal distribution).
@@
 The plot of an object of class \code{"fitdist"} provides 
-four classical goodness-of-fit plots (\cite{Cullen99}) 
-presented on Figures \ref{groundbeef:comp} and \ref{endo:comp}: 
+four classical goodness-of-fit plots \citep{Cullen99} 
+presented on Figure \ref{groundbeef:comp}: 
 \begin{itemize}
@@
 
-For CDF, Q-Q and P-P plots, the probability plotting position is defined by default using the Hazen's rule, with probability points of the empirical distribution defined as
- \code{(1:n - 0.5)/n}, as recommended by Blom \cite{Blom}.
+For CDF, Q-Q and P-P plots, the probability plotting position is defined by default using Hazen's rule, with probability points of the empirical distribution calculated as
+ \code{(1:n - 0.5)/n}, as recommended by \cite{Blom}.
 This plotting position can be easily changed (see the reference 
-manual~\cite{fitdistrplus} for details).
+manual for details \citep{fitdistrplus}).
 % using the arguments 
@@
 further arguments to customize the plot (see the reference 
-manual~\cite{fitdistrplus} for lists of arguments that may be specific to each plot). 
+manual for lists of arguments that may be specific to each plot \citep{fitdistrplus}). 
 In the following example, we compare the fit of
@@
 The density plot and the CDF plot may be considered as the basic classical goodness-of-fits plots.
-% ca te va mieux reformule comme ca?
 The two other plots are complementary and can be very informative in some cases.
@@
 the lack-of-fit at the distribution center.
-As an example in Figure~\ref{groundbeef:comp}, none of the three fitted distributions correctly describes
+In the present example (in )Figure~\ref{groundbeef:comp}), none of the three fitted distributions correctly describes
 the center of the distribution, but the Weibull and gamma distributions 
@@
 
-The data set named  \code{endosulfan} is now used to illustrate other features of 
+The data set named  \code{endosulfan} will now be used to illustrate other features of 
 the \pkg{fitdistrplus} package. This data set contains
@@
 tested on Australian and non-Australian laboratory-species 
-(arthropods, fish or nonarthropod invertebrates, see \cite{Hose04}). 
+%(arthropods, fish or nonarthropod invertebrates)
+\citep{Hose04}. 
 <<datendosulfan, echo=TRUE, eval=FALSE>>=
@@
 to characterize the species sensitivity distribution (SSD) for a pollutant. A low percentile
-of the fitted distribution, generally the 5$\%$ percentile, is calculated and named 
+of the fitted distribution, generally the 5$\%$ percentile, is then calculated and named 
 the hazardous concentration 5$\%$ (HC5). It is interpreted as the value of the
-pollutant concentration protecting 95$\%$ of the species (\cite{Posthuma2010}).
+pollutant concentration protecting 95$\%$ of the species \citep{Posthuma2010}.
 But the fit of a lognormal or a loglogistic distribution to the whole \code{endosulfan} data set is rather bad,
@@
 It can thus help to find correct initial values for the distribution parameters in non trivial cases, by iterative calls if necessary (see the reference
-manual~\cite{fitdistrplus} for examples).
+manual for examples\citep{fitdistrplus}).
 },%----
@@
 When fitting continuous distributions, three  goodness-of-fit statistics are classicaly considered: 
-Cramer-von Mises, Kolmogorov-Smirnov and Anderson-Darling statistics \cite{Stephens86}.
+Cramer-von Mises, Kolmogorov-Smirnov and Anderson-Darling statistics \citep{Stephens86}.
 Naming $x_{i}$ the $n$ observations of a continuous variable $X$ arranged in an ascending order,
@@
 They can be computed using the function \code{gofstat}
-as defined by Stephens \cite{Stephens86}.
+as defined by Stephens \citep{Stephens86}.
+
+\clearpage
+
+<<fendo.gof.print, echo=TRUE, fig=FALSE>>=
+gofstat(list(fendo.ln, fendo.ll, fendo.P, fendo.B))
+@
 
@@
   \begin{center}
-    \caption{Goodness-of-fit statistics as defined by Stephens \cite{Stephens86}.}
+    \caption{Goodness-of-fit statistics as defined by Stephens \citep{Stephens86}.}
     \begin{tabular}{lll}
@@
 
-<<fendo.gof.print, echo=TRUE, fig=FALSE>>=
-gofstat(list(fendo.ln, fendo.ll, fendo.P, fendo.B))
-@
-
-\newpage
+%\newpage
 As giving  more weight to distribution tails, 
-Anderson-Darling statistics is of special interest when it matters to equally 
+the Anderson-Darling statistic is of special interest when it matters to equally 
 emphasize the tails as well as the main body of a distribution.
-This is often the case in risk assessment, e.g. \cite{Cullen99,Vose10}. For this reason, 
+This is often the case in risk assessment \citep{Cullen99,Vose10}. For this reason, 
 this statistics is often used to select the best distribution among those fitted.
@@
 continuous distributions (see Section~\ref{otherdata} and the reference 
-manual~\cite{fitdistrplus} for examples).
+manual for examples \citep{fitdistrplus}).
 
@@
 The uncertainty in the parameters of the fitted distribution may be estimated by parametric or nonparametric
-bootstraps using the \code{boodist} function for non-censored data, see \cite{efrontibshirani94}.
+bootstraps using the \code{boodist} function for non-censored data \citep{efrontibshirani94}.
 This function returns the bootstrapped values of parameters in a 
@@
 It can be directly used within a second-order Monte Carlo
-simulation framework, especially within the package \pkg{mc2d} (\cite{mc2d}). One could refer to Pouillot and Delignette-Muller
-(\cite{Pouillot10}) for an introduction to the use of \pkg{mc2d} 
+simulation framework, especially within the package \pkg{mc2d} \citep{mc2d}. One could refer to \cite{Pouillot10} for an introduction to the use of \pkg{mc2d} 
 and \pkg{fitdistrplus} packages in the context of quantitative 
@@
 is the maximum goodness-of-fit estimation method also called
-minimum distance estimation method \cite{Stephens86,actuarJSS}. 
+minimum distance estimation method \citep{Stephens86,actuarJSS}. 
 In this package this method is proposed with
@@
   \begin{center}
-    \caption{Modified Anderson-Darling statistics as defined by Luceno \cite{Luceno06}.}
+    \caption{Modified Anderson-Darling statistics as defined by \cite{Luceno06}.}
     \begin{tabular}{lll}
@@
 more weight to data at one tail of the distribution.
-Going back to the previous example from ecotoxicology, 
-instead of trying to find a less classical distribution to correctly
-fit the empirical distribution especially on its left tail, one could
-consider the fit of the classical lognormal distribution, but minimizing a
-goodness-of-fit distance giving more weight to the left tail of the empirical distribution,
-so as to correctly estimate the 5$\%$ percentile. 
-In the following example of \code{endosulfan} data set, we use left tail 
-Anderson-Darling distances of first or second order
+In the previous example from ecotoxicology, 
+we used a non classical distribution (the Burr distribution) to correctly
+fit the empirical distribution especially on its left tail. 
+In order to correctly estimate the 5$\%$ percentile, we could
+also consider the fit of the classical lognormal distribution, but minimizing a
+goodness-of-fit distance giving more weight to the left tail of the empirical distribution. 
+In what follows, the left tail 
+Anderson-Darling distances of first or second order are used to fit a lognormal
+to \code{endosulfan} data set
 (see Figure~\ref{plotfitMGE}).
@@
   \caption{Comparison of a lognormal distribution fitted by MLE
-  and by MGE using  two different goodness-of-fit distances : left-tail Anderson-Darling and left-tail Anderson Darling of seond order (example on the \code{endosulfan} data set)
+  and by MGE using  two different goodness-of-fit distances : left-tail Anderson-Darling and left-tail Anderson Darling of second order (example with the \code{endosulfan} data set)
   as provided by the \code{cdfcomp} function, with CDF values
@@
 %%%%%%%%%%
-Another method commonly used to fit parametric distribution is the moment matching estimation (MME). 
+The moment matching estimation (MME) is another method commonly used to fit parametric 
+distributions \citep{Vose10}. 
 It consists in finding the value of the parameter $\theta$  that equalizes the first theoretical 
@@
  $x_{i}$ the $n$ observations of variable $X$.
-For moments of order greater or equal than 2, it may  also be relevant to match centered moments as
+For moments of order greater than or equal to 2, it may  also be relevant to match centered moments as
 given by Equation (\ref{centmoments}).
@@
   \end{equation}
-This method can be performed by setting the argument \code{method} to \code{"mme"} in the call to fitdist.
+This method can be performed by setting the argument \code{method} to \code{"mme"} in the call to \code{fitdist}.
 The estimate is computed by a closed-form formula for the following distributions:
@@
 characterized by two parameters, these parameters are estimated by matching theoretical and observed 
-means and variances (see e.g. \cite{Vose10}).
+means and variances \citep{Vose10}.
 For other distributions, the equation of moments is solved numerically using 
@@
 sum of squared differences between observed and theoretical moments (see the \pkg{fitdistrplus} reference
-manual~\cite{fitdistrplus}
-for technical details).
+manual
+for technical details \citep{fitdistrplus}).
 
 A classical data set from the Danish insurance industry 
-published in \cite{mcneil97} is used to illustrate this method.
+published in \cite{mcneil97} will be used to illustrate this method.
 In \pkg{fitdistrplus}, the data set is stored in \code{danishuni} for the univariate version and contains
-the losse amounts collected at Copenhagen Reinsurance between 1980 and 1990.
-In actuarial science, it is standard to consider positive heavy-tailed distribution and have 
-a special focus on the right-tail of the distribution.
-In this numerical experiment,  the lognormal distribution is first considered and then the Pareto type II distribution (\cite{Klugmanetal09}).
-
-The lognormal distribution on \code{danishuni} data set is 
-fitted by matching moments implemented as a closed-form formula.
-On the left-hand graph of Figure \ref{fig:danish:mme}, we compare the fitted distribution function between
- MME and MLE method.
-The moment matching estimation provides a more cautious estimation of the insurance risk 
+the loss amounts collected at Copenhagen Reinsurance between 1980 and 1990.
+In actuarial science, it is standard to consider positive heavy-tailed distributions and have 
+a special focus on the right-tail of the distributions.
+In this numerical experiment,  the lognormal distribution is first considered and then the Pareto type II 
+distribution \citep{Klugmanetal09}.
+
+The lognormal distribution is 
+fitted to \code{danishuni} data set by matching moments implemented as a closed-form formula.
+On the left-hand graph of Figure \ref{fig:danish:mme}, the fitted distribution functions obtained 
+using the moment matching estimation (MME) and  maximum likelihood estimation (MLE) methods are compared.
+The MME method provides a more cautious estimation of the insurance risk 
 as the MME-fitted distribution function (resp. MLE-fitted) underestimates (overestimates) 
@@
 
+\setkeys{Gin}{width=0.9\textwidth} %default
+\begin{figure}[htb!]
+  \centering
+  %%% R code
+<<danishmmeplot, echo=FALSE, fig=TRUE, width=7, height=3.5, eps=FALSE>>=
+library(actuar)
+fdanish.P.MLE <- fitdist(danishuni$Loss, "pareto", 
+start=c(shape=10, scale=10), lower=2+1e-6, upper=Inf)
+memp <- function(x, order) sum(x^order)/length(x)
+fdanish.P.MME <- fitdist(danishuni$Loss, "pareto", method="mme", order=1:2, 
+memp="memp", start=c(shape=10, scale=10), lower=c(2+1e-6,2+1e-6), 
+upper=c(Inf,Inf))
+par(mfrow=c(1, 2))
+cdfcomp(list(fdanish.ln.MLE, fdanish.ln.MME), 
+        legend=c("lognormal MLE", "lognormal MME"), main="Fitting a lognormal distribution",
+        xlogscale=TRUE, datapch=".")
+cdfcomp(list(fdanish.P.MLE, fdanish.P.MME), 
+        legend=c("Pareto MLE", "Pareto MME"), main="Fitting a Pareto distribution",
+        xlogscale=TRUE, datapch=".")
+@
+\caption{Comparison between MME and MLE when fitting a lognormal or a Pareto distribution 
+to loss data from the \code{danishuni} data set}
+\label{fig:danish:mme}
+\end{figure}
+
+% je trouve nos legendes de graph trop longues?
+%
+
 In a second time, a Pareto type-II distribution, which gives more weight to 
@@
 We use the implementation of the \pkg{actuar} package providing raw and centered moments
-for that distribution (in addition to \code{d}, \code{p}, \code{q} and \code{r} functions,
-see \cite{actuar12}).
+for that distribution (in addition to \code{d}, \code{p}, \code{q} and \code{r} functions \citep{actuar12}.
 Fitting a heavy-tailed distribution for which the first and the second moments do not exist for
 certain values of the shape parameter requires some cautiousness.
-This is carried out by providing for the optimization process 
+This is carried out by providing, for the optimization process, 
 a lower and an upper bound for each parameter.
-Our call below immediately calls the L-BFGS-B optimization method in \code{optim}, since this quasi-Newton
+The code below calls the L-BFGS-B optimization method in \code{optim}, since this quasi-Newton
 allows box constraints\footnote{That is what the B stands for.}.
-Note that a function (called \code{memp} in our example) 
-for computing the empirical raw moment to \code{fitdist} has to be passed, 
+Note that a function 
+for computing the empirical raw moment (called \code{memp} in our example) has to be passed to \code{fitdist} , 
 since the user may choose either Equation (\ref{moments}) or (\ref{centmoments}).
@@
 
-\setkeys{Gin}{width=0.9\textwidth} %default
-
-\begin{figure}[htb!]
-  \centering
-  %%% R code
-<<danishmmeplot, echo=FALSE, fig=TRUE, width=7, height=3.5, eps=FALSE>>=
-par(mfrow=c(1, 2))
-cdfcomp(list(fdanish.ln.MLE, fdanish.ln.MME), 
-        legend=c("lognormal MLE", "lognormal MME"), main="Fitting a lognormal distribution",
-        xlogscale=TRUE, datapch=".")
-cdfcomp(list(fdanish.P.MLE, fdanish.P.MME), 
-        legend=c("Pareto MLE", "Pareto MME"), main="Fitting a Pareto distribution",
-        xlogscale=TRUE, datapch=".")
-@
-\caption{Comparison between MME and MLE when fitting a lognormal or a Pareto distribution 
-to loss data from the \code{danishuni} data set}
-\label{fig:danish:mme}
-\end{figure}
-
-% je trouve nos legendes de graph trop longues?
-%
-
-As shown on Figure \ref{fig:danish:mme}, matching moments and maximum likelihood fits
+As shown on Figure \ref{fig:danish:mme}, MME and MLE fits
 are far less distant (when looking at the right-tail) for the Pareto distribution 
 than for the lognormal distribution on this data set. 
-Furthermore for these two distributions, the MME method better fits the
+Furthermore, for these two distributions, the MME method better fits the
 right-tail of the distribution, which seems logical since empirical moments 
 are influenced by large observed values.
-In the previous traces, we give the values of goodness-of-fit statistics.
+In the previous traces, we gave the values of goodness-of-fit statistics.
 Whatever the statistic considered, the MLE-fitted lognormal always provides the best fit 
@@
 estimations are 
-certainly the most commonly used method for fitting distributions (\cite{Cullen99}). 
+certainly the most commonly used method for fitting distributions \citep{Cullen99}. 
 Keeping in mind that these two methods may produce very different results,
 the user should be aware of its great sensitivity to outliers
-when choosing the matching moment estimation (MME). This may be seen as an advantage in our example if the
+when choosing the moment matching estimation. This may be seen as an advantage in our example if the
 objective is to better describe the right tail of the distribution, but it may be seen as a
@@
 
-Quantile matching estimation is performed by setting the argument \code{method}
+Quantile matching estimation (QME) is performed by setting the argument \code{method}
 to \code{"qme"} in the call to \code{fitdist} and adding an argument \code{probs}
@@
 by matching probabilities $(p_1= 1/3, p_2=2/3)$ and $(p_1= 8/10, p_2=9/10)$.
-As expected, the second QME fit gives more weight to the right-tail of the distribution, despite
-we do not choose the Pareto type-II distribution.
+As expected, the second QME fit gives more weight to the right-tail of the distribution.
+%, despite we do not choose the Pareto type-II distribution.
 Compared to the maximum likelihood estimation, the second QME fit is also more conservative, whereas
@@
 
-%%%%%%%%%%%%% J'EN SUIS LA !!!!!!!!!!!!!!
-
 \subsection[Other types of data]{Fitting distributions to other types of data}
@@
 a limit of quantification, or more generally to uncertainty bounds around the observation.
-Right-censored data are also commonly encountered with survival data \cite{kleinmoeschberger03}.
+Right-censored data are also commonly encountered with survival data \citep{kleinmoeschberger03}.
 A data set may thus contain right-, left-, or interval-censored data, or may be a mixture
@@
 in both cases may lead to biased results. A more recommended approach to correctly model
-such data is based upon maximum likelihood \cite{kleinmoeschberger03,helsel05}.
-
-Censored data can thus contain left-censored, right-censored and interval-censored values, 
+such data is based upon maximum likelihood \citep{kleinmoeschberger03,helsel05}.
+
+Censored data may thus contain left-censored, right-censored and interval-censored values, 
 with several lower and upper bounds. 
@@
 To illustrate the use of package \pkg{fitdistrplus} to fit distributions
-to censored continous data, we  use another
+to censored continous data, we  will use another
 data set from ecotoxicology, included in our package and named  \code{salinity}. 
@@
 Central Victoria, Australia
-(see \cite{kefford07}).
+\citep{kefford07}.
 
@@
 the empirical distribution can be plotted using the \code{plotdistcens} function.
-By default, this function uses the Expectation-Maximization approach of Turnbull \cite{Turnbull74} to compute the overall
+By default, this function uses the Expectation-Maximization approach of \cite{Turnbull74} to compute the overall
 empirical cdf curve with optional confidence intervals, by calls to \code{survfit}
@@
 A less rigorous
-but sometimes more illustrative plot to see the real nature of censored data can be 
+but sometimes more illustrative plot can be 
 obtained by fixing the argument \code{Turnbull} to \code{FALSE} in the call to \code{plotdistcens}
 (see Figure~\ref{plotdistcens} for an example and the help page of Function
-\code{plotdistcens} for details).
+\code{plotdistcens} for details). This plot enables to see the real nature of censored data, as
+points and intervals.
 
@@
 and F the cumulative distribution function of the parametric distribution 
-(\cite{kleinmoeschberger03,helsel05}).
+\citep{kleinmoeschberger03,helsel05}.
 
@@
 
-To our knowledge, computations of goodness-of-fit statistics have not yet been 
+Computations of goodness-of-fit statistics have not yet been 
 developed for fits using censored data but the quality of fit 
@@
 Function \code{bootdistcens} is the equivalent of \code{bootdist} for censored data, except that it 
-only proposes nonparametric bootstrap, as it is not obvious to simulate censoring within a parametric 
-bootstrap resampling procedure. The generic function \code{quantile} can also be applied, as for 
-continuous non-censored data,
-to an object of class \code{"fitdistcens"} or \code{"bootdistcens"}.
+only proposes nonparametric bootstrap. Indeed, it is not obvious to simulate censoring within a parametric 
+bootstrap resampling procedure. The generic function \code{quantile} can also be applied
+to an object of class \code{"fitdistcens"} or \code{"bootdistcens"}, as for 
+continuous non-censored data.
 
@@
 It reports numbers of \emph{Toxocara cati} parasites present in digestive tract,
-from a random sampling of feral cats living on Kerguelen island (\cite{Fromont01}). We  use it 
+from a random sampling of feral cats living on Kerguelen island \citep{Fromont01}. We  will use it 
 to illustrate the case of discrete data.
@@
 two goodness-of-fit plots comparing empirical and theoretical distributions
-in pdf and in cd f(Figure~\ref{plotdiscfitP}). Function \code{cdfcomp} can also be used to compare several plots
+in density and in  CDF (Figure~\ref{plotdiscfitP}). Function \code{cdfcomp} can also be used to compare several plots
 to the same data set, as follows for the previous fits (Figure~\ref{plotdiscfit}). 
@@
   \caption{Plot of the fit of a discrete distribution (Poisson distribution
-  fitted to numbers of \emph{Toxocara cati} parasites from the \code{toxocara} data set)}
+  fitted to numbers of \emph{Toxocara cati} parasites from the \code{toxocara} data set}
   \label{plotdiscfitP}
@@
 by the argument \code{chisqbreaks} or cells automatically defined from the data in order 
-to reach roughly the same number of observations per cell, roughly equal to the argument
-\code{meancount}, or sligthly more if there are some ties. 
+to reach roughly the same number of observations per cell. This number is roughly equal to the argument
+\code{meancount}, or sligthly greater if there are some ties. 
 The choice to define cells from the empirical distribution (data), and not from the
@@
 are both omitted, \code{meancount} is fixed in order to obtain roughly $(4n)^{2/5}$ cells, 
-with $n$ the length of the data set~\cite{Vose10}. Using this default option with 
-the fits of the Poisson and the negative binomial distribution to \code{toxocara} data set 
-are compared as follows, 
+with $n$ the length of the data set \citep{Vose10}. Using this default option the two previous fits are 
+compared as follows, 
 giving the preference
@@
 molecular biology, genomics, bioinformatics, hydraulics, mechanics, financial and actuarial mathematics 
-or operations research. Indeed, this package is already used a lot by practionners and academics 
-for simple MLE fits in \cite{distrModJSS,jaloustreetal11,saketal11,kochetal12,marquetouxetal12,nordan12,
-scholletal12,suuronenetal12,varoetal12,mandletal13}, for MLE fits and goodness-of-fit statistics in 
-\cite{garcia12,tarnczi11,bagariaetal12,benavidesetal12,breitbach12,poduje12,Pouillot10}, for MLE fits 
-and bootstrap in  \cite{croucheretal12,meheustetal12,orellanoetal12,telloetal12,hoelzeretal12}, for MLE fits, 
-bootstrap and goodness-of-fit statistics in \cite{rebuge12}, for MME fit in \cite{luangkesornetal12}, for 
-censored MLE and bootstrap in \cite{lehaetal11,poulliotetal12,jongenburgeretal12,commeauetal12,pouillotetal11}, 
-for graphic analysing in \cite{anandetal12,rosa12}, for grouped-data fitting methods in  \cite{fusteinercostafreda12} 
-and more generally in  \cite{eling12,busschaertetal10,brooksetal11,aktassjostrand11,eikhermann12,bakos11}.
+or operations research. Indeed, this package is already used by a lot of practionners and academics 
+for simple MLE fits \citep{distrModJSS,jaloustreetal11,saketal11,kochetal12,marquetouxetal12,nordan12,
+scholletal12,suuronenetal12,varoetal12,mandletal13}, for MLE fits and goodness-of-fit statistics 
+\citep{garcia12,tarnczi11,bagariaetal12,benavidesetal12,breitbach12,poduje12,Pouillot10}, 
+for MLE fits 
+and bootstrap \citep{croucheretal12,meheustetal12,orellanoetal12,telloetal12,hoelzeretal12}, 
+for MLE fits, 
+bootstrap and goodness-of-fit statistics \citep{rebuge12}, 
+for MME fit \citep{luangkesornetal12}, 
+for 
+censored MLE and bootstrap \citep{lehaetal11,poulliotetal12,jongenburgeretal12,commeauetal12,pouillotetal11}, 
+for graphic analysing in \citep{anandetal12,rosa12}, 
+for grouped-data fitting methods \citep{fusteinercostafreda12} 
+or more generally \citep{eling12,busschaertetal10,brooksetal11,aktassjostrand11,eikhermann12,bakos11}.
 Many extensions of this package are planned in the future: 

### 2013-04-17 JSS/fitdistrplus/paper2JSS.Rnw
fix pb with moments.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@245 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 where $m_{2}$, $m_{3}$, $m_{4}$ denote empirical moments defined by
-$m_{r}=\frac{1}{n}\sum_{i=1}^n(x_{i}-\overline{x})^{r}$, with
+$m_{k}=\frac{1}{n}\sum_{i=1}^n(x_{i}-\overline{x})^{k}$, with
 $x_{i}$ the $n$ observations of variable $x$ and $\overline{x}$ their mean value.
@@
 distributions \citep{Vose10}. 
-It consists in finding the value of the parameter $\theta$  that equalizes the first theoretical 
-raw moments of the parametric distribution  to the empirical moments (Equation~\ref{moments})
+MME consists in finding the value of the parameter $\theta$  that equalizes the first theoretical 
+raw moments of the parametric distribution  to the corresponding empirical raw moments as in Equation (\ref{moments})
   \begin{equation}
@@
  $x_{i}$ the $n$ observations of variable $X$.
-For moments of order greater than or equal to 2, it may  also be relevant to match centered moments as
-given by Equation (\ref{centmoments}).
+For moments of order greater than or equal to 2, it may  also be relevant to match centered moments.
+Therefore, we match the moments given in Equation (\ref{centmoments})
   \begin{equation}
     \label{centmoments}
-    E\left((X-E(X))^{k}|\theta\right)=\frac{1}{n}\sum_{i=1}^{n}(x_{i}-\bar x_n)^{k} 
+    E(X\vert \theta) = \overline{x} ~,~
+    E\left((X-E(X))^{k}|\theta\right)=m_k, \text{ for }  k=2,\ldots,d,
   \end{equation}
+where $m_k$ denotes the empirical centered moments.
 This method can be performed by setting the argument \code{method} to \code{"mme"} in the call to \code{fitdist}.
@@
 allows box constraints\footnote{That is what the B stands for.}.
-Note that a function 
-for computing the empirical raw moment (called \code{memp} in our example) has to be passed to \code{fitdist} , 
-since the user may choose either Equation (\ref{moments}) or (\ref{centmoments}).
+We choose match moments defined in Equation (\ref{moments}), and so 
+a function for computing the empirical raw moment (called \code{memp} in our example) is passed to \code{fitdist}.
+For two-parameter distributions (i.e. $d=2$),  Equations (\ref{moments}) and (\ref{centmoments}) are equivalent.
 
@@
 legend=c("MLE", "QME(1/3, 2/3)", "QME(8/10, 9/10)"), 
-main="Fitting a lognormal distribution", xlogscale=TRUE, datapch=".")
+main="Fitting a lognormal distribution", xlogscale=TRUE, datapch=20)
 @
@@
 %, despite we do not choose the Pareto type-II distribution.
-Compared to the maximum likelihood estimation, the second QME fit is also more conservative, whereas
-the first QME fit is less conservative.
+Compared to the maximum likelihood estimation, the second QME fit best suits the right-tail
+of the distribution, whereas the first QME fit best models the body of the distribution.
 The quantile matching estimation is of particular interest when we need a focus 
@@
         legend=c("MLE", "QME(1/3, 2/3)", "QME(8/10, 9/10)"), main="Fitting a lognormal distribution",
-        xlogscale=TRUE, datapch=".")
+        xlogscale=TRUE, datapch=20)
 @

### 2013-04-17 JSS/fitdistrplus/paper2JSS.Rnw
last version, ready for submission
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@246 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 does not enable maximum likelihood estimation with this type of data. Some packages 
-deal with censored data, especially survival data \citep{survival,hiranoetal94,jordan05},
+can be used to work with censored data, especially survival data \citep{survival,hiranoetal94,jordan05},
 but those packages generally focused on specific models, enabling the fit of  a 
@@
 To help the user in this choice, we developed functions to plot and
-characterise the empirical distribution.
+characterize the empirical distribution.
 
@@
 (see \code{?Distributions}). Others
-may be found in various packages (see the CRAN task view: Probability Distributions). 
-%%%%%%%%%%% PEUT-ON METTRE UNE REF POUR LA CRAN TASK VIEW?
+may be found in various packages (see the CRAN task view: Probability Distributions at \url{http://cran.r-project.org/web/views/Distributions.html}). 
 Distributions not found in any package must be implemented by the user as
@@
 the lack-of-fit at the distribution center.
-In the present example (in )Figure~\ref{groundbeef:comp}), none of the three fitted distributions correctly describes
+In the present example (in Figure~\ref{groundbeef:comp}), none of the three fitted distributions correctly describes
 the center of the distribution, but the Weibull and gamma distributions 
@@
 pollutant concentration protecting 95$\%$ of the species \citep{Posthuma2010}.
-But the fit of a lognormal or a loglogistic distribution to the whole \code{endosulfan} data set is rather bad,
+But the fit of a lognormal or a loglogistic distribution to the whole \code{endosulfan} data set is rather bad (Figure~\ref{endo:comp}),
 especially due to a minority of very high values. 
@@
 values for the distribution parameters have to be supplied
-in the argument \code{start} as a named list  with initial values for each parameter 
+in the argument \code{start}, as a named list  with initial values for each parameter 
 (as they appear in the \code{d}, \code{p}, \code{q} functions). 
@@
 It can thus help to find correct initial values for the distribution parameters in non trivial cases, by iterative calls if necessary (see the reference
-manual for examples\citep{fitdistrplus}).
-},%----
+manual for examples \citep{fitdistrplus}).
+}, %----
 various distributions can be fitted and graphically
 compared. On this example, the function \code{cdfcomp} can be used to report CDF values in a logscale so as to emphasize discrepancies on the tail of
-interest while defining an HC5 value.
+interest while defining an HC5 value (Figure~\ref{endo:comp}).
 
@@
 empirical quantile for comparison.
+
+\newpage
+
 %%% R code
@@
 
-The uncertainty in the parameters of the fitted distribution may be estimated by parametric or nonparametric
+The uncertainty in the parameters of the fitted distribution can be estimated by parametric or nonparametric
 bootstraps using the \code{boodist} function for non-censored data \citep{efrontibshirani94}.
-This function returns the bootstrapped values of parameters in a 
-S3 class object which may be plotted to visualize the bootstrap region. 
+This function returns the bootstrapped values of parameters in an 
+S3 class object which can be plotted to visualize the bootstrap region. 
 The medians and the 95 percent confidence intervals of parameters (2.5 and 97.5 percentiles) are printed 
@@
 of the bootstrapped values of parameters providing
-a representation of the joint uncertainty distribution of the fitted parameters (see Figure~\ref{fig:bootstrap}).
+a representation of the joint uncertainty distribution of the fitted parameters.
 Below is an example of the use of the \code{bootdist} function with the previous fit of the Burr distribution to the
-\code{endosulfan} data set. 
+\code{endosulfan} data set (Figure~\ref{fig:bootstrap}). 
 %%% R code
@@
 
+\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 MME consists in finding the value of the parameter $\theta$  that equalizes the first theoretical 
-raw moments of the parametric distribution  to the corresponding empirical raw moments as in Equation (\ref{moments})
+raw moments of the parametric distribution  to the corresponding empirical raw moments as in Equation (\ref{moments}):
   \begin{equation}
@@
 For moments of order greater than or equal to 2, it may  also be relevant to match centered moments.
-Therefore, we match the moments given in Equation (\ref{centmoments})
+Therefore, we match the moments given in Equation (\ref{centmoments}):
   \begin{equation}
@@
         main="Fitting a lognormal distribution",
-        xlogscale=TRUE, datapch=".")
+        xlogscale=TRUE, datapch=20)
 @
@@
         legend=c("lognormal MLE", "lognormal MME"), main="Fitting a lognormal distribution",
-        xlogscale=TRUE, datapch=".")
+        xlogscale=TRUE, datapch=20)
 cdfcomp(list(fdanish.P.MLE, fdanish.P.MME), 
         legend=c("Pareto MLE", "Pareto MME"), main="Fitting a Pareto distribution",
-        xlogscale=TRUE, datapch=".")
+        xlogscale=TRUE, datapch=20)
 @
@@
 
-In a second time, a Pareto type-II distribution, which gives more weight to 
+In a second time, a Pareto distribution, which gives more weight to 
 the right-tail of the distribution, is fitted.
-As the lognormal distribution, the Pareto type-II has two parameters, which
+As the lognormal distribution, the Pareto has two parameters, which
 allows a fair comparison.
@@
 a function for computing the empirical raw moment (called \code{memp} in our example) is passed to \code{fitdist}.
-For two-parameter distributions (i.e. $d=2$),  Equations (\ref{moments}) and (\ref{centmoments}) are equivalent.
+For two-parameter distributions (i.e. $d=2$),  
+Equations (\ref{moments}) and (\ref{centmoments}) are equivalent.
 
@@
 Fitting of a parametric distribution may also be done by matching theoretical quantiles
-of the parametric distributions (for specified probabilities) against the empirical quantiles (\cite{Tse2009}).
-Equation (\ref{quantiles}) below is thus very similar to Equations (\ref{moments}) and (\ref{centmoments})
+of the parametric distributions (for specified probabilities) against the empirical quantiles (\cite{Tse2009}). The equality of theoretical and empirical qunatiles
+is expressed by
+Equation (\ref{quantiles}) below, which is very similar to Equations (\ref{moments}) and (\ref{centmoments}):
   \begin{equation}
@@
 %%%%%%%%%%
-Each time a numerical minimization (or maximization) is carried out in the \code{fitdistrplus} package, 
+Each time a numerical minimization is carried out in the \code{fitdistrplus} package, 
 the \code{optim} function of the \pkg{stats} package  is used by default with the 
@@
 interesting to change some options of the \code{optim} function or to use another optimization
-function than \code{optim} to maximize the likelihood or to minimize a squared difference.
+function than \code{optim} to minimize the objective function.
 The argument \code{optim.method} can be used in the call to \code{fitdist} or
-\code{fitdistcens}. It will internally be passed to \code{mledist} and to \code{optim}
+\code{fitdistcens}. It will internally be passed to \code{mledist},
+\code{mmedist}, \code{mgedist} or \code{qmedist}, and to \code{optim}
 (see \code{?optim} for details about the different algorithms available).
@@
 
-It is also possible to use another function than \code{optim} to maximize the likelihood by
+It is also possible to use another function than \code{optim} to minimize the objective function by
 specifying by the argument \code{custom.optim} in the call to \code{fitdist}. 
@@
 @
+
+%\newpage
 
@@
 with several lower and upper bounds. 
-Such data must be coded into a dataframe with two columns, respectively named \code{left} 
+Before their use in package \pkg{fitdistrplus}, 
+such data must be coded into a dataframe with two columns, 
+respectively named \code{left} 
 and \code{right}, describing each observed value as an interval.
@@
 As \code{fitdist}, \code{fitdistcens} returns the results of the fit of any parametric distribution
-to a data set as a S3 class object that can be easily printed,
+to a data set as an S3 class object that can be easily printed,
 summarized or plotted.
@@
 These functions will provide somewhat different graphs
-and statistics, taking into account the discrete type of the modeled variable.
+and statistics, taking into account the discrete nature of the modeled variable.
 The discrete nature of the variable is automatically recognized when a classical distribution is fitted to data
@@
 The \code{toxocara} data set included in the package corresponds to the observation of such a discrete variable. 
-It reports numbers of \emph{Toxocara cati} parasites present in digestive tract,
+Numbers of \emph{Toxocara cati} parasites present in digestive tract are reported
 from a random sampling of feral cats living on Kerguelen island \citep{Fromont01}. We  will use it 

### 2013-05-10 JSS/fitdistrplus/paper2JSS.Rnw
few changes to respect JSS requirement (to continue and check)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@248 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 producing S4 objects of class \code{"mle"} for which generic methods are 
-implemented, e.g. \code{confint} and \code{logLik}.
+implemented, e.g., \code{confint} and \code{logLik}.
 In \pkg{fitdistrplus}, we chose to use the standard S3 class system for its understanding by most \proglang{R} users.
@@
 <<par4vignette, echo=FALSE, results=hide>>=
-options(digits=4, SweaveHooks=list(fig=function() par(mar=c(5.1, 4.1, 1.1, 2.1))))
+options(digits=4, prompt="R>", 
+        SweaveHooks=list(fig=function() par(mar=c(5.1, 4.1, 1.1, 2.1))))
 set.seed(1234)
@@
 <<datgroundbeef, echo=TRUE>>=
-library(fitdistrplus)
-data(groundbeef)
+library("fitdistrplus")
+data("groundbeef")
 str(groundbeef)
@@
 function and the histogram, which can be obtained with the \code{plotdist} function of the \pkg{fitdistrplus} package.
-This function provides two plots (see Figure 1): the left-hand plot is the histogram (on a density scale) 
+This function provides two plots (see Figure~\ref{plotdistcont}): the left-hand plot is the histogram (on a density scale) 
 and the right-hand  plot the empirical cumulative distribution function (CDF).
@@
   \caption{Histogram and CDF plots of an empirical distribution for a continuous variable 
-  (serving size from the \code{groundbeef} data set) as provided by the \code{plotdist} function}
+  (serving size from the \code{groundbeef} data set) as provided by the \code{plotdist} function.}
   \label{plotdistcont}
@@
   \caption{Skewness-kurtosis plot for a continuous variable 
-  (serving size from the \code{groundbeef} data set) as provided by the \code{descdist} function}
+  (serving size from the \code{groundbeef} data set) as provided by the \code{descdist} function.}
   \label{Cullenplotcont}
@@
 Classical distributions are already defined in that way in the \pkg{stats} package,
-e.g. \code{dnorm}, \code{pnorm} and \code{qnorm} for the normal distribution
+e.g., \code{dnorm}, \code{pnorm} and \code{qnorm} for the normal distribution
 (see \code{?Distributions}). Others
@@
 used in the names of \code{d}, \code{p}, \code{q} functions
-(e.g. \code{"norm"} for the normal distribution) 
+(e.g., \code{"norm"} for the normal distribution) 
 or by the density function itself, from which
-the root name is extracted (e.g. \code{dnorm} for the normal distribution).
+the root name is extracted (e.g., \code{dnorm} for the normal distribution).
 Numerical results returned by the \code{fitdist} function are 
@@
 four classical goodness-of-fit plots \citep{Cullen99} 
-presented on Figure \ref{groundbeef:comp}: 
+presented on Figure~\ref{groundbeef:comp}: 
 \begin{itemize}
@@
   distributions fitted to serving sizes from the \code{groundbeef} data set)  
-  as provided by functions \code{denscomp}, \code{qqcomp}, \code{cdfcomp} and \code{ppcomp}}
+  as provided by functions \code{denscomp}, \code{qqcomp}, \code{cdfcomp} and \code{ppcomp}.}
 \label{groundbeef:comp}  
@@
 <<datendosulfan, echo=TRUE, eval=FALSE>>=
-data(endosulfan)
+data("endosulfan")
 str(endosulfan)
@@
 <<datendosulfan2, echo=FALSE, eval=TRUE>>=
-data(endosulfan)
+data("endosulfan")
 str(endosulfan, width = 70, strict.width = "cut")
@@
 fendo.ln <- fitdist(ATV, "lnorm")
-library(actuar)
+library("actuar")
 fendo.ll <- fitdist(ATV, "llogis", start=list(shape=1, scale=500))
@@
   (\code{endosulfan} data set) as provided by the \code{cdfcomp} function, with CDF values
-  in a logscale to emphasize discrepancies on the left tail}
+  in a logscale to emphasize discrepancies on the left tail.}
 \label{endo:comp}  
@@
 The purpose of goodness-of-fit statistics aims to measure the distance between the
-fitted parametric distribution and the empirical distribution: e.g. 
+fitted parametric distribution and the empirical distribution: e.g., 
 the distance between the fitted cumulative distribution function $F$ 
@@
 <<fendo.gof.print, echo=TRUE, fig=FALSE>>=
-gofstat(list(fendo.ln, fendo.ll, fendo.P, fendo.B))
+gofstat(list(fendo.ln, fendo.ll, fendo.P, fendo.B),
+        fitnames = c("lnorm","llogis","Pareto","Burr"))
 @
@@
 Kolmogorov-Smirnov ones, does not take into account the complexity of 
-the model (i.e. parameter number). It is not
+the model (i.e., parameter number). It is not
 a problem when compared distributions are characterized by the same number of parameters, but
@@
 characterized by three parameters (example on the \code{endosulfan} data set) as provided 
-by the plot of an object of class \code{"bootdist"}}
+by the plot of an object of class \code{"bootdist"}.}
   \label{fig:bootstrap}
@@
  %%% R code
-<<fitATV.lnorm.quantile, echo=TRUE>>=
+<<fitATV.lnorm.quantile, echo=TRUE,results=hide>>=
 quantile(bendo.B, probs = 0.05)
 @
+\begin{footnotesize}
+<<fitATV.lnorm.quantileb, echo=FALSE>>=
+quantile(bendo.B, probs = 0.05)
+@
+\end{footnotesize}
 
@@
   as provided by the \code{cdfcomp} function, with CDF values
-  in a logscale to emphasize discrepancies on the left tail}
+  in a logscale to emphasize discrepancies on the left tail.}
   \label{plotfitMGE}
@@
 MME consists in finding the value of the parameter $\theta$  that equalizes the first theoretical 
-raw moments of the parametric distribution  to the corresponding empirical raw moments as in Equation (\ref{moments}):
+raw moments of the parametric distribution  to the corresponding empirical raw moments as in Equation~(\ref{moments}):
   \begin{equation}
@@
 For moments of order greater than or equal to 2, it may  also be relevant to match centered moments.
-Therefore, we match the moments given in Equation (\ref{centmoments}):
+Therefore, we match the moments given in Equation~(\ref{centmoments}):
   \begin{equation}
@@
 fitted to \code{danishuni} data set by matching moments implemented as a closed-form formula.
-On the left-hand graph of Figure \ref{fig:danish:mme}, the fitted distribution functions obtained 
+On the left-hand graph of Figure~\ref{fig:danish:mme}, the fitted distribution functions obtained 
 using the moment matching estimation (MME) and  maximum likelihood estimation (MLE) methods are compared.
@@
 <<danish.mme, echo=TRUE, fig=FALSE>>=
-data(danishuni)
+data("danishuni")
 str(danishuni)
@@
 <<danishmmeplot, echo=FALSE, fig=TRUE, width=7, height=3.5, eps=FALSE>>=
-library(actuar)
+library("actuar")
 fdanish.P.MLE <- fitdist(danishuni$Loss, "pareto", 
@@
 \caption{Comparison between MME and MLE when fitting a lognormal or a Pareto distribution 
-to loss data from the \code{danishuni} data set}
+to loss data from the \code{danishuni} data set.}
 \label{fig:danish:mme}
@@
 allows box constraints\footnote{That is what the B stands for.}.
-We choose match moments defined in Equation (\ref{moments}), and so 
+We choose match moments defined in Equation~(\ref{moments}), and so 
 a function for computing the empirical raw moment (called \code{memp} in our example) is passed to \code{fitdist}.
-For two-parameter distributions (i.e. $d=2$),  
-Equations (\ref{moments}) and (\ref{centmoments}) are equivalent.
+For two-parameter distributions (i.e., $d=2$),  
+Equations~(\ref{moments}) and (\ref{centmoments}) are equivalent.
 
 <<danish.mme.pareto, echo=TRUE, fig=FALSE>>=
-library(actuar)
+library("actuar")
 fdanish.P.MLE <- fitdist(danishuni$Loss, "pareto", 
@@
 memp <- function(x, order) sum(x^order)/length(x)
-fdanish.P.MME <- fitdist(danishuni$Loss, "pareto", method="mme", order=1:2, 
-memp="memp", start=c(shape=10, scale=10), lower=c(2+1e-6,2+1e-6), 
-upper=c(Inf,Inf))
+fdanish.P.MME <- fitdist(danishuni$Loss, "pareto", method="mme", 
+order=1:2, memp="memp", start=c(shape=10, scale=10), 
+lower=c(2+1e-6,2+1e-6), upper=c(Inf,Inf))
 cdfcomp(list(fdanish.P.MLE, fdanish.P.MME), 
-legend=c("Pareto MLE", "Pareto MME"), main="Fitting a Pareto distribution",
-xlogscale=TRUE, datapch=".")
-gofstat(list(fdanish.ln.MLE, fdanish.P.MLE, fdanish.ln.MME, fdanish.P.MME))
-@
-
-As shown on Figure \ref{fig:danish:mme}, MME and MLE fits
+legend=c("Pareto MLE", "Pareto MME"), 
+main="Fitting a Pareto distribution", xlogscale=TRUE, datapch=".")
+gofstat(list(fdanish.ln.MLE, fdanish.P.MLE, 
+fdanish.ln.MME, fdanish.P.MME),
+fitnames = c("lnorm.mle","Pareto.mle","lnorm.mle","Pareto.mme"))
+@
+
+As shown on Figure~\ref{fig:danish:mme}, MME and MLE fits
 are far less distant (when looking at the right-tail) for the Pareto distribution 
@@
 is expressed by
-Equation (\ref{quantiles}) below, which is very similar to Equations (\ref{moments}) and (\ref{centmoments}):
+Equation~(\ref{quantiles}) below, which is very similar to Equations~(\ref{moments}) and (\ref{centmoments}):
   \begin{equation}
@@
 The quantile matching estimation is of particular interest when we need a focus 
-around particular quantiles, e.g. $p=99.5\%$ in the Solvency II insurance context or
+around particular quantiles, e.g., $p=99.5\%$ in the Solvency II insurance context or
 $p=5\%$ for the HC5 estimation in the ecotoxicology context.
@@
 @
-\caption{Comparison between QME and MLE when fitting a lognormal distribution to loss data from the \code{danishuni} data set}
+\caption{Comparison between QME and MLE when fitting a lognormal distribution to loss data from the \code{danishuni} data set.}
 \label{fig:danish:qme}
@@
 <<optimmethod.gamma, echo=TRUE>>=
-data(groundbeef)
+data("groundbeef")
 fNM <- fitdist(groundbeef$serving, "gamma", optim.method="Nelder-Mead")
@@
 <<datsalinity, echo=TRUE>>=
-data(salinity)
+data("salinity")
 str(salinity)
@@
    the \code{salinity} data set) as ordered points 
-   and intervals}
+   and intervals.}
   \label{plotdistcens}
@@
     \label{likelihoodC}
-    L(\theta)=\prod_{i=1}^{N_{nonC}} f(x_{i}|\theta)
-    \times \prod_{j=1}^{N_{leftC}} F(x^{upper}_{j}|\theta)
-    \times \prod_{k=1}^{N_{rightC}} (1- F(x^{lower}_{k}|\theta))
+    \begin{array}{lll}
+    L(\theta)&=&\prod_{i=1}^{N_{nonC}} f(x_{i}|\theta)
+    \times \prod_{j=1}^{N_{leftC}} F(x^{upper}_{j}|\theta) \\
+    & & \times \prod_{k=1}^{N_{rightC}} (1- F(x^{lower}_{k}|\theta))
     \times \prod_{m=1}^{N_{intC}} (F(x^{upper}_{m}|\theta)- F(x^{lower}_{j}|\theta))
+    \end{array}
   \end{equation}
@@
   to censored data: LC50 values from 
-   the \code{salinity} data set}
+   the \code{salinity} data set.}
   \label{cdfcompcens}
@@
 <<dattoxocara, echo=TRUE>>=
-data(toxocara)
+data("toxocara")
 str(toxocara)
@@
   \caption{Plot of the fit of a discrete distribution (Poisson distribution
-  fitted to numbers of \emph{Toxocara cati} parasites from the \code{toxocara} data set}
+  fitted to numbers of \emph{Toxocara cati} parasites from the \code{toxocara} data set.}
   \label{plotdiscfitP}
@@
 <<fittoxocara.poisnbinom.echo, echo=TRUE, fig=FALSE>>=
-cdfcomp(list(ftoxo.P,ftoxo.nb),legendtext=c("Poisson", "negative binomial"))
+cdfcomp(list(ftoxo.P,ftoxo.nb),
+legendtext=c("Poisson", "negative binomial"))
 @
@@
 <<fittoxocarapoisnbinomplot, echo=FALSE, results=hide, fig=TRUE, width=4, height=4, eps=FALSE>>=
-cdfcomp(list(ftoxo.P,ftoxo.nb),legendtext=c("Poisson", "negative binomial"))
+cdfcomp(list(ftoxo.P,ftoxo.nb),
+        legendtext=c("Poisson", "negative binomial"))
 @
   \caption{Comparison of the fits of a negative binomial and a Poisson distribution
-   to numbers of \emph{Toxocara cati} parasites from the \code{toxocara} data set)}
+   to numbers of \emph{Toxocara cati} parasites from the \code{toxocara} data set).}
   \label{plotdiscfit}
@@
 <<fittoxocara.poisnbinom.gof, echo=TRUE, fig=FALSE>>=
-gofstat(list(ftoxo.P,ftoxo.nb))
+gofstat(list(ftoxo.P,ftoxo.nb),fitnames = c("Poisson","negative binomial"))
 @
@@
 especially concerning goodness-of-fit evaluation and fitting methods. We will also enlarge the choice of 
-fitting methods for non-censored data, by proposing new goodness-of-fit distances (e.g. distances 
+fitting methods for non-censored data, by proposing new goodness-of-fit distances (e.g., distances 
 based on quantiles) for maximum goodness-of-fit estimation 
-and new types of moments (e.g. limited expected values) for moment matching estimation. At last, we will
+and new types of moments (e.g., limited expected values) for moment matching estimation. At last, we will
 consider the case of multivariate distribution fitting.

### 2013-05-15 JSS/fitdistrplus/paper2JSS.Rnw
take into account pre-screening.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@250 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 @
-\begin{footnotesize}
+\begin{small}
 <<fitATV.lnorm.quantileb, echo=FALSE>>=
@@
 @
-\end{footnotesize}
+\end{small}
 
@@
 <<fittoxocara.poisnbinom.gof, echo=TRUE, fig=FALSE>>=
-gofstat(list(ftoxo.P,ftoxo.nb),fitnames = c("Poisson","negative binomial"))
+gofstat(list(ftoxo.P,ftoxo.nb),
+        fitnames = c("Poisson","negative binomial"))
 @

### 2013-05-15 JSS/fitdistrplus/paper2JSS.R

@@ -1,0 +1,188 @@
+#1
+options(digits=4, prompt="R>", 
+        SweaveHooks=list(fig=function() par(mar=c(5.1, 4.1, 1.1, 2.1))))
+set.seed(1234)
+
+#2
+library("fitdistrplus")
+data("groundbeef")
+str(groundbeef)
+
+#3
+plotdist(groundbeef$serving)
+
+#4
+descdist(groundbeef$serving, boot=1000)
+
+#5
+fw <- fitdist(groundbeef$serving, "weibull")
+summary(fw)
+
+#6
+fg <- fitdist(groundbeef$serving,"gamma")
+fln <- fitdist(groundbeef$serving,"lnorm")
+par(mfrow=c(2, 2))
+denscomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+qqcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+cdfcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+ppcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+
+#7
+data("endosulfan")
+str(endosulfan)
+
+#8
+ATV <-endosulfan$ATV
+fendo.ln <- fitdist(ATV, "lnorm")
+library("actuar")
+fendo.ll <- fitdist(ATV, "llogis", start=list(shape=1, scale=500))
+fendo.P <- fitdist(ATV, "pareto", start=list(shape=1, scale=500))
+fendo.B <- fitdist(ATV, "burr", start=list(shape1=0.3, shape2=1, rate=1))
+cdfcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), xlogscale = TRUE,
+        ylogscale = TRUE, legendtext = 
+          c("lognormal", "loglogistic", "Pareto", "Burr"))
+
+#9
+quantile(fendo.B, probs = 0.05)
+quantile(ATV, probs = 0.05)
+
+#10
+gofstat(list(fendo.ln, fendo.ll, fendo.P, fendo.B),
+        fitnames = c("lnorm","llogis","Pareto","Burr"))
+
+#11
+bendo.B <- bootdist(fendo.B, niter=1001)
+summary(bendo.B)
+plot(bendo.B)
+
+#12
+quantile(bendo.B, probs = 0.05)
+
+#13
[…]

### 2013-05-17 JSS/fitdistrplus/paper2JSS.Rnw
minor corrections
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@251 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 fdanish.ln.MME, fdanish.P.MME),
-fitnames = c("lnorm.mle","Pareto.mle","lnorm.mle","Pareto.mme"))
+fitnames = c("lnorm.mle","Pareto.mle","lnorm.mme","Pareto.mme"))
 @

### 2013-05-17 JSS/fitdistrplus/paper2JSS.R

@@ -38,6 +38,7 @@ library("actuar")
 fendo.ll <- fitdist(ATV, "llogis", start=list(shape=1, scale=500))
 fendo.P <- fitdist(ATV, "pareto", start=list(shape=1, scale=500))
 fendo.B <- fitdist(ATV, "burr", start=list(shape1=0.3, shape2=1, rate=1))
+par(mfrow=c(1, 1))
 cdfcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), xlogscale = TRUE,
         ylogscale = TRUE, legendtext = 
           c("lognormal", "loglogistic", "Pareto", "Burr"))
@@ -92,7 +93,7 @@ memp <- function(x, order) sum(x^order)/length(x)
 fdanish.P.MME <- fitdist(danishuni$Loss, "pareto", method="mme", order=1:2, 
                          memp="memp", start=c(shape=10, scale=10), lower=c(2+1e-6,2+1e-6), 
                          upper=c(Inf,Inf))
-par(mfrow=c(1, 2))
+par(mfrow=c(1,2))
 cdfcomp(list(fdanish.ln.MLE, fdanish.ln.MME), 
         legend=c("lognormal MLE", "lognormal MME"), main="Fitting a lognormal distribution",
         xlogscale=TRUE, datapch=20)
@@ -101,12 +102,13 @@ cdfcomp(list(fdanish.P.MLE, fdanish.P.MME),
         xlogscale=TRUE, datapch=20)
 
 #17
+par(mfrow=c(1, 1))
 cdfcomp(list(fdanish.P.MLE, fdanish.P.MME), 
         legend=c("Pareto MLE", "Pareto MME"), 
         main="Fitting a Pareto distribution", xlogscale=TRUE, datapch=".")
 gofstat(list(fdanish.ln.MLE, fdanish.P.MLE, 
              fdanish.ln.MME, fdanish.P.MME),
-        fitnames = c("lnorm.mle","Pareto.mle","lnorm.mle","Pareto.mme"))
+        fitnames = c("lnorm.mle","Pareto.mle","lnorm.mme","Pareto.mme"))
 
 #18
 fdanish.ln.QME1 <- fitdist(danishuni$Loss, "lnorm", method="qme", 
@@ -174,6 +176,7 @@ str(toxocara)
 #28
 (ftoxo.P <- fitdist(toxocara$number, "pois"))
[…]

### 2013-06-21 share/rencontresR2013/fitdistrplus280613.Rnw
presentation for French R days in Lyon
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@252 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{beamer}

\mode<presentation>
{
  %\usetheme{Boadilla}
 % \setbeamercovered{transparent}
  %\setbeamertemplate{navigation symbols}{}

    %---- themes
  	\usetheme{Antibes}
  	%\usetheme{Singapore}
  	%\usetheme{Warsaw}

	%nice theme for math fonts
	\usefonttheme{professionalfonts}

	%color theme
	\usecolortheme{lily}

	  %for math characters
  %	\usefonttheme[onlylarge]{structuresmallcapsserif}
	%\usefonttheme[onlysmall]{structurebold}

	%remove navigation symbols
	\setbeamertemplate{navigation symbols}{}

	%dark blue for normal text
%	\definecolor{MyDarkBlue}{rgb}{0.0235,0.1569,0.4745}
%	\setbeamercolor{normal text}{fg = MyDarkBlue}

	%white for background text color
	\setbeamercolor{structure}{bg = white}

	%---- headline
	% black and white for headline and footline
	\setbeamercolor{section in head/foot}{bg = white, fg=black}
	\setbeamercolor{subsection in head/foot}{bg = white, fg=black}
	\setbeamercolor{title in head/foot}{bg = white, fg=black}

	%What do you want in your header? by default, there is the current subsection

	%if you want to insert a complete navigation	
	%	\setbeamertemplate{headline}
	%	{%--
	%		\begin{beamercolorbox}{section in head/foot}
	%			\vskip2pt\insertnavigation{\paperwidth}\vskip2pt
	%		\end{beamercolorbox}
	%	}%--

	%---- footline	
	%add frame number and logo
	\setbeamertemplate{footline}
	{%--
	   %you can adjust the left and right white space with leftskip and rightskip
	\begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,
	    leftskip=.15cm,rightskip=.15cm plus1fil]{title in head/foot}
	     \insertframenumber/\inserttotalframenumber
	     \hspace{0.5cm}	%you can use different horizontal spacing!
	     \insertshortdate
	     \hspace{0.5cm}
	     Rencontres R 2013
	     \hfill
	\end{beamercolorbox}
  	}%--

	%% No footline on title slide
	\defbeamertemplate{footline}{title slide}{%
	\relax
	}

	%---- misc
  %\setbeamercovered{transparent}
  % or whatever (possibly just delete it)

}

\usepackage[francais]{babel}
\usepackage[latin1]{inputenc}
\usepackage{beamerthemesplit}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{pdfcolmk}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{url}

% packages de Christophe en plus ?
\usepackage{wrapfig}
\usepackage{amsmath,amsfonts,amssymb,latexsym,qsymbols}
\usepackage{multirow}

\title[fitdistrplus : nouvelles fonctionnalités]% 
{Nouvelles fonctionnalités du package \texttt{fitdistrplus}}

\author[M.L. Delignette-Muller et C. Dutang]
{\textbf{Marie Laure Delignette-Muller}\\Université de Lyon \\
UMR CNRS 5558, Laboratoire de Biométrie et Biologie Evolutive\\
VetAgro Sup, Campus vétérinaire de Lyon\\et\\
           \textbf{Christophe Dutang} \\Université de Strasbourg \\
           UMR CNRS 7501, Institut de Recherche Mathématique Avancée}

\date[26_27 juin 2013]{Deuxièmes rencontres R, Lyon, 27-28 juin 2013}

\begin{document}
\SweaveOpts{concordance=TRUE}
\setkeys{Gin}{width=0.5\textwidth}
% une valeur inférieure au 0.8 par défaut réduit la taille des figures

\begin{frame}[fragile]
\titlepage
\begin{center}
\includegraphics[width=1cm]{Rlogo_2.jpg}
\end{center}
\end{frame}

  %DIA 2
\begin{frame}[fragile]
\frametitle{Présentation du package \texttt{fitdistrplus}}
\begin{center}
\url{http://cran.r-project.org/package=fitdistrplus}\\
\url{http://riskassessment.r-forge.r-project.org/}
\end{center}
Package pour \textbf{l'ajustement de distributions
paramétriques à des données univariées},\\ proposant diverses fonctions pour :
  \begin{itemize}
\item le choix de distributions candidates pour décrire les données,
\item l'ajustement de chacune des distributions candidates aux données,
\item la comparaison des ajustements en vue de choisir la distribution la plus adaptée,
\item le calcul, par bootstrap, de l'incertitude sur les paramètres estimés de la distribution choisie.
\end{itemize}
\end{frame}

% DIA 3
\begin{frame}
\frametitle{Quelques spécificités du package}
\begin{itemize}
\item \textbf{Différentes méthodes d'estimation des paramètres} sont proposées :
  \begin{itemize}
  \item méthode du \textcolor{blue}{maximum de vraisemblance (MLE)}, 
  \item méthode des \textcolor{blue}{moments (MME)}, 
  \item méthode des \textcolor{blue}{quantiles (QME)}, 
  \item méthode de \textcolor{blue}{minimisation d'une statistique d'ajustement (MGE)} (8 distances proposées).
  \end{itemize}
\item Prise en compte 
    \begin{itemize}
  \item des données discrètes 
    \item et des \textbf{données censurées} quel que
soit le type de censures \\(à droite, à gauche ou par intervalle).
  \end{itemize}
\item Possibilité de modifier l'algorithme d'optimisation \\(par défaut \texttt{optim}).
\end{itemize}
\end{frame}

% DIA 4
\begin{frame}
\frametitle{Utilisation actuelle du package}
\textbf{Package développé initialement pour l'appréciation quantitative du risque}\\ 
%(projet Rforge
%\url{http://riskassessment.r-forge.r-project.org/})\\
notamment pour permettre
la caractérisation hiérarchique de la variabilité et de l'incertitude dans le cadre de simulations de 
Monte carlo à deux dimensions (package \texttt{mc2d}).\\
\textbf{Package généraliste simple à utiliser}, \\utilisé dans des domaines variés :
\begin{itemize}
\item risque alimentaire, 
\item épidémiologie, 
\item biologie moléculaire, 
%\item génomique, 
\item bioinformatique, 
\item mathématiques financières et actuarielles, \dots
\end{itemize}
\textcolor{blue}{$\rightarrow$ les retours des utilisateurs nous ont incité à développer de 
\textbf{nouvelles fonctionnalités}}.
\end{frame}

% DIA 5
\begin{frame}[fragile]
\frametitle{Comparaison visuelle de plusieurs ajustements (\texttt{cdfcomp})}
\begin{footnotesize}
<<echo=FALSE,results=hide>>=
rm(list=ls(all=TRUE))
set.seed(1234) 
library(fitdistrplus)
options(digits=3)
@
\setkeys{Gin}{width=0.9\textwidth}
%<<cdfgroundbeef, echo=TRUE, fig=TRUE, width=7, height=4, eps=FALSE>>=
%data(groundbeef)
%fw <- fitdist(groundbeef$serving, "weibull")
%fg <- fitdist(groundbeef$serving,"gamma")
%fln <- fitdist(groundbeef$serving,"lnorm")
%cdfcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
%@
<<cdfgroundbeefecho, echo=TRUE, fig=FALSE>>=
data(groundbeef)
fw <- fitdist(groundbeef$serving, "weibull")
fg <- fitdist(groundbeef$serving,"gamma")
fln <- fitdist(groundbeef$serving,"lnorm")
cdfcomp(list(fw,fln,fg), 
        legendtext=c("Weibull", "lognormal", "gamma"))
@
<<cdfgroundbeeffig, echo=FALSE, fig=TRUE, width=7, height=4, eps=FALSE>>=
cdfcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"),
        fitlty=1,lwd=2)
@
\end{footnotesize}
\end{frame}

% DIA 6
\begin{frame}[fragile]
\frametitle{Comparaison visuelle de plusieurs ajustements : \\représentation en densité de probabilité (\texttt{denscomp})}
\begin{footnotesize}
\setkeys{Gin}{width=0.9\textwidth}
<<densgroundbeefecho, echo=TRUE, fig=FALSE>>=
denscomp(list(fw,fln,fg), 
         legendtext=c("Weibull", "lognormal", "gamma"))
@
<<densgroundbeeffig, echo=FALSE, fig=TRUE, width=7, height=4.5, eps=FALSE>>=
denscomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"),
        fitlty=1,lwd=2)
@
\end{footnotesize}
\end{frame}

% DIA 7
%\begin{frame}[fragile]
%\frametitle{Comparaison visuelle de plusieurs ajustements : P-P plot (\texttt{ppcomp})
%ou en Q-Q plot (\texttt{qqcomp})}
%\begin{footnotesize}
%\setkeys{Gin}{width=0.9\textwidth}
%<<ppqqgroundbeefecho, echo=TRUE, fig=FALSE>>=
%ppcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
%qqcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
%@
%<<ppqqgroundbeeffig, echo=FALSE, fig=TRUE, width=7, height=4, eps=FALSE>>=
%par(mfrow=c(1,2))
%par(mar=c(4,4,1,1))
%ppcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"),
%       fitpch=16)
%qqcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"),
%       fitpch=16)
%@
%\end{footnotesize}
%\end{frame}

% DIA 7
\begin{frame}[fragile]
\frametitle{Comparaison visuelle de plusieurs ajustements :\\ P-P plot (\texttt{ppcomp})}
\begin{footnotesize}
\setkeys{Gin}{width=0.9\textwidth}
<<ppgroundbeefecho, echo=TRUE, fig=FALSE>>=
ppcomp(list(fw,fln,fg), 
       legendtext=c("Weibull", "lognormal", "gamma"))
@
<<ppgroundbeeffig, echo=FALSE, fig=TRUE, width=7, height=4.5, eps=FALSE>>=
ppcomp(list(fw,fln,fg), 
       legendtext=c("Weibull", "lognormal", "gamma"),
       fitpch=16)
@
\end{footnotesize}
\end{frame}

% DIA 8
\begin{frame}[fragile]
\frametitle{Comparaison visuelle de plusieurs ajustements : \\
Q-Q plot (\texttt{qqcomp})}
\begin{footnotesize}
\setkeys{Gin}{width=0.9\textwidth}
<<qqgroundbeefecho, echo=TRUE, fig=FALSE>>=
qqcomp(list(fw,fln,fg), 
       legendtext=c("Weibull", "lognormal", "gamma"))
@
<<qqgroundbeeffig, echo=FALSE, fig=TRUE, width=7, height=4.5, eps=FALSE>>=
qqcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"),
       fitpch=16)
@
\end{footnotesize}
\end{frame}

% DIA 9
\begin{frame}[fragile]
\frametitle{Comparaison numérique de plusieurs ajustements}
\begin{itemize}
\item \textbf{Statistiques d'ajustement}, \\
Kolmogorov-Smirnov, Cramér-von Mises et Anderson-Darling,\\
\item et \textbf{critères d'information} , AIC et BIC.
\end{itemize}
\begin{footnotesize}
\setkeys{Gin}{width=0.9\textwidth}
<<ppqqgroundbeefecho, echo=TRUE, fig=FALSE>>=
gofstat(list(fw,fln,fg),
        fitnames = c("Weibull", "lognormal", "gamma"))
@
\end{footnotesize}
\end{frame}

% DIA 10 
\begin{frame}[fragile]
\frametitle{Calcul de quantiles à partir d'une loi ajustée : \\ exemple sur données censurées en écotoxicologie}
\begin{footnotesize}
<<salinityecho, echo=TRUE, fig=FALSE>>=
data(salinity)
f <- fitdistcens(salinity, "lnorm")
quantile(f,probs = c(0.05, 0.20))
@
\begin{center}
\setkeys{Gin}{width=0.7\textwidth}
<<salinityfig, echo=FALSE, fig=TRUE, width=6, height=3.5, eps=FALSE>>=
par(mar=c(4,4,1,1))
cdfcompcens(f,lwd=2,addlegend=FALSE)
probs <- c(0.05,0.2)
q <- quantile(f,probs = probs)$quantiles
abline(h=probs[1],col="red",lty=2,lwd=2)
abline(h=probs[2],col="red",lty=3,lwd=2)
abline(v=q[1],col="red",lty=2,lwd=2)
abline(v=q[2],col="red",lty=3,lwd=2)
points(q,probs,col="red",pch=16)
text(2,0.05,"5%",pos=3,col="red")
text(2,0.2,"20%",pos=3,col="red")
@
\end{center}
\end{footnotesize}
\end{frame}

% DIA 11 
\begin{frame}[fragile]
\frametitle{Bootstrap}
\begin{footnotesize}
<<salinitybootecho, echo=TRUE, fig=FALSE>>=
b <- bootdistcens(f,niter=1001) 
summary(b)
plot(b)
@
\begin{center}
\setkeys{Gin}{width=0.45\textwidth}
<<salinitybootfig, echo=FALSE, fig=TRUE, width=6, height=5, eps=FALSE>>=
par(mar=c(4,4,1,1))
plot(b,pch=16)
@
\end{center}\end{footnotesize}
\end{frame}

% DIA 12
\begin{frame}[fragile]
\frametitle{Intervalles de confiance bootstrap sur les quantiles}
\begin{footnotesize}
<<salinityICecho, echo=TRUE, fig=FALSE>>=
cdfcompcens(f)
q <- quantile(b, probs = seq(0,1,length=101))
points(q$quantCI[1,],q$probs,type="l",lty=2,col="red")
points(q$quantCI[2,],q$probs,type="l",lty=2,col="red")
@
\begin{center}
\setkeys{Gin}{width=0.9\textwidth}
<<salinityICfig, echo=FALSE, fig=TRUE, width=7, height=4, eps=FALSE>>=
par(mar=c(4,4,1,1))
cdfcompcens(f,lwd=2,,addlegend=FALSE)
probs = seq(0,1,length=101)
q <- quantile(b, probs = probs)
points(q$quantCI[1,],q$probs,type="l",lty=2,col="red",lwd=2)
points(q$quantCI[2,],q$probs,type="l",lty=2,col="red",lwd=2)
arrows(q$quantCI[1,70],probs[70],q$quantCI[2,70],probs[70],
       col="red",lwd=2,code=3,length=0.1)
text((q$quantCI[1,70]+q$quantCI[2,70])/2,probs[70],"IC 95%",col="red",pos=3)
@
\end{center}\end{footnotesize}
\end{frame}

% DIA 13 
\begin{frame}[fragile]
\frametitle{Perspectives}
Nous envisageons 
\begin{itemize}
\item d'\textbf{étendre aux données censurées certaines méthodes} disponibles
actuellement sur les données non censurées,
\item d'\textbf{élargir le choix de distances} utilisables pour l'ajustement par minimisation de ces distances et pour l'évaluation de la qualité de l'ajustement \\(distances basées sur les quantiles),
\item d'\textbf{élargir le choix des moments} pour l'ajustement par la méthode des moments \\(``L-moment'' et ``TL-moment''),
\item aborder l'\textbf{ajustement de distributions multivariées}.
\end{itemize}
\textcolor{blue}{N'hésitez pas à nous envoyer vos retours ou idées en vue de l'amélioration de ce package.}
\end{frame}

\end{document}

### 2013-06-21 share/rencontresR2013/Sweave/fitdistrplus280613.Rnw
update.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@254 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2013-06-21 share/rencontresR2013/fitdistrplus280613.Rnw
@@
-\documentclass{beamer}
+\documentclass[9pt, hide notes]{beamer}
 
@@
 {
-  %\usetheme{Boadilla}
- % \setbeamercovered{transparent}
-  %\setbeamertemplate{navigation symbols}{}
 
@@
   	\usetheme{Antibes}
+    %\usetheme{Boadilla}
   	%\usetheme{Singapore}
@@
 	%if you want to insert a complete navigation	
-	%	\setbeamertemplate{headline}
-	%	{%--
-	%		\begin{beamercolorbox}{section in head/foot}
-	%			\vskip2pt\insertnavigation{\paperwidth}\vskip2pt
-	%		\end{beamercolorbox}
-	%	}%--
+		\setbeamertemplate{headline}
+		{%--
+			\begin{beamercolorbox}{section in head/foot}
+				\vskip2pt\insertnavigation{\paperwidth}\vskip2pt
+			\end{beamercolorbox}
+		}%--
 
@@
 	     \hfill
+       M.L. Delignette-Muller et C. Dutang
 	\end{beamercolorbox}
@@
 \usepackage[latin1]{inputenc}
-\usepackage{beamerthemesplit}
+%\usepackage{beamerthemesplit}
 \usepackage{graphicx}
@@
 
-\title[fitdistrplus : nouvelles fonctionnalités]% 
-{Nouvelles fonctionnalités du package \texttt{fitdistrplus}}
+\title[fitdistrplus : nouvelles fonctionnalit\'es]% 
+{Nouvelles fonctionnalit\'es du package \texttt{fitdistrplus}}
 
 \author[M.L. Delignette-Muller et C. Dutang]
-{\textbf{Marie Laure Delignette-Muller}\\Université de Lyon \\
-UMR CNRS 5558, Laboratoire de Biométrie et Biologie Evolutive\\
-VetAgro Sup, Campus vétérinaire de Lyon\\et\\
-           \textbf{Christophe Dutang} \\Université de Strasbourg \\
-           UMR CNRS 7501, Institut de Recherche Mathématique Avancée}
+{\textbf{Marie Laure Delignette-Muller}\\ Universit\'e de Lyon \\
+UMR CNRS 5558, Laboratoire de Biom\'etrie et Biologie Evolutive\\
+VetAgro Sup, Campus v\'et\'erinaire de Lyon\\et\\
+           \textbf{Christophe Dutang} \\ Universit\'e de Strasbourg \\
+           UMR CNRS 7501, Institut de Recherche Math\'ematique Avanc\'ee}
 
-\date[26_27 juin 2013]{Deuxièmes rencontres R, Lyon, 27-28 juin 2013}
+\date[26-27 juin 2013]{Deuxi\`emes rencontres R, Lyon, 27-28 juin 2013}
 
@@
 \setkeys{Gin}{width=0.5\textwidth}
-% une valeur inférieure au 0.8 par défaut réduit la taille des figures
+% une valeur inf?rieure au 0.8 par d?faut r?duit la taille des figures
 
@@
 \begin{frame}[fragile]
-\frametitle{Présentation du package \texttt{fitdistrplus}}
+\frametitle{Pr\'esentation du package \texttt{fitdistrplus}}
 \begin{center}
@@
 Package pour \textbf{l'ajustement de distributions
-paramétriques à des données univariées},\\ proposant diverses fonctions pour :
+param\'etriques \`a des donn\'ees univari\'ees},\\ proposant diverses fonctions pour :
   \begin{itemize}
-\item le choix de distributions candidates pour décrire les données,
-\item l'ajustement de chacune des distributions candidates aux données,
-\item la comparaison des ajustements en vue de choisir la distribution la plus adaptée,
-\item le calcul, par bootstrap, de l'incertitude sur les paramètres estimés de la distribution choisie.
+\item le choix de distributions candidates pour d\'ecrire les donn\'ees,
+\item l'ajustement de chacune des distributions candidates aux donn\'ees,
+\item la comparaison des ajustements en vue de choisir la distribution la plus adapt\'ee,
+\item le calcul, par bootstrap, de l'incertitude sur les param\`etres estim\'es de la distribution choisie.
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Quelques spécificités du package}
+\frametitle{Quelques sp\'ecificit\'es du package}
 \begin{itemize}
-\item \textbf{Différentes méthodes d'estimation des paramètres} sont proposées :
+\item \textbf{Diff\'erentes m\'ethodes d'estimation des param\`etres} sont propos\'ees :
   \begin{itemize}
-  \item méthode du \textcolor{blue}{maximum de vraisemblance (MLE)}, 
-  \item méthode des \textcolor{blue}{moments (MME)}, 
-  \item méthode des \textcolor{blue}{quantiles (QME)}, 
-  \item méthode de \textcolor{blue}{minimisation d'une statistique d'ajustement (MGE)} (8 distances proposées).
+  \item m\'ethode du \textcolor{blue}{maximum de vraisemblance (MLE)}, 
+  \item m\'ethode des \textcolor{blue}{moments (MME)}, 
+  \item m\'ethode des \textcolor{blue}{quantiles (QME)}, 
+  \item m\'ethode de \textcolor{blue}{minimisation d'une statistique d'ajustement (MGE)} (8 distances propos\'ees).
   \end{itemize}
@@
     \begin{itemize}
-  \item des données discrètes 
-    \item et des \textbf{données censurées} quel que
-soit le type de censures \\(à droite, à gauche ou par intervalle).
+  \item des donn\'ees discr\`etes 
+    \item et des \textbf{donn\'ees censur\'ees} quel que
+soit le type de censures \\(\`a droite, \`a gauche ou par intervalle).
   \end{itemize}
-\item Possibilité de modifier l'algorithme d'optimisation \\(par défaut \texttt{optim}).
+\item Possibilit\'e de modifier l'algorithme d'optimisation \\(par d\'efaut \texttt{optim}).
 \end{itemize}
@@
 \frametitle{Utilisation actuelle du package}
-\textbf{Package développé initialement pour l'appréciation quantitative du risque}\\ 
+\textbf{Package d\'evelopp\'e initialement pour l'appr\'eciation quantitative du risque}\\ 
 %(projet Rforge
@@
 notamment pour permettre
-la caractérisation hiérarchique de la variabilité et de l'incertitude dans le cadre de simulations de 
-Monte carlo à deux dimensions (package \texttt{mc2d}).\\
-\textbf{Package généraliste simple à utiliser}, \\utilisé dans des domaines variés :
+la caract\'erisation hi\'erarchique de la variabilit\'e et de l'incertitude dans le cadre de simulations de 
+Monte carlo \`a deux dimensions (package \texttt{mc2d}).\\
+\textbf{Package g\'en\'eraliste simple \`a utiliser}, \\utilis\'e dans des domaines vari\'es :
 \begin{itemize}
 \item risque alimentaire, 
-\item épidémiologie, 
-\item biologie moléculaire, 
-%\item génomique, 
+\item \'epid\'emiologie, 
+\item biologie mol\'eculaire, 
+%\item g\'enomique, 
 \item bioinformatique, 
-\item mathématiques financières et actuarielles, \dots
+\item math\'ematiques financi\`eres et actuarielles, \dots
 \end{itemize}
-\textcolor{blue}{$\rightarrow$ les retours des utilisateurs nous ont incité à développer de 
-\textbf{nouvelles fonctionnalités}}.
+\textcolor{blue}{$\rightarrow$ les retours des utilisateurs nous ont incit\'e \`a d\'evelopper de 
+\textbf{nouvelles fonctionnalit\'es}}.
 \end{frame}
@@
 \begin{frame}[fragile]
-\frametitle{Comparaison visuelle de plusieurs ajustements : \\représentation en densité de probabilité (\texttt{denscomp})}
+\frametitle{Comparaison visuelle de plusieurs ajustements : \\repr\'esentation en densit\'e de probabilit\'e (\texttt{denscomp})}
 \begin{footnotesize}
@@
 \begin{frame}[fragile]
-\frametitle{Comparaison numérique de plusieurs ajustements}
+\frametitle{Comparaison num\'erique de plusieurs ajustements}
 \begin{itemize}
 \item \textbf{Statistiques d'ajustement}, \\
-Kolmogorov-Smirnov, Cramér-von Mises et Anderson-Darling,\\
-\item et \textbf{critères d'information} , AIC et BIC.
+Kolmogorov-Smirnov, Cram\'er-von Mises et Anderson-Darling,\\
+\item et \textbf{crit\`eres d'information} , AIC et BIC.
 \end{itemize}
@@
 \begin{frame}[fragile]
-\frametitle{Calcul de quantiles à partir d'une loi ajustée : \\ exemple sur données censurées en écotoxicologie}
+\frametitle{Calcul de quantiles \`a partir d'une loi ajust\'ee : \\ exemple sur donn\'ees censur\'ees en \'ecotoxicologie}
 \begin{footnotesize}
@@
 \begin{itemize}
-\item d'\textbf{étendre aux données censurées certaines méthodes} disponibles
-actuellement sur les données non censurées,
-\item d'\textbf{élargir le choix de distances} utilisables pour l'ajustement par minimisation de ces distances et pour l'évaluation de la qualité de l'ajustement \\(distances basées sur les quantiles),
-\item d'\textbf{élargir le choix des moments} pour l'ajustement par la méthode des moments \\(``L-moment'' et ``TL-moment''),
-\item aborder l'\textbf{ajustement de distributions multivariées}.
+\item d'\textbf{\'etendre aux donn\'ees censur\'ees certaines m\'ethodes} disponibles
+actuellement sur les donn\'ees non censur\'ees,
+\item d'\textbf{\'elargir le choix de distances} utilisables pour l'ajustement par minimisation de ces distances et pour l'\'evaluation de la qualit\'e de l'ajustement \\(distances bas\'ees sur les quantiles),
+\item d'\textbf{\'elargir le choix des moments} pour l'ajustement par la m\'ethode des moments \\(``L-moment'' et ``TL-moment''),
+\item aborder l'\textbf{ajustement de distributions multivari\'ees}.
 \end{itemize}
-\textcolor{blue}{N'hésitez pas à nous envoyer vos retours ou idées en vue de l'amélioration de ce package.}
+\textcolor{blue}{N'h\'esitez pas ? nous envoyer vos retours ou id\'ees en vue de l'am\'elioration de ce package.}
 \end{frame}

### 2013-10-08 pkg/mc2d/R/pert.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@257 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,12 +1,12 @@
 #<<BEGIN>>
 dpert <- function(x,min=-1,mode=0,max=1,shape=4,log=FALSE)
-#TITLE The Pert Distribution
+#TITLE The (Modified) PERT Distribution
 #NAME pert
 #KEYWORDS distribution
 #DESCRIPTION
 #Density, distribution function, quantile function and random generation
-#for the pert distribution with minimum equal to \samp{min}, mode equal to \samp{mode}
-#and maximum equal to \samp{max}.
+#for the PERT (\emph{aka} Beta PERT) distribution with minimum equals to \samp{min}, mode equals to \samp{mode}
+#and maximum equals to \samp{max}. 
 #INPUTS
 #{x,q}<<Vector of quantiles.>>
 #{p}<<Vector of probabilities.>>
@@ -15,22 +15,29 @@ dpert <- function(x,min=-1,mode=0,max=1,shape=4,log=FALSE)
 #{min}<<Vector of minima.>>
 #{mode}<<Vector of modes.>>
 #{max}<<Vector of maxima.>>
-#{shape}<<Vector of scaling parameters.>>
+#{shape}<<Vector of scaling parameters. Default value: 4.>>
 #{log, log.p}<<Logical; if \samp{TRUE}, probabilities \samp{p} are given as \samp{log(p)}.>>
 #{lower.tail}<<Logical; if \samp{TRUE} (default), probabilities are \samp{P[X <= x]}, otherwise, \samp{P[X > x]}.>>
 #DETAILS
-#The Pert distribution is a special case of the Beta distribution specified by the following parameters.
-#Given:
-#\deqn{\mu=\frac{min+max+shape\times mode}{shape+2}}{mu = (min + max + shape * mode)/(shape + 2)}
-#the values of \eqn{\alpha_{1}}{shape1} and \eqn{\alpha_{2}}{shape2} are
-#\deqn{\alpha_{1}=\frac{(\mu-min)(2\times mode-min-max)}{(mode-\mu)(max-min)}}{shape1=(mu - min)*(2 mode-min-max)/((mode-mu)*(max - min)}
+#The PERT distribution is a beta distribution extended to the domain \samp{[min, max]} with mean 
[…]

### 2013-10-08 pkg/mc2d/R/pert.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@258 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -144,6 +144,11 @@ rpert <- function(n,min=-1,mode=0,max=1,shape=4)
 { if(n < 0) stop("integer(n) can not be negative in rpert")
   #other pathologic cases for n treated in rbeta
   
+  min <- as.vector(min)
+  mode <- as.vector(mode)
+  max <- as.vector(max)
+  shape <- as.vector(shape)
+  
   a1 <- 1+shape*(mode-min)/(max-min)  
   a2 <- 1+shape*(max-mode)/(max-min)
   

### 2013-10-09 pkg/mc2d/R/pert.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@259 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -36,7 +36,7 @@ dpert <- function(x,min=-1,mode=0,max=1,shape=4,log=FALSE)
 #of the min, max and mode of a random variable in a smooth parametric distribution. 
 #REFERENCE
 #Vose D. Risk Analysis - A Quantitative Guide (John Wiley and Sons, 2000).  
-#The equations used in these functions are from 
+#The formulas used in these functions are from 
 #\href{http://www.vosesoftware.com/ModelRiskHelp/Distributions/Continuous_distributions/Modified_PERT_equations.htm}{Vose software}.
 #AUTHOR Regis Pouillot and Matthew Wiener
 #EXAMPLE

### 2014-02-05 JSS/fitdistrplus/paper2JSS.Rnw
new version
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@263 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 function using the \code{optim} function. 
+No distinction between parameter role (e.g. a scale %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+parameter) is made, as our talk focus 
+on parameter estimation from a general point-of-view. %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 In some cases, other estimation methods could be prefered, 
@@
 and support different types of data. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 The \pkg{distrMod} package \citep{distrModJSS} provides an object-oriented (S4)  
-implementation of probability models for non-censored data
-and includes distribution fitting procedures for a given minimization criterion.
-This package is based on the \proglang{R}-shipped function \code{mle} of the \pkg{stats4} package
-producing S4 objects of class \code{"mle"} for which generic methods are 
-implemented, e.g., \code{confint} and \code{logLik}.
+implementation of probability models and includes distribution fitting procedures  
+for a given minimization criterion. 
+This criterion is a user-supplied function 
+which is sufficiently flexible to handle censored data, yet not in a trivial way,
+see Example M4 of the \pkg{distrMod} vignette.
+The fitting functions \code{MLEstimator} and \code{MDEstimator} return an S4 class 
+for which a coercion method to class mle is provided so that the respective 
+functionalities (e.g. \code{confint} and \code{logLik}) from package \pkg{stats4} 
+are available, too.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 In \pkg{fitdistrplus}, we chose to use the standard S3 class system for its understanding by most \proglang{R} users.
@@
 @
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In addition to the ecotoxicology context, the \code{quantile} generic function is also attractive in
+the actuarial--financial context. In fact, the value-at-risk $VAR_\alpha$ is defined as the $1-\alpha$-quantile
+of the loss distribution and can be computed with \code{quantile} on a \code{"fitdist"} object.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 a special focus on the right-tail of the distributions.
-In this numerical experiment,  the lognormal distribution is first considered and then the Pareto type II 
-distribution \citep{Klugmanetal09}.
+In this numerical experiment, we choose classic actuarial 
+distributions for loss modelling: the lognormal distribution and the Pareto type II 
+distribution, see e.g. \citep{Klugmanetal09}.
 
@@
 Furthermore, for these two distributions, the MME method better fits the 
-right-tail of the distribution, which seems logical since empirical moments 
+right-tail of the distribution from a visual point of view. 
+This seems logical since empirical moments 
 are influenced by large observed values.
@@
 of the distribution, whereas the first QME fit best models the body of the distribution.
-The quantile matching estimation is of particular interest when we need a focus 
+The quantile matching estimation is of particular interest when we need to focus 
 around particular quantiles, e.g., $p=99.5\%$ in the Solvency II insurance context or
@@
 (see \code{?optim} for details about the different algorithms available).
+
+Even if no error is raised when computing the optimization, changing the algorithm 
+is of particular interest to enforce bounds on some parameters. For instance, a 
+volatility parameter $\sigma$ is strictly positive $\sigma>0$ and a probability 
+ parameter $p$ lies in $p\in [0,1]$. This is possible by using arguments \code{lower} 
+ and/or \code{upper}, for which their use automatically forces 
+ \code{optim.method="L-BFGS-B"}.
 
@@
 or operations research. Indeed, this package is already used by a lot of practionners and academics 
-for simple MLE fits \citep{distrModJSS,jaloustreetal11,saketal11,kochetal12,marquetouxetal12,nordan12,
-scholletal12,suuronenetal12,varoetal12,mandletal13}, for MLE fits and goodness-of-fit statistics 
-\citep{garcia12,tarnczi11,bagariaetal12,benavidesetal12,breitbach12,poduje12,Pouillot10}, 
-for MLE fits 
-and bootstrap \citep{croucheretal12,meheustetal12,orellanoetal12,telloetal12,hoelzeretal12}, 
-for MLE fits, 
-bootstrap and goodness-of-fit statistics \citep{rebuge12}, 
-for MME fit \citep{luangkesornetal12}, 
-for 
-censored MLE and bootstrap \citep{lehaetal11,poulliotetal12,jongenburgeretal12,commeauetal12,pouillotetal11}, 
+for simple MLE fits \citep{jaloustreetal11,saketal11,kochetal12,marquetouxetal12,nordan12,
+scholletal12,suuronenetal12,varoetal12,mandletal13,mala2013,mala2013publi,nadarajahbakar2013,simoetal13,
+westphalfitch13,wayland13,vianaetal13,wu13,staggeetal13,fiorellietal13,mukhopadhyay13,tikoleetal13,
+voigtetal14}, 
+for MLE fits and goodness-of-fit statistics 
+\citep{tarnczi11,garcia12,bagariaetal12,benavidesetal12,breitbach12,poduje12,Pouillot10,vaninsky13}, 
+for MLE fits and bootstrap \citep{croucheretal12,meheustetal12,orellanoetal12,telloetal12,hoelzeretal12,
+prosseretal13,Zhang2013,lebaut13,Rigaux2014}, 
+for MLE fits, bootstrap and goodness-of-fit statistics \citep{rebuge12,larrasetal13}, 
+for MME fit \citep{luangkesornetal12,callauetal13,satoetal13}, 
+for censored MLE and bootstrap \citep{lehaetal11,pouillotetal11,poulliotetal12,jongenburgeretal12,
+commeauetal12,Kingetal2013,contrerasetal2013,duarte13,feng13}, 
 for graphic analysing in \citep{anandetal12,rosa12}, 
 for grouped-data fitting methods \citep{fusteinercostafreda12} 
-or more generally \citep{eling12,busschaertetal10,brooksetal11,aktassjostrand11,eikhermann12,bakos11}.
+or more generally \citep{busschaertetal10,brooksetal11,aktassjostrand11,bakos11,eling12,eikhermann12,
+sosaetal2013,srinivasanetal2013,meyeretal13,Guillier2013471,Daelmanetal13,gustafssonetal13,
+eiketal13,Wu2:2013,drakeetal2014}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The \pkg{fitdistrplus} package is also complementary with the \pkg{distrMod} package, \citep{distrModJSS}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 Many extensions of this package are planned in the future: 
@@
 his very valuable comments on the first version of this paper.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The authors gratefully acknowledges the two anonymous referees and the Editor for useful and constructive
+comments. The remaining errors, of course, should be attributed to the authors alone.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2014-02-05 share/logs.R
monthly download from cran-logs.rstudio.com.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@264 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,96 @@
+
+#----------------------------------------------------------------------------------------
+# download
+
+  # Here's an easy way to get all the URLs in R
+  # http://cran-logs.rstudio.com/
+  start <- as.Date('2012-10-01')
+  today <- as.Date('2014-02-05')
+
+  all_days <- seq(start, today, by = 'day')
+
+  year <- as.POSIXlt(all_days)$year + 1900
+  urls <- paste0('http://cran-logs.rstudio.com/', year, '/', all_days, '.csv.gz')
+
+setwd("~/Documents/recherche-enseignement/data/CRANlogs")
+# dir.create("CRANlogs")
+for(i in 1:length(urls))
+	download.file(urls[i], destfile=paste("./CRANlogs/", all_days[i], ".csv.gz", sep=""))
+
+
+#----------------------------------------------------------------------------------------
+# count hits by month
+
+setwd("~/Documents/recherche-enseignement/data/CRANlogs/")
+
+
+nbdaysbymonth <- c(31,28,31,30,31,30,31,31,30,31,30,31)
+names(nbdaysbymonth) <- c("janvier","février", "mars","avril","mai","juin","juillet",
+"août","septembre","octobre", "novembre","décembre")
+monthfile <- c("1-jan", "2-feb", "3-mar", "4-apr", "5-may", "6-jun", "7-jul", "8-aug", "9-sep", "10-oct", "11-nov", "12-dec")
+
+startdate <- as.Date('2012-10-01')
+startdate <- as.Date('2013-12-01')
+enddate <- as.Date("2014-01-30") 
+
+while(startdate < enddate)
+{
+	alldays <- seq(startdate, by = 'day', length= nbdaysbymonth[months(startdate)])
+	startdate <- tail(alldays, 1)+1
+	logfiles <- paste0(alldays, '.csv.gz')
+	
+	idata <- NULL
+	for(i in logfiles)
+	{
+	temp <- read.csv(gzfile(i))
+	idata <- c(idata, as.character(temp[,"package"]))
+	}
+	
+	
+	#----------
+	# merge
+	
+	logcsvfile <- paste(substr(alldays[1], 1, 4), "-", 
+	monthfile[names(nbdaysbymonth) == months(alldays[1])], ".csv", sep="")
+	print(logcsvfile)
+	
[…]

### 2014-02-06 JSS/fitdistrplus/paper2JSS.Rnw
new version of the paper.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@267 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 function using the \code{optim} function. 
-No distinction between parameter role (e.g. a scale %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-parameter) is made, as our talk focus 
-on parameter estimation from a general point-of-view. %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+No distinction between parameters with different roles (e.g. main parameter and nuisance parameter)
+is made, as our talk focuses 
+on parameter estimation from a general point-of-view. 
 In some cases, other estimation methods could be prefered, 
@@
 can be used to work with censored data, especially survival data \citep{survival,hiranoetal94,jordan05},
-but those packages generally focused on specific models, enabling the fit of  a 
+but those packages generally focus on specific models, enabling the fit of  a 
 restricted set of distributions. 
-A third objective was thus to provide \proglang{R} users a function 
+A third objective is thus to provide \proglang{R} users a function 
 to estimate univariate distribution parameters from right-, left- and interval-censored data.
@@
 
-This manuscript reviews the various features of version 1.0-1 of \pkg{fitdistrplus}. The package
+This manuscript reviews the various features of version 1.0-2 of \pkg{fitdistrplus}. The package
 is available from the Comprehensive \proglang{R} Archive Network at \url{http://cran.r-project.org/package=fitdistrplus}.
@@
 First of all, it is common to start with plots of the empirical distribution 
-function and the histogram, which can be obtained with the \code{plotdist} function of the \pkg{fitdistrplus} package.
-This function provides two plots (see Figure~\ref{plotdistcont}): the left-hand plot is the histogram (on a density scale) 
+function and the histogram (or density plot), which can be obtained with the \code{plotdist} function of the \pkg{fitdistrplus} package.
+This function provides two plots (see Figure~\ref{plotdistcont}): the left-hand plot is by default the histogram 
+on a density scale
+(or density plot of both, according to values of arguments \code{histo} and \code{demp})
 and the right-hand  plot the empirical cumulative distribution function (CDF).
@@
 <<figgroundbeef.echo, echo=TRUE, fig=FALSE, eval=FALSE>>=
-plotdist(groundbeef$serving)
+plotdist(groundbeef$serving, histo = TRUE, demp = TRUE)
 @
@@
 <<figgroundbeefplot, echo=FALSE, fig=TRUE, width=10, height=5, eps=FALSE>>=
-plotdist(groundbeef$serving)
+plotdist(groundbeef$serving, histo = TRUE, demp = TRUE)
 @
@@
 skewness-kurtosis plot. 
+Nevertheless, the user needs to know that skewness and kurtosis,  like all higher moments, have a very high variance.
+This is a problem which cannot be completely solved by the use of bootstrap.
+The skewness-kurtosis plot should then be regarded as indicative only.
 The properties of the random variable should be considered, notably its expected value and its range,
@@
 \citep{Hose04}. 
-<<datendosulfan, echo=TRUE, eval=FALSE>>=
-data("endosulfan")
-str(endosulfan)
-@
-<<datendosulfan2, echo=FALSE, eval=TRUE>>=
-data("endosulfan")
-str(endosulfan, width = 70, strict.width = "cut")
-@
-
+%<<datendosulfan, echo=TRUE, eval=FALSE>>=
+%data("endosulfan")
+%str(endosulfan)
+%@
+%<<datendosulfan2, echo=FALSE, eval=TRUE>>=
+%data("endosulfan")
+%str(endosulfan, width = 70, strict.width = "cut")
+%@
 In ecotoxicology, a
@@
 <<fitendo.echo, echo=TRUE, fig=FALSE>>=
+data("endosulfan")
 ATV <-endosulfan$ATV
@@
 eiketal13,Wu2:2013,drakeetal2014}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The \pkg{fitdistrplus} package is complementary with the \pkg{distrMod} package \citep{distrModJSS}. 
+\pkg{distrMod} provides an even more flexible way to estimate distribution parameters but its use requires
+a greater initial investment to learn how to manipulate the \texttt{S4} classes and methods
+developed in the \texttt{distr}-family packages.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The \pkg{fitdistrplus} package is also complementary with the \pkg{distrMod} package, \citep{distrModJSS}.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Many extensions of this package are planned in the future: 
+Many extensions of the \pkg{fitdistrplus} package are planned in the future: 
 we target to extend to censored data some methods for the moment only available for non-censored data, 

### 2014-02-10 JSS/fitdistrplus/paper2JSS.Rnw
before resubmit
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@272 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 or operations research. Indeed, this package is already used by a lot of practionners and academics 
-for simple MLE fits \citep{jaloustreetal11,saketal11,kochetal12,marquetouxetal12,nordan12,
-scholletal12,suuronenetal12,varoetal12,mandletal13,mala2013,mala2013publi,nadarajahbakar2013,simoetal13,
-westphalfitch13,wayland13,vianaetal13,wu13,staggeetal13,fiorellietal13,mukhopadhyay13,tikoleetal13,
+for simple MLE fits \citep{jaloustreetal11,saketal11,kochetal12,marquetouxetal12,
+scholletal12,suuronenetal12,varoetal12,mandletal13,mala2013publi,nadarajahbakar2013,simoetal13,
+westphalfitch13,wayland13,vianaetal13,wu13,staggeetal13,fiorellietal13,tikoleetal13,
 voigtetal14}, 
 for MLE fits and goodness-of-fit statistics 
-\citep{tarnczi11,garcia12,bagariaetal12,benavidesetal12,breitbach12,poduje12,Pouillot10,vaninsky13}, 
+\citep{tarnczi11,bagariaetal12,benavidesetal12,breitbach12,Pouillot10,vaninsky13}, 
 for MLE fits and bootstrap \citep{croucheretal12,meheustetal12,orellanoetal12,telloetal12,hoelzeretal12,
-prosseretal13,Zhang2013,lebaut13,Rigaux2014}, 
-for MLE fits, bootstrap and goodness-of-fit statistics \citep{rebuge12,larrasetal13}, 
+prosseretal13,Zhang2013,Rigaux2014}, 
+for MLE fits, bootstrap and goodness-of-fit statistics \citep{larrasetal13}, 
 for MME fit \citep{luangkesornetal12,callauetal13,satoetal13}, 
-for censored MLE and bootstrap \citep{lehaetal11,pouillotetal11,poulliotetal12,jongenburgeretal12,
-commeauetal12,Kingetal2013,contrerasetal2013,duarte13,feng13}, 
-for graphic analysing in \citep{anandetal12,rosa12}, 
+for censored MLE and bootstrap \citep{lehaetal11,poulliotetal12,jongenburgeretal12,
+commeauetal12,contrerasetal2013}, 
+for graphic analysing in \citep{anandetal12}, 
 for grouped-data fitting methods \citep{fusteinercostafreda12} 
-or more generally \citep{busschaertetal10,brooksetal11,aktassjostrand11,bakos11,eling12,eikhermann12,
-sosaetal2013,srinivasanetal2013,meyeretal13,Guillier2013471,Daelmanetal13,gustafssonetal13,
+or more generally \citep{busschaertetal10,eling12,
+sosaetal2013,srinivasanetal2013,meyeretal13,Guillier2013471,Daelmanetal13,
 eiketal13,Wu2:2013,drakeetal2014}.

### 2014-02-11 JSS/fitdistrplus/paper2JSS.Rnw
new definition of authors, maintainer and contributors in the Description file and minor changes in paper2JSS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@273 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 distributions for loss modelling: the lognormal distribution and the Pareto type II 
-distribution, see e.g. \citep{Klugmanetal09}.
+distribution \citep{Klugmanetal09}.
 

### 2014-02-13 JSS/fitdistrplus/paper2JSS.R
version submitted to JSS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@274 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -9,7 +9,7 @@ data("groundbeef")
 str(groundbeef)
 
 #3
-plotdist(groundbeef$serving)
+plotdist(groundbeef$serving, histo = TRUE, demp = TRUE)
 
 #4
 descdist(groundbeef$serving, boot=1000)
@@ -29,9 +29,6 @@ ppcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
 
 #7
 data("endosulfan")
-str(endosulfan)
-
-#8
 ATV <-endosulfan$ATV
 fendo.ln <- fitdist(ATV, "lnorm")
 library("actuar")
@@ -43,23 +40,23 @@ cdfcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), xlogscale = TRUE,
         ylogscale = TRUE, legendtext = 
           c("lognormal", "loglogistic", "Pareto", "Burr"))
 
-#9
+#8
 quantile(fendo.B, probs = 0.05)
 quantile(ATV, probs = 0.05)
 
-#10
+#9
 gofstat(list(fendo.ln, fendo.ll, fendo.P, fendo.B),
         fitnames = c("lnorm","llogis","Pareto","Burr"))
 
-#11
+#10
 bendo.B <- bootdist(fendo.B, niter=1001)
 summary(bendo.B)
 plot(bendo.B)
 
-#12
+#11
 quantile(bendo.B, probs = 0.05)
 
-#13
+#12
 fendo.ln.ADL <- fitdist(ATV,"lnorm",method="mge",gof="ADL")
 fendo.ln.AD2L <- fitdist(ATV,"lnorm",method="mge",gof="AD2L")
 cdfcomp(list(fendo.ln, fendo.ln.ADL, fendo.ln.AD2L),
@@ -67,7 +64,7 @@ cdfcomp(list(fendo.ln, fendo.ln.ADL, fendo.ln.AD2L),
         main = "Fitting a lognormal distribution",xlegend = "bottomright",
         legendtext = c("MLE","Left-tail AD", "Left-tail AD 2nd order"))
 
-#14
+#13
 (HC5.estimates <- c(
   empirical = as.numeric(quantile(ATV, probs=0.05)),
   Burr = as.numeric(quantile(fendo.B, probs=0.05)$quantiles),
@@ -75,7 +72,7 @@ cdfcomp(list(fendo.ln, fendo.ln.ADL, fendo.ln.AD2L),
[…]

### 2014-09-17 pkg/documair/inst/sumf.f
14_09_17: mise a jour de documair et placement de rbmn
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@283 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,5 @@
+      subroutine sumf(a, b, res)
+      integer a,b
+      double precision  res
+      res = (a+b)*1.0
+      end

### 2014-09-17 pkg/rbmn/R/auxi.code.r

@@ -1,0 +1,504 @@
+
+#<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<
+aux0 <- function(az)
+#TITLE returns a convenient function
+#DESCRIPTION \samp{sqrt(1-sum(az^2))}.\cr Not intended for the
+# final user
+#DETAILS
+#KEYWORDS 
+#INPUTS
+#{az} << numerical vector.>>
+#[INPUTS]
+#VALUE
+# sqrt(1-sum(az^2))
+#EXAMPLE
+#REFERENCE
+#SEE ALSO
+#CALLING
+#COMMENT
+#FUTURE
+#AUTHOR J.-B. Denis
+#CREATED 12_01_13
+#REVISED 12_01_18
+#--------------------------------------------
+{
+  sqrt(1-sum(az^2));
+}
+#>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
+
+#<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<
+aux1 <- function(rho,inverse=FALSE)
+#TITLE returns the generating matrix of a sequential chain
+#DESCRIPTION From a vector of length \samp{n-1} of correlations
+# computes the generating matrix \samp{n x n} of a sequential
+# chain with the root in first position (centred and normalized).\cr
+# This function is not intented for the end user: much better
+# to use \samp{chain2gema} which allows also non zero expectations
+# and non unity variances.
+#DETAILS
+#KEYWORDS 
+#INPUTS
+#{rho} << Vector of the \samp{n-1} correlations.>>
+#[INPUTS]
+#{inverse} <<Indicates if the order of rows and columns must be reverted.>>
+#VALUE
+# a square matrix, say \samp{L} such
+# that the vector \samp{L\%*\%W} has identical
+# distribution with the chain; \samp{W} being
+# a white noise.
+#EXAMPLE
+#REFERENCE
+#SEE ALSO
+#CALLING
+#COMMENT
+#FUTURE
+#AUTHOR J.-B. Denis
+#CREATED 12_01_13
+#REVISED 12_02_06
+#--------------------------------------------
+{
+  # no check
+  # completing the vector
+  rho <- c(0,rho);
+  # numbers of nodes
+  nn <- length(rho);
+  # defining an auxiliary function
+  ax <- function(h,k) {
[…]

### 2014-09-17 pkg/rbmn/R/chain.code.r

@@ -1,0 +1,1078 @@
+
+#<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<
+generate8chain <- function(rnn=c(3,7),proo=0.5,rcor=c(-1,1),
+                         rmu=c(0,0),rsig=c(0,1),
+                         nona=sssform3names(max(rnn)),check=rbmn0$check$v)
+#TITLE generation of a /chain/ /nbn/
+#DESCRIPTION [randomly] generates a /chain/ /nbn/ for the structure
+# and for the regressions coefficients.
+#DETAILS
+# Proposed ranges can be a unique value, implying no randomness
+# in the value.\cr
+# Roots are placed according to \samp{proo} probabilities, then collider
+# are placed in between with uniform probability on the possibles nodes.
+#KEYWORDS 
+#INPUTS
+#[INPUTS]
+#{rnn} <<Range of the number of nodes.>>
+#{proo} <<Probabilit[y|ies] that the successive and acceptable nodes be colliders. Can be a vector.>>
+#{rcor} <<Range of the correlations between neighbour nodes.>>
+#{rmu}  <<Range of the expectations.>>
+#{rsig} <<Range of the standard deviations.>>
+#{nona} <<Proposed names for the maximum number of nodes, only the
+#         necessary first ones will be used.>>
+#{check} << Must a check of the arguments be performed?>>
+#VALUE
+# A /chain/ coding list is returned.
+#EXAMPLE
+# set.seed(1234);
+# print8chain(generate8chain(check=TRUE));
+# print8chain(generate8chain(check=TRUE));
+# print8chain(generate8chain(rnn=10,rcor=0.5,check=TRUE));
+# print8chain(generate8chain(rnn=10,rcor=0.5,check=TRUE));
+#REFERENCE
+#SEE ALSO
+#CALLING
+#COMMENT
+#FUTURE split the function into the generation of the structure and
+# the generation of the regression coefficients.
+#AUTHOR J.-B. Denis
+#CREATED 13_07_03
+#REVISED 13_07_03
+#--------------------------------------------
+{
+  # checking
+  if (check) {
+    sssobject9(rnn,"integer",c(1,2),c(1,Inf),mensaje="'rnn' is not acceptable");
[…]

### 2014-09-17 pkg/rbmn/R/misc.code.r

@@ -1,0 +1,285 @@
+
+#<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<
+bn7nb <- function(n,label=FALSE,check=rbmn0$check$v)
+#TITLE number of distinct Bayesian networks
+#DESCRIPTION
+# returns the number of distinct Bayesian networks
+# having \samp{n} (labelled or not) nodes. Non labelled nodes means
+# that nodes are exchangeable: \samp{A -> B} is identical to
+# \samp{A <- B}.
+#DETAILS
+# When not labelled nodes, the results were proposed by Sloane in 'the on line
+# encyclopedy of integer sequences' (http://oeis.org/A003087).
+# For labelled nodes, just the application of the recursive formula of Robinson
+# (rounding errors are possible for great values of \samp{n}). 
+#KEYWORDS 
+#INPUTS
+#{n} << number of nodes. Must be less or equal to 18.>>
+#[INPUTS]
+#{label} <<Indicates if the nodes must be considered as labelled or not.>>
+#{check} <<Must argument be checked?>>
+#VALUE
+# Number of Bayesian networks
+#EXAMPLE
+# bn7nb(5,FALSE,TRUE)
+# bn7nb(5,TRUE,TRUE);
+#REFERENCE
+#SEE ALSO
+#CALLING
+#COMMENT
+#FUTURE
+#AUTHOR J.-B. Denis
+#CREATED 12_06_13
+#REVISED 13_04_29
+#--------------------------------------------
+{
+  # checking
+  if (check) {
+    sssobject9(n,"integer",1);
+    sssobject9(label,"logical",1);
+  }
+  if (label) {
+    n <- max(0,round(n));
+    if (n > 18) {
+      ssserreur(n,"'n' is too large",w=TRUE);
+      return(NA);
+    }
+    res <- vector("numeric",19);
+    res[1+ 0] <- 1;
+    res[1+ 1] <- 1;
+    res[1+ 2] <- 2;
+    res[1+ 3] <- 6;
+    res[1+ 4] <- 31;
+    res[1+ 5] <- 302;
+    res[1+ 6] <- 5984;
+    res[1+ 7] <- 243668;
+    res[1+ 8] <- 20286025;
+    res[1+ 9] <- 3424938010;
+    res[1+10] <- 1165948612902;
[…]

### 2014-09-17 pkg/rbmn/R/nbnmi.code.r

@@ -1,0 +1,752 @@
+
+#<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<
+topo9 <- function(nbn)
+#TITLE checks a topological order
+#DESCRIPTION checks that the supposed \samp{nbn} has got
+# a topological order. If not return some indications about
+# the inconsistencies.
+#DETAILS
+# I am not completely certain about the validity of the
+# proposed algorithm...
+#KEYWORDS 
+#INPUTS
+#{nbn} << The \samp{nbn} object to check.>>  
+#[INPUTS]
+#VALUE
+# \samp{character(0)} if true or a \samp{character} containing 
+# the nodes involved in some cycle or, descendants of nodes involved
+# in some cycle.
+#EXAMPLE
+# rbmn0 <- complete8rbmn0(rbmn0);
+# topo9(rbmn0$nbn1$v);
+# topo9(rbmn0$adja1$v);
+#REFERENCE
+#SEE ALSO
+#CALLING
+#COMMENT
+#FUTURE
+#AUTHOR J.-B. Denis
+#CREATED 14_08_13
+#REVISED 14_08_13
+#--------------------------------------------
+{
+  # initial checking
+  if (!is.list(nbn)) {
+    return("A /nbn/ must be a list");
+  }
+  # getting the node numbers
+  nbno <- length(nbn);
+  # corner case
+  if (nbno==0) { return(character(0));}
+  # checking the existence of node names
+  nano <- names(nbn);
+  if (length(unique(nano)) < nbno) {
+    return("A /nbn/ must be a named list with different names");
+  }
+  # initialization
+  innod <- character(0);
+  # getting the root nodes
+  for (nn in nano) {
+    if (length(nbn[[nn]]$parents)==0) {
+      innod <- c(innod,nn);
+    }
+  }
+  if (length(innod)==0) {
+    return(nano);
+  }
+  # getting the children of the root nodes
+  newnod <- 1;
+  while (newnod > 0) {
+    newnod <- 0;
+    # trying every non included node
+    for (nn in setdiff(nano,innod)) {
+      # are all its parents already included?
+      ppa <- nbn[[nn]]$parents;
+      if (length(innod)==length(union(innod,ppa))) {
+        # it can be included
+        newnod <- 1;
+        innod <- c(innod,nn);
+      }
+    }
+  }
+  # are all nodes included
[…]

### 2014-09-17 pkg/rbmn/inst/nbnmi.test.r

@@ -1,0 +1,40 @@
+#
+# 14_08_13
+#
+debug <- FALSE;
+# loading the package
+if (debug) {
+  source8file(Sys.glob("*.code.r"));
+  voir <- 1;
+  pau <- TRUE;
+} else {
+  library(rbmn);
+  voir <- NA;
+  pau <- FALSE;
+}
+rbmn0 <- complete8rbmn0(rbmn0);
+#
+## testing the topological ordering
+#
+sssform3display(topo9(list(A=NULL,B=NULL,C=list(parents=c("A","B")))),pau=pau);
+sssform3display(topo9(list(A=list(parents="B"),B=NULL,C=list(parents=c("A","B")))),pau=pau);
+sssform3display(topo9(list(A=list(parents="B"),
+                        B=list(parents="C"),
+                        C=list(parents=c("A","B")))),pau=pau);
+sssform3display(topo9(list(A=list(parents="B"),
+                        B=list(parents="C"),
+                        C=list(parents=c("C")))),pau=pau);
+sssform3display(topo9(list(A=list(parents=character(0)),
+                        B=list(parents="C"),
+                        C=list(parents=c("B")))),pau=pau);
+sssform3display(topo9(list(A=list(parents=character(0)),
+                        B=list(parents="D"),
+                        C=list(parents="B"),
+                        D=list(parents="C"))),pau=pau);
+sssform3display(topo9(list(A=list(parents="C"),
+                        B=list(parents="A"),
+                        C=list(parents="B"),
+                        D=list(parents="C"))),pau=pau);
+#
+## finished
+sssform3title("nbnmi.test.r finished at the end",8);

### 2014-09-17 pkg/rbmn/inst/normal.code.r

@@ -1,0 +1,818 @@
+
+#<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<
+check8mn <- function(mn)
+#TITLE checks a /mn/ object
+#DESCRIPTION checks the consistency of \samp{mn} as a /mn/ object,
+# returns some clues if inconsistent.
+#DETAILS
+# Looking a the code of this function provides a way to know which
+# are the requirements of a /mn/ object. But rouhghly speaking, it is a list 
+# with two (or three) names components. \samp{$mu} for the expectation (when 
+# possible covariables are nought); \samp{gamma} for the (conditional)
+# variance matrix (and \samp{rho} for the regression coefficients of
+# the covariables). All this structure must be consistently named or dimnamed.
+#KEYWORDS 
+#INPUTS
+#{mn} << The \samp{mn} object to check.>>  
+#[INPUTS]
+#VALUE
+# \samp{character(0)} or a \samp{character} containing some clue
+# about the discovered inconsistency.
+#EXAMPLE
+# rbmn0 <- complete8rbmn0(rbmn0);  
+# check8mn(rbmn0$mn1$v);
+# check8mn(rbmn0$adja1$v);
+#REFERENCE
+#SEE ALSO
+#CALLING
+#COMMENT
+#FUTURE
+#AUTHOR J.-B. Denis
+#CREATED 14_08_18
+#REVISED 14_08_19
+#--------------------------------------------
+{
+  res <- character(0);
+  # initial checking
+  che <- sssobject9(mn,"list",c(2,3),fatal=FALSE);
+  if (length(che) > 0) {
+    return(c(che,"A /mn/ must be a list of length 2 or 3."));
+  }
+  lino <- names(mn);
+  if (is.null(lino)) { return("A /mn/ must be a named list.");}
+  if (!("mu" %in% lino)) { return("A /mn/ must have a component '$mu'");}
+  if (!("gamma" %in% lino)) { return("A /mn/ must have a component '$gamma'");}
+  che <- sssobject9(mn$mu,"numeric",-1,fatal=FALSE);
+  # checking $mu
+  if (length(che)>0) {
+    return(c(che,"'mn$mu' must be a numerical vector"));
+  }
+  nbno <- length(mn$mu);
+  nano <- names(mn$mu);
+  if (is.null(nano)) {
[…]

### 2014-09-17 pkg/rbmn/inst/test.nbn.r

@@ -1,0 +1,93 @@
+#
+# 12_07_27 13_01_28 13_04_24 13_04_26 14_08_21
+#
+library(rbmn);
+#
+#
+## getting the data set
+data(boco)
+print(head(boco));
+#
+## estimating a simple model
+mod <- rbmn0$nbn5$v;
+est <- estimate8nbn(rbmn0$nbn5$v,boco);
+print8nbn(mod);
+print8nbn(est);
+#
+## getting the main forms of a Bayesian network
+nbnmod <- mod;
+print8nbn(nbnmod);
+mnmod <- nbn2mn(mod);
+print8mn(mnmod);
+gemamod <- nbn2gema(mod);
+print8gema(gemamod);
+#
+stop("fini");
+#
+source("auxi.code.r");
+source("proba.code.r");
+source("nbn.code.r");
+source("interface.code.r");
+source("rbsb0.code.r");
+#
+#
+nbn1 <- read8nbn("test.nbn.code.dat1");
+form3titre("Initial /nbn/");
+print8nbn(nbn1);
+mn1 <- gema2mn(nbn2gema(nbn1));
+print8mn(mn1);
+print(round(var2pre(mn1$gamma),3));
+#
+set.seed(4321);
+nn <- 3; nbn1 <- generate8nbn(rnn=nn);
+ny <- 100; don <- as.data.frame(matrix(round(100*runif(nn*ny)),ncol=nn,
+                                       dimnames=list(NULL,names(nbn1))));
+form3titre("Initial /nbn/");
+print8nbn(nbn1);
+nbn2 <- estimate8nbn(nbn1,don);
+form3titre("Estimated /nbn/");
+print8nbn(nbn2);
+# testing with /bnlearn/
+library(bnlearn);
+dag <- nbn2bnfit(nbn1,TRUE);
+bnf <- bn.fit(dag,data=don);
+nbn3 <- bnfit2nbn(bnf);
+form3titre("Estimated /nbn/ by /bnlearn/");
+print8nbn(nbn3);
+pause("estimates OK ?");
+#
+set.seed(4321);
+nbn1 <- generate8nbn(rnn=5);
+form3titre("Initial /nbn/");
+print8nbn(nbn1);
+form3titre("Its R-matrix");
+rm1 <- nbn2rmatrix(nbn1);
+print(rm1);
[…]

### 2014-09-17 pkg/rbmn/inst/test.stars.r

@@ -1,0 +1,74 @@
+#
+# 13_11_28 13_11_29 13_12_01 13_12_02 13_12_04
+# 13_12_05
+#
+library(rbsa);
+library(rbmn);
+source("stars.code.r");
+xx1 <- create8star();
+print(xx1);
+vv1 <- create8star(quoi="vari");
+print(vv1);
+print(xx1%*%t(xx1));
+pause("regarde");
+#
+for (ni in c(0,1,3,10,20,30)) {
+  for (no in c(0,1,3,7,15,50,100)) {
+    xx <- create8star(rep(1,ni),rep(1,no),
+                      rep(1,ni),1,rep(1,no),
+                      "gema");
+    vv <- create8star(rep(1,ni),rep(1,no),
+                      rep(1,ni),1,rep(1,no),
+                      "vari");
+    dd <- sum((xx%*%t(xx)-vv)^2);
+    cat(ni,no,dd,"\n");
+  }
+}
+#
+form3display(vv1);
+form3display(round(solve(vv1),3));
+dia1 <- svd(vv1);
+form3display(dia1$d);
+form3display(round(dia1$u,4));
+#
+vv2 <- create8star(rho=rep(0.4,4),
+                   lambda=rep(0.5,2),
+                   quoi="vari");
+form3display(vv2);
+form3display(round(solve(vv2),3));
+dia2 <- svd(vv2);
+form3display(dia2$d);
+form3display(round(dia2$u,4));
+#
+vv3 <- create8star(rho=rep(0.4,4),
+                   lambda=c(0.2,0.4,0.6),
+                   quoi="vari");
+form3display(vv3);
+form3display(round(solve(vv3),3));
+dia3 <- svd(vv3);
+form3display(dia3$d);
+form3display(round(dia3$u,4));
+#
+vv4 <- create8star(rho=(1:4)/10,
+                   lambda=c(0.2,0.4,0.6),
+                   quoi="vari");
+form3display(vv4);
+form3display(round(solve(vv4),3));
+dia4 <- svd(vv4);
+form3display(dia4$d);
+form3display(round(dia4$u,4));
+#
+for (rho in c(0.1,0.2,0.3)) {
+  for (lambda in c(0.1,0.2,0.3)) {
+    vv5 <- create8star(rho=rho,
+                       lambda=lambda,
+                       quoi="vari");
+    cat("\n\n\n rho =",rho," et lambda =",lambda,"\n\n");
+    form3display(vv5);
[…]

### 2014-09-17 pkg/rbmn/tests/chain.test.r.R

@@ -1,0 +1,25 @@
+#
+# 14_08_24
+#
+debug <- FALSE;
+# loading the package
+if (debug) {
+  source8file(Sys.glob("*.code.r"));
+  voir <- 1;
+  pau <- TRUE;
+} else {
+  library(rbmn);
+  voir <- NA;
+  pau <- FALSE;
+}
+rbmn0 <- complete8rbmn0(rbmn0);
+#
+## testing read8chain
+#
+print8chain(rbmn0$chain1$v);
+write8chain(rbmn0$chain1$v,file="toto.txt");
+print8chain(read8chain("toto.txt"));
+unlink("toto.txt");
+#
+## finished
+sssform3title("chain.test.r finished at the end",8);

### 2014-09-17 pkg/rbmn/tests/misc.test.r.R

@@ -1,0 +1,22 @@
+#
+# 14_04_07 14_08_11
+#
+# testing misc.code.r functions
+#
+debug <- FALSE;
+# loading the package
+if (debug) {
+  source8file(Sys.glob("*.code.r"));
+} else {
+  library(rbmn);
+}
+rbmn0 <- complete8rbmn0(rbmn0);
+#
+aa1 <- rbmn0$adja2$v;
+sssform3title("Adjacency Matrix");
+print(aa1);
+sssform3title("Its raw dissimilarity expression");
+print(dissim4adja(aa1,nature="d"));
+sssform3title("Its symetricized dissimilarity expression");
+print(dissim4adja(aa1));
+

### 2014-09-17 pkg/rbmn/R/objects.code.r
14_09_17: mise a jour de documair et placement de rbmn
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@283 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,264 @@
+#
+# defining constants for the /rbmn/ package
+#
+#<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<
+rbmn0 <- 
+#TITLE list of the /rbmn/ constants
+#DESCRIPTION
+# Just a list of constants defined with a name, a definition
+# and a value. They can be modified by the user. Values
+# definitions and names can be obtained with the function
+# \samp{sssdisplay8k} of /rbsa/ package.
+#DETAILS
+# It is a named list, one component for each constant.
+# A sublist is associated to each constant with two components: \samp{$d}
+# for the definition and \samp{$v} for the value. Be aware that the value
+# can be any object (vector, list, matrix, function,...)
+#KEYWORDS misc
+#INPUTS
+#[INPUTS]
+#VALUE
+# A list see the description and details to be able to know its contents.
+# (it is a self-documented object.)
+#EXAMPLE
+# names(rbmn0);
+# rbmn0$check;
+# rbmn0$chain1;
+#REFERENCE
+#SEE ALSO
+#CALLING
+#COMMENT
+#FUTURE
+#AUTHOR J.-B. Denis
+#CREATED 13_07_03
+#REVISED 14_09_16
+#--------------------------------------------
+{
+  # initialization
+  set.seed(1234);
+  #
+  #  <0> general monitoring
+  #
+  l0 <- list(
+             check=list(v=TRUE,
+                        d="Must systematic checking be done in the functions"
+                       )
+            );
+  #
+  #  <1> DEFINING SOME INITIAL OBJECTS
+  #
+  l1 <- list(
+         chain1=list(v=list(names = c("A", "B", "C"),
+                            mu    = c(  0,   0,   0),
+                            sigma = c(  1,   2,   3),
+                            corre = c(0.5, 0.5), 
+                            roots = c("A"), colliders = NULL
+                           ),
+                     d="chain1 as /chain/"),
+         chain2=list(v=list(names = letters[1:5],
+                            mu    = 1:5,
+                            sigma = 5:1,
+                            corre = rep(1/3,4), 
+                            roots = c("a","d"), colliders = c("c")
+                           ),
+                     d="chain2 as /chain/"),
+         chain3=list(v=list(names=LETTERS[1:10],
[…]

### 2014-09-17 pkg/rbmn/R/rrbsa.code.r

@@ -1,0 +1,4630 @@
+
+#<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<
+sssform3title <- function(tit,empha=3,indent=2+2*empha,imp=TRUE)
+#TITLE  prints or prepares a title
+#DESCRIPTION
+# prints or prepares the character string \samp{tit}
+# with more or less emphasis.
+# This function is a shortcut of the hidden function \samp{sssform3titre},
+# with some specialized calls.
+#DETAILS
+#KEYWORDS print
+#INPUTS
+#{tit}<<the title to print (just one line)>>
+#[INPUTS]
+#{empha} << Level of emphasizing.\cr
+#          (0): single line without carriage return\cr
+#          (1): single line\cr
+#          (2): underlined\cr
+#          (3): underlined and overlined\cr
+#          (4): (2) + 1 line before\cr
+#          (5): (3) + 1 line after\cr
+#          (6): (2) + 2 lines before and after\cr
+#          (7): corners + 1 line before and after (plus surrounding)\cr
+#          (8): box + 1 lines before and after (plus surrounding)\cr>>
+#{indent} << Number of spaces to introduce before the title>>
+#{imp} << Printing is performed and nothing is returned.
+#                If FALSE, the character string is returned 
+#                (including possible new lines)>>
+#VALUE
+# either nothing or a character string according to imp
+#EXAMPLE
+# for (ii in 0:8) {sssform3title("Some Title",ii,imp=TRUE)};
+#REFERENCE
+#SEE ALSO
+#CALLING {sssform3repeat}
+#COMMENT
+#FUTURE
+#AUTHOR J.-B. Denis
+#CREATED 07_09_24
+#REVISED 10_09_15
+#--------------------------------------------
+{
+# adjusting
+empha <- round(max(0,min(8,empha)));
+# preparing the argument for sssform3titre
+if (length(tit)>1) { tit <- paste(tit,collapse=" ");}
+if (empha == 0) { tit <- paste0("<",tit,">")}
+if (empha == 1) { tit <- paste0("(*)",tit,"(*)")}
+sbef <- round(max(indent,0));
+caret <- (empha != 0);
+saft=""; lbef <- 0; laft <- 0;
+box <- "no";
+if (empha == 2) { box <- "un";}
+if (empha == 3) { box <- "unov";}
+if (empha == 4) { box <- "unov"; lbef <- 1;}
[…]

### 2014-09-17 pkg/rbmn/inst/make-rbmn.r

@@ -1,0 +1,15 @@
+#
+# 14_05_20 14_05_21 14_05_22 14_08_29 14_09_02
+# 14_09_15
+#
+library(documair); library(rbsa);
+#
+resu <- build8pkg("rbmn",".","../resu",what="pzl");
+#
+if (void9(resu)) {
+  form3title("'make-rbmn.r' finished its work!",8);
+} else {
+  print(resu);
+  stop("ERREUR");
+}
+

### 2014-09-17 pkg/rbmn/inst/nbnst.test.r

@@ -1,0 +1,98 @@
+#
+# 13_04_19 13_04_20 13_04_21 13_04_24 13_04_25
+# 14_08_15 14_08_16
+#
+debug <- FALSE;
+# loading the package
+if (debug) {
+  source8file(Sys.glob("*.code.r"));
+  voir <- 1;
+} else {
+  library(rbmn);
+  voir <- NA;
+}
+rbmn0 <- complete8rbmn0(rbmn0);
+#
+library(bnlearn);
+#
+print8nbn(rbmn0$nbn5$v);
+data(boco);
+print8nbn(estimate8nbn(rbmn0$nbn5$v,boco));
+ssspause("estimation",top=voir);
+sarc5 <- rbmn0$arc5$v;
+print8nbn(estimate8constrained7nbn(rbmn0$nbn5$v,sarc5,boco));
+ssspause("constrained estimation",top=voir);
+#
+set.seed(4321);
+nbnc <- generate8nbn(3,nona=paste("c",1:3,sep=""));
+nbnr <- generate8nbn(4,nona=paste("r",1:4,sep=""));
+nbn1 <- crossed4nbn1nbn(nbnr,nbnc);
+nn <- length(nbn1);
+ny <- 100; don <- as.data.frame(matrix(round(100*runif(nn*ny)),ncol=nn,
+                                       dimnames=list(NULL,names(nbn1))));
+sssform3title("Initial /nbn/");
+print8nbn(nbn1);
+nite <- 20; eps <- 10^-3;
+sarc <- vector("list",length(nbnr)+length(nbnc));
+sarc[1:length(nbnr)] <- arcs4nbn1nbn(nbnr,nbnc,"a2");
+sarc[length(nbnr)+(1:length(nbnc))] <- arcs4nbn1nbn(nbnr,nbnc);
+nbn2 <- estimate8constrained7nbn(nbn1,sarc,don,
+                                imp=TRUE,nite=nite,eps=eps);
+print8nbn(nbn2);
+ssspause("constrained estimates OK (integrated)?",top=voir);
+#
+#
+if (debug) {
+  nbn1 <- read8nbn("perso/test.nbn.code.dat4");
+  nn <- length(nbn1);
+  ny <- 10; don <- as.data.frame(matrix(round(100*runif(nn*ny)),ncol=nn,
+                                         dimnames=list(NULL,names(nbn1))));
+  sssform3title("Initial /nbn/");
+  print8nbn(nbn1);
[…]

### 2014-09-17 pkg/rbmn/inst/normal.test.r

@@ -1,0 +1,39 @@
+#
+# 14_08_19
+#
+debug <- FALSE;
+# loading the package
+if (debug) {
+  source8file(Sys.glob("*.code.r"));
+  voir <- 1;
+  pau <- TRUE;
+} else {
+  library(rbmn);
+  voir <- NA;
+  pau <- FALSE;
+}
+rbmn0 <- complete8rbmn0(rbmn0);
+#
+## testing the simulation
+set.seed(1234);
+sisi1 <- simulate8mn(rbmn0$mn1$v,5);
+sssform3display(sisi1,pau=debug);
+loi <- list(mu=c(D=2,E=4),
+            rho=matrix(1:6,2,dimnames=list(LETTERS[4:5],
+                                          LETTERS[1:3])),
+            gamma=matrix(c(1,1,1,2),2,dimnames=list(LETTERS[4:5],LETTERS[4:5])));
+cova <- matrix(runif(36),12,dimnames=list(NULL,LETTERS[1:3]));
+sisi2 <- simulate8gmn(loi,cova,12);
+sssform3display(sisi2,pau=debug);
+#
+## testing print8mn
+print8mn(rbmn0$mn1$v);
+loi <- list(mu=c(D=2,E=4),
+            rho=matrix(1:6,2,dimnames=list(LETTERS[4:5],
+                                          LETTERS[1:3])),
+            gamma=matrix(c(1,1,1,2),2,dimnames=list(LETTERS[4:5],LETTERS[4:5])));
+print8mn(loi);
+ssspause("printing",top=voir);
+#
+## finished
+sssform3title("normal.test.r finished at the end",8);

### 2014-09-17 pkg/rbmn/inst/stars.code.r

@@ -1,0 +1,96 @@
+
+#<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<
+create8star <- function(rho=rep(0.5,2),
+                        lambda=rep(0.4,4),
+                        sigi=rep(1,length(rho)),
+                        sigc=sqrt(1-sum(rho^2)),
+                        sigo=sqrt(1-lambda^2),
+                        quoi="gema",check=rbmn0$check$v)
+#TITLE returns a star /rbmn/
+#DESCRIPTION 
+# A star /rbmn/ is a /rbmn/ comprising a centre node to which enter 
+# \samp{ni=length(rho)} independent input nodes and from which
+# come out \samp{no=length(lambda)} output nodes. This
+# gives roughly the configuration of a star, the centre node being the center
+# of the star and the other nodes some kind of radiuses around it.
+#DETAILS
+#KEYWORDS 
+#INPUTS
+#[INPUTS]
+#{rho}<< Gives the regression coefficients of each input nodes,
+# which are the parents of the center node.>>
+#{lambda}<< Gives the regression coefficients of the centre node for each output node.>>
+#{sigi} << Vector of the standard deviation of the input nodes.>>
+#{sigc} << Scalar giving the conditional standard deviation of the center node.
+#  The default value ensures a marginal variance of 1.>>
+#{sigo} << Vector giving the conditional standard deviation of the output nodes.
+#  The default value ensures a marginal variance of 1.>>
+#{quoi}<< Indicates the desired type of output. The different possibilities 
+# are \samp{gema} for a generating matrix, \samp{vari} for the variance
+# matrix.>>
+#{check} << Must a check of the arguments be performed?>>
+#VALUE
+# The resulting structure with standard names.
+#EXAMPLE
+# create8star(check=TRUE);
+# create8star(rho=rep(0.2,10),check=TRUE);
+# create8star(rho=rep(0,10),quoi="vari",check=TRUE);
+# create8star(rho=rep(1,4),sigc=0,quoi="vari",check=TRUE);
+#REFERENCE
+#SEE ALSO
+#CALLING
+#COMMENT
+#FUTURE
+#AUTHOR J.-B. Denis
[…]

### 2014-09-17 pkg/rbmn/inst/test.idea.r

@@ -1,0 +1,81 @@
+#
+# 13_04_19 13_04_20 13_04_21 13_04_24 13_04_25
+#
+library(rbsb);
+source("auxi.code.r");
+source("proba.code.r");
+source("nbn.code.r");
+source("idea.code.r");
+source("crossed.code.r");
+#
+set.seed(4321);
+nbnc <- generate8nbn(3,nona=paste("c",1:3,sep=""));
+nbnr <- generate8nbn(4,nona=paste("r",1:4,sep=""));
+nbn1 <- nbn2crossed(nbnr,nbnc);
+nn <- length(nbn1);
+ny <- 100; don <- as.data.frame(matrix(round(100*runif(nn*ny)),ncol=nn,
+                                       dimnames=list(NULL,names(nbn1))));
+form3title("Initial /nbn/");
+print8nbn(nbn1);
+nite <- 20; eps <- 10^-3;
+sarc <- vector("list",length(nbnr)+length(nbnc));
+sarc[bf(nbnr)] <- nbn2crossed4arcs(nbnr,nbnc,TRUE);
+sarc[length(nbnr)+bf(nbnc)] <- nbn2crossed4arcs(nbnr,nbnc,FALSE);
+nbn2 <- estimate8constrainednbn(nbn1,sarc,don,
+                                imp=TRUE,nite=nite,eps=eps);
+print8nbn(nbn2);
+pause("constrained estimates OK (integrated)?");
+#
+#
+nbn1 <- read8nbn("test.nbn.code.dat4");
+nn <- length(nbn1);
+ny <- 10; don <- as.data.frame(matrix(round(100*runif(nn*ny)),ncol=nn,
+                                       dimnames=list(NULL,names(nbn1))));
+form3title("Initial /nbn/");
+print8nbn(nbn1);
+nite <- 10; nbne <- nbn1;
+for (ite in bc(nite)) {
+  aa1 <- matrix(c("A","C",
+                  "B","D"),ncol=2);
+  nbne <- estimate8arcs(nbne,aa1,don);
+  aa2 <- matrix(c("A","B",
+                  "C","D"),ncol=2);
+  nbne <- estimate8arcs(nbne,aa2,don);
+  form3title(paste("Estimated /nbn/ with",ite,"iteration"));
+  print8nbn(nbne);
+  pause(ite);
+}
[…]

### 2014-09-17 pkg/rbmn/tests/crossed.test.r.R

@@ -1,0 +1,67 @@
+#
+# 14_04_02 14_08_12 14_08_14
+#
+debug <- FALSE;
+# loading the package
+if (debug) {
+  source8file(Sys.glob("*.code.r"));
+  voir <- 1;
+} else {
+  library(rbmn);
+  voir <- NA;
+}
+rbmn0 <- complete8rbmn0(rbmn0);
+#
+library(bnlearn);
+library(igraph);
+#
+gg <- function(nbn) {
+  dag <- nbn2bnfit(nbn$nbn,TRUE);
+  dag.ig <- igraph.from.graphNEL(as.graphNEL(dag));
+  V(dag.ig)$label <- V(dag.ig)$name;
+  plot(dag.ig,layout=nbn$posi[V(dag.ig)$label,]);
+}
+#
+pdf("crossed.test.pdf");
+par(mfrow=c(1,1));
+nbn1 <- generate8grid7crossed(2,2);
+print8nbn(nbn1$nbn);
+gg(nbn1);
+#
+nbn2 <- generate8grid7crossed(rhorow=matrix(1:2,2,3));
+print(nbn2$cova2vari);
+print(nbn2$row2row);
+print(nbn2$col2col);
+print(nbn2$vari2cova);
+ssspause("arcs",top=voir);
+print8nbn(nbn2$nbn);
+gg(nbn2);
+#
+nbn3 <- generate8grid7crossed(rhorow=matrix(1:2,2,3),
+                              rhocol=matrix(10*(1:2),2,3),
+                              xrhorow=matrix(100*(1:3),3),
+                              xrhocol=matrix(1000*(1:3),3));
+print8nbn(nbn3$nbn);
+gg(nbn3);
+#
+nbn4 <- generate8grid7crossed(nrow=4,ncol=5,xrow=2,xcol=4);
+print8nbn(nbn4$nbn);
+gg(nbn4);
+dev.off();
+#
+for (nro in c(2,4,7,12)) {
+  for (nco in c(2,4,7,12)) {
+    for (xro in c(0,2,7)) {
+      for (xco in c(0,2,7)) {
+        uuu <- generate8grid7crossed(nro,nco,xro,xco);
+        cat(nro,nco,xro,xco,"\n");
+        checo <- check8nbn(uuu$nbn);
+        if (length(checo)>0) {
+          print(checo);
+          stop("va pas !");
+        }
+      }
+    }
+  }
+}
+print8nbn(uuu$nbn);

### 2014-09-19 JSS/fitdistrplus/paper2JSS.R
last version of the JSS paper I had sent to JSS on July
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@284 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,6 +1,6 @@
 #1
-options(digits=4, prompt="R>", 
-        SweaveHooks=list(fig=function() par(mar=c(5.1, 4.1, 1.1, 2.1))))
+options(digits = 4, prompt = "R>", 
+        SweaveHooks = list(fig=function() par(mar = c(5.1, 4.1, 1.1, 2.1))))
 set.seed(1234)
 
 #2
@@ -12,30 +12,30 @@ str(groundbeef)
 plotdist(groundbeef$serving, histo = TRUE, demp = TRUE)
 
 #4
-descdist(groundbeef$serving, boot=1000)
+descdist(groundbeef$serving, boot = 1000)
 
 #5
 fw <- fitdist(groundbeef$serving, "weibull")
 summary(fw)
 
 #6
-fg <- fitdist(groundbeef$serving,"gamma")
-fln <- fitdist(groundbeef$serving,"lnorm")
-par(mfrow=c(2, 2))
-denscomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
-qqcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
-cdfcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
-ppcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+fg <- fitdist(groundbeef$serving, "gamma")
+fln <- fitdist(groundbeef$serving, "lnorm")
+par(mfrow = c(2, 2))
+denscomp(list(fw, fln, fg), legendtext = c("Weibull", "lognormal", "gamma"))
+qqcomp(list(fw, fln, fg), legendtext = c("Weibull", "lognormal", "gamma"))
+cdfcomp(list(fw, fln, fg), legendtext = c("Weibull", "lognormal", "gamma"))
+ppcomp(list(fw, fln, fg), legendtext = c("Weibull", "lognormal", "gamma"))
 
 #7
 data("endosulfan")
-ATV <-endosulfan$ATV
+ATV <- endosulfan$ATV
 fendo.ln <- fitdist(ATV, "lnorm")
 library("actuar")
-fendo.ll <- fitdist(ATV, "llogis", start=list(shape=1, scale=500))
-fendo.P <- fitdist(ATV, "pareto", start=list(shape=1, scale=500))
[…]

### 2014-09-19 JSS/fitdistrplus/paper2JSS.Rnw
version submitted to JSS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@274 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
last version of the JSS paper I had sent to JSS on July
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@284 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 function using the \code{optim} function. 
-No distinction between parameters with different roles (e.g. main parameter and nuisance parameter)
-is made, as our talk focuses 
+No distinction between parameters with different roles (e.g., main parameter and nuisance parameter)
+is made, as this paper focuses 
 on parameter estimation from a general point-of-view. 
@@
 for which a coercion method to class mle is provided so that the respective 
-functionalities (e.g. \code{confint} and \code{logLik}) from package \pkg{stats4} 
+functionalities (e.g., \code{confint} and \code{logLik}) from package \pkg{stats4} 
 are available, too.
@@
 The development version of the package is located at \proglang{R}-forge as one package of the 
-project ``Risk Assessment with R''
+project ``Risk Assessment with \proglang{R}''
 (\url{http://r-forge.r-project.org/projects/riskassessment/}).
@@
 <<par4vignette, echo=FALSE, results=hide>>=
-options(digits=4, prompt="R>", 
+options(digits = 4, prompt="R> ", 
         SweaveHooks=list(fig=function() par(mar=c(5.1, 4.1, 1.1, 2.1))))
@@
 <<descgroundbeef.echo, echo=TRUE, fig=FALSE, eval=FALSE>>=
-descdist(groundbeef$serving, boot=1000)
+descdist(groundbeef$serving, boot = 1000)
 @
@@
 <<descgroundbeefplot, echo=FALSE, fig=TRUE, width=6, height=6, eps=FALSE>>=
-descdist(groundbeef$serving, boot=1000)
+descdist(groundbeef$serving, boot = 1000)
 @
@@
 \end{figure}
+
+\newpage
 
@@
 <<fitgroundbeef.echo, echo=TRUE, fig=FALSE>>=
-fg <- fitdist(groundbeef$serving,"gamma")
-fln <- fitdist(groundbeef$serving,"lnorm")
-par(mfrow=c(2, 2))
-denscomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
-qqcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
-cdfcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
-ppcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+fg <- fitdist(groundbeef$serving, "gamma")
+fln <- fitdist(groundbeef$serving, "lnorm")
+par(mfrow = c(2, 2))
+plot.legend <- c("Weibull", "lognormal", "gamma")
+denscomp(list(fw, fln, fg), legendtext = plot.legend)
+qqcomp(list(fw, fln, fg), legendtext = plot.legend)
+cdfcomp(list(fw, fln, fg), legendtext = plot.legend)
+ppcomp(list(fw, fln, fg), legendtext = plot.legend)
 @
@@
 par(mfrow=c(2, 2))
-denscomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
-qqcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
-cdfcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
-ppcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
+denscomp(list(fw, fln, fg), legendtext=c("Weibull", "lognormal", "gamma"))
+qqcomp(list(fw, fln, fg), legendtext=c("Weibull", "lognormal", "gamma"))
+cdfcomp(list(fw, fln, fg), legendtext=c("Weibull", "lognormal", "gamma"))
+ppcomp(list(fw, fln, fg), legendtext=c("Weibull", "lognormal", "gamma"))
 @ 
@@
 
-The density plot and the CDF plot may be considered as the basic classical goodness-of-fits plots.
+The density plot and the CDF plot may be considered as the basic classical goodness-of-fit plots.
 The two other plots are complementary and can be very informative in some cases.
@@
 tested on Australian and non-Australian laboratory-species 
-%(arthropods, fish or nonarthropod invertebrates)
 \citep{Hose04}. 
-%<<datendosulfan, echo=TRUE, eval=FALSE>>=
-%data("endosulfan")
-%str(endosulfan)
-%@
-%<<datendosulfan2, echo=FALSE, eval=TRUE>>=
-%data("endosulfan")
-%str(endosulfan, width = 70, strict.width = "cut")
-%@
 In ecotoxicology, a
@@
 library("actuar")
-fendo.ll <- fitdist(ATV, "llogis", start=list(shape=1, scale=500))
-fendo.P <- fitdist(ATV, "pareto", start=list(shape=1, scale=500))
-fendo.B <- fitdist(ATV, "burr", start=list(shape1=0.3, shape2=1, rate=1))
-cdfcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), xlogscale = TRUE,
-      ylogscale = TRUE, legendtext = 
-          c("lognormal", "loglogistic", "Pareto", "Burr"))
+fendo.ll <- fitdist(ATV, "llogis", start = list(shape = 1, scale = 500))
+fendo.P <- fitdist(ATV, "pareto", start = list(shape = 1, scale = 500))
+fendo.B <- fitdist(ATV, "burr", start = list(shape1 = 0.3, shape2 = 1, 
+  rate = 1))
+cdfcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), 
+  xlogscale = TRUE, ylogscale = TRUE, 
+  legendtext = c("lognormal", "loglogistic", "Pareto", "Burr"))
 @
@@
 empirical quantile for comparison.
-
-\newpage
 
@@
 
-\clearpage
-
 <<fendo.gof.print, echo=TRUE, fig=FALSE>>=
 gofstat(list(fendo.ln, fendo.ll, fendo.P, fendo.B),
-        fitnames = c("lnorm","llogis","Pareto","Burr"))
+  fitnames = c("lnorm", "llogis", "Pareto", "Burr"))
 @
@@
   \begin{center}
-    \caption{Goodness-of-fit statistics as defined by Stephens \citep{Stephens86}.}
     \begin{tabular}{lll}
@@
     \end{tabular}
+    \caption{Goodness-of-fit statistics as defined by Stephens \citep{Stephens86}.}
     \label{tabKSCvMAD}
@@
 <<fitBurr.boot.echo, echo=TRUE, results=hide, fig=FALSE>>=
-bendo.B <- bootdist(fendo.B, niter=1001)
+bendo.B <- bootdist(fendo.B, niter = 1001)
 @
@@
 
-\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \begin{center}
-    \caption{Modified Anderson-Darling statistics as defined by \cite{Luceno06}.}
     \begin{tabular}{lll}
@@
     \end{tabular}
+    \caption{Modified Anderson-Darling statistics as defined by \cite{Luceno06}.}
     \label{modifiedAD}
@@
 <<mge.gofcomp.echo, echo=TRUE, fig=FALSE>>=
-fendo.ln.ADL <- fitdist(ATV,"lnorm",method="mge",gof="ADL")
-fendo.ln.AD2L <- fitdist(ATV,"lnorm",method="mge",gof="AD2L")
+fendo.ln.ADL <- fitdist(ATV, "lnorm", method = "mge", gof = "ADL")
+fendo.ln.AD2L <- fitdist(ATV, "lnorm", method = "mge", gof = "AD2L")
 cdfcomp(list(fendo.ln, fendo.ln.ADL, fendo.ln.AD2L), 
   xlogscale = TRUE, ylogscale = TRUE,
-main = "Fitting a lognormal distribution",xlegend = "bottomright",
+  main = "Fitting a lognormal distribution",
+  xlegend = "bottomright", 
   legendtext = c("MLE","Left-tail AD", "Left-tail AD 2nd order"))
@@
 (HC5.estimates <- c(
-empirical = as.numeric(quantile(ATV, probs=0.05)),
-Burr = as.numeric(quantile(fendo.B, probs=0.05)$quantiles),
-lognormal_MLE = as.numeric(quantile(fendo.ln, probs=0.05)$quantiles),
-lognormal_AD2 = as.numeric(quantile(fendo.ln.ADL, probs=0.05)$quantiles),
-lognormal_AD2L = as.numeric(quantile(fendo.ln.AD2L,probs=0.05)$quantiles)))
+  empirical = as.numeric(quantile(ATV, probs = 0.05)), 
+  Burr = as.numeric(quantile(fendo.B, probs = 0.05)$quantiles), 
+  lognormal_MLE = as.numeric(quantile(fendo.ln, probs = 0.05)$quantiles), 
+  lognormal_AD2 = as.numeric(quantile(fendo.ln.ADL, 
+    probs = 0.05)$quantiles), 
+  lognormal_AD2L = as.numeric(quantile(fendo.ln.AD2L, 
+    probs = 0.05)$quantiles)))
 @
@@
 fdanish.ln.MLE <- fitdist(danishuni$Loss, "lnorm")
-fdanish.ln.MME <- fitdist(danishuni$Loss, "lnorm", method="mme", order=1:2)
+fdanish.ln.MME <- fitdist(danishuni$Loss, "lnorm", method = "mme", 
+  order = 1:2)
 cdfcomp(list(fdanish.ln.MLE, fdanish.ln.MME), 
-        legend=c("lognormal MLE", "lognormal MME"), 
-        main="Fitting a lognormal distribution",
-        xlogscale=TRUE, datapch=20)
+  legend = c("lognormal MLE", "lognormal MME"), 
+  main = "Fitting a lognormal distribution", 
+  xlogscale = TRUE, datapch = 20)
 @
@@
 fdanish.P.MLE <- fitdist(danishuni$Loss, "pareto", 
-start=c(shape=10, scale=10), lower=2+1e-6, upper=Inf)
+  start=c(shape=10, scale=10), lower = 2+1e-6, upper = Inf)
 memp <- function(x, order) sum(x^order)/length(x)
-fdanish.P.MME <- fitdist(danishuni$Loss, "pareto", method="mme", order=1:2, 
+fdanish.P.MME <- fitdist(danishuni$Loss, "pareto", 
+    method="mme", order=1:2, 
 memp="memp", start=c(shape=10, scale=10), lower=c(2+1e-6,2+1e-6), 
@@
 fdanish.P.MLE <- fitdist(danishuni$Loss, "pareto",  
-start=c(shape=10, scale=10), lower=2+1e-6, upper=Inf)
+  start = c(shape = 10, scale = 10), lower = 2+1e-6, upper = Inf)
 memp <- function(x, order) sum(x^order)/length(x)
-fdanish.P.MME <- fitdist(danishuni$Loss, "pareto", method="mme", 
-order=1:2, memp="memp", start=c(shape=10, scale=10), 
-lower=c(2+1e-6,2+1e-6), upper=c(Inf,Inf))
+fdanish.P.MME <- fitdist(danishuni$Loss, "pareto", method = "mme", 
+  order = 1:2, memp = "memp", start = c(shape = 10, scale = 10), 
+  lower = c(2+1e-6, 2+1e-6), upper = c(Inf, Inf))
 cdfcomp(list(fdanish.P.MLE, fdanish.P.MME), 
-legend=c("Pareto MLE", "Pareto MME"), 
-main="Fitting a Pareto distribution", xlogscale=TRUE, datapch=".")
+  legend = c("Pareto MLE", "Pareto MME"), 
+  main = "Fitting a Pareto distribution", 
+  xlogscale = TRUE, datapch = ".")
 gofstat(list(fdanish.ln.MLE, fdanish.P.MLE, 
   fdanish.ln.MME, fdanish.P.MME),
-fitnames = c("lnorm.mle","Pareto.mle","lnorm.mme","Pareto.mme"))
+  fitnames = c("lnorm.mle", "Pareto.mle", "lnorm.mme", "Pareto.mme"))
 @
@@
 <<danish.qme.echo, echo=TRUE, fig=FALSE>>=
-fdanish.ln.QME1 <- fitdist(danishuni$Loss, "lnorm", method="qme", 
-probs=c(1/3, 2/3))
-fdanish.ln.QME2 <- fitdist(danishuni$Loss, "lnorm", method="qme", 
-probs=c(8/10, 9/10))
+fdanish.ln.QME1 <- fitdist(danishuni$Loss, "lnorm", method = "qme", 
+  probs = c(1/3, 2/3))
+fdanish.ln.QME2 <- fitdist(danishuni$Loss, "lnorm", method = "qme", 
+  probs = c(8/10, 9/10))
 cdfcomp(list(fdanish.ln.MLE, fdanish.ln.QME1, fdanish.ln.QME2), 
-legend=c("MLE", "QME(1/3, 2/3)", "QME(8/10, 9/10)"), 
-main="Fitting a lognormal distribution", xlogscale=TRUE, datapch=20)
+  legend = c("MLE", "QME(1/3, 2/3)", "QME(8/10, 9/10)"), 
+  main = "Fitting a lognormal distribution", 
+  xlogscale = TRUE, datapch = 20)
 @
@@
 \end{figure}
+
+\clearpage
 
@@
 data("groundbeef")
-fNM <- fitdist(groundbeef$serving, "gamma", optim.method="Nelder-Mead")
-fBFGS <- fitdist(groundbeef$serving, "gamma", optim.method="BFGS") 
-fSANN <- fitdist(groundbeef$serving, "gamma", optim.method="SANN")
-fCG <- try(fitdist(groundbeef$serving, "gamma", optim.method="CG", 
-    control=list(maxit=10000)))
+fNM <- fitdist(groundbeef$serving, "gamma", optim.method = "Nelder-Mead")
+fBFGS <- fitdist(groundbeef$serving, "gamma", optim.method = "BFGS") 
+fSANN <- fitdist(groundbeef$serving, "gamma", optim.method = "SANN")
+fCG <- try(fitdist(groundbeef$serving, "gamma", optim.method = "CG", 
+  control = list(maxit = 10000)))
 if(class(fCG) == "try-error")
-  fCG <- list(estimate=NA)
+  fCG <- list(estimate = NA)
 @
@@
    require(rgenoud)
-   res <- genoud(fn, starting.values=par, ...)        
-   standardres <- c(res, convergence=0)
+   res <- genoud(fn, starting.values = par, ...)        
+   standardres <- c(res, convergence = 0)
    return(standardres)
@@
 <<optimmethod.customgenoud.fitdist, echo=TRUE, eval=TRUE>>=
-fgenoud <- mledist(groundbeef$serving, "gamma", custom.optim= mygenoud, 
-    nvars=2, max.generations=10, Domains=cbind(c(0,0), c(10,10)), 
-    boundary.enforcement=1, hessian=TRUE, print.level=0, P9=10)
+fgenoud <- mledist(groundbeef$serving, "gamma", custom.optim = mygenoud, 
+  nvars = 2, max.generations = 10, Domains = cbind(c(0,0), c(10,10)), 
+  boundary.enforcement = 1, hessian = TRUE, print.level = 0, P9 = 10)
 cbind(NM = fNM$estimate,
@@
 fsal.ln <- fitdistcens(salinity, "lnorm")
-fsal.ll <- fitdistcens(salinity, "llogis", start=list(shape=5, scale=40))
+fsal.ll <- fitdistcens(salinity, "llogis",
+  start = list(shape = 5, scale = 40))
 summary(fsal.ln)
@@
 cdfcompcens(list(fsal.ln, fsal.ll), 
-            legendtext=c("lognormal", "loglogistic "))
+  legendtext = c("lognormal", "loglogistic "))
 @
@@
   \caption{Plot of the fit of a discrete distribution (Poisson distribution
-  fitted to numbers of \emph{Toxocara cati} parasites from the \code{toxocara} data set.}
+  fitted to numbers of \emph{Toxocara cati} parasites from the \code{toxocara} data set).}
   \label{plotdiscfitP}
@@
 <<fittoxocara.poisnbinom.echo, echo=TRUE, fig=FALSE>>=
-cdfcomp(list(ftoxo.P,ftoxo.nb),
-legendtext=c("Poisson", "negative binomial"))
+cdfcomp(list(ftoxo.P, ftoxo.nb), 
+  legendtext = c("Poisson", "negative binomial"))
 @
@@
 <<fittoxocarapoisnbinomplot, echo=FALSE, results=hide, fig=TRUE, width=4, height=4, eps=FALSE>>=
-cdfcomp(list(ftoxo.P,ftoxo.nb),
-        legendtext=c("Poisson", "negative binomial"))
+cdfcomp(list(ftoxo.P, ftoxo.nb), 
+        legendtext = c("Poisson", "negative binomial"))
 @
   \caption{Comparison of the fits of a negative binomial and a Poisson distribution
-   to numbers of \emph{Toxocara cati} parasites from the \code{toxocara} data set).}
+   to numbers of \emph{Toxocara cati} parasites from the \code{toxocara} data set.}
   \label{plotdiscfit}
@@
 <<fittoxocara.poisnbinom.gof, echo=TRUE, fig=FALSE>>=
-gofstat(list(ftoxo.P,ftoxo.nb),
-        fitnames = c("Poisson","negative binomial"))
+gofstat(list(ftoxo.P, ftoxo.nb), 
+  fitnames = c("Poisson", "negative binomial"))
 @
@@
 %\bibliographystyle{plain}
-\bibliography{fitdistrplus,paperscitingfitdistrplus}
+\bibliography{jssfitdistrplus}
 

### 2014-09-19 pkg/mc2d/R/mcstoc.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@285 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -219,9 +219,14 @@ mcstoc <- function(func=runif, type=c("V","U","VU","0"), ..., nsv=ndvar(), nsu=n
 
     if(largsd != 0){
       argsd <- mapply(LAFUNC, argsd, typemc, SIMPLIFY=FALSE)
-#print(argsd)
+      #print(argsd)
       }
-
+#####################################
+# If lhs or rtrunc, redefine the function to draw random variables 
+#########
+  #keep a copy of original function
+  funcorigin <- func
+  
   if(lhs || rtrunc){                                             #define good function for the random sampling
     distr <- as.character(match.call()$func)                     #retrieve the name of the function
     distr <- substr(distr, 2, 1000)                              #remove the r
@@ -238,13 +243,13 @@ mcstoc <- function(func=runif, type=c("V","U","VU","0"), ..., nsv=ndvar(), nsu=n
           lmax <- max(length(linf),length(lsup))
           if(any(rep(linf, length.out = lmax) >= rep(lsup, length.out = lmax))) stop("linf should be < lsup")  #recycle vectors
           argsd$linf <- argsd$lsup <- argsd[[nsample]] <- NULL
-          
+          #find the p of the limit
           pinf <- as.vector(do.call(pfun,c(list(q=linf),argsd),quote=TRUE))
           psup <- as.vector(do.call(pfun,c(list(q=lsup),argsd),quote=TRUE))
-
+          #sample uniformely between the limits
           if(!lhs) lesp <- runif(nnfin,min=pinf,max=psup)
           else     lesp <- lhs(distr="runif", nsv=dimf[1], nsu=dimf[2], nvariates=dimf[3], min=pinf, max=psup) 
-
+          #get the q
           data <- do.call(qfun,c(list(p=lesp),argsd))
           data[pinf==0 & data > lsup] <- NaN          #ex: rtrunc("lnorm",10,linf=-2,lsup=-1)
           data[psup==1 & data < linf] <- NaN          #ex: rtrunc("unif",10,linf=2,lsup=4,max=1)
@@ -253,28 +258,31 @@ mcstoc <- function(func=runif, type=c("V","U","VU","0"), ..., nsv=ndvar(), nsu=n
           #Two tests for extreme situations. None Catch all possibilities. THe error is first to avoid the warning
           if(any(data <= linf | data > lsup, na.rm=TRUE)) stop("Error in rtrunc: some values are not in the expected range (maybe due to rounding errors)")
[…]

### 2014-09-19 pkg/mc2d/R/mcstoc.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@286 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -250,7 +250,7 @@ mcstoc <- function(func=runif, type=c("V","U","VU","0"), ..., nsv=ndvar(), nsu=n
           if(!lhs) lesp <- runif(nnfin,min=pinf,max=psup)
           else     lesp <- lhs(distr="runif", nsv=dimf[1], nsu=dimf[2], nvariates=dimf[3], min=pinf, max=psup) 
           #get the q
-          data <- do.call(qfun,c(list(p=lesp),argsd))
+          data <- (do.call(qfun,c(list(p=lesp),argsd)))[1:nnfin]
           data[pinf==0 & data > lsup] <- NaN          #ex: rtrunc("lnorm",10,linf=-2,lsup=-1)
           data[psup==1 & data < linf] <- NaN          #ex: rtrunc("unif",10,linf=2,lsup=4,max=1)
           data[is.na(linf) | is.na(lsup)] <- NaN      #ex: rtrunc("norm",10,sd=-2)

### 2017-07-27 share/summary-NPMLE.Rmd
add new files for NPMLE
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@517 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

---
title: "On the estimation of cdf/sf for censored data"
author: "Christophe Dutang"
date: "26/07/2017"
header-includes:
   - \usepackage{amsmath,amsfonts}
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Non-censored data
Consider observations $x_1,\dots,x_n$ as realizations of random variables $X_1,\dots,X_n$.
The empirical cdf is defined as
$$
F_n(x) = \frac{1}{n}\sum_{i=1}^n 1_{X_i\leq x}
$$
In particular $F_n(x_j)=R_j/n$ where $R_j$ is the rank of the $j$th observation.
The empirical survival function  is defined as
$$
S_n(x) = \frac{1}{n}\sum_{i=1}^n 1_{X_i> x}
$$
```{r}
xorig <- rexp(25)
plot(ecdf(xorig))
```

# Right-censored data
Consider observations $t_1,\dots,t_n$ and censoring dummies $d_1,\dots,d_n$ such that
$$
d_i=0\Rightarrow t_i = x_i\leq c_i,
d_i=1\Rightarrow t_i = c_i < x_i
$$
where $c_i$'s are censored values. In other words, when $d_i=1$, we know that $x_i>c_i$.
This can be represented as interval values
$$
d_i=0\Rightarrow [t_i,t_i],
d_i=1\Rightarrow (t_i, +\infty).
$$
The survival function is estimated (generally) by Kaplan-Meier estimator
$$
S_n(x) = \prod_{i, t_i < x} \frac{n_i-d_i}{n_i},
$$
where $n_i$ is the number of at-risk individuals at time $t_i$ and $d_i$ 
the number of non-censored deaths at $t_i$.
We get back to usual estimation of previous section if there is no censoring.

Consider a distribution function $F$ with density/mass proba function $f$.
The likelihood function to observe $(t_i,d_i)_i$ is
$$
L(F,f) = \prod_{i=1}^nf(t_i)^{1-d_i} (1-F(t_i))^{d_i}.
$$
Kaplan-Meier is known to maximize this likelihood function and is called
sometimes NPMLE, non parametric ML Estimator.
In R, we can do this via `Surv` in `survival` package.

# Left-censored data
Consider observations $t_1,\dots,t_n$ and censoring dummies $d_1,\dots,d_n$ such that
$$
d_i=0\Rightarrow t_i = x_i\geq c_i,
d_i=1\Rightarrow t_i = c_i > x_i
$$
where $c_i$'s are censored values. In other words, when $d_i=1$, we know that $x_i<c_i$.
This can be represented as interval values
$$
d_i=0\Rightarrow [t_i,t_i],
d_i=1\Rightarrow (0,t_i).
$$

# Interval-censored data

Source: https://www.r-project.org/conferences/useR-2010/tutorials/Fay_1.pdf

Consider observations $(l_i,r_i]$  such that the true observation $x_i$ lies 
within this interval.
If there is no censoring, we observe $[l_i,l_i]$.
Right-censoring is obtained with $u_i=\infty$ and left-censoring with $l_i=0$.
The likelihood function for cdf $F$ is
$$
L(F) = \prod_{i=1}^n(F(r_i)-F(l_i)).
$$
Consider $t_0=0<t_1<\dots<t_m<t_{m+1}=\infty$ with $t_1,\dots,t_m$ unique 
observations times among $l_i,r_i$'s values.
We want to estimate $F(t_i)-F(t_{i-1})=p(t_i)$ with observations.

The NPMLE is the Turnbull EM algorithm given below
1. Initialize probability vector estimating $p(t_i)$
$$
\widehat{p} =(\widehat{p}_0,\dots,\widehat{p}_{m+1}) 
= (\frac{1}{m+1},\dots,\frac{1}{m+1})
$$
2. Expectation step: Iterate for all $i$, we 
estimate $P(t_{j-1}<X_i\leq t_j \vert l_i<X_i\leq r_i)$ by 
$$
\widehat{p}_{i}(t_j) \leftarrow \frac{\widehat{p}_j 1_{t_j \in (l_i,r_i]}}{\sum_{k, t_k \in (l_i,r_i]}\widehat{p}_k}, j=0,\dots,m+1
$$
We obtain a vector $\widehat{p}_i$ for all individuals.  
3. Maximization step: average over individuals
$$
\widehat{p} \leftarrow 1/n \sum_{i=1}^n\widehat{p}_i.
$$
4. If not converge, go to step 2.

The estimator of the cdf is the cumsum of $\widehat p$.
If there is no censoring, the binary dummy $1_{t_j \in (l_i,r_i]$ equals
one only once, and $\widehat{p}_i$ has a single one component leading to 
$\widehat{p}=(1/n,\dots,1/n)$.
In R, we can do this via `Surv` in `survival` package.

### 2017-07-27 share/compare_NPMLE.R

@@ -1,0 +1,112 @@
+require(fitdistrplus)
+require(interval)
+require(npsurv)
+require(Icens)
+options(digits = 3)
+
+# par(mfrow = c(2,2))
+# d1 Exemple de la publi de Guolin sur lequel npsurv donne un résultat différent de tous les autres
+d1 <- data.frame(left = c(1, 2, 3, 4, 3, 7), right = c(2, 5, 3, 7, 8, 9))
+# d3 Exemple où toute les méthodes diffèrent et où on voit bien le problème des masses proches de 0
+d3 <- data.frame(left = c(-1.4, 1.18, -1.4, 2, -1.4, 0),
+                 right = c(1, 1.18, 2, NA, 0, 2))
+# d4 exemple sur lequel toutes les méthodes s'accordent sauf Turnbull de survival et PGM de Icens
+data(fluazinam)
+d4 <- as.data.frame(log10(fluazinam))
+# d5 exemple sans censure sur lequel toutes les méthodes font ce qu'il faut sauf PGM de Icens
+r <- rnorm(20)
+d5 <- data.frame(left = r, right = r)
+# d6 exemple de Fay 2010 sur lequel toutes les méthodes s'accordent sauf Turnbull de survival (avec très petites masses non à 0)
+data(bcos)
+d6 <- subset(bcos, treatment == "Rad")[,1:2]
+d6$right[is.infinite(d6$right)] <- NA
+# d7 exemple avec censures à droite
+data(salinity)
+d7 <- log10(salinity)
+
+
+listd <- list(d1 = d1,  d3 = d3,  d6 = d6, d7 = d7)
+for (i in 1:4)
+{
+  d <- listd[[i]]
+  ## transformation des NA en -Inf ou Inf necessaire pour certains packages
+  db <- d 
+  db$left[is.na(db$left)] <- -Inf
+  db$right[is.na(db$right)] <- Inf
+  
+  e1 <- VEM(db) # A Vertex-exchange-method in $D$-optimal Design Theory , D. Bohning, Metrika, 1986.
+  e1$numiter
+  # trial to analyse the difference with d1
+  e1bis <- VEM(db, tol = 1e-10, tolbis = 1e-10) # A Vertex-exchange-method in $D$-optimal Design Theory , D. Bohning, Metrika, 1986.
+  e1bis$numiter
+  
+  # plot(e1)
+  e2 <- EM(db) # Turnbull 1976
+  # plot(e2)
[…]

### 2017-07-27 share/functions4fitdistrplusv3.R
add new files for NPMLE
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@517 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,284 @@
+require(npsurv)
+require(fitdistrplus)
+#require(interval)
+options(digits = 3)
+################## CHOIX du JEU DE DONNEES ###########################################
+# exemple avec uniquement censures par intervalle
+# cf. papier Guolin 2008
+# DATA 1 : npsurv ne converge pas vers les mêmes si que survival et interval
+d1 <- data.frame(left = c(1, 2, 3, 4, 3, 7), right = c(2, 5, 3, 7, 8, 9))
+
+# exemple avec tous types de censures et censures à gauche mises à très faible valeur
+# pour que icfit tourne
+# DATA 2 npsurv ne visualise pas les valeurs négatives (il suppose qu'on est sur de la survie)
+# mais les calcule - sur cet exemple survival, interval et npsurv sont cohérents
+data(smokedfish)
+d2 <- as.data.frame(log10(smokedfish))
+# minleft  <- min(d2$left[!is.na(d2$left)])
+# d2$left[is.na(d2$left)] <- minleft - 1
+
+
+# exemple avec censures à droite uniquement
+# exemple où les algos de Turnbull de survival et du package interval diffèrent un chouilla
+# DATA 4 survival, interval et npsurv sont cohérents à un chouilla près
+data(fluazinam)
+d4 <- as.data.frame(log10(fluazinam))
+
+## Exemple sans censure
+# DATA 5
+r <- rnorm(20)
+d5 <- data.frame(left = r, right = r)
+
+
+# exemple de Fay (JSS)
+# DATA 6 survival, interval et npsurv sont cohérents
+#data(bcos)
+#d6 <- subset(bcos, treatment == "Rad")[,1:2]
+#d6$right[is.infinite(d6$right)] <- NA
+
+# exemple salinity
+data(salinity)
+d7 <- log10(salinity)
+
+################# FONCTIONS ########################
+#### Function to plot the empirical CDF from censored data ###################
+CDFplotcens <- function(d, plot = TRUE, add.Turnbull = TRUE)
+# if add.Turnbull == TRUE add of the plot provided by survival
+# output : vectors with Q and corresponding P values  for GOF stat
+{
+  db <- d
+  db$left[is.na(db$left)] <- -Inf
[…]

### 2017-07-28 share/NPMLE.Rmd
A summary of what I did on NPMLE, to assemble with the file written by Christophe
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@518 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

---
title: A point on the nonparametric maximum likelihood estimation (NPMLE) of the empirical
  cumulative distribution function (ECDF) for censored data
author: "Marie Laure Delignette-Muller"
date: "21 juillet 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 3)
require(fitdistrplus)
require(Icens)
require(interval)
require(npsurv)
```

# NPMLE : a difficult problem

In the current version of our package, the empirical cumulative distribution function (ECDF) is estimated
using the Turnbull algorithm as implemented in the survival package. This estimation was shown to
be problematic (see Gentleman and Geyer 1994,
Fay 2010, ...) and indeed is not directly usable to build PP plots or QQ plots.
In nonparametric maximum likelihood estimation (NPMLE) algorithms, the firts step is the definition of the set of disjoint intervals (some with null width)
under which the ECDF may change : this set is called the **maximal intersection intervals** (Wang 2008) or the 
maximal cliques of the data (Liu 2011) or the equivalence classes (Guolin 2008) or the Turnbull
intervals or the innermost intervals (Fay 2010). These maximal intersection intervals are defined by each left bounds immediately
followed by a right bounds once the bounds are ordered. Interval of null width correspond to values of non 
censored data or to a left bound exactly at the same value of a rigth bound.

To understand this definition let us take a simple example with left, right, interval censored data and non censored data. In the following we can visualize the bounds as a black plain line for left bound and a green dotted line for right bound 
(non censored data coded with the same value for left and right bound were plotted in pink plain lines) added at the Turnbull
plot provided in the survival package.

```{r}
d1 <- data.frame(left = c(1, 2, 3, 4, 3, 7), right = c(2, 5, 3, 7, 8, 9))
d <- d1
(survdata <- Surv(time = d$left, time2 = d$right, type="interval2"))
survfitted <- survfit(Surv(left, right, type="interval2") ~ 1, data = d)
plot(survfitted$time, 1- survfitted$surv, type = "s", col = "red", xlim = c(min(d$left), max(d$right)))
points(survfitted$time, 1 - survfitted$surv, col = "red", pch = 16)
abline(v = d$left, col = "black")
abline(v = d$right, col = "green", lty = 2)
noncensored <- d$left[d$left == d$right]
abline(v = noncensored, col = "pink")
```

The second step of an NPMLE algorithm aims at estimating the **mass of each maximal intersection interval**. 
Many algorithms were proposed for this purpose and it seems to be a tricky problem. The result of the algorithm 
is thus a list of **maximal intersection intervals with their corresponding mass** from which it i easy to compute an ECDF.
The Turnbull algorithm implemented in the survival package does not provide the intervals but only the survival function
at each mid-point of an interval. Latest algorithms provided in R packages Icens, interval and npsurv provide the intervals
and propose to represent by rectangles the areas where the function (survival or ECDF) is not unique. This type of
representation (representational non-uniqueness) was
introduced by Gentleman and Vandal (2002). Let us compare the representation provided by the survival package (red line)
and the one provided by the Icens package for the Turnbull algorithm on the previous example.

```{r}
fit <- EM(d)
plot(fit)
lines(survfitted$time, 1- survfitted$surv, type = "s", col = "red", xlim = c(min(d$left), max(d$right)))
points(survfitted$time, 1 - survfitted$surv, col = "red", pch = 16)
abline(v = d$left, col = "black")
abline(v = d$right, col = "green", lty = 2)
noncensored <- d$left[d$left == d$right]
abline(v = noncensored, col = "pink")
```

The package interval gives the same type of representation with a linear interpolation within each
rectangle but only as a survival curve (it is not possible to plot it directly in ECDF).

```{r}
fit <- icfit(Surv(left, right, type ="interval2") ~ 1, data = d)
plot(fit)
abline(v = d$left, col = "lightgrey", lty = 2)
abline(v = d$right, col = "lightgrey", lty = 2)

```

The package npsurv also gives the same type of representation in survival plot with different options of plot.

```{r}
par(mfrow = c(2, 2))
par(mar = c(4,4,0.5,0.5))
f <- npsurv(d, verb = 0)$f
plot(f, style = "box")
abline(v = d$left, col = "lightgrey", lty = 2)
abline(v = d$right, col = "lightgrey", lty = 2)
plot(f, style = "uniform")
abline(v = d$left, col = "lightgrey", lty = 2)
abline(v = d$right, col = "lightgrey", lty = 2)
plot(f, style = "midpoint")
abline(v = d$left, col = "lightgrey", lty = 2)
abline(v = d$right, col = "lightgrey", lty = 2)

```

The different algorithms differ in their estimation of the mass corresponding to each maximal intersection interval. To visualize that we will use the following function on different data sets.

```{r}
compareNPMLE <- function(d)
{
  # d is a data set with censored values coded by NA as in fitdistrplus
  ## transformation of NA in -Inf or Inf necessary for other packages than survival
  db <- d 
  db$left[is.na(db$left)] <- -Inf
  db$right[is.na(db$right)] <- Inf

  e1 <- VEM(db) # A Vertex-exchange-method in $D$-optimal Design Theory , D. Bohning, Metrika, 1986.
  e1$numiter
  # trial to analyse the difference with d1
  e1bis <- VEM(db, tol = 1e-10, tolbis = 1e-10) # A Vertex-exchange-method in $D$-optimal Design Theory , D. Bohning, Metrika, 1986.
  e1bis$numiter

  # plot(e1)
  e2 <- EM(db) # Turnbull 1976
  # plot(e2)
  e3 <- PGM(db) # Some Algorithmic Aspects of the Theory of Optimal Designs, C.-F. Wu, 1978, Annals.
  # plot(e3)
  e4 <- EMICM(db) # A hybrid algorithm for computation of the nonparametric maximum likelihood 
  # estimator from censored data, J. A. Wellner and Y. Zhan, 1997, JASA.
  # plot(e4)
  #e5 <- ISDM(db) # An Algorithm for Computing the Nonparametric MLE of a Mixing Distribution, 
  # Lesperance, Mary L. and Kalbfleisch, John D., JASA, 1992
  # plot(e5)
  # Methode trop lente sur les gros jeux de donnes !!!!!!!!!!!!!!!!!!!!

  mass.Icens.VEM <- e1$pf
  mass.Icens.EM <- e2$pf
  mass.Icens.PGM <- e3$pf
  mass.Icens.EMICM <- e4$pf
  #mass.Icens.ISDM <- e5$pf

  # pour avoir les bornes des intervalles d'equivalence
  # e1$intmap

  # comparatif avec algo de npsurv
  f <- npsurv(db, verb = 0)$f
  mass.npsurv <- f$p
  # pour avoir les bornes des intervalles d'equivalence
  # rbind(f$left, f$right)

  # npsurv(db, verb = 0, tol = 1e-10)$numiter
  # npsurv(db, verb = 0, tol = 1e-12)$f

  # comparatif avec Turnbull de survival
  survdata <- Surv(time = d$left, time2 = d$right, type="interval2")
  survfitted <- survfit(Surv(left, right, type="interval2") ~ 1, data = d)
  s <- survfitted$surv
  ns <- length(s)
  savant <- c(1, s[-ns])
  mass.survival <- savant - s

  # comparatif avec icfit
  fit <- icfit(Surv(left, right, type ="interval2") ~ 1, data = db)
  mass.interval <- fit$pf
  names(mass.interval) <- NULL

  print("mass.Icens.VEM")
  print(mass.Icens.VEM)
  print("mass.Icens.EM")
  print(mass.Icens.EM)
  print("mass.Icens.PGM")
  print(mass.Icens.PGM)
  print("mass.Icens.EMICM")
  print(mass.Icens.EMICM)
#  mass.Icens.ISDM
  print("mass.survival")
  print(mass.survival)
  print("mass.interval")
  print(mass.interval)
  print("message of icfit concerning Kuhn-Tucker conditions")
  print(fit$message)
  print("mass.npsurv")
  print(mass.npsurv)

  par(mfrow = c(2,2))
  par(mar = c(4,4,0.5,0.5))
  plot(e1, surv = TRUE, main = "")
#  points(survfitted$time, survfitted$surv, col = "red", pch = 16)
  lines(survfitted$time, survfitted$surv, type = "s", col = "red", lty = 2)
  legend("bottomleft", legend = "VEM(Icens) in green and survival in red", bty = "n")

  plot(f)
#  points(survfitted$time, survfitted$surv, col = "red", pch = 16)
  lines(survfitted$time, survfitted$surv, type = "s", col = "red", lty = 2)
  legend("bottomleft", legend = "npsurv in blue and survival in red", bty = "n")

  plot(fit)
#  points(survfitted$time, survfitted$surv, col = "red", pch = 16)
  lines(survfitted$time, survfitted$surv, type = "s", col = "red", lty = 2)
  legend("bottomleft", legend = "interval in grey and survival in red", bty = "n")
}
```

Let us use this function on the previous example :

```{r, fig.width = 8, fig.height = 5}
compareNPMLE(d)
```

 We can see that the constrained Newton method (CNM) implemented in the npsurv package
 gives a different mass estimation than the other methods. Now let us look at the differences 
 between algorithms on various data sets.

```{r, fig.width = 8, fig.height = 5}
d3 <- data.frame(left = c(-1.4, 1.18, -1.4, 2, -1.4, 0),
                 right = c(1, 1.18, 2, NA, 0, 2))
compareNPMLE(d3)
```

In this example we can see differences between the algorithms by the way they polish the ECDF
function, by putting some masses to zero.

```{r, fig.width = 8, fig.height = 5}
data(bcos)
d6 <- subset(bcos, treatment == "Rad")[,1:2]
d6$right[is.infinite(d6$right)] <- NA
compareNPMLE(d6)
```

In this example,  
differences are small and not visible on the plots.

```{r, fig.width = 8, fig.height = 5}
data(salinity)
d7 <- log10(salinity)
compareNPMLE(d7)
```

In this example, differences are stronger, and appear also between packages interval and Icens.

**I did not understand all about the check of the Kuhn-Tucker conditions (see Gentleman and Geyer) but
I would like to ask Wang if this condition is checked in npsurv and if not if it is a problem.**

Using the npsurv package, I have implemented functions to plot the ECDF with rectangles. I could not use 
the plot function proposed in npsurv as it plots only positive values, considering the variable is a survival time, and dose not offer the possibility to plot an ECDF function instead of a survival function.
Below is the new plot fucntion that could be included in the plotdistcens function of fitdistrplus.

```{r}
CDFplotcens <- function(d, plot = TRUE, add.Turnbull = TRUE)
# if add.Turnbull == TRUE add of the plot provided by survival
# output : vectors with Q and corresponding P values  for GOF stat
{
  db <- d
  db$left[is.na(db$left)] <- -Inf
  db$right[is.na(db$right)] <- Inf
  f <- npsurv(db)$f
  bounds <- c(f$right, f$left)
  finitebounds <- bounds[is.finite(bounds)]
  upper <- max(finitebounds)
  lower <- min(finitebounds)
  width <- upper - lower
  xlimup <- upper + width * 0.2 
  xlimlow <- lower - width * 0.2
  xlimupinf <- upper + width * 100 
  xlimlowinf <- lower - width * 100
  ## mettre la possibilité de rentrer des xlim et ylim par l'utilisateur bien sûr
  k <- length(f$left)
  Fnpsurv <- cumsum(f$p) # DONNER UN AUTRE NOM CAR RACCOURCI DE FALSE !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

  ## calul des points points pour Q et P dnas les GOF stat et graph
  Fbefore <- c(0, Fnpsurv[-k])
  Fmid <- (Fbefore + Fnpsurv)/2
  df <- data.frame(left = f$left, right = f$right, Fmid = Fmid)
  # restriction to equivalence intervals without Inf
  dfb <- subset(df, (left != - Inf) & (right != Inf) ) 
  dfb$xmid <- (dfb$left + dfb$right) / 2

  if (plot)
  {
    ### tracé de la ligne à droite des zones indéterminées
    dright <- c(f$left[1], rep(f$right, rep(2,k)), f$right[k]) 
    Fright <- rep(c(0,Fnpsurv), rep(2,k+1))
    plot(dright, Fright, type = "l", xlim = c(xlimlow, xlimup), xlab = "x", ylab = "F")
    ### tracé de la ligne à gauche des zones indéterminées
    dleft = rep(c(f$left,f$right[k]), rep(2,k+1))
    Fleft = c(0,rep(Fnpsurv, rep(2,k)),1)
    lines(dleft, Fleft)
    ### Ajout des rectangles
    Fc <- c(0, Fnpsurv, 1)
    for(i in 1:k) {
      if(f$right[i] - f$left[i] > 0) 
        rect(f$left[i], Fc[i], f$right[i], Fc[i+1], border = "black",
             col = "lightgrey")
    }
    abline(h = 0, lty = 2, col = "grey")
    abline(h = 1, lty = 2, col = "grey")
    # ajout des rectangles à droite et à gauche si besoin
    if(f$right[k] == Inf) rect(f$left[k], Fnpsurv[k - 1], xlimupinf, 1, border = "black",
                               col = "lightgrey")
    if(f$left[1] == - Inf) rect(xlimlowinf, Fnpsurv[1], f$right[1], 0, border = "black",
                                col = "lightgrey")

    #### Ajout des points qu'on prendrait en compte pour les GOF graph et stat
    points(dfb$xmid, dfb$Fmid, pch = 16, col = "blue")

    #### Ajout de l'ancienne présentation pour voir
    if (add.Turnbull)
    {
      (survdata <- Surv(time = d$left, time2 = d$right, type="interval2"))
      survfitted <- survfit(Surv(left, right, type="interval2") ~ 1, data = d)
      lines(survfitted$time, 1- survfitted$surv, type = "s", col = "red")
    }
  }
  return(list(Qi = dfb$xmid, Pi = dfb$Fmid))
}

```

Let us use test it on different data set  including one (d2) with left censored values
with which it was not possible to use icfit (icfit dosen not admit values at  -Inf).
The Turnbull plot provided by survival as added just compare it to the new representation.

```{r}
data(smokedfish)
d2 <- as.data.frame(log10(smokedfish))
data(fluazinam)
d4 <- as.data.frame(log10(fluazinam))
par(mfrow = c(2,2))
par(mar = c(4, 4, 0.5, 0.5))
listd <- list(d1= d1, d2 = d2, d4 = d4, d7 = d7)
for (i in 1:length(listd))
{
  CDFplotcens(listd[[i]], add.Turnbull = TRUE)  
}
```

At last, I have implemented, for the moment in a unique function, 
code that could be used in fitdistrplus in cdfcompcens and in new functions
qqcompcens and ppcompcens.

```{r}
#### Function to plot GOF graphs from a fit on censored data ########################
#### It will certainly be difficult to plot more than one fit on the same plot ######
#### as it is proposed for non censored data ########################################
plotfitcens <- function(fti, CDFplot = TRUE, PPplot = TRUE, QQplot = TRUE, add.midpoints = TRUE)
  # output : vectors with Q and corresponding P values  for GOF stat
{
  d <- fti$censdata
  db <- d
  db$left[is.na(db$left)] <- -Inf
  db$right[is.na(db$right)] <- Inf
  f <- npsurv(db)$f
  bounds <- c(f$right, f$left)
  finitebounds <- bounds[is.finite(bounds)]
  upper <- max(finitebounds)
  lower <- min(finitebounds)
  width <- upper - lower
  xlimup <- upper + width * 0.2 
  xlimlow <- lower - width * 0.2
  xmax <- xlimup # pour le plots ensuite
  xmin <- xlimlow
  xlimupinf <- upper + width * 100 
  xlimlowinf <- lower - width * 100
  ## mettre la possibilité de rentrer des xlim et ylim par l'utilisateur bien sûr
  k <- length(f$left)
  Fnpsurv <- cumsum(f$p)

  ## calul des points points pour Q et P dans les GOF stat et graph
  Fbefore <- c(0, Fnpsurv[-k])
  Fmid <- (Fbefore + Fnpsurv)/2
  df <- data.frame(left = f$left, right = f$right, Fmid = Fmid)
  dfb <- subset(df, (left != - Inf) & (right != Inf) ) # without Inf
  dfb$xmid <- (dfb$left + dfb$right) / 2

  Qi <- dfb$xmid # dim k if no Inf, k-1 or k-2
  Pi <-  dfb$Fmid

  # Definition of vertices of each rectangle
  Qi.left <- df$left # dim k
  Qi.left4plot <- Qi.left
  if (Qi.left4plot[1] == - Inf) Qi.left4plot[1] <- xlimlowinf
  Qi.right <- df$right
  Qi.right4plot <- Qi.right
  if (Qi.right4plot[k] == Inf) Qi.right4plot[k] <- xlimupinf
  Pi.low <- Fbefore
  Pi.up <- Fnpsurv

  # recuperation des parametres ajustees et du nom de la distribution ajustee
  para <- c(as.list(fti$estimate), as.list(fti$fix.arg))
  distname <- fti$distname
  pdistname <- paste("p", distname, sep="")
  qdistname <- paste("q", distname, sep="")

  if (CDFplot)
  {
    plot(0, 0, type = "n", xlim = c(xlimlow, xlimup), ylim = c(0,1), xlab = "x", ylab = "F")
    for(i in 1:k) {
        rect(xleft = Qi.left4plot, ybottom = Pi.low, xright = Qi.right4plot, ytop = Pi.up, 
             border = "black", col = "lightgrey")
    }
    abline(h = 0, lty = 2, col = "grey")
    abline(h = 1, lty = 2, col = "grey")

    ### Add of the two lines (only one is need if the rectangles are plotted)
    # the one at right
    dright <- c(f$left[1], rep(f$right, rep(2,k)), f$right[k]) 
    Fright <- rep(c(0,Fnpsurv), rep(2,k+1))
    lines(dright, Fright, type = "l", xlim = c(xlimlow, xlimup), xlab = "x", ylab = "F")
    ### the one at left
    dleft = rep(c(f$left,f$right[k]), rep(2,k+1))
    Fleft = c(0,rep(Fnpsurv, rep(2,k)),1)
    lines(dleft, Fleft)

    #### Add of the mid points
    if (add.midpoints) points(Qi, Pi, pch = 16, col = "black")

    ### Ajout de la courbe théo ajustée
    sfin <- seq(xmin, xmax, by=(xmax-xmin)/100)
    theopfin <- do.call(pdistname, c(list(q=sfin), as.list(para)))
    lines(sfin, theopfin, col = "red")
  } ############################# END CDFplot

  if (PPplot)
  {
    plot(0, 0, type = "n", main = "PPplot", xlim = c(0,1), ylim = c(0, 1))

    # plot of rectangles
    Pitheo.low <- do.call(pdistname, c(list(q=Qi.left), as.list(para)))
    Pitheo.up <- do.call(pdistname, c(list(q=Qi.right), as.list(para)))
    rect(xleft = Pitheo.low, ybottom = Pi.low, xright = Pitheo.up, ytop = Pi.up, 
           border = "black", col = "lightgrey")
    abline(0,1)

    # add of mid moints
    if (add.midpoints)
    {
      Pitheo <- do.call(pdistname, c(list(q = Qi), as.list(para)))
      points(Pitheo, Pi, pch = 16, col = "black")
    }
  } ############# END of PPplot

  if (QQplot)
  {
    plot(0, 0, type = "n", main = "QQplot", xlim = c(xmin, xmax), ylim = c(xmin, xmax))

    # plot of rectangles
    Qitheo.left <- do.call(qdistname, c(list(p = Pi.low), as.list(para)))
    Qitheo.right <- do.call(qdistname, c(list(p = Pi.up), as.list(para)))
    Qitheo.left4plot <- Qitheo.left
    if (Qitheo.left4plot[1] == - Inf) Qitheo.left4plot[1] <- xlimlowinf
    Qitheo.right4plot <- Qitheo.right
    if (Qitheo.right4plot[k] == Inf) Qitheo.right4plot[k] <- xlimupinf
    rect(xleft = Qitheo.left4plot, ybottom = Qi.left4plot, xright = Qitheo.right4plot, ytop = Qi.right4plot, 
           border = "black", col = "lightgrey")
    abline(0,1)

    # add of mid moints
    if (add.midpoints)
    {
      Qitheo <- do.call(qdistname, c(list(p = Pi), as.list(para)))
      points(Qitheo, Qi, pch = 16, col = "black")
    }
  }
}
```

Look at those new plots on fits of a normal distribution on different data sets.

```{r, fig.width = 8, fig.height = 5}
d <- d2
f <- fitdistcens(d, "norm") 
par(mfrow = c(2,2))
par(mar = c(4,4,2, 0.5))
plotfitcens(f)

d <- d4
f <- fitdistcens(d, "norm")
plotfitcens(f)

d <- d7
f <- fitdistcens(d, "norm")
plotfitcens(f)

```

The last tested data set is a non censored data set on which we can compare plots to those provided using cdfcomp, ppcomp and qqcomp

```{r, fig.width = 8, fig.height = 5}
set.seed(1234) 
r <- rnorm(20)
d5 <- data.frame(left = r, right = r)
d <- d5
par(mfrow = c(2, 3))
par(mar = c(4,4,2, 0.5))
f <- fitdistcens(d, "norm")
plotfitcens(f)

fbis <- fitdist(d$left, "norm")
cdfcomp(fbis)
ppcomp(fbis)
qqcomp(fbis)

```

### 2018-07-03 share/Delignette_slides4R2108-v2.Rmd
presentation for R 2018
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@563 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

---
title: "New goodness-of-fit plots for censored data in the package **fitdistrplus**"
author: |
 | M.L. Delignette-Muller (LBBE, Lyon)
 | C. Dutang (CEREMADE, Paris) and 
 | A. Siberchicot (LBBE, Lyon)
date: "July 6th 2018"
output:
  beamer_presentation:
    theme: "Antibes"
    toc: true
    fig_width: 8
    fig_height: 6
    fig_caption: false
---
```{r, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
options(digits = 3)
require(fitdistrplus)
set.seed(123)
source("plotdistcensmod4R2018.R")
addbounds <- function(d)
{
  xbounds <- c(d$left, d$right)
  xboundsnotNA <- xbounds[!is.na(xbounds)]
  abline(v = xboundsnotNA, col = "grey")
}
addTurnbull <- function(d, conf.int = FALSE)
{
  (survdata <- Surv(time = d$left, time2 = d$right, type="interval2"))
  survfitted <- survfit(Surv(left, right, type="interval2") ~ 1, data = d)
  points(survfitted$time, 1 - survfitted$surv, col = "orange", type = "s", 
         lwd = 5, lty = 2)
  if (conf.int == TRUE)
  {
      points(survfitted$time, 1 - survfitted$lower, col = "orange", type = "s", 
             lwd = 2, lty = 3)
      points(survfitted$time, 1 - survfitted$upper, col = "orange", type = "s", 
             lwd = 2, lty = 3)
  }
}
```

# Goodness-of-fit plots for non censored data

## General presentation of the package fitdistrplus
A package to help the **fit of parametric distributions** to 
univariate discrete our continuous data.

* stable version 1.0-9 on **CRAN** (first release in 2009): [https://cran.r-project.org/web/packages/fitdistrplus](https://cran.r-project.org/web/packages/fitdistrplus)
* version 1.0-10 in development on **Rforge** (soon on CRAN): [https://rdrr.io/rforge/fitdistrplus/](https://rdrr.io/rforge/fitdistrplus/)
* Delignette-Muller, M. L., & Dutang, C. (2015). fitdistrplus: An R package for fitting distributions. **Journal of Statistical Software**, 64(4), 1-34. (311 citations in scholar google)

The package provides, among others, tools to
assess/compare the goodness-of-fit of various distributions on a same data set.

**Left, right and interval censored** continuous data can be handled.

We will present you
**new goodness-of-fit plots for such data** available in version 1.0-10.

## An example with non censored data
```{r}
r <- rweibull(100, shape = 3, scale = 1)
fw <- fitdist(r, "weibull")
fg <- fitdist(r, "gamma")
fl <- fitdist(r, "lnorm")
gofstat(list(fw, fg, fl), 
        fitnames = c("Weibull", "gamma", "lnorm"))
```

## Goodness-of-fit plot in density plot
```{r}
denscomp(list(fw, fg, fl), demp = TRUE, fitlty = 1)
```

## Goodness-of-fit plot in CDF plot
```{r}
cdfcomp(list(fw, fg, fl), fitlty = 1)
```

## The Q-Q plot which emphasizes differences at tails
```{r}
qqcomp(list(fw, fg, fl))
```

## The P-P plot which emphasizes differences in the center
```{r}
ppcomp(list(fw, fg, fl))
```

# Representation of the ECDF for censored data

## How to represent an ECDF from censored data ?

**A first toy example**
```{r, echo = FALSE}
d0 <- data.frame(left = c(NA, 2, 4, 7, 10), right = c(1, 2.5, 6, 8, NA))
d <- d0
```
```{r}
d
```
```{r, echo = FALSE, fig.height=2}
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue") 
```

## The CDF plot from the package survival
```{r, echo = FALSE,fig.height=5}
par(mar = c(3, 4, 0.1, 0.1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue")
addbounds(d)
plotdistcens(d, NPMLE.method = "Turnbull", lwd = 5, col = "orange")
addbounds(d)
```

The Turnbull plot implemented in the package survival and used in former versions 
of fitdistrplus.

## The Turnbull plot with confidence intervals
```{r, echo = FALSE}
par(mar = c(3, 4, 0.1, 0.1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue")
addbounds(d)
plotdistcens(d, NPMLE.method = "Turnbull", Turnbull.confint = TRUE, lwd = 5, col = "orange")
addbounds(d)
```

## A new plot from the package npsurv (Wang plot)
```{r, echo = FALSE}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue")
addbounds(d)
plotdistcens(d, NPMLE = TRUE, lwd = 5)
addTurnbull(d)
addbounds(d)
legend("right", lty = c(1,2), lwd = c(5,5), col = c("black", "orange"),
       legend = c("Wang", "Turnbull"), bty = "n", cex = 2)
```

## A second toy example with overlapping intervals
```{r, echo = FALSE}
d0b <- data.frame(left = c(NA, 2, 4, 6, 10), right = c(1, 3, 7, 8, NA))
d <- d0b
```
```{r, echo = FALSE}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue")
addbounds(d)
plotdistcens(d, NPMLE = TRUE, lwd = 5)
addTurnbull(d)
addbounds(d)
legend("right", lty = c(1,2), lwd = c(5,5), col = c("black", "orange"),
       legend = c("Wang", "Turnbull"), bty = "n", cex = 2)
```

## The two steps of an NPMLE algorithm (Non Parametric Maximum Likelihood Estimation)

* Identification of **equivalence classes** (also named **Turnbull intervals** or **maximal intersection intervals** or **innermost intervals** or **maximal cliques** of the data) 
= set of points/intervals under which the ECDF may change (each region between each left bound immediately followed by a right bound). The NPMLE is unique only up to these equivalence classes (non uniqueness represented by grey rectangles).

* Assign a **probability mass** to each equivalence class (may be 0).

Various algorithms implemented in the packages **Icens**, **interval** and **npsurv** (more or less performant and not all handling left censored data).

## A third toy example including non censored data
```{r, echo = FALSE}
d1 <- data.frame(left = c(1, 2, 3, 4, 3, 7), right = c(2, 5, 3, 7, 8, NA))
d <- d1
```
```{r, echo = FALSE}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue")
addbounds(d)
plotdistcens(d, NPMLE = TRUE, lwd = 5)
addTurnbull(d)
addbounds(d)
legend("right", lty = c(1,2), lwd = c(5,5), col = c("black", "orange"),
       legend = c("Wang", "Turnbull"), bty = "n", cex = 2)
```

## A 4th toy example
```{r, echo = FALSE}
d3 <- data.frame(left = c(-1.4, -1.4, 2, -1.4, 0),
                 right = c(1, 2, NA, 0, 2))
d <- d3
```
```{r, echo = FALSE}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue")
addbounds(d)
plotdistcens(d, NPMLE = TRUE, lwd = 5)
addTurnbull(d)
addbounds(d)
legend("right", lty = c(1,2), lwd = c(5,5), col = c("black", "orange"),
       legend = c("Wang", "Turnbull"), bty = "n", cex = 2)
```

## The 4th toy example with the add of a non censored obs.
```{r, echo = FALSE}
d3b <- data.frame(left = c(1.18, -1.4, -1.4, 2, -1.4, 0),
                 right = c(1.18, 1, 2, NA, 0, 2))
d <- d3b
```
```{r, echo = FALSE}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue")
addbounds(d)
plotdistcens(d, NPMLE = TRUE, lwd = 5)
addTurnbull(d)
addbounds(d)
legend("right", lty = c(1,2), lwd = c(5,5), col = c("black", "orange"),
       legend = c("Wang", "Turnbull"), bty = "n", cex = 2)
```

# New CDF, Q-Q and P-P plots implemented for censored data

## Use of cdfcompcens() to assess the fit of 3 distributions on data smokedfish
```{r, echo = FALSE}
data(smokedfish)
d <-  log10(smokedfish)
fitsfn <- fitdistcens(d,"norm")
fitsfl <- fitdistcens(d,"logis")
dgumbel <- function(x,a,b) 1/b*exp((a-x)/b)*exp(-exp((a-x)/b))
pgumbel <- function(q,a,b) exp(-exp((a-q)/b))
qgumbel <- function(p,a,b) a-b*log(-log(p))
fitsfg<-fitdistcens(d,"gumbel",start=list(a=-3,b=3))
```
```{r, echo = FALSE}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 2, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 0.15)
# addbounds(d)
cdfcompcens(list(fitsfn,fitsfl,fitsfg), fitlty = 1)
# addbounds(d)
```

## Use of qqcompcens() for one distribution
```{r, echo = FALSE, fig.height = 6, fig.width = 10}
par(mfrow = c(1, 2))
par(mar = c(4, 4, 2, 1))
cdfcompcens(fitsfn, fitlty = 1)
qqcompcens(fitsfn)
```

## Use of ppcompcens() for one distribution
```{r, echo = FALSE, fig.height = 6, fig.width = 10}
par(mfrow = c(1, 2))
par(mar = c(4, 4, 2, 1))
cdfcompcens(fitsfn, fitlty = 1)
ppcompcens(fitsfn)
```

## Q-Q plots and P-P plot for the 3 distributions
```{r, echo = FALSE, fig.height = 6, fig.width = 10}
par(mfrow = c(1, 2))
par(mar = c(4, 4, 2, 1))
qqcompcens(list(fitsfn,fitsfl,fitsfg))
ppcompcens(list(fitsfn,fitsfl,fitsfg))
```

## An alternative presentation of the Q-Q plot for the 3 dist. 
```{r, echo = FALSE, fig.height = 7, fig.width = 8}
par(mfrow = c(2,2))
par(mar = c(4, 4, 2, 1))
qqcompcens(fitsfn)
qqcompcens(fitsfl)
qqcompcens(fitsfg)
```

Will be soon implemented in the plotstyle ggplot.

## How to use of these new goodness-of-fit plots ?

Example of code:
```{r, eval = FALSE}
data(smokedfish)
d <-  log10(smokedfish)
# Plot of the NPMLE CDF on censored data
plotdistcens(d)
# Two MLE fits
fitsfn <- fitdistcens(d,"norm")
fitsfl <- fitdistcens(d,"logis")
# Three goodness-of-fit plots for one fit
plot(fitsfn)
# Goodness-of-fit plots for one or more fits
cdfcompcens(list(fitsfn,fitsfl))
qqcompcens(list(fitsfn,fitsfl))
ppcompcens(list(fitsfn,fitsfl))
```

We are waiting for your feedback on these new tools.

## Other main improvements of fitdistrplus

**Version 1.0-7**

* Start of a FAQ vignette continuously updated in each new version.

**Version 1.0-8**

* add of an optional use of ggplot2 in cdfcomp(), denscomp(), qqcomp() and ppcomp().

**Version 1.0-10**

* Improvement of goodness-of-fit plots for discrete distributions in denscomp().
* Add of new default starting values for distributions in actuar.

**Version 1.0-11**

* add of an optional use of ggplot2 in cdfcompcens(), denscompcens() and ppcompcens().

## References
* Turnbull BW (1974). Nonparametric estimation of a survivorship function with doubly censored data. Journal of American Statistical Association, 69, 169-173.
* Gentleman, R., & Geyer, C. J. (1994). Maximum likelihood for interval censored data: Consistency and computation. Biometrika, 81(3), 618-623.
* Wang, Y. (2008). Dimension-reduced nonparametric maximum likelihood computation for interval-censored data. Computational Statistics & Data Analysis, 52(5), 2388-2402.
* Wang, Y., & Taylor, S. M. (2013). Efficient computation of nonparametric survival functions via a hierarchical mixture formulation. Statistics and Computing, 23(6), 713-725.
* Wang, Y., & Fani, S. (2018). Nonparametric maximum likelihood computation of a U-shaped hazard function. Statistics and Computing, 28(1), 187-200.
